---
title: "Functions"
author: "Anish Bellamkonda"
tutorial:
  id: functions
output: 
  learnr::tutorial:
    progressive: true
    allow_skip: true
runtime: shiny_prerendered
description: "Tutorial for Chapter 26: Functions"
---

```{r setup, include = FALSE}
library(learnr)
library(tutorial.helpers)
knitr::opts_chunk$set(echo = FALSE)
options(tutorial.exercise.timelimit = 60, 
        tutorial.storage = "local") 
```

```{r copy-code-chunk, child = system.file("child_documents/copy_button.Rmd", package = "tutorial.helpers")}
```

```{r info-section, child = system.file("child_documents/info_section.Rmd", package = "tutorial.helpers")}
```

<!-- CHECKLIST BEFORE STARTING: -->
<!-- * Edit yaml at the top of this file -->
<!-- * Save the file as "tutorial.Rmd" in a new directory under inst/tutorials/. -->
<!-- * Load any necessary libraries for the tutorial in the first code chunk -->
<!-- * Review: https://ppbds.github.io/tutorial.helpers/articles/instructions.html -->
<!-- * Delete this and the other commented instructions below. -->

## Introduction
###


## Vector Functions
###

### Exercise 1


```{r vector-functions-1, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r vector-functions-1-hint, eval = FALSE}

```

###

### Exercise 2


```{r vector-functions-2, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r vector-functions-2-hint, eval = FALSE}

```

###

### Exercise 3


```{r vector-functions-3, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r vector-functions-3-hint, eval = FALSE}

```

###

### Exercise 4


```{r vector-functions-4, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r vector-functions-4-hint, eval = FALSE}

```

###

### Exercise 5


```{r vector-functions-5, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r vector-functions-5-hint, eval = FALSE}

```

###

### Exercise 6


```{r vector-functions-6, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r vector-functions-6-hint, eval = FALSE}

```

###




## Data Frame Functions
###

### Exercise 1


```{r data-frame-functions-1, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r data-frame-functions-1-hint, eval = FALSE}

```

###

### Exercise 2


```{r data-frame-functions-2, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r data-frame-functions-2-hint, eval = FALSE}

```

###

### Exercise 3


```{r data-frame-functions-3, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r data-frame-functions-3-hint, eval = FALSE}

```

###

### Exercise 4


```{r data-frame-functions-4, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r data-frame-functions-4-hint, eval = FALSE}

```

###

### Exercise 5


```{r data-frame-functions-5, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r data-frame-functions-5-hint, eval = FALSE}

```

###

### Exercise 6


```{r data-frame-functions-6, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r data-frame-functions-6-hint, eval = FALSE}

```

###

### Exercise 7


```{r data-frame-functions-7, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r data-frame-functions-7-hint, eval = FALSE}

```

###



## Plot Functions
###

### Exercise 1


```{r plot-functions-1, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r plot-functions-1-hint, eval = FALSE}

```

###

### Exercise 2


```{r plot-functions-2, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r plot-functions-2-hint, eval = FALSE}

```

###

### Exercise 3


```{r plot-functions-3, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r plot-functions-3-hint, eval = FALSE}

```

###

### Exercise 4


```{r plot-functions-4, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r plot-functions-4-hint, eval = FALSE}

```

###

### Exercise 5


```{r plot-functions-5, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r plot-functions-5-hint, eval = FALSE}

```

###

### Exercise 6


```{r plot-functions-6, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r plot-functions-6-hint, eval = FALSE}

```

###

### Exercise 7


```{r plot-functions-7, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r plot-functions-7-hint, eval = FALSE}

```

###




## Style
###

### Exercise 1


```{r style-1, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r style-1-hint, eval = FALSE}

```

###

### Exercise 2


```{r style-2, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r style-2-hint, eval = FALSE}

```

###

### Exercise 3


```{r style-3, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r style-3-hint, eval = FALSE}

```

###

### Exercise 4


```{r style-4, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r style-4-hint, eval = FALSE}

```

###

### Exercise 5


```{r style-5, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r style-5-hint, eval = FALSE}

```

###

### Exercise 6


```{r style-6, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r style-6-hint, eval = FALSE}

```

###

### Exercise 7


```{r style-7, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r style-7-hint, eval = FALSE}

```

###




## Summary
###

<!-- Two to four sentences which bring the lessons of the tutorial together for the student. What do they know now that they did not know before? How does this tutorial connect to other tutorials? OK if this is very similar to the Introduction. You made a promise as to what they would learn. You (we hope!) kept that promise.-->

```{r download-answers, child = system.file("child_documents/download_answers.Rmd", package = "tutorial.helpers")}
```
