---
title: "Functions"
author: "Anish Bellamkonda"
tutorial:
  id: functions
output: 
  learnr::tutorial:
    progressive: true
    allow_skip: true
runtime: shiny_prerendered
description: "Tutorial for Chapter 26: Functions"
---

```{r setup, include = FALSE}
library(learnr)
library(tutorial.helpers)
library(tidyverse)
library(nycflights13)
knitr::opts_chunk$set(echo = FALSE)
options(tutorial.exercise.timelimit = 60, 
        tutorial.storage = "local") 
```

```{r copy-code-chunk, child = system.file("child_documents/copy_button.Rmd", package = "tutorial.helpers")}
```

```{r info-section, child = system.file("child_documents/info_section.Rmd", package = "tutorial.helpers")}
```

<!-- CHECKLIST BEFORE STARTING: -->
<!-- * Edit yaml at the top of this file -->
<!-- * Save the file as "tutorial.Rmd" in a new directory under inst/tutorials/. -->
<!-- * Load any necessary libraries for the tutorial in the first code chunk -->
<!-- * Review: https://ppbds.github.io/tutorial.helpers/articles/instructions.html -->
<!-- * Delete this and the other commented instructions below. -->

## Introduction
###


## Vector Functions
###

### Exercise 1

We’ll begin with vector functions: functions that take one or more vectors and return a vector result. Let's first give you an example on why you need functions instead of copy and pasting commands.

Create a variable `df` and assign `tibble()` to it. We will have 4 columns: `a`, `b`, `c`,`d`. For all four columns assign `rnorm(5)`. 

```{r vector-functions-1, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r vector-functions-1-hint, eval = FALSE}

```

###

### Exercise 2


```{r vector-functions-2, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r vector-functions-2-hint, eval = FALSE}

```

###

### Exercise 3


```{r vector-functions-3, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r vector-functions-3-hint, eval = FALSE}

```

###

### Exercise 4


```{r vector-functions-4, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r vector-functions-4-hint, eval = FALSE}

```

###

### Exercise 5


```{r vector-functions-5, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r vector-functions-5-hint, eval = FALSE}

```

###

### Exercise 6


```{r vector-functions-6, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r vector-functions-6-hint, eval = FALSE}

```

###




## Data Frame Functions
###

### Exercise 1


```{r data-frame-functions-1, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r data-frame-functions-1-hint, eval = FALSE}

```

###

### Exercise 2


```{r data-frame-functions-2, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r data-frame-functions-2-hint, eval = FALSE}

```

###

### Exercise 3


```{r data-frame-functions-3, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r data-frame-functions-3-hint, eval = FALSE}

```

###

### Exercise 4


```{r data-frame-functions-4, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r data-frame-functions-4-hint, eval = FALSE}

```

###

### Exercise 5


```{r data-frame-functions-5, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r data-frame-functions-5-hint, eval = FALSE}

```

###

### Exercise 6


```{r data-frame-functions-6, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r data-frame-functions-6-hint, eval = FALSE}

```

###

### Exercise 7


```{r data-frame-functions-7, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r data-frame-functions-7-hint, eval = FALSE}

```

###



## Plot Functions
###

### Exercise 1


```{r plot-functions-1, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r plot-functions-1-hint, eval = FALSE}

```

###

### Exercise 2


```{r plot-functions-2, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r plot-functions-2-hint, eval = FALSE}

```

###

### Exercise 3


```{r plot-functions-3, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r plot-functions-3-hint, eval = FALSE}

```

###

### Exercise 4


```{r plot-functions-4, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r plot-functions-4-hint, eval = FALSE}

```

###

### Exercise 5


```{r plot-functions-5, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r plot-functions-5-hint, eval = FALSE}

```

###

### Exercise 6


```{r plot-functions-6, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r plot-functions-6-hint, eval = FALSE}

```

###

### Exercise 7


```{r plot-functions-7, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r plot-functions-7-hint, eval = FALSE}

```

###




## Style
###

R doesn’t care what your function or arguments are called but the names make a big difference for humans. Ideally, the name of your function will be short, but clearly evoke what the function does. That’s hard! But it’s better to be clear than short, as RStudio’s autocomplete makes it easy to type long names.

### Exercise 1

Generally, function names should be verbs, and arguments should be nouns. However, there are some exceptions. It is acceptable to use nouns if the function computes a well-known noun (e.g., "mean()" is preferred over "compute_mean()") or if it accesses a property of an object (e.g., "coef()" is preferred over "get_coefficients()"). Trust your judgment and feel free to rename a function if you come up with a better name later on.

In next few exercises we will show you different examples for the names of functions. Below is one:

```{r}
f()
```

This name is too short and doesn't mean anything but just the word `f`.

###

### Exercise 2

Below is another example:

```{r}
my_awesome_function()
```

This name is not a verb and also is not descriptive at the same time  like the function above.

###

### Exercise 3

Below is another example:

```{r style-3, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r style-3-hint, eval = FALSE}
impute_missing()
collapse_years()
```

###

Even though this function name is clear, it is still little to long for us to type again and again.

### Exercise 4

R also doesn’t care about how you use white space in your functions but future readers will. Additionally, `function()` should always be followed by squiggly brackets `({})`, and the contents should be indented by an additional two spaces. This makes it easier to see the hierarchy in your code by skimming the left-hand margin. 

Here is a good example of good indentation, which helps us see the hierarchy in the code.

```{r}
density <- function(color, facets, binwidth = 0.1) {
diamonds |> 
  ggplot(aes(x = carat, y = after_stat(density), color = {{ color }})) +
  geom_freqpoly(binwidth = binwidth) +
  facet_wrap(vars({{ facets }}))
}
```

###

Missing extra two spaces

### Exercise 5

Below is a example of having two extra spaces:

```{r}
density <- function(color, facets, binwidth = 0.1) {
  diamonds |> 
  ggplot(aes(x = carat, y = after_stat(density), color = {{ color }})) +
  geom_freqpoly(binwidth = binwidth) +
  facet_wrap(vars({{ facets }}))
}
```

###

In the code above the pipe is indented incorrectly. As you can see we recommend putting extra spaces inside of `{{ }}`. This makes it very obvious that something unusual is happening.

## Summary
###

<!-- Two to four sentences which bring the lessons of the tutorial together for the student. What do they know now that they did not know before? How does this tutorial connect to other tutorials? OK if this is very similar to the Introduction. You made a promise as to what they would learn. You (we hope!) kept that promise.-->

```{r download-answers, child = system.file("child_documents/download_answers.Rmd", package = "tutorial.helpers")}
```
