---
title: Logical vectors
author: David Kane, Zayan Farooq, and Rajarshi Mandal
tutorial:
  id: logical-vectors
output:
  learnr::tutorial:
    progressive: yes
    allow_skip: yes
runtime: shiny_prerendered
description: 'Tutorial for Chapter 13: Logical vectors'
---


```{r setup, include = FALSE}
library(learnr)
library(tutorial.helpers)
library(tidyverse)
library(nycflights13)

library(socviz)
library(ggthemes)

knitr::opts_chunk$set(echo = FALSE)
options(tutorial.exercise.timelimit = 60, 
        tutorial.storage = "local") 
```

```{r copy-code-chunk, child = system.file("child_documents/copy_button.Rmd", package = "tutorial.helpers")}
```

```{r info-section, child = system.file("child_documents/info_section.Rmd", package = "tutorial.helpers")}
```

## Introduction
### 

This tutorial covers [Chapter 13: Logical vectors](https://r4ds.hadley.nz/logical-vectors.html) from [*R for Data Science (2e)*](https://r4ds.hadley.nz/) by Hadley Wickham, Mine Çetinkaya-Rundel, and Garrett Grolemund. You will learn how to create logical vectors with `>`, `<`, `<=`, `=>`, `==`, `!=`, and `is.na()`, how to combine them with `!`, `&`, and `|`, and how to summarize them with `any()`, `all()`, `sum()`, and `mean()`. You will also learn about the powerful `if_else()` and `case_when()` functions that allow you to return values depending on the value of a logical vector.

## Comparisons
### 

Logical vectors are the simplest type of vector because each element can only be one of three possible values: `TRUE`, `FALSE`, and `NA`. It's relatively rare to find logical vectors in your raw data, but you'll create and manipulate them in almost every analysis.

### Exercise 1

Load the **tidyverse** package

```{r comparisons-1, exercise = TRUE}

```

```{r comparisons-1-hint, eval = FALSE}
library(...)
```

```{r, include=FALSE}
library(tidyverse)
```


### 

Most of the functions you'll learn about in this chapter are provided by base R, so we don't need the tidyverse, but we'll still load it so we can use `mutate()`, `filter()`, and others to work with data frames.

### Exercise 2

Load the **nycflights13** package.

```{r comparisons-2, exercise = TRUE}

```

```{r comparisons-2-hint, eval = FALSE}
library(...)
```

```{r, include=FALSE}
library(nycflights13)
```

### 

You can find more information about the package with `help(package = "nycflights13")`.

### Exercise 3

Type `flights` and hit "Run Code".

```{r comparisons-3, exercise = TRUE}

```

### 

A very common way to create a logical vector is via a numeric comparison with `<`, `<=`, `>`, `>=`, `!=`, and `==`.

### Exercise 4


Pipe `flights` into the `filter()` function. Within `filter()`, add `dep_time > 600` to only look at flights scheduled to depart after 6:00 AM.

```{r comparisons-4, exercise = TRUE}

```

<button onclick="transfer_code(this)">Copy previous code</button>

```{r comparisons-4-hint, eval = FALSE}
flights |> 
  filter(...)
```

```{r, include=FALSE}
flights |> 
  filter(dep_time > 600)
```


### 

In addition to `&` and `|`, R also has `&&` and `||`. Don't use them in **dplyr** functions! These are called short-circuiting operators and only ever return a single `TRUE` or `FALSE`. They're mainly used for programming, not data science.

### Exercise 5

Use the same code as above and add `& dep_time < 2000` to the call to `filter()`. This will look at flights scheduled to depart after 6:00 AM and before 8:00 PM.

```{r comparisons-5, exercise = TRUE}

```

<button onclick="transfer_code(this)">Copy previous code</button>

```{r comparisons-5-hint, eval = FALSE}
flights |> 
  filter(dep_time > 600 & ...)
```

```{r, include=FALSE}
flights |> 
  filter(dep_time > 600 & dep_time < 2000)
```

### 

An easy way to avoid the problem of getting your `==`'s and `|`'s in the right order is to use `%in%`. `x %in% y` returns a logical vector the same length as `x` that is `TRUE` whenever a value in `x` is anywhere in `y` .

### Exercise 6

Use the same code as above and add `& abs(arr_delay) < 20` to the call to `filter()`. This will filter the data out so that the absolute value of the arrival delay is less then 20 minutes.

```{r comparisons-6, exercise = TRUE}

```

<button onclick="transfer_code(this)">Copy previous code</button>

```{r comparisons-6-hint, eval = FALSE}
flights |> 
  filter(dep_time > 600 & dep_time < 2000 & abs(arr_delay) < ...)
```

```{r, include=FALSE}
flights |> 
  filter(dep_time > 600 & dep_time < 2000 & abs(arr_delay) < 20)
```

### 

You can always explicitly create the underlying logical variables with `mutate()`

### Exercise 7


Start a new pipe with `flights`. Pipe `flights` into the `mutate()` function. Within `mutate`, use `daytime = dep_time > 600 & dep_time < 2000`. This will create a new column called `daytime` which is `TRUE` when the `dep_time` is between 6:00 AM and 8:00 PM.

```{r comparisons-7, exercise = TRUE}

```

```{r comparisons-7-hint, eval = FALSE}
flights |> 
  mutate(daytime = dep_time > ... & dep_time < ...)
```

```{r, include=FALSE}
flights |> 
  mutate(daytime = dep_time > 600 & dep_time < 2000)
```

###

Just remember that any manipulation we do to a free-floating vector, you can do to a variable inside a data frame with `mutate()` and friends.

### Exercise 8

Use the same code as above and add `, approx_ontime = abs(arr_delay) < 20` to the `mutate()` function. This will create a new column named `approx_ontime` which is true when the `arr_delay` is less than 20 minutes.

```{r comparisons-8, exercise = TRUE}

```

<button onclick="transfer_code(this)">Copy previous code</button>

```{r comparisons-8-hint, eval = FALSE}
flights |> 
  mutate(daytime = dep_time > 600 & dep_time < 2000,
         approx_ontime = abs(arr_delay) < ...)
```

```{r, include=FALSE}
flights |> 
  mutate(daytime = dep_time > 600 & dep_time < 2000,
         approx_ontime = abs(arr_delay) < 20)
```

### 

### Exercise 9

Copy previous code and add `,.keep = "used"` to the `mutate()` function. This keeps only the new columns and the columns used to create them in the tibble and drops all the unused columns.

```{r comparisons-9, exercise = TRUE}

```

<button onclick="transfer_code(this)">Copy previous code</button>

```{r comparisons-9-hint, eval = FALSE}
flights |> 
  mutate(daytime = dep_time > 600 & dep_time < 2000,
         approx_ontime = abs(arr_delay) < 20,
         .keep = ...)
```

```{r, include=FALSE}
flights |> 
  mutate(daytime = dep_time > 600 & dep_time < 2000,
         approx_ontime = abs(arr_delay) < 20,
         .keep = "used")
```

### 

### Exercise 10

Copy previous code and pipe it into `filter()`. Within `filter()` add `daytime & approx_ontime`. This will keep only the rows that meet the specified conditions.

```{r comparisons-10, exercise = TRUE}

```

<button onclick="transfer_code(this)">Copy previous code</button>

```{r comparisons-10-hint, eval = FALSE}
flights |> 
  mutate(daytime = dep_time > 600 & dep_time < 2000,
         approx_ontime = abs(arr_delay) < 20,
         .keep = "used")|>
  filter(daytime & ...)
```

```{r, include=FALSE}
flights |> 
  mutate(daytime = dep_time > 600 & dep_time < 2000,
         approx_ontime = abs(arr_delay) < 20,
         .keep = "used") |>
  filter(daytime & approx_ontime)
```

### 

So far, we’ve mostly created logical variables transiently within `filter()` — they are computed, used, and then thrown away. For example, the previous `filter()` finds all daytime departures that arrive roughly on time.

### Exercise 11

Set `x` to `c(1 / 49 * 49, sqrt(2) ^ 2)` using `<-`. Then, click "Run Code".

```{r comparisons-11, exercise = TRUE}

```

```{r comparisons-11-hint, eval = FALSE}
x <- c(..., ...)
```

```{r, include=FALSE}
x <- c(1 / 49 * 49, sqrt(2) ^ 2)
```

###

Beware of using `==` with numbers. For example, it looks like this vector contains the numbers `1` and `2`. But if you test them for equality, you get `FALSE`.

### Exercise 12

Copy previous code and run `x == c(1,2)`. The expected output is `FALSE`.

```{r comparisons-12, exercise = TRUE}

```

<button onclick="transfer_code(this)">Copy previous code</button>

```{r comparisons-12-hint, eval = FALSE}
x <- c(1 / 49 * 49, sqrt(2) ^ 2)
x == c(..., ...)
```

```{r, include=FALSE}
x <- c(1 / 49 * 49, sqrt(2) ^ 2)
x == c(1, 2)
```

### 

What's going on? Computers store numbers with a fixed number of decimal places so there's no way to exactly represent `1/49` or `sqrt(2)` and subsequent computations will be very slightly off.

### Exercise 13

We can see the exact values by calling `print()` with the digits argument. Copy previous code and run `print(x, digits = 16)`. This will run the values of `x` with 16 digits.

```{r comparisons-13, exercise = TRUE}

```

<button onclick="transfer_code(this)">Copy previous code</button>

```{r comparisons-13-hint, eval = FALSE}
x <- c(1 / 49 * 49, sqrt(2) ^ 2)
x == c(1, 2)
print(x, digits = ...)
```

```{r, include=FALSE}
x <- c(1 / 49 * 49, sqrt(2) ^ 2)
x == c(1, 2)
print(x, digits = 16)
```

### 

You can see why R defaults to rounding these numbers; they really are very close to what you expect. Now that you've seen why `==` is failing, what can you do about it? One option is to use `dplyr::near()` which ignores small differences:

### Exercise 14

Copy the previous code and type `near()` on the next line. Within `near()`, add `x, c(1,2)`. This should come out as `TRUE` for both.

```{r comparisons-14, exercise = TRUE}

```

<button onclick="transfer_code(this)">Copy previous code</button>

```{r comparisons-14-hint, eval = FALSE}
x <- c(1 / 49 * 49, sqrt(2) ^ 2)
x == c(1, 2)
print(x, digits = 16)
near(x, c(..., ...))
```

```{r, include=FALSE}
x <- c(1 / 49 * 49, sqrt(2) ^ 2)
x == c(1, 2)
print(x, digits = 16)
near(x, c(1, 2))
```

### 

Missing values represent the unknown so they are “contagious”: almost any operation involving an unknown value will also be unknown.

### Exercise 15

Run `NA > 5` and `10 == NA`. They should both come out as `NA`

```{r comparisons-15, exercise = TRUE}

```

<button onclick="transfer_code(this)">Copy previous code</button>

```{r comparisons-15-hint, eval = FALSE}
NA > ...
... == NA
```

```{r, include=FALSE}
NA > 5
10 == NA
```

### 

R normally calls print for you (i.e. `x` is a shortcut for `print(x)`), but calling it explicitly is useful if you want to provide other arguments.

### Exercise 16

Now, run `NA == NA`.This should also come out as `NA`

```{r comparisons-16, exercise = TRUE}

```

<button onclick="transfer_code(this)">Copy previous code</button>

```{r, include=FALSE}
NA == NA
```

### 

That is the most confusing result. It's easiest to understand why this is true if we artificially supply a little more context.

### Exercise 17

Set `age_mary <- NA` and `age_john <- NA` on the next line. On a new line, type `age_mary == age_john` and click run code.

```{r comparisons-17, exercise = TRUE}

```

<button onclick="transfer_code(this)">Copy previous code</button>

```{r comparisons-17-hint, eval = FALSE}
age_mary <- ...
... <- NA
age_mary == age_john
```

```{r, include=FALSE}
age_mary <- NA
age_john <- NA
age_mary == age_john
```

### 

This should come out as `NA` because if both of their ages are unknown, then we can't know if they are the same age.

### Exercise 18

Start a new pipe with `flights`. Pipe that into the `filter()` function. Within `filter()`, add `dep_time == NA`. This will attempt to find all the rows where `dep_time` is missing.

```{r comparisons-18, exercise = TRUE}

```

<button onclick="transfer_code(this)">Copy previous code</button>

```{r comparisons-18-hint, eval = FALSE}
flights|>
  filter(dep_time == ...)
```

```{r, include=FALSE}
flights|>
  filter(dep_time == NA)
```

### 

The following code doesn’t work because `dep_time == NA` will yield `NA` for every single row, and `filter()` automatically drops missing values. Instead we'll need a new tool: `is.na()`.

### Exercise 19

Type `is.na()` into the code chunk. Within `is.na()`, add `c(TRUE, NA, FALSE)`.

```{r comparisons-19, exercise = TRUE}

```

```{r comparisons-19-hint, eval = FALSE}
is.na(c(TRUE, NA, ...))
```

```{r, include=FALSE}
is.na(c(TRUE, NA, FALSE))
```

### 

You will see the output `FALSE` for `TRUE` and `FALSE`, but `TRUE` for `NA`. What happens when there are characters or numbers in the input vector?

### Exercise 20

Type `is.na()` into the code chunk. Within `is.na()`, add `c(1, NA, 'b')`.

```{r comparisons-20, exercise = TRUE}

```

```{r comparisons-20-hint, eval = FALSE}
is.na(c(1, NA, ...))
```

```{r, include=FALSE}
is.na(c(1, NA, 'b'))
```

### 

You will see the output `True` for `NA` and `FALSE` for `1` and `b`. This is because `is.na()` can work with any type of vector and returns `TRUE` for missing values and `FALSE` for everything else..

### Exercise 21

Start a new pipe with `flights`. Pipe `flights` into `filter()`. Within `filter()`, add `is.na(dep_time)`. This will find all the rows with a missing `dep_time`.

```{r comparisons-21, exercise = TRUE}

```

```{r comparisons-21-hint, eval = FALSE}
flights |> 
  filter(is.na(...))
```

```{r, include=FALSE}
flights |> 
  filter(is.na(dep_time))
```

### 

`is.na()` can also be useful in `arrange()`. `arrange()` usually puts all the missing values at the end but you can override this default by first sorting with `is.na()`.

### Exercise 22

Start a new pipe with `flights`. Pipe it into the `filter()` function. Within `filter()`, add `month == 1, day == 1`.

```{r comparisons-22, exercise = TRUE}

```

```{r comparisons-22-hint, eval = FALSE}
flights|>
  filter(month == ..., day == ...)
```

```{r, include=FALSE}
flights |> 
  filter(month == 1, day == 1)
```

### 

The `==` operator is a comparison operator in R that checks if two values are equal. It returns `TRUE` if the values are equal and `FALSE` otherwise.

### Exercise 23

Continue the pipe with `arrange()`. Within `arrange()`, add `dep_time`. This should arrange `dep_time` from least to greatest.

```{r comparisons-23, exercise = TRUE}

```

<button onclick="transfer_code(this)">Copy previous code</button>

```{r comparisons-23-hint, eval = FALSE}
flights|>
  filter(month == 1, day == 1) |>
  arrange(...)
```

```{r, include=FALSE}
flights |> 
  filter(month == 1, day == 1) |>
  arrange(dep_time)
```

### 

The `arrange()` function is used to order rows in a data set based on a column. It allows you to sort the data set in either ascending or descending order.

### Exercise 24

Within `arrange()`, put `is.na(dep_time)`.

```{r comparisons-24, exercise = TRUE}

```

<button onclick="transfer_code(this)">Copy previous code</button>

```{r comparisons-24-hint, eval = FALSE}
flights|>
  filter(month == 1, day == 1)|>
  arrange(is.na(...))
```

```{r, include=FALSE}
flights |> 
  filter(month == 1, day == 1) |>
  arrange(is.na(dep_time))
```

### 

This checks if the `dep_time` column has missing values (`NA`). It returns a logical vector with `TRUE` for rows where `dep_time` is `NA` and `FALSE` otherwise.

### Exercise 25

Within `arrange()`, put `desc(is.na(dep_time)), dep_time`

```{r comparisons-25, exercise = TRUE}

```

<button onclick="transfer_code(this)">Copy previous code</button>

```{r comparisons-25-hint, eval = FALSE}
flights|>
  filter(month == 1, day == 1)|>
  arrange(desc(is.na(...)), ...)
```

```{r, include=FALSE}
flights |> 
  filter(month == 1, day == 1) |>
  arrange(desc(is.na(dep_time)), dep_time)
```

### 

The `desc()` function is used to create a descending order of the logical vector obtained from `is.na(dep_time)`. This means that rows with missing `dep_time` values (`NA`) will appear first in the data frame. We will discuss missing values further in the Missing Values tutorial.



## Boolean algebra
### 

Once you have multiple logical vectors, you can combine them together using Boolean algebra. In R, `&` is “and”, `|` is “or”, `!` is “not”, and `xor()` is exclusive or. For example, `df |> filter(!is.na(x))` finds all rows where `x` is not missing and `df |> filter(x < -10 | x > 0)` finds all rows where `x` is smaller than `-10` or bigger than `0`. Figure 13.1 shows the complete set of Boolean operations and how they work.

### Exercise 1

The rules for missing values in Boolean algebra are a little tricky to explain because they seem inconsistent at first glance. Click "Run Code".

```{r boolean-algebra-1, exercise = TRUE}
df <- tibble(x = c(TRUE, FALSE, NA))

df |> 
  mutate(
    and = x & NA,
    or = x | NA
  )
```

###

To understand what’s going on, think about `NA | TRUE` (`NA` or `TRUE`). A missing value in a logical vector means that the value could either be `TRUE` or `FALSE`. `TRUE | TRUE` and `FALSE | TRUE` are both `TRUE` because at least one of them is `TRUE`. `NA | TRUE` must also be `TRUE` because `NA` can either be `TRUE` or `FALSE`. However, `NA | FALSE` is `NA` because we don’t know if `NA` is `TRUE` or `FALSE`. Similar reasoning applies with `NA & FALSE`.

### Exercise 2

Start a new pipe with `flights` and pipe it into `filter()`. Within `filter()`, add `month == 11 | month == 12`.

```{r boolean-algebra-2, exercise = TRUE}

```

```{r boolean-algebra-2-hint, eval = FALSE}
flights |>
  filter(month == ... | month == ...)
```

```{r, include=FALSE}
flights |>
  filter(month == 11 | month == 12)
```

###

Note that the order of operations doesn’t work like English. To prove this, we can take the previous code that finds all flights that departed in November or December. You might be tempted to write it like you’d say in English: “Find all flights that departed in November or December.”

### Exercise 3

Now, try replacing `month == 11 | month == 12` with `month == 11 | 12`. This should do the same thing as before. (Spoiler Alert: It doesn't!)

```{r boolean-algebra-3, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r boolean-algebra-3-hint, eval = FALSE}
flights |>
  filter(month == ... | ...)
```

```{r, include=FALSE}
flights |>
  filter(month == 11 | 12)
```

###

The code does not return an error, but it doesn’t seem to have worked either. What happened? Here, R first evaluates `month == 11` creating a logical vector, which we call `nov`. It now computes `nov | 12`. When you use a number with a logical operator it converts everything apart from `0` to `TRUE`, so this is equivalent to `nov | TRUE` which will always be `TRUE`, so every row will be selected.

### Exercise 4

Start a new pipe with `flights`. Pipe `flights` into the `mutate()` function. Within `mutate()`, set `nov = month == 11`.

```{r boolean-algebra-4, exercise = TRUE}

```

```{r boolean-algebra-4-hint, eval = FALSE}
flights|>
  mutate(nov = month == ...)
```

```{r, include=FALSE}
flights|>
  mutate(nov = month == 11)
```

###

`nov = month == 11` creates a new variable called `nov` and assigns it to the Boolean `TRUE` for rows where the month is equal to `11`, and `FALSE` otherwise. The expression `month == 11` checks the equality between the month column and the value `11`, resulting in a logical vector.

### Exercise 5

Copy previous code and add `final = nov | 12, .keep = "used"` to the `mutate()` function. `final = nov | 12` results in the variable `final` being `TRUE` if `nov` is `TRUE` or if the value `12` is considered as `TRUE`. The `.keep = "used"` argument keeps only the new variables, and the variables that were used to make them in the tibble.

```{r boolean-algebra-5, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r boolean-algebra-5-hint, eval = FALSE}
flights|>
 mutate(nov = month == 11,
        final = nov | ...,
        .keep = ...)
```

```{r, include=FALSE}
flights|>
 mutate(nov = month == 11,
        final = nov | 12,
        .keep = "used")
```

###

As well as & and |, R also has && and ||. Don’t use them in dplyr functions! These are called short-circuiting operators and only ever return a single TRUE or FALSE. They’re important for programming, not data science.

### Exercise 6

An easy way to avoid the problem of getting your `==`s and `|`s in the right order is to use `%in%`. `x %in% y` returns a logical vector the same length as `x` that is `TRUE` whenever a value in `x` is anywhere in `y`. `x %in% y` returns `TRUE` if anything in `x` is in `y`. For example, type `1:12 %in% c(1,5,11)` and click "Run Code". 

```{r boolean-algebra-6, exercise = TRUE}

```

```{r boolean-algebra-6-hint, eval = FALSE}
1:12 %in% c(..., ..., ...)
```

```{r, include=FALSE}
1:12 %in% c(1, 5, 11)
```

###

This code assigns a vector containing all the numbers from `1` through `11` to `x`, and `1`,`5`, and `11` to `y`.

### Exercise 7

To find all of the flights in November and December, start a new pipe with `flights`. Pipe it into `filter()`. Within `filter()`, add `month %in% c(11,12)`.

```{r boolean-algebra-7, exercise = TRUE}

```

```{r boolean-algebra-7-hint, eval = FALSE}
flights |>
  filter(month %in% c(..., ...))
```

```{r, include=FALSE}
flights |>
  filter(month %in% c(11, 12))
```

###

Fun Fact: `%in%` obeys different rules for `NA` than `==`, as `NA %in% NA` is `TRUE` while `NA == NA` is `NA`.

### Exercise 8

Type `c(1, 2, NA) == NA` and click "Run code".

```{r boolean-algebra-8, exercise = TRUE}

```

```{r boolean-algebra-8-hint, eval = FALSE}
c(1, 2, NA) == ...
```

```{r, include=FALSE}
c(1, 2, NA) == NA
```

###

This should output `#> [1] NA NA NA` because when comparing any value (including `NA`) with `NA` using the `==` operator, the result will be `NA`. This is because the value of `NA` represents unknown information, so the result of most comparisons involving `NA` are also unknown.

### Exercise 9

Copy previous code but instead of using `==`, use `%in%`. This checks to see if any value on the left side (`x`), is on the rights side (`y`).

```{r boolean-algebra-9, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r boolean-algebra-9-hint, eval = FALSE}
c(1, 2, NA) %in% ...
```

```{r, include=FALSE}
c(1, 2, NA) %in% NA
```

###

This should output `#> [1] FALSE FALSE  TRUE` because `NA` is the only variable on both sides.

### Exercise 10

Start a new pipe with `flights`. Pipe it into `filter()`. Within `filter()`, add `dep_time %in% c(NA, 0800)` to find all rows where `dep_time` is `NA`, of `0800`.

```{r boolean-algebra-10, exercise = TRUE}

```

```{r boolean-algebra-10-hint, eval = FALSE}
flights |> 
  filter(dep_time %in% c(..., ...))
```

```{r, include=FALSE}
flights |> 
  filter(dep_time %in% c(NA, 0800))
```

###

This should result in a tibble where the only values in the `dep_time` column are `NA` and `0800`.

## Summaries
###

Next, we will describe some useful techniques for summarizing logical vectors. In addition to functions that only work with logical vectors, you can also use functions that work with numeric vectors.

### Exercise 1

Start a pipe with `flights`. Pipe it into `summarize()`. Within `summarize()`, add `all_delayed = all(dep_delay <= 60)`. This will make a new column named `all_delayed` that is `TRUE` when `dep_delay` is less than or equal to `60` minutes.

```{r summaries-1, exercise = TRUE}

```

```{r summaries-1-hint, eval = FALSE}
flights |>
  summarize(all_delayed = all(dep_delay <= ...))
```

```{r, include=FALSE}
flights |>
  summarize(all_delayed = all(dep_delay <= 60))
```

###

There are two main logical summary functions: `any()` and `all()`. `any(x)` is the equivalent of `|`; it’ll return `TRUE` if there are any `TRUE`’s in `x`. `all(x)` is equivalent of `&`; it’ll return `TRUE` only if all values of `x` are `TRUE`’s.

### Exercise 2

Copy the previous code. Within `all()`, add `na.rm = TRUE` separated with a `,`. When `na.rm` is set to `TRUE`, it removes all `NA` values. `na.rm` is short for `na.remove`.

```{r summaries-2, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r summaries-2-hint, eval = FALSE}
flights |> 
  summarize(all_delayed = all(dep_delay <= 60, na.rm = ...))
```

```{r, include=FALSE}
flights |> 
  summarize(all_delayed = all(dep_delay <= 60, na.rm = TRUE))
```

###

Next, we will use `all()` and `any()` to find out if every flight was delayed on departure by at most an hour or if any flights were delayed on arrival by five hours or more. Using the `group_by()` will allow us to do that by day.

### Exercise 3

Copy the previous code. Within `summarize()`, add `any_long_delay = any(arr_delay >= 300)`. This will make a new column named `any_long_delay` that is `TRUE` when `arr_delay` is greater than or equal to `300` minutes.

```{r summaries-3, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r summaries-3-hint, eval = FALSE}
flights |> 
  summarize(
    all_delayed = all(dep_delay <= 60, na.rm = TRUE),
    any_long_delay = any(arr_delay >= ...))
```

```{r, include=FALSE}
flights |> 
  summarize(
    all_delayed = all(dep_delay <= 60, na.rm = TRUE),
    any_long_delay = any(arr_delay >= 300))
```

###

Like all summary functions, `any()` and `all()` will return `NA` if there are any missing values present. As usual, you can make the missing values go away with `na.rm = TRUE`.

### Exercise 4

Copy the previous code. Add the `na.rm = TRUE` argument in `any()`.

```{r summaries-4, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r summaries-4-hint, eval = FALSE}
flights |> 
  summarize(
    all_delayed = all(dep_delay <= 60, na.rm = TRUE),
    any_long_delay = any(arr_delay >= 300), ...)
```

```{r, include=FALSE}
flights |> 
  summarize(
    all_delayed = all(dep_delay <= 60, na.rm = TRUE),
    any_long_delay = any(arr_delay >= 300), na.rm = TRUE)
```

###

`>=` is used when you want to find a variable that that is greater than or equal to a number. Conversely, you could use `<=` to find a variable that that is less than or equal to a number. Finally, `==` can be used to find a variable that is exactly equal to a number.

### Exercise 5

Copy the previous code. Add `.by = c(year, month)` to the `summarize()` function. Make sure to separate the arguments with commas. The added code will make the tibble have one row for each variation of `month` and `year`.

```{r summaries-5, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r summaries-5-hint, eval = FALSE}
flights |> 
  summarize(
    all_delayed = all(dep_delay <= 60, na.rm = TRUE),
    any_long_delay = any(arr_delay >= 300), na.rm = TRUE,
    .by = c(..., ...))
```

```{r, include=FALSE}
flights |> 
  summarize(
    all_delayed = all(dep_delay <= 60, na.rm = TRUE),
    any_long_delay = any(arr_delay >= 300), na.rm = TRUE,
    .by = c(year, month))
```

###

However, in most cases, `any()` and `all()` are a little too crude, and it would be helpful to get a little more detail about how many values are `TRUE` and `FALSE`. This is what numeric summaries are for.

### Exercise 6

Add `day` as a new argument in `.by()`.

```{r summaries-6, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r summaries-6-hint, eval = FALSE}
flights |>
  summarize(
    all_delayed = all(dep_delay <= 60, na.rm = TRUE),
    any_long_delay = any(arr_delay >= 300, na.rm = TRUE),
    .by = c(year, month, ...))
```

```{r, include=FALSE}
flights |> 
  summarize(
    all_delayed = all(dep_delay <= 60, na.rm = TRUE),
    any_long_delay = any(arr_delay >= 300), na.rm = TRUE,
    .by = c(year, month, day))
```

###

This allows us to find out if every flight was delayed on departure by at most an hour and the number of flights that were delayed on arrival by five hours or more. This will make the resulting tibble have one row for each combination of `year`, `month`, and `day`. Essentially, we will have one row for each day of the year.

### Exercise 7

Start a new pipe with `flights`. Pipe it into `summarize()`. Within `summarize()`, add `all_delayed = dep_delay <= 60`.

```{r summaries-7, exercise = TRUE}

```

```{r summaries-7-hint, eval = FALSE}
flights |> 
  summarize(
    all_delayed = dep_delay <= ...)
```

```{r, include=FALSE}
flights |> 
  summarize(
    all_delayed = dep_delay <= 60)
```

###

This should create a column called `all_delayed` that is `TRUE` when `dep_delay` is less than or equal to `60`.

### Exercise 8

Copy the previous code. Add `dep_delay <= 60` within `mean()`.

```{r summaries-8, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r summaries-8-hint, eval = FALSE}
flights |> 
  summarize(
    all_delayed = ...(dep_delay <= 60))
```

```{r, include=FALSE}
flights |> 
  summarize(
    all_delayed = mean(dep_delay <= 60))
```

###

This should create a column called `all_delayed` that is the mean of the rows where `dep_delay` is less than `60`

### Exercise 9

Copy the previous code. Add `na.rm = TRUE` within `mean()`.

```{r summaries-9, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r summaries-9-hint, eval = FALSE}
flights|>
  summarize(all_delayed = mean(dep_delay <= 60,...))
```

```{r, include=FALSE}
flights |> 
  summarize(
    all_delayed = mean(dep_delay <= 60, na.rm = TRUE))
```

###

Setting `na.rm` equal to `TRUE` removes all `NA` values.

### Exercise 10

Copy the previous code and add `any_long_delay = arr_delay >= 300` as a new argument in `summarize()`. Again, make sure to separate the arguments with commas.

```{r summaries-10, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r summaries-10-hint, eval = FALSE}
flights|>
  summarize(all_delayed = mean(dep_delay <= 60, na.rm = TRUE),
            any_long_delay = ...)
```

```{r, include=FALSE}
flights |> 
  summarize(
    all_delayed = mean(dep_delay <= 60, na.rm = TRUE),
    any_long_delay = arr_delay >= 300)
```

###

### Exercise 11

Copy the previous code and add `arr_delay >= 300` within `sum()`.

```{r summaries-11, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r summaries-11-hint, eval = FALSE}
flights |> 
  summarize(
    all_delayed = mean(dep_delay <= 60, na.rm = TRUE),
    any_long_delay = ...(arr_delay >= 300))
```

```{r, include=FALSE}
flights |> 
  summarize(
    all_delayed = mean(dep_delay <= 60, na.rm = TRUE),
    any_long_delay = sum(arr_delay >= 300))
```

###

The `sum()` function gives the number of `TRUE`'s of whatever is inside it.

### Exercise 12

Copy the previous code and add `na.rm = TRUE` within `sum()`.

```{r summaries-12, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r summaries-12-hint, eval = FALSE}
flights|>
  summarize(all_delayed = mean(dep_delay <= 60, na.rm = TRUE),
            any_long_delay = sum(arr_delay >= 300, ...))
```

```{r, include=FALSE}
flights |> 
  summarize(
    all_delayed = mean(dep_delay <= 60, na.rm = TRUE),
    any_long_delay = sum(arr_delay >= 300, na.rm = TRUE))
```

###

When you use a logical vector in a numeric context, `TRUE` becomes `1` and `FALSE` becomes `0`. This makes `sum()` and `mean()` very useful with logical vectors because `sum(x)` gives the number of `TRUE`s and `mean(x)` gives the proportion of `TRUE`s (`mean()` is `sum()` divided by `length()`).

### Exercise 13

Copy the previous code and add `.by = c(year, month)` as a new argument within `summarize()`.

```{r summaries-13, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r summaries-13-hint, eval = FALSE}
flights|>
  summarize(all_delayed = mean(dep_delay <= 60, na.rm = TRUE),
            any_long_delay = sum(arr_delay >= 300, na.rm = TRUE),
            .by = c(..., ...))
```

```{r, include=FALSE}
flights |> 
  summarize(
    all_delayed = mean(dep_delay <= 60, na.rm = TRUE),
    any_long_delay = sum(arr_delay >= 300, na.rm = TRUE),
    .by = c(year, month))
```

###

Always put lists of values within `c()`; otherwise, the code will result in an error.

### Exercise 14

Copy previous code and add `day` to the `.by()` argument within `summarize()` to get rows for every single day of the year. Make sure to separate the arguments with commas!

```{r summaries-14, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r summaries-14-hint, eval = FALSE}
flights |>  
  summarize(
    all_delayed = mean(dep_delay <= 60, na.rm = TRUE),
    any_long_delay = sum(arr_delay >= 300, na.rm = TRUE),
    .by = c(year, month, ...))
```

```{r, include=FALSE}
flights |> 
  summarize(
    all_delayed = mean(dep_delay <= 60, na.rm = TRUE),
    any_long_delay = sum(arr_delay >= 300, na.rm = TRUE),
    .by = c(year, month, day))
```

###

This allows for us to see the proportion of flights that were delayed on departure by at most 1 hour (`60` minutes) and the number of flights that were delayed on arrival by 5 hours (`300` minutes) or more.

### Exercise 15

Start a new pipe with `flights`. Pipe it into `filter`. Within `filter()`, add `arr_delay > 0` as the argument.

```{r summaries-15, exercise = TRUE}

```

```{r summaries-15-hint, eval = FALSE}
flights |>
  filter(arr_delay > ...)
```

```{r, include=FALSE}
flights |>
  filter(arr_delay > 0)
```

###

There’s one final use for logical vectors in summaries: you can use a logical vector to filter a single variable to a subset of interest. This makes use of the base `[` (pronounced subset) operator.

### Exercise 16

Continue the pipe with `summarize()`. Within `summarize()`, add the argument `behind = mean(arr_delay)`.

```{r summaries-16, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r summaries-16-hint, eval = FALSE}
flights |>
  filter(arr_delay > 0) |>
  summarize(behind = mean(...))
```

```{r, include=FALSE}
flights |>
  filter(arr_delay > 0) |>
  summarize(behind = mean(arr_delay))
```

###

This will look at the average delay only for flights that were actually delayed. We did this by first filtering the flights, then calculating the average delay.

### Exercise 17

Copy the previous code and add the argument `n = n()` within `summarize()`. Make sure to separate arguments within a function with comma's.

```{r summaries-17, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r summaries-17-hint, eval = FALSE}
flights |>
  filter(arr_delay > 0)|>
  summarize(behind = mean(arr_delay),
            n = ...())
```

```{r, include=FALSE}
flights |>
  filter(arr_delay > 0) |>
  summarize(behind = mean(arr_delay),
            n = n())
```

###

`n = n()` is used to calculate the count of observations where `arr_delay` is greater than `0`.

### Exercise 18

Add `.by = c(year, month, day)` to have one row for each day of the year.

```{r summaries-18, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r summaries-18-hint, eval = FALSE}
flights |>
  filter(arr_delay > 0) |>
  summarize(behind = mean(arr_delay),
            n = n(),
            .by = c(..., ..., ...))
```

```{r, include=FALSE}
flights |>
  filter(arr_delay > 0) |>
  summarize(behind = mean(arr_delay),
            n = n(),
            .by = c(year, month, day))
```

###

This looks at the average delay just for flights that were actually delayed. We did this by first filtering the flights and then calculating the average delay:

This works, but what if we wanted to also compute the average delay for flights that arrived early? We’d need to perform a separate filter step, and then figure out how to combine the two data frames together. Alternatively, you could use `[` to perform an inline filtering: `arr_delay[arr_delay > 0]` will yield only the positive arrival delays.

### Exercise 19

Press "Run Code". 

```{r summaries-19, exercise = TRUE}
flights|> 
  summarize(
    behind = mean(arr_delay[arr_delay > 0], na.rm = TRUE),
    ahead = mean(arr_delay[arr_delay < 0], na.rm = TRUE),
    n = n(),
    .by = c(year, month, day)
  )
```

###

Also note the difference in the group size: in the first chunk `n()` gives the number of delayed flights per day; in the second, `n()` gives the total number of flights.

## Conditional transformations
### 

### Exercise 1

Let’s begin with a simple example of labeling a numeric vector as either `“+ve”` (positive) or `“-ve”` (negative). Assign `x` to the vector `c(-3:3, NA)` and press "Run Code".

```{r conditional-transfor-1, exercise = TRUE}

```


```{r conditional-transfor-1-hint, eval = FALSE}
x <- c(-3:3, ...)
```

```{r, include=FALSE}
x <- c(-3:3, NA)
```
###

### Exercise 2

Copy the previous code. Then, type `if_else(x > 0, "+ve", "-ve")`. This should output `#> [1] "-ve" "-ve" "-ve" "-ve" "+ve" "+ve" "+ve" NA`.

```{r conditional-transfor-2, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r conditional-transfor-2-hint, eval = FALSE}
x <- c(-3:3, NA)
if_else(x > ..., "+ve", "-ve")
```

```{r, include=FALSE}
x <- c(-3:3, NA)
if_else(x > 0, "+ve", "-ve")
```

###

### Exercise 3

There’s an optional fourth argument, `missing`, which will be used if the input is `NA`. In this example, we can add the string `"???"` as an argument to `if_else()`.

```{r conditional-transfor-3, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r conditional-transfor-3-hint, eval = FALSE}
x <- c(-3:3, NA)
if_else(x > 0, "+ve", "-ve", ...)
```

```{r, include=FALSE}
x <- c(-3:3, NA)
if_else(x > 0, "+ve", "-ve", "???")
```

###

The output should be `#> [1] "-ve" "-ve" "-ve" "-ve" "+ve" "+ve" "+ve" "???"`

### Exercise 4

You can also use vectors for the the `TRUE` and `FALSE` arguments. Let us attempt to create a minimal implementation of `abs()`. Below the code assigning a vector to `x`, type `if_else(x < 0, -x, x)`. Click "Run Code". 

```{r conditional-transfor-4, exercise = TRUE}
x <- c(-3:3, NA)
```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r conditional-transfor-4-hint, eval = FALSE}
x <- c(-3:3, NA)
if_else(x < ..., -x, x)
```

```{r, include=FALSE}
x <- c(-3:3, NA)
if_else(x < 0, -x, x)
```

###

This should output `#> [1]  3  2  1  0  1  2  3 NA`. So far, all the arguments have used the same vectors, but you can also mix and match!

### Exercise 5

For example, you could implement a simple version of `coalesce()`. Below the code assigning vectors to `x1` and `y1`, type `if_else(is.na(x1), y1, x1)` and click "Run Code".

```{r conditional-transfor-5, exercise = TRUE}
x1 <- c(NA, 1, 2, NA)
y1 <- c(3, NA, 4, 6)
```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r conditional-transfor-5-hint, eval = FALSE}
x1 <- c(NA, 1, 2, NA)
y1 <- c(3, NA, 4, 6)
if_else(is.na(...), y1, x1)
```

```{r, include=FALSE}
x1 <- c(NA, 1, 2, NA)
y1 <- c(3, NA, 4, 6)
if_else(is.na(x1), y1, x1)
```

###

The output should be `#> [1] 3 1 2 6`. You might have noticed a small infelicity in our labeling example above: zero is neither positive nor negative. We could resolve this by adding an additional `if_else()`.

### Exercise 6

Below the provided code, type `if_else(x == 0, "0", if_else(x < 0, "-ve", "+ve"), "???")` and click "Run Code".

```{r conditional-transfor-6, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r conditional-transfor-6-hint, eval = FALSE}
x <- c(-3:3, NA)
if_else(x == 0, "0", ...(x < 0, "-ve", "+ve"), "???")
```

```{r, include=FALSE}
x <- c(-3:3, NA)
if_else(x == 0, "0", if_else(x < 0, "-ve", "+ve"), "???")
```

###

This is already a little hard to read, and you can imagine it would only get harder if you have more conditions. One solution is to switch to `dplyr::case_when()`.

### Exercise 7

The code below should have the same output as above, but instead of using `if_else`, we are using `case_when()`. Click "Run Code".

```{r conditional-transfor-7, exercise = TRUE}
x <- c(-3:3, NA)
case_when(
  x == 0   ~ "0",
  x < 0    ~ "-ve", 
  x > 0    ~ "+ve",
  is.na(x) ~ "???"
)
```

###

dplyr’s `case_when()` is inspired by SQL’s `CASE` statement and provides a flexible way of performing different computations for different conditions. It has a special syntax that unfortunately looks like nothing else you’ll use in the tidyverse. It takes pairs that look like `condition ~ output`. `condition` must be a logical vector; when it’s `TRUE`, output will be used.

### Exercise 8

To explain how `case_when()` works, let’s explore some simpler cases. If none of the cases match, the output gets an `NA`. Click "Run Code".

```{r conditional-transfor-8, exercise = TRUE}
x <- c(-3:3, NA)
case_when(
  x < 0 ~ "-ve",
  x > 0 ~ "+ve"
)
```

###

The output should be `#> [1] "-ve" "-ve" "-ve" NA    "+ve" "+ve" "+ve" NA`.

### Exercise 9

If you want to create a “default”/catch all value, use TRUE on the left hand side. Copy the previous code and add `TRUE ~ "???"` as a new argument. Make sure to separate the arguments with a comma.

```{r conditional-transfor-9, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r conditional-transfor-9-hint, eval = FALSE}
x <- c(-3:3, NA)
case_when(
  x < 0 ~ "-ve",
  x > 0 ~ "+ve",
  ... ~ "???"
)
```

```{r, include=FALSE}
x <- c(-3:3, NA)
case_when(
  x < 0 ~ "-ve",
  x > 0 ~ "+ve",
  TRUE ~ "???"
)
```

###

This should output `#> [1] "-ve" "-ve" "-ve" "???" "+ve" "+ve" "+ve" "???"`.

### Exercise 10

Also, note that if multiple conditions match, only the first will be used. Click "Run Code".

```{r conditional-transfor-10, exercise = TRUE}
x <- c(-3:3, NA)
case_when(
  x > 0 ~ "+ve",
  x > 2 ~ "big"
)
```

###

This should output `#> [1] NA    NA    NA    NA    "+ve" "+ve" "+ve" NA` because it used the first argument. Just like with `if_else()` you can use variables on both sides of the`~` and you can mix and match variables as needed for your problem.

### Exercise 11

For example, we could use `case_when()` to provide some human readable labels for the arrival delay. Your gratitude for having the code written for you is much appreciated.

```{r conditional-transfor-11, exercise = TRUE}
flights |> 
  mutate(
    status = case_when(
      is.na(arr_delay)      ~ "cancelled",
      arr_delay < -30       ~ "very early",
      arr_delay < -15       ~ "early",
      abs(arr_delay) <= 15  ~ "on time",
      arr_delay < 60        ~ "late",
      arr_delay < Inf       ~ "very late",
    ),
    .keep = "used"
  )
```

###

Be wary when writing this sort of complex `case_when()` statement.

### Exercise 12

Note that both `if_else()` and `case_when()` require compatible types in the output. If they’re not compatible, you’ll see errors. To demonstrate, click "Run Code".

```{r conditional-transfor-12, exercise = TRUE}
if_else(TRUE, "a", 1)
```

###

This should output `#> Error in if_else(): #> ! Can't combine true <character> and false <double>.`

## Public Project
### 

<!-- Missing: Questions about the two libraries (socviz and ggtheme). Also -->

We will be making a plot about the gap in life expactancy in the US. 

```{r fig.width = 10}
final_plot <- 
  oecd_sum |> 
  drop_na() |> 
  ggplot(mapping = aes(x = year, y = diff, fill = hi_lo)) +
      geom_col() + 
      guides(fill = "none") +
      theme_minimal() +
      labs(x = NULL, 
           y = "Difference in Years",
           title = "The US Life Expectancy Gap",
           subtitle = "Difference between US and OECD average life expectancies,\n1960-2015",
           caption = "Data: OECD. After a chart by Christopher Ingraham,\nWashington Post, December 27th 2017.") +
      theme(plot.title = element_text(face = "bold",
                                      size = 20,
                                      margin = margin(b = 5)),
            plot.subtitle = element_text(face = "italic",
                                         size = 15,
                                         margin = margin(b = 10)),
            plot.caption = element_text(face = "italic",
                                        margin = margin(t = 5)),
            axis.title.y = element_text(face = "bold"),
            axis.text = element_text(face = "bold"))

final_plot
```

### Exercise 1

We will first explore the data set we will be using for the first graph. Run `oecd_sum` in the code chunk below.

```{r the-us-life-expectan-1, exercise = TRUE}

```

```{r the-us-life-expectan-1-hint-1, eval = FALSE}
oecd_sum
```

### 

The data set is from the Organisation for Economic Co-operation and Development, otherwise known as OECD. 

### Exercise 2

We will explore the `oecd_sum` data set further. Go onto this website [here](https://kjhealy.github.io/socviz/reference/oecd_sum.html). Copy and paste what the hi_lo and diff variables represent.

```{r the-us-life-expectan-2}
  question_text(NULL,
    answer(NULL, correct = TRUE),
    allow_retry = TRUE,
    try_again_button = "Edit Answer",
    incorrect = NULL,
    rows = 3)
```

### 

This dataset focuses specifically on the USA in relation to OECD stats.

### Exercise 3

We will be building a plot mapping the USA's average life expectancy compared to the OECD's from 1960 to 2015. Start by building a pipe with the data set. Use the function `select()`. Within select, input `-other, -usa`. 

```{r the-us-life-expectan-3, exercise = TRUE}

```

```{r the-us-life-expectan-3-hint-1, eval = FALSE}
oecd_sum |> 
  select(-other, -usa)
```

We do not need to know the life expectancy in OECD countries or in the US. We simply need to know the difference in life expectancy, which is already calculated for us. 

### Exercise 4

Continue the pipe with ggplot. Within `ggplot()`, set the argument `mapping` to equal `aes()`. Within `aes()`, map `year` to the x-axis and `diff` to the y-axis.

```{r the-us-life-expectan-4, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r the-us-life-expectan-4-hint-1, eval = FALSE}
oecd_sum |> 
  select(...) |> 
  ggplot(mapping = aes(x = ...,
                       y = ...))
```

### 

Remember: simply mapping variables to the x and y axis will create an empty plot. You must add a geom layer to visualize the data.

### Exercise 5

Continue building your plot. Reminder, since you are building the plot layer by layer, you must use `+`. Within `aes`, set the `fill` argument to `hi_lo`.

```{r the-us-life-expectan-5, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r the-us-life-expectan-5-hint-1, eval = FALSE}
oecd_sum |> 
  select(...) |> 
  ggplot(mapping = aes(x = ...,
                       y = ...,
                       fill = ...))
```

### 

This will allow the bars to be a different colour if the US' life expectancy is below/above the OECD's.

### Exercise 6

Add the `geom_col()` layer.

```{r the-us-life-expectan-6, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r the-us-life-expectan-6-hint-1, eval = FALSE}
oecd_sum |> 
  select(...) |> 
  ggplot(mapping = aes(x = ...,
                       y = ...,
                       fill = ...)) +
    geom_col()
```

### 

We choose to use `geom_col()` instead of `geom_bar()` since we already have the math done for how big the difference is between the US's life expectancy and OECD's in the `diff` column.

### Exercise 7

Add the layer `guides(fill = "none")`

```{r the-us-life-expectan-7, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r the-us-life-expectan-7-hint-1, eval = FALSE}
oecd_sum |> 
  select(...) |> 
  ggplot(mapping = aes(x = ...,
                       y = ...,
                       fill = ...)) +
    geom_col() +
    guides(fill = "none")
```

### 

`guides(fill = "none")` tells ggplot to drop the legend that would be generated automatically when the `fill` argument is used, when we have no need for it. The other way to accomplish this would be to do `theme(legend.position = "none")`.

### Exercise 8

Add the theme `theme_minimal()`.

```{r the-us-life-expectan-8, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r the-us-life-expectan-8-hint-1, eval = FALSE}
oecd_sum |> 
  select(...) |> 
  ggplot(mapping = aes(x = ...,
                       y = ...,
                       fill = ...)) +
    geom_col() +
    guides(fill = "none") +
    theme_minimal()
```

### 

This is part of the ggplot2 package. Explore the themes on the ggplot2 package [here](https://www.shanelynn.ie/themes-and-colours-for-r-ggplots-with-ggthemr/).

### Exercise 9

Add the `labs()` layer to give your plot the appropriate `title`, `subtitle`, `caption`, etc. Reminder: this is the plot we are trying to create:

```{r}
final_plot
```

```{r the-us-life-expectan-9, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r the-us-life-expectan-9-hint-1, eval = FALSE}
oecd_sum |> 
  select(...) |> 
  ggplot(mapping = aes(x = ...,
                       y = ...,
                       fill = ...)) +
    geom_col() +
    guides(fill = FALSE) +
    theme_minimal() +
    labs(title = ...,
         subtitle = ...,
         caption = ...,
         x = ...,
         y = ...)
```

### Exercise 10

Skim through [this documentation here](https://ggplot2-book.org/themes.html). We will be using these functions to modify the font of the graph and make it bold. What are the four element functions that can be used in `theme()`? 

```{r the-us-life-expectan-10}
question_text(NULL,
	message = "element_text(), element_line(), element_rect(), element_blank()",
	answer(NULL,
	correct = TRUE),
	allow_retry = FALSE,
	incorrect = NULL,
	rows = 6)
```

### 

We will primarily be using `element_text()`.

### Exercise 11

Summarize in your own words what the function `element_text()` does.
```{r the-us-life-expectan-11}
question_text(NULL,
	answer(NULL, correct = TRUE),
	allow_retry = TRUE,
	try_again_button = "Text for button",
	incorrect = NULL)
```

### 

Note: the only family fonts that will reliably work are "serif", "sans", and "mono". The usage of any other is not guaranteed to work.

### Exercise 12

We will now use `theme(plot.title = ...)` to change the plot title. Set the argument `face` to "bold".

```{r the-us-life-expectan-12, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r the-us-life-expectan-12-hint-1, eval = FALSE}
... +
  theme(plot.title = element_text(face ="bold")
```

### Exercise 13

Add the `size` argument within `element_text()`. Set the size to 20.

```{r the-us-life-expectan-13, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r the-us-life-expectan-13-hint-1, eval = FALSE}
... +
  theme(plot.title = element_text(face = "bold",
                                  size = 20)
```

### 

This aligns with the typical font sizes you have on your laptop.

### Exercise 14

We will adjust the margins to increase readability. Margins act as a buffer around each individual text block. Add the argument `margin` and then set it to `margin(b = 5)`

```{r the-us-life-expectan-14, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r the-us-life-expectan-14-hint-1, eval = FALSE}
... +
  theme(plot.title = element_text(face = ...,
                                  size = ...,
                                  margin = margin(b = 5))
```

### 

`b` stands for the "bottom" margin. 

### Exercise 15

We will now format the subtitle. Within the big `theme()` function, add a new argument titled `plot.subtitle`. Set it equal to element_text(). Note: there will be an error if you attempt to run this since there is no arguments within element_text().

```{r the-us-life-expectan-15, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r the-us-life-expectan-15-hint-1, eval = FALSE}
... +
  theme(plot.title = element_text(face = ...,
                                  size = ...,
                                  margin = ...),
        plot.subtitle = element_text())
```


### Exercise 16

Add the argument `face` to element_text(). Set it equal to "italic".

```{r the-us-life-expectan-16, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r the-us-life-expectan-16-hint-1, eval = FALSE}
...,
plot.subtitle = element_text(face = "italic")
```


### Exercise 17

Change the font size to 15.

```{r the-us-life-expectan-17, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r the-us-life-expectan-17-hint-1, eval = FALSE}
...,
plot.subtitle = element_text(face = "italic",
                             size = 15)
```


### Exercise 18

From here, set the bottom margin to 10 to create space between the subtitle and the graph.

```{r the-us-life-expectan-18, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r the-us-life-expectan-18-hint-1, eval = FALSE}
...,
plot.subtitle = element_text(face = "italic",
                             size = 15,
                             margin = margin(b = 10))
```

### 

The margin is measured in pixels.

### Exercise 19

We will now modify the caption to make it italic, and also create some space between the caption and the graph. Add the `plot.caption =` argument within `theme()`. Set it to element_text(). 

```{r the-us-life-expectan-19, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r the-us-life-expectan-19-hint-1, eval = FALSE}
...,
plot.subtitle = element_text(face = "italic",
                             size = 15,
                             margin = margin(b = 10)),
plot.caption = element_text()
```


### Exercise 20

Make the font italic.

```{r the-us-life-expectan-20, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r the-us-life-expectan-20-hint-1, eval = FALSE}
...,
plot.caption = element_text(face = "italic")
```

### Exercise 21

Set the top margin to 5. We do this by using `margin = margin(t = 5)`.

```{r the-us-life-expectan-21, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r the-us-life-expectan-21-hint-1, eval = FALSE}
...,
plot.caption = element_text(face = "italic",
                            margin = margin(t = 5))
```

### 

"t" stands for the top margin.

### Exercise 22

We will now simply set the y-axis title to be bold. Add axis.title.y as an argument to theme(). Set it equal to element_text(), and then face = "bold".

```{r the-us-life-expectan-22, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r the-us-life-expectan-22-hint-1, eval = FALSE}
axis.title.y = element_text(face = "bold")
```

### 

Any code that modifies the axes will have the prefix "axis" instead of "plot".

### Exercise 23

Last but not least, we will modify the axes ticks to be bold, so they are more easily readable. Add `axis.text` to within the `theme()` argument. Set it equal to `element_text()`. Make the font bold.

```{r the-us-life-expectan-23, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r the-us-life-expectan-23-hint-1, eval = FALSE}
axis.text = element_text(face = "bold"))
```

Reminder: your final graph should look like this.

```{r}
final_plot
```


## Summary
### 

This tutorial covered [Chapter 13: Logical vectors](https://r4ds.hadley.nz/logical-vectors.html) from [*R for Data Science (2e)*](https://r4ds.hadley.nz/) by Hadley Wickham, Mine Çetinkaya-Rundel, and Garrett Grolemund. You learned how to create logical vectors with `>`, `<`, `<=`, `=>`, `==`, `!=`, and `is.na()`, how to combine them with `!`, `&`, and `|`, and how to summarize them with `any()`, `all()`, `sum()`, and `mean()`. You also learnws about the powerful `if_else()` and `case_when()` functions that allow you to return values depending on the value of a logical vector.

```{r download-answers, child = system.file("child_documents/download_answers.Rmd", package = "tutorial.helpers")}
```
