---
title: Strings
author: David Kane
tutorial:
  id: strings
output:
  learnr::tutorial:
    progressive: yes
    allow_skip: yes
runtime: shiny_prerendered
description: 'Tutorial for Chapter 15: Strings'
---

```{r setup, include = FALSE}
library(learnr)
library(tutorial.helpers)
library(tidyverse)
library(babynames)
library(gghighlight)

x <- "n"
knitr::opts_chunk$set(echo = FALSE)
options(tutorial.exercise.timelimit = 60, 
        tutorial.storage = "local") 

set.seed(9)
police <- read_csv("data/philly-clean.csv",
                   col_types = cols()) |>
  slice_sample(n = 4000)

df <- tibble(name = c("Flora", "David", "Terra", NA))

df1 <- tibble(name = c("Flora", "David", "Terra", NA))

df2 <- tribble(
  ~ name, ~ fruit,
  "Carmen", "banana",
  "Carmen", "apple",
  "Marvin", "nectarine",
  "Terence", "cantaloupe",
  "Terence", "papaya",
  "Terence", "mandarin"
)

```

```{r copy-code-chunk, child = system.file("child_documents/copy_button.Rmd", package = "tutorial.helpers")}
```

```{r info-section, child = system.file("child_documents/info_section.Rmd", package = "tutorial.helpers")}
```

## Introduction
### 

This tutorial covers [Chapter 15: Strings](https://r4ds.hadley.nz/strings.html) from [*R for Data Science (2e)*](https://r4ds.hadley.nz/) by Hadley Wickham, Mine Çetinkaya-Rundel, and Garrett Grolemund. 
You will learn about some of the power of the [**stringr**](https://stringr.tidyverse.org/) package: how to create, combine, and extract strings, and about some of the challenges you might face with non-English strings. Some important functions which we will learn include:
[`str_c()`](https://stringr.tidyverse.org/reference/str_c.html),
[`str_glue()`](https://stringr.tidyverse.org/reference/str_glue.html),
[`str_flatten()`](https://stringr.tidyverse.org/reference/str_flatten.html),
[`separate_longer_delim()`](https://tidyr.tidyverse.org/reference/separate_longer_delim.html), and more.

## Creating a String
### 

We’ve created strings in passing earlier in the book but didn’t discuss the details. Firstly, you can create a string using either single quotes (') or double quotes (").

### Exercise 1

Type `this is a string` in double quotes and assign it to a variable named `string1` using `<-`.

```{r creating-a-string-1, exercise = TRUE}

```

```{r creating-a-string-1-hint-1, eval = FALSE}
string1 <- "This is a string"
```

### 

Now there’s no difference in behavior between the single quotes and double quotes, but in the interests of consistency, the [tidyverse](https://style.tidyverse.org/syntax.html#character-vectors) style guide recommends using ", unless the string contains multiple double quotes.

### Exercise 2

Now type `hello world` with one double quote, the other missing, and assign to `str2`. 

```{r creating-a-string-2, exercise = TRUE}

```

```{r creating-a-string-2-hint-1, eval = FALSE}
str2 <- "hello world
```

### 

If you forget to close a quote, you’ll see `+`: the continuation prompt. If this happens to you and you can’t figure out which quote to close, press Escape to cancel and try again.

### Exercise 3

Now type `hello "world"` in double quotes and use another pair of double quotes on `world` and assign to `str3`. 

```{r creating-a-string-3, exercise = TRUE}

```

```{r creating-a-string-3-hint-1, eval = FALSE}
str3 <- "hello "world""
```

### 

The code `str3 <- "hello "world""` doesn't work in RStudio due to a syntax error.Without the proper escape character, RStudio will interpret the second double quote as the end of the string, resulting in a syntax error.

### Exercise 4

In the previous exercise, we encountered a syntax error when trying to assign the string `str3 <- "hello "world""` in RStudio. To fix this error, we can modify the code by changing the outer double quotes to single quotes, while keeping the inner double quotes intact. This way, the double quotes will be visible within the string.

```{r creating-a-string-4, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r creating-a-string-4-hint-1, eval = FALSE}
str3 <- 'hello "world"'
```

### 

If you wanted to include single quotes in a string you could put the outer quotes as double quotes.

### Exercise 5

New create a new variable `x` and set it `n` using `<-` in a string format using double quotes. Then on a new line call x.

```{r creating-a-string-5, exercise = TRUE}

```

```{r creating-a-string-5-hint-1, eval = FALSE}
x <- "n"
x
```

### 

The reason we don't see n only without quotes is that R shows the whole string as a vector.

### Exercise 6

To view only the string and not the quotes, lets load the [`stringr`](https://stringr.tidyverse.org/) library using `library()`.

```{r creating-a-string-6, exercise = TRUE}

```

```{r creating-a-string-6-hint-1, eval = FALSE}
library(...)
```

### 

Stringr library is a part of tidyverse and we can just load tidyverse and stringr will automatically loaded but for the practice we are loading libaries.

### Exercise 7

Let's type `str_view()` and for the arguement use the varibale `x` that we set earlier.

```{r creating-a-string-7, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r creating-a-string-7-hint-1, eval = FALSE}
str_view(...)
```

### 

### Exercise 8

Let's do the same thing on single quote, so create a new variable `y` and set it to `"'"` using `<-`. Then on a new line, type `str_view()` and set `y` as arguement.

```{r creating-a-string-8, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r creating-a-string-8-hint-1, eval = FALSE}
y <- ...
str_view(...)
```

### 

### Exercise 9

The way that I demonstrated previously to include quotes in string is not the only way. To include a literal single or double quote in a string , you can use `\` to “escape” it like `"\""` which will return `'"'`. Make a string with a single quote or literal single as they call it and assign it to `single_quote`.

```{r creating-a-string-9, exercise = TRUE}

```

```{r creating-a-string-9-hint-1, eval = FALSE}
single_quote <- '\''
```

### 

If you wanted to include a literal backslash in your string, you'll need to escape it with "\\".

### Exercise 10

Let's assign

```{r creating-a-string-10, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r creating-a-string-10-hint-1, eval = FALSE}

```

### 

### Exercise 11

The printed representation of a string is not the same as the string itself because the printed representation shows the escapes. To see the raw contents of the string, we use `str_view()`

```{r creating-a-string-11, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r creating-a-string-11-hint-1, eval = FALSE}

```

### 

## Creating many strings from data
### 

Now that you’ve learned the basics of creating a string or two by “hand”, we’ll go into the details of creating strings from other strings. This will help you solve the common problem where you have some text you wrote that you want to combine with strings from a data frame.

We’ll show you how to combine strings with `str_c()` and `str_glue()` and how you can use them with `mutate()`.
 
### Exercise 1

The first function we will learn is `str_c()`. `str_c()`takes any number of vectors as arguments and returns a character vector, so run `str_c("x","y")` and see what you get.

```{r creating-many-string-1, exercise = TRUE}

```

```{r creating-many-string-1-hint-1, eval = FALSE}
str_c(...,...)
```

### 

`str_c()` is a function in the `stringr` package in R that combines multiple character vectors into a single character vector. It is similar to the `paste()` function, but it uses `tidyverse` recycling and NA rules.



### Exercise 2

Copy previous code and change the argument to `str_c("Hello", c("Precptor","Anish"))` and run it.

```{r creating-many-string-2, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r creating-many-string-2-hint-1, eval = FALSE}
str_c("Hello", c("Precptor","Anish"))
```

### 

 
Based from the result we can see that `str_c()` is vectorized, which means it can take multiple arguments and combine them element-wise. 
 
### Exercise 3

To show how we can use `str_c()` on a dataset, lets create a tibble. Create a variable named `df` and set it to `tibble(name = c("Flora","Preceptor","Terra",NA))` which create a data set with the string characters.

```{r creating-many-string-3, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r creating-many-string-3-hint-1, eval = FALSE}
... <- ...
```

### 

You might be wondering what tibble is, A tibble is a new form of a data frame in R that is part of the tidyverse library, and Tibbles print the data in a more efficient format than a data frame, showing the values of the columns, their datatype, and the size of the dataset.

### Exercise 4

Run `df` to see the data frame.

```{r creating-many-string-4, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r creating-many-string-4-hint-1, eval = FALSE}
...
```

### 

the tibble has one column named "name." The first three rows contain the names "Flora," "David," and "Terra," respectively. However, the fourth row has an NA value, which typically represents missing or unknown data. 

<!-- AB: should i use this, There are so many functions you could use for a tibble like `select()` which drops columns from a tibble, `filter()` whch drops rows based on certain conditions and more. -->


### Exercise 5

Let's modify the tibble by first piping the tibble `df` to the `mutate()` function, then in `mutate()`, we will create new column `greeting` and set it equal to `str_c("Hi ", name, "!")`.

```{r creating-many-string-5, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r creating-many-string-5-hint-1, eval = FALSE}
... |>
  mutate(... = ...("...",name,"..."))
```

### 

If you want missing values to display in another way, use [`coalesce()`](https://dplyr.tidyverse.org/reference/coalesce.html) to replace them.

### Exercise 6

Run `df` to see the modified data frame.

```{r creating-many-string-6, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r creating-many-string-6-hint-1, eval = FALSE}

```

### 

After modifying the tibble, we still have a  column called `name`. We used the `mutate()` function from the ["dplyr"](https://dplyr.tidyverse.org/) package to create a new column called `greeting`. The `str_c()` function is used to concatenate the string "Hi " with the values in the `name` column. 

### Exercise 7

If you are mixing many fixed and variable strings with str_c(), you’ll notice that you type a lot of `"`s, making it hard to see the overall goal of the code. An alternative approach is provided by the glue package via `str_glue()`.

Let's create a variable named `df1` and set it to `tibble(name = c("Flora","Preceptor","Terra",NA))` which create a data set with the string characters.

```{r creating-many-string-7, exercise = TRUE}

```

```{r creating-many-string-7-hint-1, eval = FALSE}
... <- ...
```

### 

### Exercise 8

Let's modify the tibble by first piping the tibble `df1` to the `mutate()` function, then in `mutate()`, we will create new column `greeting` and set it equal to `str_glue("Hi {name} !")`.

```{r creating-many-string-8, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r creating-many-string-8-hint-1, eval = FALSE}
df1 |>
  ...(greeting = ...())
```

### 

How `str_glue()` works is that if you give it a single string that has a special feature: anything inside `{}` will be evaluated like it’s outside of the quotes.

### Exercise 9

Run `df1` to see the data frame.

```{r creating-many-string-9, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r creating-many-string-9-hint-1, eval = FALSE}
...
```

### 

After modifying the tibble, we still have a  column called `name`. We used the `mutate()` function from the ["dplyr"](https://dplyr.tidyverse.org/) package to create a new column called `greeting`. As you can see, `str_glue()` currently converts missing values to the string "NA" unfortunately making it inconsistent with `str_c()`.

### Exercise 10

`str_c()` and `str_glue()` work well with mutate() because their output is the same length as their inputs. What if you want a function that works well with `summarize()`, i.e. something that always returns a single string? That’s the job of `str_flatten()`: it takes a character vector and combines each element of the vector into a single string.

Type `str_flatten()` and within it have a vector of strings like this `c("x","y","z")`.

```{r creating-many-string-10, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r creating-many-string-10-hint, eval = FALSE}
str_flatten(...)
```

###

One variation of `str_flatten()` is `str_flatten_comma()` is a variation designed specifically for flattening with commas. It automatically recognizes if last uses the Oxford comma and handles the special case of 2 elements.

### Exercise 11

What if you want commas after every letter. Copy the previous code and within the argument, add a comma and type a comma in string like this `","`.

```{r creating-many-string-11, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r creating-many-string-11-hint, eval = FALSE}
str_flatten(c(...),...)
```

###

The second argument which was a comma as a string as we saw, it is actually knows as a `collapse`: a string to enter between each piece. Defaults to "".

### Exercise 12

Let's fix the grammar of the string we got in previous exercise, copy the previous code and add `last = ", and "`.

```{r creating-many-string-12, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r creating-many-string-12-hint, eval = FALSE}

```

###

The `last` argument is a optional string to use in place of the final separator.

### Exercise 13

Let's now use `str_flatten()` with `summarize()`. Run `df3` to have a quick look at the data frame.

```{r creating-many-string-13, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r creating-many-string-13-hint, eval = FALSE}
df3
```

###

The tibble has two columns named "name" and "fruit". Carmen has more than one fruit to his name and Terence has more than two fruits to his name.

### Exercise 14

Start a pipe with `df2` to `group_by()` which takes `name` as arguement.

```{r creating-many-string-14, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r creating-many-string-14-hint, eval = FALSE}
df2 |>
  ...(...)
```

###

### Exercise 15

Continue the pipe from `group_by()` to `summarize()`.

```{r creating-many-string-15, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r creating-many-string-15-hint, eval = FALSE}

```

###

### Exercise 16

Within `summarize()`, set the column `fruit` to `str_flatten(fruit, ",")`.

```{r creating-many-string-16, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r creating-many-string-16-hint, eval = FALSE}

```

###






## Extracting data from strings
### 

It's very common for multiple variables to be crammed together into a single string. In this section, you’ll learn how to use four tidyr functions to extract them:

`df |> separate_longer_delim(col, delim)`
`df |> separate_longer_position(col, width)`
`df |> separate_wider_delim(col, delim, names)`
`df |> separate_wider_position(col, widths)`

## Letters
### 

In this section, we’ll introduce you to functions that allow you to work with the individual letters within a string. You’ll learn how to find the length of a string, extract substrings, and handle long strings in plots and tables.

## Non-English Text
### 

When working with non-English text, the first challenge is often the encoding. To understand what’s going on, we need to dive into how computers represent strings. In R, we can get at the underlying representation of a string using charToRaw()


## Last letter of boy names
<!-- DK: Some code here seems too hard. Or maybe it just needs a clearer explanation. -->

The **[babynames](http://hadley.github.io/babynames/)** package contains data about baby names through the years. 

### 

Using this data, let's create the folllowing graph.

```{r baby-lastnames}
boys_p <- babynames |>
  filter(sex == 'M') |>
  
  # prop is supposed to be proportion of a specific name in all male baby names
  # in that year. But I am not sure it is! For example, if you sum up prop for
  # all M in a given year, it never adds to 1. It always adds to a number less
  # than 1, but greater than 0.9. My *guess* would be that this is due to
  # babynames dropping names which have very small numbers of babies with that
  # name. If true, then prop is likely correct.
  
  select(name, year, prop) |>
  
  # Add last letter of baby names. 
  
  mutate(last_letter = stringr::str_sub(name, -1, -1)) |>
  
  # frequency is the poroportion of a certain last letter in all male baby names in that year. 
  
  summarize(frequency = sum(prop), 
            .by = c(year, last_letter)) |>
  ggplot(aes(x = year, y = frequency, colour = last_letter)) +
    geom_line() + 
  
    # only color last letters that appears more than 15 % of all boy names in any year. 
    
    gghighlight(max(frequency) > 0.15,
                label_key = last_letter) + 
    scale_x_continuous(guide = guide_axis(angle = 70),
                       breaks = c(1880, 1900, 1925, 1950,
                                  1975, 2000, 2017)) +
    scale_y_continuous(labels = scales::percent_format()) +
    theme_minimal() + 
    labs(x = NULL,
         y = NULL,
         subtitle = "Names ending with 'N' increased rapidly after 1950",
         title = "Last Letter of Boy Baby Names",
         caption = "Source: http://hadley.github.io/babynames")

boys_p
```

### Exercise 1

`glimpse()` the `babynames` data set. Pay close attention to the data type of each variable.

```{r last-letter-of-boy-n-1, exercise = TRUE}

```

### 

We could have also used `head()` function to check the first few rows of the tibble.

### Exercise 2

<!-- DK: Seems bad. We never do this! -->

To check for `NA` values, use `any()` function with argument `is.na()`, argument of which should be the dataset. 

```{r last-letter-of-boy-n-2, exercise = TRUE}

```

```{r last-letter-of-boy-n-2-hint-1, eval = FALSE}
any(is.na(...))
```

### 

This function returned `FALSE`, which indicates that there are no NA values. 

### Exercise 3

Because we are trying to analyse only boy names by last letter, start a pipe with     `babynames`,`filter()` the data set to only get the data of boy names. 

```{r last-letter-of-boy-n-3, exercise = TRUE}

```

```{r last-letter-of-boy-n-3-hint-1, eval = FALSE}
babynames |>
  filter(... == 'M')

```

```{r last-letter-of-boy-n-3-hint-2, eval = FALSE}
babynames |>
  filter(sex == 'M')

```

### Exercise 4

Select the `name`, `year`, and `prop` columns. 

```{r last-letter-of-boy-n-4, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r last-letter-of-boy-n-4-hint-1, eval = FALSE}
...  |>
  select(..., ..., ...)
```

### 

When selecting columns, we can also combine multiple select helpers. For example,
when selecting `name` and `year` columns, we can use the following code: 
`select(starts_with("na") | ends_with("ar"))`.

### Exercise 5

We want the last letters of boy names. Therefore, create a new variable names `last_letters` that should be equal to last letter of each boy name. To get the last letter of each boy name, we will make use of `str_sub()` function of **[stringr](https://stringr.tidyverse.org/)** package. 

### 

Use the `mutate()` function to add a new column: `last_letter`. Its value is the `str_sub()` function with `name`, `-1` and `-1` as its first three arguments.

```{r last-letter-of-boy-n-5, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r last-letter-of-boy-n-5-hint-1, eval = FALSE}
... |>
  mutate(last_letter = stringr::str_sub(..., ..., -1))
```

```{r last-letter-of-boy-n-5-hint-2, eval = FALSE}
babynames |>
  filter(sex == 'M')|>
  select(name, year, prop)|>
  mutate(last_letter = stringr::str_sub(name, -1, -1))
```

### 

`str_sub` takes three arguments: character vector to work on, starting character, and ending character. Because we wanted to extract last letter, or -1 element, of each name, we set the starting and ending characters as -1.


### Exercise 6

Use `summarize()` to create the variable `frequency` that is equal to `sum()` of proportions (`prop`) of each group. Also, set the `.by` argument to `c(year, last_letter)` because we want to know the value for each year/letter combination. 

```{r last-letter-of-boy-n-6, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r last-letter-of-boy-n-6-hint-1, eval = FALSE}
... |> 
  summarize(frequency = sum(prop), 
            .by = c(year, last_letter))
```

### Exercise 7

Now, create a ggplot object, and within `ggplot(aes())`, set x-axis values to `year`, y-axis values to `frequency`, and colour to `last_letter.` 

<!-- AB: not working for somereason with colour and problem happend through whole tutorial  -->

```{r last-letter-of-boy-n-7, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r last-letter-of-boy-n-7-hint-1, eval = FALSE}
...|>
  ggplot(aes(x = ..., y = ..., colour = ...)) 
```

### Exercise 8

Add `geom_line()` to the ggplot object from previous exercise.


```{r last-letter-of-boy-n-8, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r last-letter-of-boy-n-8-hint-1, eval = FALSE}
...  +
  geom_line()
```

### 

Alternatively, we could have used other chart types such as scatter chart. However, for our goal, line plot is a great option. 

### Exercise 9

Because we want to analyse last letters that are prevalent, we want to color only last letters that have been above a certainn threshold.

### 

We use 15% as the threshold. Therefore, we only want to color last letters that for certain time have been the last letter of more than 15% of all baby boys that were born during that time. 

### 

We will utilize `gghighlight` library for this purpose. Add `gghighlight()` to the ggplot object. Within this function, set `max()` `frequency` to be greater than `0.15`, and to avoid any warnings, set `label_key` equal to `last_letter`.

```{r last-letter-of-boy-n-9, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r last-letter-of-boy-n-9-hint-1, eval = FALSE}
...+ 
  gghighlight(max(...) > ...,
                label_key = ...) 
```

### 

We don't think current x-axis ticks are informative enough. Therefore, we should set custom x-axis ticks.

### Exercise 10

Add `scale_x_continuous`, and within this function, set `breaks` argument equal to numerical vector of years. You can choose any years within the data set. We preferred 1880, 1900, 1925, 1950, 1975, 2000, and 2017. 

```{r last-letter-of-boy-n-10, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r last-letter-of-boy-n-10-hint-1, eval = FALSE}
... +
  scale_x_continuous(breaks = c(1880, 1900, ..., ...,
                                ..., 2000, 2017)) 
```

### 

A rule of thumb, however, is to include ending and beginning years so that whoever sees your plot could understand the range of years that data is collected.

### Exercise 11

Moreover, we want to visualize y-axis as percentage instead of proportion. Therefore, add `scale_y_continuous`, and within this function, set `labels` argument equal to `scales::percent_format()`.

```{r last-letter-of-boy-n-11, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r last-letter-of-boy-n-11-hint-1, eval = FALSE}
... +
  scale_y_continuous(... = scales::...()) 
```

```{r last-letter-of-boy-n-11-hint-2, eval = FALSE}
... +
  scale_y_continuous(labels = scales::percent_format()) 
```

### 

`scale` library provides many functions to assist the scaling of axes. For example, `scale_y_binned()` can be used to discretize continuous position data.

### Exercise 12

Add `theme_minimal()`. 

```{r last-letter-of-boy-n-12, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r last-letter-of-boy-n-12-hint-1, eval = FALSE}
... +
  theme_minimal()
```

### 

We also could have used themes such as `theme_gray()`, `theme_void()`, and`theme_bw()`.

### Exercise 13

Finally, add `caption`, `title`, `subtitle`, and `axis` labels of your choice. Add `labs()` function to the plot. 

Reminder: This is what your plot should look like

```{r last-letter-of-boy-n-13}
boys_p
```

```{r baby-lastnames-13, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r last-letter-of-boy-n-13-hint-1, eval = FALSE}
... +
  labs(...)
```

## Mutating Strings 
### 

A string is one or more characters that are enclosed inside a pair of matching 'single' or "double quotes." We will now use various functions to manipulate strings.

### Exercise 1

Let's start with a small subset of data about police complaints in Philadelphia. Run `police` in the code chunk below to view this data.

```{r mutating-strings-1, exercise = TRUE}
police
```

### 

As wide as this data is, we will not be using most of the columns.


### Exercise 2

Start a new pipe with our data by typing `police` followed by the `|>` operator. On a new line,`select()` the `summary` column,`po_race`, `district_occurrence`, and `investigative_findings`. We will first focus on the `summary` column, but will use the other columns later in this section. 

```{r mutating-strings-2, exercise = TRUE}

```

```{r mutating-strings-2-hint-1, eval = FALSE}
police |> 
  select(...)
```

### 

Use the buttons in the bottom right corner to navigate through all the rows.

### Exercise 3

Use `mutate()` and `str_detect()` to create a new column called `race`. Each observation will either be `TRUE` or `FALSE` depending on whether or not the adjacent `summary` observation contains the string `"rac"`.

```{r mutating-strings-3, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>


```{r mutating-strings-3-hint-1, eval = FALSE}
The first argument in string detect is the column in 
which you want to detect a string, the second argument 
is the quoted string you want to detect.
```

```{r mutating-strings-3-hint-2, eval = FALSE}
... |> 
  mutate(race = str_detect(..., "..."))
```

### 

You may need to use the arrow key in the top right corner of the table to view the `race` column.

### 

This new column in `police` tells us if a complaint involved someone's race. We had to specifically look for the string `"rac"` to account for words like "race" and "racial".

### Exercise 4

Filter your pipe so that you keep rows only where the column `race` has value `TRUE`. 

```{r mutating-strings-4, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r mutating-strings-4-hint-1, eval = FALSE}
Use the function filter() with the == evaluator
```

```{r mutating-strings-4-hint-2, eval = FALSE}
... |> 
  filter(race == ...) 
```

### 

There should be 180 rows in which the `race` column has value `TRUE`.

### Exercise 5

In your pipe, use `mutate()` and `str_detect()` to create a new column called `physical` with value `TRUE` when the column `summary` contains the string `"physical"` and `FALSE` if it does not. Then filter so that you keep rows only when the column `physical` has value `TRUE`. 

```{r mutating-strings-5, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r mutating-strings-5-hint-1, eval = FALSE}
... |> 
  mutate(... = str_detect(..., "...")) |> 
  filter(... == ...)
```

```{r mutating-strings-5-hint-2, eval = FALSE}
police |> 
  select(summary, po_race, district_occurrence, investigative_findings) |>
  mutate(race = str_detect(summary, "rac")) |>
  mutate(physical = str_detect(..., "...")) |>
  filter(race == TRUE & physical == ...)
```

### 

This gives us a table with all of the complaints that involved someone's race and some kind of physical altercation. There should be 43 rows in which both the race and the physical columns have value `TRUE`. 

### Exercise 6

Copy-paste your pipe from the previous exercise and use `mutate()` and `str_sub()` to change the column `po_race` so that it only contains the first two letters of each row. The second argument should be the character position from where you want to start, and the third argument should be the character position where you want to end.

```{r mutating-strings-6, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r mutating-strings-6-hint-1, eval = FALSE}
... 
  mutate(po_race = str_sub(..., ..., ...))
```

```{r mutating-strings-6-hint-2, eval = FALSE}
... |> 
  mutate(po_race = str_sub(po_race, 1, ...))
```

### 

The "sub" in `str_sub()` stands for `substrings`. This function extracts "substrings" (parts of a string) based on the `start` and `end` arguments.

### Exercise 7

Copy-paste your pipe from the previous exercise and use `mutate()` and `str_c()` to add the string `"PA-"` in front of every string in the column `district_occurrence`. `str_c` **c**ombines two strings and its syntax is similar to `c()`. The first argument is what you want to appear first in the combined string, `"PA-"`. The second argument should be the name of the column, `district_occurrence`. 

```{r mutating-strings-7, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r mutating-strings-7-hint-1, eval = FALSE}
... |> 
  mutate(... = str_c("PA-", ...))
```

```{r mutating-strings-7-hint-2, eval = FALSE}
... |> 
  mutate(... = str_c("PA-", district_occurrence))
```

### 

A useful context for this operation might be if we were going to add analogous data from another state into this tibble, which might have overlapping district numbers. 

### 

Note that R assigned this column to a character variable. If this were a numeric variable, the command would still work. R would coerce the edited column from a numeric variable to a character variable. 


### Exercise 8

Copy-paste your pipe from the previous exercise. Use `mutate()` and `str_replace()` to replace the pattern `"Sustained "` with `"S"` in the column `investigative_findings`.  Note that we have included a space after the word "Sustained". Set the second argument to the pattern in the current vector which you want to change. Set the third argument to the replacement which you want to replace the pattern with. 

```{r mutating-strings-8, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r mutating-strings-8-hint-1, eval = FALSE}
x |> 
  select(po_race, district_occurrence, investigative_findings, summary) |> 
  mutate(race = str_detect(summary, "rac")) |> 
  filter(race == TRUE) |> 
  mutate(physical = str_detect(summary, "physical")) |> 
  filter(physical  == TRUE) |> 
  mutate(po_race = str_sub(po_race, 1, 2)) |> 
  mutate(... = str_c("PA-", district_occurrence)) |> 
  mutate(investigative_findings = str_replace(investigative_findings, "... ", "..."))
```

### Exercise 9

Copy-paste your pipe from the previous exercise. Use `mutate()` and `str_replace()` to replace the pattern `"Finding"` with `"F"` in the column `investigative_findings`. 

```{r mutating-strings-9, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r mutating-strings-9-hint-1, eval = FALSE}
x |> 
  select(po_race, district_occurrence, investigative_findings, summary) |> 
  mutate(race = str_detect(summary, "rac")) |> 
  filter(race == TRUE) |> 
  mutate(physical = str_detect(summary, "physical")) |> 
  filter(physical  == TRUE) |> 
  mutate(po_race = str_sub(po_race, 1, 2)) |> 
  mutate(... = str_c("PA-", district_occurrence)) |> 
  mutate(investigative_findings = str_replace(investigative_findings, "Sustained ", "S")) |> 
  mutate(... = str_replace(..., "...", "..."))

```

## Summary
### 

This tutorial covered [Chapter 15: Strings](https://r4ds.hadley.nz/strings.html) from [*R for Data Science (2e)*](https://r4ds.hadley.nz/) by Hadley Wickham, Mine Çetinkaya-Rundel, and Garrett Grolemund. 
You learned about some of the power of the [**stringr**](https://stringr.tidyverse.org/) package: how to create, combine, and extract strings, and about some of the challenges you might face with non-English strings. Some important functions which we learned include:
[`str_c()`](https://stringr.tidyverse.org/reference/str_c.html),
[`str_glue()`](https://stringr.tidyverse.org/reference/str_glue.html),
[`str_flatten()`](https://stringr.tidyverse.org/reference/str_flatten.html),
[`separate_longer_delim()`](https://tidyr.tidyverse.org/reference/separate_longer_delim.html), and more.

```{r download-answers, child = system.file("child_documents/download_answers.Rmd", package = "tutorial.helpers")}
```
