---
title: Exploratory data analysis
author: David Kane and Srihith Garlapati
tutorial:
  id: exploratory-data-analysis
output:
  learnr::tutorial:
    progressive: yes
    allow_skip: yes
runtime: shiny_prerendered
description: 'Tutorial for Chapter 11: Exploratory data analysis'
---

```{r setup, include = FALSE}
library(learnr)
library(tutorial.helpers)

library(tidyverse)
library(tidymodels)
library(hexbin)

knitr::opts_chunk$set(echo = FALSE)
options(tutorial.exercise.timelimit = 60, 
        tutorial.storage = "local") 

smaller <- diamonds |> 
  filter(carat < 3)

p1 <- nycflights13::flights |> 
  mutate(
    cancelled = is.na(dep_time),
    sched_hour = sched_dep_time %/% 100,
    sched_min = sched_dep_time %% 100,
    sched_dep_time = sched_hour + (sched_min / 60)
  ) |> 
  ggplot(aes(x = sched_dep_time)) + 
  geom_freqpoly(aes(color = cancelled), binwidth = 1/4) + 
  labs(title = "Frequency of Cancelled and Non-cancelled Flights for Scheduled Flights",
       subtitle = "There are a lot of non-cancelled flights at 3 PM",
       x = "Scheduled Departure Time",
       y = "Frequency",
       color = "Cancelled or Non-Cancelled")

p2 <- mpg |> 
  ggplot(aes(x = hwy, y = fct_reorder(class, hwy, median))) +
  geom_boxplot() + 
  labs(
    title = "Variation of Highway Mileage over Classes", 
    subtitle = "The compact and midsized cars are tied for the highest highway mileage.",
    x = "Highway Mileage",
    y = "Car Classes"
  )

p3 <- diamonds |> 
  count(color, cut) |>  
  ggplot(aes(x = color, y = cut)) +
  geom_tile(aes(fill = n)) +
  labs(
    title = "Covariation between Diamond Cut and the Color of the Diamond",
    subtitle = "There are more ideally cut diamonds than other cut types of diamonds",
    x = "Color of diamond",
    y = "Cut of diamond"
  )

p4 <- smaller |> 
  ggplot( aes(x = carat, y = price)) + 
  geom_boxplot(aes(group = cut_width(carat, 0.1))) +
  labs(title = "Covariation between Number of Carats and Price of Diamonds",
       subtitle = "As the number of carats of a diamond increases the price also increases",
       x = "Number of Carats",
       y = "Price of the Diamond")
  
diamonds3 <- diamonds |>
  mutate(
    log_price = log(price),
    log_carat = log(carat)
  )

diamonds_fit <- linear_reg() |>
  fit(log_price ~ log_carat, data = diamonds3)

diamonds_aug <- augment(diamonds_fit, new_data = diamonds3) |>
  mutate(.resid = exp(.resid))

p5 <- diamonds_aug |> 
  ggplot(aes(x = cut, y = .resid)) + 
  geom_boxplot() + 
  labs(
    title = "Prediction Model between Carat and Residual Price",
    subtitle = "As the number of carat increases the Residual Price decreases",
    x = "Number of Carats",
    y = "Residual Price"
  )

```

<!-- SG: Add labs() to this area -->


```{r copy-code-chunk, child = system.file("child_documents/copy_button.Rmd", package = "tutorial.helpers")}
```

```{r info-section, child = system.file("child_documents/info_section.Rmd", package = "tutorial.helpers")}
```

<!-- The below links should be dropped at least once, if not twice. -->

<!-- https://ggplot2.tidyverse.org/ -->
<!-- https://ggplot2-book.org/ -->
<!-- https://socviz.co/ -->
<!-- https://clauswilke.com/dataviz/ -->

<!-- DK: Always load the libraries. -->


## Introduction
### 


This tutorial covers [Chapter 11: Exploratory data analysis](https://r4ds.hadley.nz/eda.html) from [*R for Data Science (2e)*](https://r4ds.hadley.nz/) by Hadley Wickham, Mine Çetinkaya-Rundel, and Garrett Grolemund. We will be making use of the [**ggplot2**](https://ggplot2.tidyverse.org/) package and the [**dplyr**](https://dplyr.tidyverse.org/) to learn about Variation and Patterns. Some key commands that we will be making use of is [`coord_cartesian()`](https://ggplot2.tidyverse.org/reference/coord_cartesian.html) which sets limits on the coordinate system and will zoom the plot, [`geom_freqpoly()`](https://ggplot2.tidyverse.org/reference/geom_histogram.html) which visualizes the distribution of a single continuous variable by dividing the x axis into bins and counts the number of observations in each bin, [`fct_reorder()`](https://forcats.tidyverse.org/reference/fct_reorder.html), [`geom_tile()`](https://ggplot2.tidyverse.org/reference/geom_tile.html), and [`geom_count()`](https://ggplot2.tidyverse.org/reference/geom_count.html) which is a variant of `geom_point()` that counts the number of observations at each location, that maps the count to point area. 

## Variation
### 

In this section, we will learn about variation or basically the tendency of values of a variable to change from measurement to measurement. 

### Exercise 1

Load in the **tidyverse** package using `library()`.

```{r variation-1, exercise = TRUE}

```

```{r variation-1-hint-1, eval = FALSE}
library(...)
```

### 

In this tutorial, we’ll combine what you’ve learned about the **dplyr** package and the **ggplot2** package to interactively ask questions, answer them with data, and then ask new questions.


### Exercise 2

Type in `diamonds` and hit "Run Code"

```{r variation-2, exercise = TRUE}

```

```{r variation-2-hint-1, eval = FALSE}
diamonds
```

### 

The `diamonds` dataset contains the prices and other attributes of almost 54,000 diamonds, such as the `price`, `carat`, `cut`, `color`, and so on.

### Exercise 3

With a new pipeline, pipe `diamonds` to `ggplot()`.

```{r variation-3, exercise = TRUE}

```

```{r variation-3-hint-1, eval = FALSE}
diamonds |> 
  ggplot()
```

Note that this plot is blank.

### 

You can see variation easily in real life; if you measure any continuous variable twice, you will get two different results. Each of your measurements will include a small amount of error that varies from measurement to measurement.

### Exercise 4

Within `ggplot()`, using `aes()`, set the `x` argument to `carat`.

```{r variation-4, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r variation-4-hint-1, eval = FALSE}
diamonds |> 
  ggplot(aes(x = ...))
```

### 

Variables can also vary if you measure across different subjects (e.g., the eye colors of different people) or at different times (e.g., the energy levels of an electron at different moments). 

### Exercise 5

Add `geom_histogram()` to the pipeline.

```{r variation-5, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r variation-5-hint-1, eval = FALSE}
... + 
  geom_histogram()
```

### 

Let's visualize the distribution of the variable's values using the `binwidth` argument to `geom_histogram()`.

### Exercise 6

Within `geom_histogram()`, add the `binwidth` argument and set it equal to `0.5`. 

```{r variation-6, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r variation-6-hint-1, eval = FALSE}
... +
  geom_histogram(binwidth = ...)
```

### 

This plot is a way to visualize variation. Let's now take a look at the distribution of carat for smaller diamonds.

### Exercise 7

Create a new variable called `smaller` and set it equal to `diamonds`. 

```{r variation-7, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r variation-7-hint-1, eval = FALSE}
smaller <- ...
```

### 

Some valuable questions to ask yourself when dealing with variation are the following:

- Which values are the most common? Why?
- Which values are rare? Why? Does that match your expectations?
- Can you see any unusual patterns? What might explain them?

### Exercise 8

We are still creating the `smaller` tibble. Copy your previous code and create a pipeline with `diamonds`. Pipe `diamonds` with `filter()` and filter out `carat` that are less than `3`. 

```{r variation-8, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r variation-8-hint-1, eval = FALSE}
smaller <- diamonds |> 
  filter(carat < ...)
```

### 

*“There are no routine statistical questions, only questionable statistical routines.” — Sir David Cox*

### Exercise 9

With a new pipeline, pipe `smaller` to `ggplot()`. 

```{r variation-9, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r variation-9-hint-1, eval = FALSE}
smaller |> ggplot()
```

### 

EDA, or exploratory data analysis, is an iterative cycle. You:

1. Generate questions about your data.

2. Search for answers by visualizing, transforming, and modelling your data.

3. Use what you learn to refine your questions and/or generate new questions.

### Exercise 10

Within `ggplot()`, using `aes()`, set the `x` argument to `carat`. 

```{r variation-10, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r variation-10-hint-1, eval = FALSE}
smaller |> 
  ggplot(aes(x = ...))
```

### 

EDA is not a formal process with a strict set of rules. More than anything, EDA is a state of mind. During the initial phases of EDA you should feel free to investigate every idea that occurs to you. Some of these ideas will pan out, and some will be dead ends.

### Exercise 11

Add `geom_histogram()` to the pipeline.

```{r variation-11, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r variation-11-hint-1, eval = FALSE}
... + 
  geom_histogram()
```

### 

This histogram suggests several interesting questions:

- Why are there more diamonds at whole carats and common fractions of carats?

- Why are there more diamonds slightly to the right of each peak than there are slightly to the left of each peak?

### Exercise 12

Within `geom_histogram()`, set the `binwidth` argument to `0.01`. 

```{r variation-12, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r variation-12-hint-1, eval = FALSE}
... +
  geom_histogram(binwidth = ...)
```

### 

This visualization also revealed clusters, which suggest that subgroups exist in your data. To understand the subgroups, ask:

- How are the observations within each subgroup similar to each other?

- How are the observations in separate clusters different from each other?

- How can you explain or describe the clusters?

- Why might the appearance of clusters be misleading?

### Exercise 13

Let's observe and take note of unusual values in the dataset. With a new pipeline, pipe `diamonds` to `ggplot()`. 

```{r variation-13, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r variation-13-hint-1, eval = FALSE}
diamonds |> 
  ggplot()
```

### 

Outliers are observations that are unusual; data points that don’t seem to fit the pattern.

### Exercise 14

Within `ggplot()`, using `aes()`, set the `x` argument to `y`. 

```{r variation-14, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r variation-14-hint-1, eval = FALSE}
diamonds |> 
  ggplot(aes(x = ...))
```

### 

Sometimes outliers are data entry errors, sometimes they are simply values at the extremes that happened to be observed in this data collection, and other times they suggest important new discoveries.

### Exercise 15

Add `geom_histogram()` to the pipeline

```{r variation-15, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r variation-15-hint-1, eval = FALSE}
... +
  geom_histogram()
```

### 

When you have a lot of data, outliers are sometimes difficult to see in a histogram.

### Exercise 16

Within `geom_histogram()`, add the `binwidth` argument and set it equal to `0.5`. 

```{r variation-16, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r variation-16-hint-1, eval = FALSE}
... +
  geom_histogram(binwidth = ...)
```

### 

There are so many observations in the common bins that the rare bins are very short, making it very difficult to see them (although maybe if you stare intently at 0 you’ll spot something). To make it easy to see the unusual values, we need to zoom to small values of the y-axis with `coord_cartesian()`.

### Exercise 17

Add `coord_cartesian()` to the pipeline. Within `coord_cartesian()`, add the `ylim` argument and set it equal to `c(0, 50)`. 

```{r variation-17, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r variation-17-hint-1, eval = FALSE}
... + 
  coord_cartesian(ylim = ...)
```

### 

The function, `coord_cartesian()`, also has an `xlim` argument for when you need to zoom into the x-axis. ggplot2 also has `xlim()` and `ylim()` functions that work slightly differently: they throw away the data outside the limits. This allows us to see that there are three unusual values: `0`, `~30`, and `~60`. 

### Exercise 18

Pipe `diamonds` to `filter()` and filter `y` to be less than `3`.


```{r variation-18, exercise = TRUE}

```


```{r variation-18-hint-1, eval = FALSE}
diamonds |> 
  filter(y < ...)
```

```{r, include=FALSE}
diamonds |> 
  filter(y < 3)
```


### 

The `y` variable measures one of the three dimensions of these diamonds, in `mm`. We know that diamonds can’t have a width of 0 mm, so these values must be incorrect. 


### Exercise 19

Modify the pipe by adding, to the `filter()` command,  `|` and `y` greater than `20`.

```{r variation-19, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r variation-19-hint-1, eval = FALSE}
diamonds |> 
  filter(y < 3 | y > ...)
```

```{r, include=FALSE}
diamonds |> 
  filter(y < 3 | y > 20)
```


### 

EDA is an important part of any data analysis, even if the primary research questions are handed to you on a platter, because you always need to investigate the quality of your data.

### Exercise 20

Add `select()` to the pipeline. Within `select()`, select the variables `price`, `x`, `y`, and `z`.

```{r variation-20, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r variation-20-hint-1, eval = FALSE}
... |> 
  select(price, x, ...)
```

### Exercise 21

Add `arrange()` to the pipeline. Within `arrange()`, add the variable `y`  

```{r variation-21, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r variation-21-hint-1, eval = FALSE}
... |> 
  arrange(...)
```

This arranges the data by the width of the diamonds. By doing EDA, we have discovered missing data that was coded as `0`, which we never would have found by simply searching for `NAs`. Going forward we might choose to re-code these values as `NAs` in order to prevent misleading calculations. We might also suspect that measurements of `32mm` and `59mm` are implausible: those diamonds are over an inch long, but don’t cost hundreds of thousands of dollars!


## Unusual Values
### 

In this section we will learn about the ways to deal with unusual values in a dataset.

### 

One way is to drop the entire row with the strange values.

### 

By the end of this section we will be creating the plot below:

```{r}
p1 <- nycflights13::flights |> 
  mutate(
    cancelled = is.na(dep_time),
    sched_hour = sched_dep_time %/% 100,
    sched_min = sched_dep_time %% 100,
    sched_dep_time = sched_hour + (sched_min / 60)
  ) |> 
  ggplot(aes(x = sched_dep_time)) + 
  geom_freqpoly(aes(color = cancelled), binwidth = 1/4) + 
  labs(title = "Frequency of Cancelled and Non-cancelled Flights for Scheduled Flights",
       subtitle = "There are a lot of non-cancelled flights at 3 PM",
       x = "Scheduled Departure Time",
       y = "Frequency",
       color = "Cancelled or Non-Cancelled")

p1
```

### Exercise 1

With a new pipeline, pipe `diamonds` to `filter()`. Within `filter()`, type `between(y, 3, 20)`. 

```{r unusual-values-1, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r unusual-values-1-hint-1, eval = FALSE}

```

### 

This just keeps the rows with the normal values and drops all of the rows with the strange values. We don’t recommend this option because one invalid value doesn’t imply that all the other values for that observation are also invalid. Additionally, if you have low quality data, by the time that you’ve applied this approach to every variable you might find that you don’t have any data left!

### 

### Exercise 2

Create a new variable called `diamonds2` and set it equal to `diamonds`.


```{r unusual-values-2, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r unusual-values-2-hint-1, eval = FALSE}
diamonds2 <- ...
```

### 

The function `if_else()` is a vectorized *if-else*. Compared to the base R equivalent, `ifelse()`, this function allows you to handle missing values in the `condition` with `missing` and always takes `true`, `false`, and `missing` into account when determining what the output type should be.

### Exercise 3

Copy your previous code and pipe `diamonds` to `mutate()`. Within `mutate()`, create a new variable, `y`, and set it equal to `if_else(y < 3 | y > 20, NA, y)`.

```{r unusual-values-3, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r unusual-values-3-hint-1, eval = FALSE}
diamonds2 <- diamonds |> 
  mutate(y_new = ...)
```

### 

Here we replaced the unusual values with missing values. Here, we use the `mutate()` function to replace the variable with a modified copy. We used the `if_else()` function to replace unusual values with `NA`:

### Exercise 4

Add `ggplot()` to the pipeline.

```{r unusual-values-4, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r unusual-values-4-hint-1, eval = FALSE}
... |> 
  ggplot()
```

### Exercise 5

Within `ggplot()`, using `aes()`, set the `x` argument to `x` and the `y` argument to `y_new`.

```{r unusual-values-5, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r unusual-values-5-hint-1, eval = FALSE}
... +
  ggplot(aes(x = ..., y = ...))
```

### Exercise 6

Add `geom_point()` to the pipeline.

```{r unusual-values-6, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r unusual-values-6-hint-1, eval = FALSE}
... +
  geom_point()
```

### 

It’s not obvious where you should plot missing values, so **ggplot2** doesn’t include them in the plot, but it does warn that they’ve been removed. To suppress that warning, set `na.rm = TRUE`.

### Exercise 7

Within `geom_point()`, add the `na.rm` argument and set it equal to `TRUE`. 

```{r unusual-values-7, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r unusual-values-7-hint-1, eval = FALSE}
... + 
  geom_point(na.rm = ...)
```

### 

### Exercise 8

Add a `title`, `subtitle`, `x` axis title, and a `y` axis title using `labs()`. 

```{r unusual-values-8, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r unusual-values-8-hint-1, eval = FALSE}
... + 
  labs(
    title = ...,
    subtitle = ...,
    x = ...,
    y = ...
    )
```

### 

Other times you want to understand what makes observations with missing values different to observations with recorded values. For example, in `nycflights13::flights`, missing values in the `dep_time` variable indicate that the flight was cancelled. So you might want to compare the scheduled departure times for cancelled and non-cancelled times. You can do this by making a new variable, using `is.na()` to check if `dep_time` is missing. Let's make a plot using `geom_freqpoly()`. 

### Exercise 9

With a new pipeline, pipe `nycflights13::flights` to `mutate()`. Within `mutate()`, create a new variable called `cancelled` and set it equal to `is.na(dep_time)`.

```{r unusual-values-9, exercise = TRUE}

```

```{r unusual-values-9-hint-1, eval = FALSE}
nycflights13::flights |> 
  mutate(
    cancelled = ...
  )
```

### 

In the dataset `nycflights13::flights1`, missing values in the `dep_time` variable indicate that the flight was cancelled. So you might want to compare the scheduled departure times for cancelled and non-cancelled times. You can do this by making a new variable, using `is.na()` to check if `dep_time` is missing.


### Exercise 10

Within `mutate()`, create the `sched_hour` variable and set it equal to `sched_dep_time %/% 100`. 

```{r unusual-values-10, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r unusual-values-10-hint-1, eval = FALSE}
nycflights13::flights |> 
  mutate(
    cancelled = is.na(dep_time),
    sched_hour = ...
  )
```

### 

Data cleaning is just one application of EDA: you ask questions about whether your data meets your expectations or not. To do data cleaning, you’ll need to deploy all the tools of EDA: visualization, transformation, and modelling.

### Exercise 11

Within `mutate()`, create the `sched_min` variable and set it equal to `sched_dep_time %% 100`.

```{r unusual-values-11, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r unusual-values-11-hint-1, eval = FALSE}
nycflights13::flights |> 
  mutate(
    cancelled = is.na(dep_time),
    sched_hour = sched_dep_time %/% 100,
    sched_min = ...
  )
```

### 

*“Far better an approximate answer to the right question, which is often vague, than an exact answer to the wrong question, which can always be made precise.” — John Tukey*

### Exercise 12

Within `mutate()`, create the `sched_dep_time` variable and set it equal to `sched_hour` plus `(sched_min / 60)`.

```{r unusual-values-12, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r unusual-values-12-hint-1, eval = FALSE}
nycflights13::flights |> 
  mutate(
    cancelled = is.na(dep_time),
    sched_hour = sched_dep_time %/% 100,
    sched_min = sched_dep_time %% 100,
    sched_dep_time = ...
  )
```

### 

EDA is fundamentally a creative process. And like most creative processes, the key to asking quality questions is to generate a large quantity of questions. It is difficult to ask revealing questions at the start of your analysis because you do not know what insights can be gleaned from your dataset.

### Exercise 13

Add `ggplot()` to the pipeline.

```{r unusual-values-13, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r unusual-values-13-hint-1, eval = FALSE}
... + 
  ggplot()
```

### 

Each new question that you ask will expose you to a new aspect of your data and increase your chance of making a discovery. You can quickly drill down into the most interesting parts of your data—and develop a set of thought-provoking questions—if you follow up each question with a new question based on what you find.

### Exercise 14

Within `ggplot()`, using `aes()`, set the `x` argument to the `sched_dep_time` variable. 

```{r unusual-values-14, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r unusual-values-14-hint-1, eval = FALSE}
... + 
  ggplot(aes(x = ...))
```

### 

There is no rule about which questions you should ask to guide your research. However, two types of questions will always be useful for making discoveries within your data. You can loosely word these questions as:

1. What type of variation occurs within my variables?

2. What type of covariation occurs between my variables?

### Exercise 15

Add `geom_freqpoly()` to the pipeline.

```{r unusual-values-15, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r unusual-values-15-hint-1, eval = FALSE}
... + 
  geom_freqpoly()
```

### Exercise 16

Within `geom_freqpoly()`, using `aes()`, set the `color` argument equal to the `cancelled` variable.

```{r unusual-values-16, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r unusual-values-16-hint-1, eval = FALSE}
... + 
  geom_freqpoly(aes(color = ...))
```

### Exercise 17

Within `geom_freqpoly()`, add the `binwidth` argument and set it equal to `1/4`. 

```{r unusual-values-17, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r unusual-values-17-hint-1, eval = FALSE}
... +
  geom_freqpoly(aes(color = cancelled), binwidth = ...)
```

### Exercise 18

Add a `title`, `subtitle`, `x` axis title, and a `y` axis title by adding `labs()` to the pipeline. 

```{r unusual-values-18, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r unusual-values-18-hint-1, eval = FALSE}
... + 
  labs(
    title = ...,
    subtitle = ...,
    x = ...,
    y = ...
  )
```

### 

However this plot isn’t great because there are many more non-cancelled flights than cancelled flights. In the next section we’ll explore some techniques for improving this comparison.

### 

Reminder: This is what our graph should look like

```{r}
p1

```

## Covariation
### 

In this section we will be learning how to describe the behavior between variables.

### 

By the end of this section we will have made this graph:
```{r}
p2 <- mpg |> 
  ggplot(aes(x = hwy, y = fct_reorder(class, hwy, median))) +
  geom_boxplot() + 
  labs(
    title = "Variation of Highway Mileage over Classes", 
    subtitle = "The compact and midsized cars are tied for the highest highway mileage.",
    x = "Highway Mileage",
    y = "Car Classes"
  )

p2
```

### Exercise 1

With a new pipeline, pipe `diamonds` to `ggplot()`.

```{r covariation-1, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r covariation-1-hint-1, eval = FALSE}
diamonds |> 
  ...
```

### 

Covariation is the tendency for the values of two or more variables to vary together in a related way. The best way to spot covariation is to visualize the relationship between two or more variables.

### Exercise 2

Within `ggplot()`, using `aes()`, set the `x` argument to the variable `price`. 

```{r covariation-2, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r covariation-2-hint-1, eval = FALSE}
diamonds |> 
  ggplot(aes(x = ...))
```

### 

Every variable has its own pattern of variation, which can reveal interesting information about how that it varies between measurements on the same observation as well as across observations. The best way to understand that pattern is to visualize the distribution of the variable’s values.

### Exercise 3

Add `geom_freqpoly()` to the pipeline.

```{r covariation-3, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r covariation-3-hint-1, eval = FALSE}
... + 
  geom_freqpoly()
```

### 

The function, `geom_freqpoly()`, visualizes the distribution of a single continuous variable by dividing the x-axis into bins and counting the number of observations in each bin. Frequency polygons display he counts with lines. Frequency polygons are more suitable when you want to compare the distribution across the levels of a categorical variable.

### Exercise 4

Within `geom_freqpoly()`, using `aes()`, set the `color` argument to the `cut` variable. 

```{r covariation-4, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r covariation-4-hint-1, eval = FALSE}
... + 
  geom_freqpoly(aes(color = ...))
```

### 

Note that ggplot2 uses an ordered color scale for `cut` because it’s defined as an ordered factor variable in the data.

### Exercise 5

Within `geom_freqpoly()`, add the `binwidth` argument and set it equal to `500`. 

```{r covariation-5, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r covariation-5-hint-1, eval = FALSE}
... + 
  geom_freqpoly(aes(color = cut), binwidth = ...)
```

### Exercise 6

Within `geom_freqpoly()`, add the `linewidth` argument and set it equal to `0.75`. 

```{r covariation-6, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r covariation-6-hint-1, eval = FALSE}
... + 
  geom_freqpoly(aes(color = cut), binwidth = 500, linewidth = ...)
```

### 

The default appearance of `geom_freqpoly()` is not that useful here because the height, determined by the overall count, differs so much across `cut`s, making it hard to see the differences in the shapes of their distributions. To make the comparison easier we need to swap what is displayed on the y-axis. Instead of displaying count, we’ll display the **density**, which is the count standardized so that the area under each frequency polygon is one.

### Exercise 7

Within `ggplot()`, within `aes()`, set the `y` argument to `after_stat(density)`. 

```{r covariation-7, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r covariation-7-hint-1, eval = FALSE}
diamonds |> 
  ggplot(aes(x = price, y = ...) + 
  geom_freqpoly(aes(color = cut), binwidth = 500, linewidth = 0.75)
```

### 

Note that we’re mapping the `density` to `y`, but since `density` is not a variable in the `diamonds` dataset, we need to first calculate it. We use the `after_stat()` function to do so.

### Exercise 8

Add a `title`, `subtitle`, `x` axis title, a `y` axis title, and a legend title (`color`) by adding `labs()` to the pipeline. 

```{r covariation-8, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r covariation-8-hint-1, eval = FALSE}
... + 
  labs(
    title = ...,
    subtitle = ...,
    x = ...,
    y = ...
  )
```

### 

There’s something rather surprising about this plot - it appears that fair diamonds (the lowest quality) have the highest average price! But maybe that’s because frequency polygons are a little hard to interpret - there’s a lot going on in this plot. A visually simpler plot for exploring this relationship is using side-by-side boxplots.

### Exercise 9

With a new pipeline, pipe `diamonds` to `ggplot()`. 

```{r covariation-9, exercise = TRUE}

```

```{r covariation-9-hint-1, eval = FALSE}
diamonds |> 
  ggplot()
```

### 

Note that this plot is blank.

### Exercise 10

Within `ggplot()`, using `aes()`, set the `x` argument to the `cut` variable. 

```{r covariation-10, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r covariation-10-hint-1, eval = FALSE}
diamonds |> 
  ggplot(aes(x = ...))
```

### 

Note that this plot should only have an x-axis labelled.

### Exercise 11

Within `ggplot()`, using `aes()`, set the `y` argument to the `price` variable. 

```{r covariation-11, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r covariation-11-hint-1, eval = FALSE}
diamonds |> 
  ggplot(aes(x = cut, y = ...))
```

### Exercise 12

Add `geom_boxplot()` to the pipeline.

```{r covariation-12, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r covariation-12-hint-1, eval = FALSE}
... +
  geom_boxplot()
```

### 

We see much less information about the distribution, but the boxplots are much more compact so we can more easily compare them (and fit more on one plot). It supports the counter-intuitive finding that better quality diamonds are typically cheaper! In the exercises, you’ll be challenged to figure out why.

### Exercise 13

Add a `title`, `subtitle`, `x` axis title, and a `y` axis title by adding `labs()` to the pipeline. 

```{r covariation-13, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r covariation-13-hint-1, eval = FALSE}
... + 
  labs(
    title = ...,
    subtitle = ...,
    x = ...,
    y = ...
  )
```

### 

`cut` is an ordered factor: fair is worse than good, which is worse than very good and so on. Many categorical variables don’t have such an intrinsic order, so you might want to reorder them to make a more informative display. One way to do that is with `fct_reorder()`. Let's use a new dataset.

### Exercise 14

Type in `mpg` and hit "Run Code". 

```{r covariation-14, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r covariation-14-hint-1, eval = FALSE}
mpg
```

### 

This dataset contains a subset of the fuel economy data that the EPA makes available on https://fueleconomy.gov/. It contains only models which had a new release every year between 1999 and 2008 - this was used as a proxy for the popularity of the car.

### Exercise 15

With a new pipeline, pipe `mpg` to `ggplot()`. 

```{r covariation-15, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r covariation-15-hint-1, eval = FALSE}
mpg |> 
  ggplot()
```

### 

Note that this plot is blank. 

### Exercise 16

Within `ggplot()`, using `aes()`, set the `x` argument to the `class` variable.

```{r covariation-16, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r covariation-16-hint-1, eval = FALSE}
mpg |> 
  ggplot(aes(x = ...))
```

### Exercise 17

Within `ggplot()`, using `aes()`, set the `y` argument to `hwy`.

```{r covariation-17, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r covariation-17-hint-1, eval = FALSE}
mpg |> 
  ggplot(aes(x = class, y = ...))
```

### Exercise 18

Add `geom_boxplot()` to the pipeline.

```{r covariation-18, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r covariation-18-hint-1, eval = FALSE}
... + 
  geom_boxplot()
```

### 

The trend is pretty hard to see, so to make it easier on the eyes, we can reorder `class` based on the `median` value of `hwy`. 

### Exercise 19

Within `ggplot()`, within `aes()`, set the `x` argument to `fct_reorder(class, hwy, median)`.

```{r covariation-19, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r covariation-19-hint-1, eval = FALSE}

```

### 

The function, `fct_reorder()`, is useful for 1d displays where the factor is mapped to position and the function `fct_reorder2()` is for 2d displays where the factor is mapped to a non-position aesthetic. Since we have long variable names, `geom_boxplot()` will work better if you flip it `90°`. You can do that by exchanging the `x` and `y` aesthetic mappings.

### Exercise 20

Within `ggplot()`, within `aes()`, swap the aesthetic mappings of `x` and `y`.

```{r covariation-20, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r covariation-20-hint-1, eval = FALSE}
mpg |> 
  ggplot(aes(x = ..., y = ...)) + 
  ...
```

### 

If you have a small dataset, it’s sometimes useful to use `geom_jitter()` to avoid overplotting to more easily see the relationship between a continuous and categorical variable. The [**ggbeeswarm**](https://r-charts.com/distribution/ggbeeswarm/) package provides a number of methods similar to `geom_jitter()`.

### Exercise 21

Add a `title`, `subtitle`, `x` axis title, and a `y` axis title by adding `labs()` to the pipeline. 

```{r covariation-21, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r covariation-21-hint-1, eval = FALSE}
... + 
  labs(
    title = ...,
    subtitle = ...,
    x = ...,
    y = ...
  )
```

### 

Reminder this is what our plot should look like: 

```{r}
p2
```

## Covariation with Two Categorical Variables
### 

In this section we will learn how to visualize the covariation between categorical variables. 

### 

By the end of this section, we will be creating this plot:

```{r}

p3 <- diamonds |> 
  count(color, cut) |>  
  ggplot(aes(x = color, y = cut)) +
  geom_tile(aes(fill = n)) +
  labs(
    title = "Covariation between Diamond Cut and the Color of the Diamond",
    subtitle = "There are more ideally cut diamonds than other cut types of diamonds",
    x = "Color of diamond",
    y = "Cut of diamond"
  )

p3

```

### Exercise 1

With a new pipeline, pipe the `diamonds` dataset to `ggplot()`.

```{r covariation-with-two-1, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r covariation-with-two-1-hint-1, eval = FALSE}
diamonds |> 
  ggplot()
```

### 

Note that this plot is empty.

### Exercise 2

Within `ggplot()`, using `aes()`, set the `x` argument to the `cut` variable.

```{r covariation-with-two-2, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r covariation-with-two-2-hint-1, eval = FALSE}
diamonds |> 
  ggplot(aes(x = ...))
```

### 

To visualize the covariation between categorical variables, you’ll need to count the number of observations for each combination of levels of these categorical variables.

### Exercise 3

Within `ggplot()`, using `aes()`, set the `y` argument to the `color` variable.

```{r covariation-with-two-3, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r covariation-with-two-3-hint-1, eval = FALSE}
diamonds |> 
  ggplot(aes(x = cut, y = ...))
```

### 

One way to visualize two categorical variables is by using the function `geom_count()`. This function counts the number of observations at each location.

### Exercise 4

Add `geom_count()` to the pipeline.

```{r covariation-with-two-4, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r covariation-with-two-4-hint-1, eval = FALSE}
... + 
  geom_count()
```

### 

The size of each circle in the plot displays how many observations occurred at each combination of values. Covariation will appear as a strong correlation between specific x values and specific y values.

### Exercise 5

Let's make our plot fancy by adding a `title`, `subtitle`, `x` axis title, a `y` axis title by adding `labs()` to the pipeline.

```{r covariation-with-two-5, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r covariation-with-two-5-hint-1, eval = FALSE}
... + 
  labs(title = ...,
       subtitle = ..., 
       x = ...,
       y = ...)
```

### 

Another approach for exploring the relationship between these variables is computing the counts with **dplyr**.

### Exercise 6

With a new pipeline, pipe the `diamonds` dataset to `count()`. Within `count` add the variables `color` and `cut`. 

```{r covariation-with-two-6, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r covariation-with-two-6-hint-1, eval = FALSE}
diamonds |> 
  count(color, ...)
```

### 

Now, let's visualize this with `geom_tile()` and the fill aesthetic.

### Exercise 7

Add `ggplot()` to the pipeline.

```{r covariation-with-two-7, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r covariation-with-two-7-hint-1, eval = FALSE}
... +
  ggplot()
```

### 

Note that this plot is empty.

### Exercise 8

Within `ggplot()`, using `aes()`, set the `x` argument to the `color` variable.

```{r covariation-with-two-8, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r covariation-with-two-8-hint-1, eval = FALSE}
... +
  ggplot(aes(x = ...))
```

### Exercise 9

Within `ggplot()`, using `aes()`, set the `y` argument to the `cut` variable. 

```{r covariation-with-two-9, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r covariation-with-two-9-hint-1, eval = FALSE}
... + 
  ggplot(aes(x = color, y = ...))
```

### 

If the categorical variables are unordered, you might want to use the [**seriation**](https://cran.r-project.org/web/packages/seriation/index.html) package to simultaneously reorder the rows and columns in order to more clearly reveal interesting patterns.

### Exercise 10

Add `geom_tile()` to the pipeline.

```{r covariation-with-two-10, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r covariation-with-two-10-hint-1, eval = FALSE}
... + 
  geom_tile()
```

### Exercise 11

Within `geom_tile()`, using `aes()`, set the `fill` argument to the `n` variable.

```{r covariation-with-two-11, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r covariation-with-two-11-hint-1, eval = FALSE}
... + 
  geom_tile(aes(fill = ...))
```

### 

For larger plots, you might want to try the [**heatmaply**](https://cran.r-project.org/web/packages/heatmaply/vignettes/heatmaply.html) package, which creates interactive plots.

### Exercise 12

Add a `title`, `subtitle`, `x` axis title, and a `y` axis title by adding `labs()` to the pipeline.

```{r covariation-with-two-12, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r covariation-with-two-12-hint-1, eval = FALSE}
... + 
  labs(title = ...,
       subtitle = ...,
       x = ...,
       y = ...)

```

### 

Reminder this is what your plot should look like: 

```{r}
p3

```

## Covariation with 2 Numerical Variables
### 

In this section, we will learn how to visualize covariation between two numerical variables. 

### 

By the end of this section, we will have created this plot:

```{r}
p4 <- smaller |> 
  ggplot( aes(x = carat, y = price)) + 
  geom_boxplot(aes(group = cut_width(carat, 0.1))) +
  labs(title = "Covariation between Number of Carats and Price of Diamonds",
       subtitle = "As the number of carats of a diamond increases the price also increases",
       x = "Number of Carats",
       y = "Price of the Diamond")

p4

```

### Exercise 1

Type in `smaller` and hit "Run Code".

```{r covariation-with-2-n-1, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r covariation-with-2-n-1-hint-1, eval = FALSE}
smaller
```

### 

This dataframe shows diamonds that only have less than 3 carats. 

### Exercise 2

With a new pipeline, pipe the `smaller` dataset to `ggplot()`. 

```{r covariation-with-2-n-2, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r covariation-with-2-n-2-hint-1, eval = FALSE}
smaller |> 
  ggplot()
```

### 

Note that this plot is empty.

### Exercise 3

Within `ggplot()`, using `aes()`, set the `x` argument to the `carat` variable.

```{r covariation-with-2-n-3, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r covariation-with-2-n-3-hint-1, eval = FALSE}
smaller |> 
  ggplot(aes(x = ...))
```

### 

### Exercise 4

Within `ggplot()`, using `aes()`, set the `y` argument to the `price` variable.

```{r covariation-with-2-n-4, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r covariation-with-2-n-4-hint-1, eval = FALSE}
smaller |> 
  ggplot(aes(x = carat, y = ...))
```

### 

Scatterplots become less useful as the size of your dataset grows, because points begin to overplot, and pile up into areas of uniform black, making it hard to judge differences in the density of the data across the 2-dimensional space as well as making it hard to spot the trend.

### Exercise 5

Add `geom_point()` to the pipeline.

```{r covariation-with-2-n-5, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r covariation-with-2-n-5-hint-1, eval = FALSE}
... + 
  geom_point()
```

### 

Let's use the `alpha` aesthetic to add transparency to the plot. 

### Exercise 6

Within `geom_point()`, set the `alpha` argument to `0.01`. 

```{r covariation-with-2-n-6, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r covariation-with-2-n-6-hint-1, eval = FALSE}
... + 
  geom_point(alpha = 0.01)
```

### 

But using transparency can be challenging for very large datasets. Another solution is to use bin. Previously you used `geom_histogram()` and `geom_freqpoly()` to bin in one dimension. Now you’ll learn how to use `geom_bin2d()` and `geom_hex()` to bin in two dimensions.

### Exercise 7

Replace `geom_point()` with `geom_bin2d()`.

```{r covariation-with-2-n-7, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r covariation-with-2-n-7-hint-1, eval = FALSE}
... + 
  geom_bin2d()
```

### 

geom_bin2d() and geom_hex() divide the coordinate plane into 2d bins and then use a fill color to display how many points fall into each bin. geom_bin2d() creates rectangular bins.


### Exercise 8

Load in the `hexbin` package using `library()`. 

```{r covariation-with-2-n-8, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r covariation-with-2-n-8-hint-1, eval = FALSE}
library(...)
```

### 

The function `geom_hex()` creates hexagonal bins. You will need to install the [**hexbin**](https://cran.r-project.org/web/packages/hexbin/index.html) package to use `geom_hex()`.

### Exercise 9

Replace `geom_bin2d()` in the pipeline with `geom_hex()`. 

```{r covariation-with-2-n-9, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r covariation-with-2-n-9-hint-1, eval = FALSE}
smaller |> 
  ggplot(aes(x = carat, y = price)) +
  ...
```

### 

Another option is to bin one continuous variable so it acts like a categorical variable. Then you can use one of the techniques for visualizing the combination of a categorical and a continuous variable that you learned about. For example, you could bin carat and then for each group, display a boxplot.


### Exercise 10

Replace `geom_hex()` with `geom_boxplot()` in the previous pipeline.

```{r covariation-with-2-n-10, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r covariation-with-2-n-10-hint-1, eval = FALSE}
... +
  geom_boxplot()
```

### Exercise 11

Within `geom_boxplot()`, using `aes()`, set the `group` argument to `cut_width(carat, 0.1)`. 

```{r covariation-with-2-n-11, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r covariation-with-2-n-11-hint-1, eval = FALSE}
... +
  geom_boxplot(aes(group = ...))
```

### 

cut_width(x, width), as used above, divides x into bins of width width. By default, boxplots look roughly the same (apart from number of outliers) regardless of how many observations there are, so it’s difficult to tell that each boxplot summaries a different number of points.

### Exercise 12

Add a `title`, `subtitle`, `x` axis title, and a `y` axis title by adding `labs()` to the pipeline. 

```{r covariation-with-2-n-12, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r covariation-with-2-n-12-hint-1, eval = FALSE}
... + 
  labs(
    title = ...,
    subtitle = ...,
    x = ...,
    y = ...,
  )
```

### 

This is a reminder that our plot should look like this:

```{r}
p4
```

## Patterns and Models
### 

In this section, we will discuss how to find and display patterns in your datasets. 

### 

By the end of this section, we will be creating this plot:

```{r}
p5 <- diamonds_aug |> 
  ggplot(aes(x = cut, y = .resid)) + 
  geom_boxplot() + 
  labs(
    title = "Prediction Model between Carat and Residual Price",
    subtitle = "As the number of carat increases the Residual Price decreases",
    x = "Number of Carats",
    y = "Residual Price"
  )

p5
```

### Exercise 1

First, let's load in the `tidymodels` package by using `library()`.

```{r patterns-and-models-1, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r patterns-and-models-1-hint-1, eval = FALSE}
library(...)
```

### 

If a systematic relationship exists between two variables it will appear as a pattern in the data. If you spot a pattern, you can ask yourself the following questions:

Could this pattern be due to coincidence (i.e. random chance)?

How can you describe the relationship implied by the pattern?

How strong is the relationship implied by the pattern?

What other variables might affect the relationship?

Does the relationship change if you look at individual subgroups of the data?

### Exercise 2

Create a new variable `diamonds3` and set it equal to `diamonds`. 

```{r patterns-and-models-2, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r patterns-and-models-2-hint-1, eval = FALSE}
diamonds3 <- diamonds
```

### 

Patterns in your data provide clues about relationships, i.e., they reveal covariation. If you think of variation as a phenomenon that creates uncertainty, covariation is a phenomenon that reduces it.

### Exercise 3

Using your previous code, pipe `diamonds` to `mutate()`. Within `mutate()`, create a new variable `log_price` and set it equal to `log(price)`. 

```{r patterns-and-models-3, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r patterns-and-models-3-hint-1, eval = FALSE}
diamonds3 <- diamonds |> 
  mutate(
    log_price = ...
  )
```

### 

If two variables covary, you can use the values of one variable to make better predictions about the values of the second. If the covariation is due to a causal relationship (a special case), then you can use the value of one variable to control the value of the second.

### Exercise 4

Within `mutate()`, create another variable called `log_carat` and set it equal to `log(carat)`. 

```{r patterns-and-models-4, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r patterns-and-models-4-hint-1, eval = FALSE}
diamonds3 <- diamonds |> 
  mutate(
    log_price = log(price),
    log_carat = ...
  )
```

### 

We will be using this dataset to make a model that predicts `price` from`carat` and then computes the residuals (the difference between the predicted value and the actual value).

### Exercise 5

Create a new variable called `diamonds_fit` and set it equal to `linear_reg()`. 

```{r patterns-and-models-5, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r patterns-and-models-5-hint-1, eval = FALSE}
diamonds_fit <- ...
```

### 

Models are a tool for extracting patterns out of data. For example, consider the diamonds data. It’s hard to understand the relationship between cut and price, because cut and carat, and carat and price are tightly related.

### Exercise 6

Using your previous code, pipe `linear_reg()` with the following code `fit(log_price ~ log_carat, data = diamonds3)`.

```{r patterns-and-models-6, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r patterns-and-models-6-hint-1, eval = FALSE}
diamonds_fit <- linear_reg() |> 
  ...
```

### 

The functon [`linear_reg()`](https://parsnip.tidymodels.org/reference/linear_reg.html) defines a model that can predict numeric values from predictors using a linear function. This function can fit regression models.There are different ways to fit this model, and the method of estimation is chosen by setting the model `engine`.

### Exercise 7

Create a new variable called `diamonds_aug` and set it equal to `augment(diamonds_fit, new_data = diamonds3)`. 

```{r patterns-and-models-7, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r patterns-and-models-7-hint-1, eval = FALSE}
diamonds_aug <- ...
```

### 

The residuals give us a view of the price of the diamond, once the effect of carat has been removed.

### Exercise 8

Using your previous code, pipe `augment(diamonds_fit, new_data = diamonds3)` with `mutate()`. Within `mutate()`, create a new variable `.resid` and set it equal to `exp(.resid)`. 

```{r patterns-and-models-8, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r patterns-and-models-8-hint-1, eval = FALSE}
diamonds_aug <- augment(diamonds_fit, new_data = diamonds) |>
  ...
```

### 

Note that instead of using the raw values of `price` and `carat`, we log transform them first, and fit a model to the log-transformed values. Then, we exponentiate the residuals to put them back in the scale of raw prices.

### Exercise 9

With a new pipeline, pipe `diamonds_aug` with `ggplot()`.

```{r patterns-and-models-9, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r patterns-and-models-9-hint-1, eval = FALSE}
diamonds_aug |> 
  ggplot()
```

### 

Note that this plot is empty

### Exercise 10

Within `ggplot()`, using `aes()`, set the `x` argument to the `carat` variable.

```{r patterns-and-models-10, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r patterns-and-models-10-hint-1, eval = FALSE}
diamonds_aug |> 
  ggplot(aes(x = ...))
```

### Exercise 11

Within `ggplot()`, using `aes()`, set the `y` argument to the `.resid` variable.

```{r patterns-and-models-11, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r patterns-and-models-11-hint-1, eval = FALSE}
diamonds_aug |> 
  ggplot(aes(x = carat, y = ...))
```

### Exercise 12

Add `geom_point()` to the pipeline.

```{r patterns-and-models-12, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r patterns-and-models-12-hint-1, eval = FALSE}
... +
  geom_point()
```

### 

Once you’ve removed the strong relationship between carat and price, you can see what you expect in the relationship between cut and price: relative to their size, better quality diamonds are more expensive.

### Exercise 13

Replace `geom_point()` from the previous pipeline and change it to `geom_boxplot()`. 

```{r patterns-and-models-13, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r patterns-and-models-13-hint-1, eval = FALSE}

```

### 

A boxplot is the best way to see the relationship between the type of diamond cut and the price of the diamond. 

### Exercise 14

Add a `title`, `subtitle`, `x` axis title, and a `y` axis title by adding `labs()` to the pipeline.

```{r patterns-and-models-14, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r patterns-and-models-14-hint-1, eval = FALSE}
... +
  labs(
    title = ...,
    subtitle = ...,
    x = ...,
    y = ...
  )
```

### 

Reminder this is what our plot should look like.

```{r}
p5
```

## Summary
### 

This tutorial covered [Chapter 11: Exploratory data analysis](https://r4ds.hadley.nz/eda.html) from [*R for Data Science (2e)*](https://r4ds.hadley.nz/) by Hadley Wickham, Mine Çetinkaya-Rundel, and Garrett Grolemund. We made use of the [**ggplot2**](https://ggplot2.tidyverse.org/) package and learned about Variation and Patterns. Some key commands that we used are [`coord_cartesian()`](https://ggplot2.tidyverse.org/reference/coord_cartesian.html) which sets limits on the coordinate system and will zoom the plot, [`geom_freqpoly()`](https://ggplot2.tidyverse.org/reference/geom_histogram.html) which visualizes the distribution of a single continuous variable by dividing the x axis into bins and counts the number of observations in each bin, and [`geom_count()`](https://ggplot2.tidyverse.org/reference/geom_count.html) which is a variant of `geom_point()` that counts the number of observations at each location, that maps the count to point area. 



```{r download-answers, child = system.file("child_documents/download_answers.Rmd", package = "tutorial.helpers")}
```
