---
title: Hierarchical data
author: David Kane
tutorial:
  id: hierarchical-data
output:
  learnr::tutorial:
    progressive: yes
    allow_skip: yes
runtime: shiny_prerendered
description: 'Tutorial for Chapter 24: Hierarchical data'
---

```{r setup, include = FALSE}
library(learnr)
library(tutorial.helpers)
library(tidyverse)

knitr::opts_chunk$set(echo = FALSE)
options(tutorial.exercise.timelimit = 60, 
        tutorial.storage = "local") 


```

```{r copy-code-chunk, child = system.file("child_documents/copy_button.Rmd", package = "tutorial.helpers")}
```

```{r info-section, child = system.file("child_documents/info_section.Rmd", package = "tutorial.helpers")}
```

## Introduction
### 

This tutorial covers [Chapter 24: Hierarchical data](https://r4ds.hadley.nz/rectangling.html) from [*R for Data Science (2e)*](https://r4ds.hadley.nz/) by Hadley Wickham, Mine Çetinkaya-Rundel, and Garrett Grolemund. You will learn how to work with non-rectanglar data using packages like [**jsonlite**](https://cran.r-project.org/web/packages/jsonlite/vignettes/json-aaquickstart.html). 


## Lists
###

A list is the object which contains various data types such as logicals, integers, characters, and even another list!


### Exercise 1

Create a list with three items `a`, `b`, and `c`. Have `a` take on the values 1, 2, and 3. Have `b` take on the values 4, 5, and 6. Have `c` take on the values  7, 8, and 9. Assign your list to an object named `my_list`. 


```{r lists-1, exercise = TRUE}

```

```{r lists-1-hint-1, eval = FALSE}
Use the c() function to create the individual vectors for a, b, and c. Use the list() function to create a list with objects a, b, and c.
```

```{r lists-1-hint-2, eval = FALSE}
list(a = c(...), b = c(...), c = c(...))
```

### Exercise 2

Now, use `str()` with the argument `my_list`. 

```{r lists-2, exercise = TRUE}
my_list <- list(a = c(1, 2, 3), b = c(4, 5, 6), c = c(7, 8, 9))

```

### Exercise 3

Let's now extract a single component `a` from `my_list` using `[ ]`. The code for the creation of `my_list` has been included for you. 

```{r lists-3, exercise = TRUE}
my_list <- list(a = c(1, 2, 3), b = c(4, 5, 6), c = c(7, 8, 9))

```

```{r lists-3-hint-1, eval = FALSE}
You need to use quotation marks because a is a character.
```

```{r lists-3-hint-2, eval = FALSE}
str(my_list[" "])
```

### Exercise 4

Now, extract the number 5 from `my_list` using `[[]]` and `[ ]`.

```{r lists-4, exercise = TRUE}
my_list <- list(a = c(1, 2, 3), b = c(4, 5, 6), c = c(7, 8, 9))

```

```{r lists-4-hint-1, eval = FALSE}
Refer to the name of the list item in the double 
brackets, and the position in that item in the single 
brackets.
```

```{r lists-4-hint-2, eval = FALSE}
my_list <- list(a = c(1, 2, 3), b = c(4, 5, 6), c = c(7, 8, 9))
my_list[["b"]][...]
```

## JSON files 
### 

An increasingly common format for sharing data is JavaScript Object Notation or JSON. 

JSON is a standard text-based format for representing structured data based on JavaScript object syntax.

### Exercise 1

Consider the contents of the file `test_1.json`:

```{r}
cat(readLines("data/test_1.json"), sep = "\n")
```

### 

Write code for reading this JSON file into R as a tibble. Use the `fromJSON()` function from the [**jsonlite**](https://cran.r-project.org/web/packages/jsonlite/vignettes/json-aaquickstart.html) package.

```{r json-files-1, exercise = TRUE}

```

```{r json-files-1-hint-1, eval = FALSE}
fromJSON("data/test_1.json")
```

### 

If there are "null" values in the file, you can specify how to encode those values with the `null` argument in `fromJSON`.


### Exercise 2

Consider the following tibble, which is saved in the environment as an object called `example_1`.

```{r}
tibble(name= c("Miguel", "Sofia", "Aya", "Cheng"), 
                student_id = 1:4, exam_1 = c(85, 94, 87, 90), 
                exam_2 = c(86, 93, 88, 91))
```

### 

Write code to convert the following tibble into JSON format. Use the `toJSON()` function from the **jsonlite** package.

```{r json-files-2, exercise = TRUE}

```

```{r json-files-2-hint-1, eval = FALSE}
toJSON(example_1)
```

### 

`toJSON()` actually has the same arguments as `fromJSON()`, meaning that you can use `null`, `na`, `pretty` and many more arguments to customize the results.

### Exercise 3

After converting from a tibble to the JSON format, it is hard to read the JSON format. To fix this, use the previous code and set the second argument `pretty` equal to `TRUE`.

```{r json-files-3, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r json-files-3-hint-1, eval = FALSE}
toJSON(example_1, pretty = ...)
```

### 

The `pretty` argument comes from the [**prettify**](https://www.rdocumentation.org/packages/papeR/versions/1.0-5/topics/prettify) package, run `??prettify` to learn more about it.


### Exercise 4

Try reading from the `json` object which has already been loaded with JSON array data. Use `fromJSON()` to display the data in a tibble.

```{r json-files-4, exercise = TRUE}

```

```{r json-files-4-hint-1, eval = FALSE}
fromJSON(json)
```

### 

As you can see, `fromJSON()` can display arrays as well, which can be useful for datasets that include matrices. You can do math operations with them as well! 

### Exercise 5

You can also write information like arrays or matrices to a JSON format using `toJSON()`. 

Display a 3x3 array in JSON format by creating 3 arrays structured like `[1, 2, 3]`, containing them in a larger array, and then passing them into `toJSON()`.

```{r json-files-5, exercise = TRUE}

```

```{r json-files-5-hint-1, eval = FALSE}
json2 <- '[[..., ..., ...], [..., ..., ...], [..., ..., ...]]'

```

```{r json-files-5-hint-2, eval = FALSE}
json2 <- '[[..., ..., ...], [..., ..., ...], [..., ..., ...]]'
toJSON(json2)
```

### 

Because JSON files can so easily store complex matrix data, they're often used in APIs and websites. This makes `fromJSON()` invaluable when you're pulling data off the Internet, as the data will often come in a JSON file. 

Vice versa, `toJSON()` is useful for when you want to send data across the Internet due to its light weight (there's very little extra formatting included in the file and it's mostly data).


## Summary
### 

This tutorial covered [Chapter 24: Hierarchical data](https://r4ds.hadley.nz/rectangling.html) from [*R for Data Science (2e)*](https://r4ds.hadley.nz/) by Hadley Wickham, Mine Çetinkaya-Rundel, and Garrett Grolemund. You learned how to work with non-rectanglar data using packages like [**jsonlite**](https://cran.r-project.org/web/packages/jsonlite/vignettes/json-aaquickstart.html).

```{r download-answers, child = system.file("child_documents/download_answers.Rmd", package = "tutorial.helpers")}
```
