---
title: Dates and times
author: David Kane
tutorial:
  id: dates-and-times
output:
  learnr::tutorial:
    progressive: yes
    allow_skip: yes
runtime: shiny_prerendered
description: 'Tutorial for Chapter 18: Dates and times'
---

```{r setup, include = FALSE}
library(learnr)
library(tutorial.helpers)
library(tidyverse)
knitr::opts_chunk$set(echo = FALSE)
options(tutorial.exercise.timelimit = 60, 
        tutorial.storage = "local") 


```

```{r copy-code-chunk, child = system.file("child_documents/copy_button.Rmd", package = "tutorial.helpers")}
```

```{r info-section, child = system.file("child_documents/info_section.Rmd", package = "tutorial.helpers")}
```

## Introduction
### 

This tutorial covers [Chapter 18: Dates and times](https://r4ds.hadley.nz/datetimes.html) from [*R for Data Science (2e)*](https://r4ds.hadley.nz/) by Hadley Wickham, Mine Çetinkaya-Rundel, and Garrett Grolemund. 

## Older material
### 

### Exercise 1

Run `today()` and `now()` in the code chunk below.

```{r older-material-1, exercise = TRUE}

```

### 

`today()` gives today's date, and `now()` gives the time down to the second.

### Exercise 2

<!-- AM: In exercise 2, have the user save `ymd(date_1)` to a new object and check the type of that variable using `typeof()` to verify that it is a proper date-time format. `date_1` is a character while `ymd(date_1)` is a double. -->
<!-- date_1 <- "2020-2-29" -->
<!-- date_1_dt <- ymd(date_1) -->
<!-- typeof(date_1_dt) -->

Now use the function `ymd()` to convert the string `date_1` below into the proper date-time format. 

```{r older-material-2, exercise = TRUE}
date_1 <- "2020-2-29"
```

```{r older-material-2-hint-1, eval = FALSE}
ymd(...)
```

```{r older-material-2-hint-2, eval = FALSE}
date_1 <- "2020-2-29"

ymd(date_1)
```

### 

`ymd()` stands for **y**ear, **m**onth, **d**ate.

### Exercise 3

Assign the output of your previous code to an object called `date_1_dt`. Because both `date_1` and `date_1_dt` appear to be strings, use `typeof()` to verify that they are different object types.

```{r older-material-3, exercise = TRUE}
date_1 <- "2020-2-29"
date_1_dt <- 
```

```{r older-material-3-hint-1, eval = FALSE}
date_1_dt <- ymd(...)

typeof(...)
typeof(...)
```

```{r older-material-3-hint-2, eval = FALSE}
date_1 <- "2020-2-29"
date_1_dt <- ymd(date_1)

typeof(date_1)
typeof(date_1_dt)
```

### 

`date_1` should be a character, and `date_1_dt` should be a double.

### Exercise 4

Now make `date_2` a date-time.  Use the function `ymd_hms()`. Use the argument `tz` to set the timezone to `"EST"`.

```{r older-material-4, exercise = TRUE}
date_2 <- "2017-01-31 20:11:59"

```

```{r older-material-4-hint-1, eval = FALSE}
date_2 <- "2017-01-31 20:11:59"
ymd_hms(..., tz = "...")
```

```{r older-material-4-hint-2, eval = FALSE}
date_2 <- "2017-01-31 20:11:59"
ymd_hms(date_2, tz = "EST")
```

### 

`hms()` stands for **h**our, **m**inute, **s**econd.

### Exercise 5

The following code below selects time-related data from the `flights` data set in the **nycflights13** package. Use `mutate()` to create a new variable `arrival`. `arrival` should be set to the function `make_datetime()` with all selected variables as its argument.

```{r older-material-5, exercise = TRUE}
flights |> 
  select(year, month, day, hour, minute)
```

```{r older-material-5-hint-1, eval = FALSE}
... |> 
  mutate(... = make_datetime(year, ..., ..., ..., ...))
```

```{r older-material-5-hint-2, eval = FALSE}
flights |> 
  select(year, month, day, hour, minute) |>
  mutate(arrival = make_datetime(year, month, day, hour, minute))

```

<!-- AM: This section needs a big rehaul: It's short, confusing, and doesn't help the user learn. -->


### Exercise 6

The function `parse_date()` allows you to parse a date. The function requires the following criteria: a four digit year,  a `-` or `/`, the month, a `-` or `/`, then the day. An example would be "2010-10-01".

### 

Use `parse_date()` to parse the following date: "2009-12-06"

```{r older-material-6, exercise = TRUE}

```

```{r older-material-6-hint-1, eval = FALSE}
parse_date("...")
```

### 

This function is very strict about it's format, there are other functions such as `format_iso_8601()` and `parse_datetime()` which can parse dates and times.

### Exercise 7

Let's have some practice with `parse_datetime()`. It requires a format where the components of a date are organized from biggest to smallest: year, month, day, hour, minute, second.

### 

Use `parse_datetime()` with the argument "2010-10-01T2010".

```{r older-material-7, exercise = TRUE}

```

```{r older-material-7-hint-1, eval = FALSE}
parse_datetime("2010-10-01T2010")
```

### 

The user can also specify which number is the month, day and year by using the `%` symbol which allows more flexibility. Type "?parse_datetime" in the console to see examples.


## Summary
### 

This tutorial covered [Chapter 18: Dates and times](https://r4ds.hadley.nz/datetimes.html) from [*R for Data Science (2e)*](https://r4ds.hadley.nz/) by Hadley Wickham, Mine Çetinkaya-Rundel, and Garrett Grolemund. 

```{r download-answers, child = system.file("child_documents/download_answers.Rmd", package = "tutorial.helpers")}
```
