---
title: "Data transformation"
tutorial:
  id: "data-transformation"
output: 
  learnr::tutorial:
    progressive: true
    allow_skip: true
runtime: shiny_prerendered
description: "Tutorial for Chapter 4: Data transformation"
---

```{r setup, include = FALSE}
library(learnr)
library(dplyr)
knitr::opts_chunk$set(echo = FALSE)
options(tutorial.exercise.timelimit = 60, 
        tutorial.storage = "local") 
```

```{r copy-code-chunk, child = system.file("child_documents/copy_button.Rmd", package = "tutorial.helpers")}
```

```{r info-section, child = system.file("child_documents/info_section.Rmd", package = "tutorial.helpers")}
```



## Introduction
###

This tutorial covers [Chapter 4: Data transformation](https://r4ds.hadley.nz/data-transform.html) from [*R for Data Science (2e)*](https://r4ds.hadley.nz/) by Hadley Wickham, Mine Çetinkaya-Rundel, and Garrett Grolemund. You will learn about the key functions from the [**dplyr**](https://dplyr.tidyverse.org/) package for working with data including `[filter()](https://dplyr.tidyverse.org/reference/filter.html)`,  `[arrange()](https://dplyr.tidyverse.org/reference/arrange.html)`, `[select()](https://dplyr.tidyverse.org/reference/select.html)`, `[mutate()](https://dplyr.tidyverse.org/reference/mutate.html)`, and `[summarize()](https://dplyr.tidyverse.org/reference/summarise.html)`.


## Rows
###

## Summary
###

This tutorial covered [Chapter 4: Data transformation](https://r4ds.hadley.nz/data-transform.html) from [*R for Data Science (2e)*](https://r4ds.hadley.nz/) by Hadley Wickham, Mine Çetinkaya-Rundel, and Garrett Grolemund. You learned about the key functions from the [**dplyr**](https://dplyr.tidyverse.org/) package for working with data including `[filter()](https://dplyr.tidyverse.org/reference/filter.html)`,  `[arrange()](https://dplyr.tidyverse.org/reference/arrange.html)`, `[select()](https://dplyr.tidyverse.org/reference/select.html)`, `[mutate()](https://dplyr.tidyverse.org/reference/mutate.html)`, and `[summarize()](https://dplyr.tidyverse.org/reference/summarise.html)`.


```{r download-answers, child = system.file("child_documents/download_answers.Rmd", package = "tutorial.helpers")}
```
