---
title: "Spreadsheets"
tutorial:
  id: "spreadsheets"
output: 
  learnr::tutorial:
    progressive: true
    allow_skip: true
runtime: shiny_prerendered
description: "Tutorial for Chapter 21: Spreadsheets"
---

```{r setup, include = FALSE}
library(learnr)
library(r4ds.tutorials)
library(tidyverse)
library(readxl)
library(writexl)
knitr::opts_chunk$set(echo = FALSE)
options(tutorial.exercise.timelimit = 60, 
        tutorial.storage = "local") 
```

```{r copy-code-chunk, child = system.file("child_documents/copy_button.Rmd", package = "tutorial.helpers")}
```

```{r info-section, child = system.file("child_documents/info_section.Rmd", package = "tutorial.helpers")}
```


## Introduction
###

This tutorial covers [Chapter 21: Spreadsheets](https://r4ds.hadley.nz/spreadsheets.html) from [*R for Data Science (2e)*](https://r4ds.hadley.nz/) by Hadley Wickham, Mine Çetinkaya-Rundel, and Garrett Grolemund.

## Excel
###

### Exercise 1

**readxl** and **writexl** are the two key packages for working with Excel spreadsheets.



```{r excel-1, exercise = TRUE}

```


```{r excel-1-hint, eval = FALSE}

```

###




## Summary
###

This tutorial covered [Chapter 21: Spreadsheets](https://r4ds.hadley.nz/spreadsheets.html) from [*R for Data Science (2e)*](https://r4ds.hadley.nz/) by Hadley Wickham, Mine Çetinkaya-Rundel, and Garrett Grolemund.

```{r download-answers, child = system.file("child_documents/download_answers.Rmd", package = "tutorial.helpers")}
```

