---
title: Factors
author: David Kane
tutorial:
  id: factors
output:
  learnr::tutorial:
    progressive: yes
    allow_skip: yes
runtime: shiny_prerendered
description: 'Tutorial for Chapter 18: Dates and times'
---

```{r setup, include = FALSE}
library(learnr)
library(tutorial.helpers)
library(tidyverse)
library(primer.data)
knitr::opts_chunk$set(echo = FALSE)
options(tutorial.exercise.timelimit = 60, 
        tutorial.storage = "local") 


```

```{r copy-code-chunk, child = system.file("child_documents/copy_button.Rmd", package = "tutorial.helpers")}
```

```{r info-section, child = system.file("child_documents/info_section.Rmd", package = "tutorial.helpers")}
```

## Introduction
### 

This tutorial covers [Chapter 17: Factors](https://r4ds.hadley.nz/factors.html) from [*R for Data Science (2e)*](https://r4ds.hadley.nz/) by Hadley Wickham, Mine Çetinkaya-Rundel, and Garrett Grolemund. 

## Older material
### 

### Exercise 1

Use `select()` to look at the `ideology` variable of the `cces` data set.

```{r older-material-1, exercise = TRUE}

```

```{r older-material-1-hint-1, eval = FALSE}
cces |> 
  select(...)
```

### 

Look underneath the column name. Notice that this is a factor (`<fctr>`) variable. 


### Exercise 2

Use `str()` with the argument `cces$ideology`. On the line below, use `levels()` with the same argument. 

```{r older-material-2, exercise = TRUE}

```

```{r older-material-2-hint-1, eval = FALSE}
str(...)
levels(...)
```

### Exercise 3

Let's now get a frequency table as a tibble. Use `fct_count()` with the argument `cces$ideology`.

```{r older-material-3, exercise = TRUE}

```

```{r older-material-3-hint-1, eval = FALSE}
fct_count(...)
```

### Exercise 4

Now use `nlevels()` with the argument `cces$ideology` to find the number of existing levels.

```{r older-material-4, exercise = TRUE}

```

```{r older-material-4-hint-1, eval = FALSE}
nlevels(...)
```

### Exercise 5

Now try ordering the levels of `cces$ideology` by frequency. Start a pipe with `cces$ideology` and use the function `fct_infreq()`. Then on the line below, use `levels()` to display the new order of the levels. 

```{r older-material-5, exercise = TRUE}

```

```{r older-material-5-hint-1, eval = FALSE}
cces$ideology |> 
  fct_infreq() 

levels(cces$ideology)
```

### 

Compare this with the old order of the levels in Exercise 2 and the frequency table in Exercise 3.

### Exercise 6

Now `filter()` `cces` to only a few ideologies: `Conservative` , `Moderate`, `Liberal` and `Not Sure`. Save your work to an object named `cces_filt`.

```{r older-material-6, exercise = TRUE}

```

```{r older-material-6-hint-1, eval = FALSE}
Use %in%  and c() within filter(). Within c(), place the ideologies in quotes. 
```

```{r older-material-6-hint-2, eval = FALSE}
cces_filt <- cces |>
  filter(... %in% c(..., ..., ..., ...))
```

The ideologies are in quotes because they are not column names, but strings within the column. 

### Exercise 7

On a new line, use `nlevels()` with the argument `cces_filt$ideology`. The code which creates `cces_filt` has been included for you. Notice that you have  the same number of levels even though you have filtered to include fewer ideologies.

```{r older-material-7, exercise = TRUE}
cces_filt <- cces |>
  filter(ideology %in% c("Conservative", "Moderate", " Liberal","Not Sure"))


```

```{r older-material-7-hint-1, eval = FALSE}
cces_filt <- cces |>
  filter(ideology %in% c("Conservative", "Moderate", " Liberal","Not Sure"))

nlevels(...)
```

<!-- I had to pipe the code from the previous exercise in order to get the desired output. The code in the hint kept the output at 6 levels.
What worked for me was cces_filt <- cces |> filter(ideology %in% c("Conservative" ...)) |> droplevels() -->

### Exercise 8

Let's drop all the unused levels from `cces_filt`. Continue the pipe started for you and add the `droplevels()` function. 


```{r older-material-8, exercise = TRUE}
cces_filt <- cces |>
  filter(ideology %in% c("Conservative", "Moderate", " Liberal","Not Sure")) 

```

```{r older-material-8-hint-1, eval = FALSE}
cces_filt <- cces |>
  filter(ideology %in% c("Conservative", "Moderate", " Liberal","Not Sure")) |> 
  droplevels()
```

### Exercise 9

On a new line, use `nlevels()` with the argument `cces_filt$ideology` to see how many levels exist after you dropped levels. The code which creates `cces_filt` has been included for you. 

```{r older-material-9, exercise = TRUE}
cces_filt <- cces |>
  filter(ideology %in% c("Conservative", "Moderate", " Liberal","Not Sure")) |> 
  droplevels()
  

```

```{r older-material-9-hint-1, eval = FALSE}
cces_filt <- cces |>
  filter(ideology %in% c("Conservative", "Moderate", " Liberal","Not Sure")) |> 
  droplevels()

nlevels(...)
```

## Summary
### 

This tutorial covered [Chapter 17: Factors](https://r4ds.hadley.nz/factors.html) from [*R for Data Science (2e)*](https://r4ds.hadley.nz/) by Hadley Wickham, Mine Çetinkaya-Rundel, and Garrett Grolemund. 

```{r download-answers, child = system.file("child_documents/download_answers.Rmd", package = "tutorial.helpers")}
```
