---
title: Factors
author: David Kane
tutorial:
  id: factors
output:
  learnr::tutorial:
    progressive: yes
    allow_skip: yes
runtime: shiny_prerendered
description: 'Tutorial for Chapter 17: Factors'
---

```{r setup, include = FALSE}
library(learnr)
library(tutorial.helpers)
library(tidyverse)

# Without thishack, we fail on GHA! cols() and col_factor() have conflicting
# versions in readr, vroom and scales. Of course, we want the readr versions. We
# could solve this with either:

# library(conflicted)
# conflict_prefer("cols", "readr") # and so on. Or with:

# library(tidymodels)
# tidymodels_prefer() # because this applies to tidyverse functions we well

# But that all seems like overkill, given that it all works for students
# regardless.

cols <- readr::cols
col_factor <- readr::col_factor

knitr::opts_chunk$set(echo = FALSE)
options(tutorial.exercise.timelimit = 60, 
        tutorial.storage = "local") 

x1 <- c("Dec", "Apr", "Jan", "Mar")

x2 <- c("Dec", "Apr", "Jam", "Mar")

month_levels <- c(
  "Jan", "Feb", "Mar", "Apr", "May", "Jun", 
  "Jul", "Aug", "Sep", "Oct", "Nov", "Dec"
)

csv <- "
month,value
Feb,12
Mar,56
Feb,14
Jan,12"
```

```{r copy-code-chunk, child = system.file("child_documents/copy_button.Rmd", package = "tutorial.helpers")}
```

```{r info-section, child = system.file("child_documents/info_section.Rmd", package = "tutorial.helpers")}
```

## Introduction
### 

This tutorial covers [Chapter 17: Factors](https://r4ds.hadley.nz/factors.html) from [*R for Data Science (2e)*](https://r4ds.hadley.nz/) by Hadley Wickham, Mine Çetinkaya-Rundel, and Garrett Grolemund. [**forcats**](https://forcats.tidyverse.org/) is the core *Tidyverse* package for working with categorical variables, called "factors" in R. Key commands include [`fct()`](https://forcats.tidyverse.org/reference/fct.html) for creating factors, [`fct_reorder()`](https://forcats.tidyverse.org/reference/fct_reorder.html) for changing the order of the levels, and [`fct_recode()`](https://forcats.tidyverse.org/reference/fct_recode.html) for recoding factors.

## Factor basics
### 

Factors are used for categorical variables, variables that have a fixed and known set of possible values. They are also useful when you want to display character vectors in a non-alphabetical order.

### Exercise 1

Load the **tidyverse** library.

```{r factor-basics-1, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r factor-basics-1-hint-1, eval = FALSE}
library(...)
```

### 

One of the nine core packages within the *Tidyverse* is [**forcats**](https://forcats.tidyverse.org/), a package dedicated to working with factors. By loading **tidyverse**, we automatically get access to **forcats** and the other "core" *Tidyverse* packages.

### Exercise 2

Look up the help page for **forcats** by entering `help(package = "forcats")` at the Console. Copy/paste the lines for the first help pages.

```{r factor-basics-2}
question_text(NULL,
	answer(NULL, correct = TRUE),
	allow_retry = TRUE,
	try_again_button = "Edit Answer",
	incorrect = NULL,
	rows = 5)
```

### 

**forcats** provides tools for dealing with **cat**egorical variables --- and it’s an anagram of teh word "factors" --- using a wide range of helpers for working with factors.

### Exercise 3

Hit "Run Code" to create the variable `x1`.

```{r factor-basics-3, exercise = TRUE}
x1 <- c("Dec", "Apr", "Jan", "Mar")
```

```{r include = FALSE}
x1 <- c("Dec", "Apr", "Jan", "Mar")
```

### 

Note that `x1` is a character variable. This can lead to all sorts of problems given that months are a good example of a categorical variable, given that there are exactly 12 possible values.

### Exercise 4

Run `sort()` on `x1`.

```{r factor-basics-4, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r factor-basics-4-hint-1, eval = FALSE}
sort(...)
```

```{r include = FALSE}
sort(x1)
```

### 

Because `x1` is a character variable, this sorts alphabetically, which is not what we want. We would prefer that the sort order correspond to the order in which months appear in the calendar.

### Exercise 5

Hit "Run Code" to create the `x2` variable, another character vector. But note the misspelling of "Jan" as "Jam".

```{r factor-basics-5, exercise = TRUE}
x2 <- c("Dec", "Apr", "Jam", "Mar")
```

```{r include = FALSE}
x2 <- c("Dec", "Apr", "Jam", "Mar")
```

### 

Because `x1` and `x2` are both character vectors, nothing will catch the contradiction between "Jan" and "Jam." Using factors will force us to notice such errors.

### Exercise 6

To create a factor you must start by creating a list of the valid "levels." Hit "Run Code" to create the `month_levels` variable.


```{r factor-basics-6, exercise = TRUE}
month_levels <- c(
  "Jan", "Feb", "Mar", "Apr", "May", "Jun", 
  "Jul", "Aug", "Sep", "Oct", "Nov", "Dec"
)
```

### 

Note that `month_levels` is just another character vector. We need it, however, to create a factor variable.

### Exercise 7

Run `factor(x1, levels = month_levels)`.

```{r factor-basics-7, exercise = TRUE}

```

```{r factor-basics-7-hint-1, eval = FALSE}
factor(x1, ... = month_levels)
```

```{r include = FALSE}
factor(x1, levels = month_levels)
```

### 

The function `factor()` is a part of base R, not **forcats**. It creates a factor variable. Notice how, in addition to the values of `x1` being printed, as they are when we print a character variable, we also see the levels, printed out in order.

### Exercise 8

Wrap `factor(x1, levels = month_levels)` within a call to `sort()`.

```{r factor-basics-8, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r factor-basics-8-hint-1, eval = FALSE}
sort(...(x1, levels = ...))
```

```{r include = FALSE}
sort(factor(x1, levels = month_levels))
```

### 

Instead of being sorted in alphabetical order, as before, the values are sorted in the order of the levels, which is almost always what we want when sorting months.


### Exercise 9

Run `factor(x2, levels = month_levels)`.


```{r factor-basics-9, exercise = TRUE}

```

```{r factor-basics-9-hint-1, eval = FALSE}
...(x2, ... = month_levels)
```

```{r include = FALSE}
factor(x2, levels = month_levels)
```

### 

Since "Jam" is not one of the levels, `factor()` coerces it to be missing, as shown with the `<NA>` symbol. One big advantage of working with factors is that you are prevented from using values which are not one of the levels.

### Exercise 10

Instead of `factor()`, we recommend using the `fct()` function from the **forcats** package, precisely because it generates an explicit error rather than a silent conversion to `NA`.

Run `fct(x2, levels = month_levels)`.


```{r factor-basics-10, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r factor-basics-10-hint-1, eval = FALSE}
...(x2, ... = month_levels)
```

### 

Notice the thorough error message. "Jam" is missing from the `levels` as defined in the `month_levels` variable.

### Exercise 11

Run `factor()` and `x1`.

```{r factor-basics-11, exercise = TRUE}

```

```{r factor-basics-11-hint-1, eval = FALSE}
factor(...)
```

```{r include = FALSE}
factor(x1)
```

### 

Because we did not provide a `levels` argument, the values for the levels will be taken from the values of the `x1` vector, sorted in alphabetical order.

### Exercise 12

Run `fct()` and `x1`.

```{r factor-basics-12, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r factor-basics-12-hint-1, eval = FALSE}
fct(...)
```

```{r include = FALSE}
fct(x1)
```

### 

Sorting alphabetically is slightly risky because not every computer will sort strings in the same way. So `forcats::fct()` orders by first appearance in the original vector.

### Exercise 13

Take the code from the previous exercise use it as an argument to the function `levels()`.

```{r factor-basics-13, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r factor-basics-13-hint-1, eval = FALSE}
levels(fct(...))
```

```{r include = FALSE}
levels(fct(x1))
```

### 

If you ever need to access the set of valid levels directly, you can do so with `levels()`.

### Exercise 14

The next exercises will be focusing on the variable `csv`. Hit "Run Code" to look at `csv`.

```{r factor-basics-14, exercise = TRUE}
csv
```

###

Note the `\n`. That signifies a new line being formed, but doesn't make `csv` easier to read.

### Exercise 15

Run `read_csv()` with `csv` set as the argument

```{r factor-basics-15, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r factor-basics-15-hint, eval = FALSE}
read_csv(...)
```

```{r, include = FALSE}
read_csv(csv)
```

###

`csv` is now much easier to read and understand. We can see what belongs in the month column and the value column, as well as variable types the columns are.

### Exercise 16

Currently, the month column is a character type variable and the value column is a double type variable. Add the arguments `col_types` to `read_csv` and set it equal to `"cc"`

```{r factor-basics-16, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r factor-basics-16-hint, eval = FALSE}
read_csv(csv, ...)
```

```{r, include = FALSE}
read_csv(csv, col_types = "cc")
```

###

`col_types = "cc"` changes the variable types of the columns to both be characters. The first `c` in `cc` corresponds to the first column, and so on.

### Exercise 17

Change the value of `col_types` from `"cc"` to `cols(month = "c")` inside of `read_csv()`

```{r factor-basics-17, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r factor-basics-17-hint, eval = FALSE}
read_csv(csv, col_types = cols(...))
```

```{r, include = FALSE}
read_csv(csv, col_types = cols(month = "c"))
```

###

### Exercise 18

Change the value of `col_types` to `cols(month = "f"))`

```{r factor-basics-18, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r factor-basics-18-hint, eval = FALSE}
read_csv(csv, col_types = cols(...))
```

```{r, include = FALSE}
read_csv(csv, col_types = cols(month = "f"))
```

###

The month variable is now a variable of factor type. Having month as a factor will allow us to perform certain actions on it later on.

### Exercise 19

Continue the current pipe to `count()`. Set the argument to `month`.

```{r factor-basics-19, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r factor-basics-19-hint, eval = FALSE}
... |>
  count(...)
```

```{r, include = FALSE}
read_csv(csv, col_types = cols(month = "f")) |> count(month)
```

###

While the tibble has all the correct information, it's not amazing to read. No one thinks of the months in those order. Luckily, this can be changed.

### Exercise 20

In `read_csv`, change the value of `col_types` to `cols(month = col_factor(month_levels))`

```{r factor-basics-20, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r factor-basics-20-hint, eval = FALSE}
read_csv(csv, col_types = ...) |>
  count(month)
```

```{r, include = FALSE}
read_csv(csv, 
         col_types = cols(month = col_factor(month_levels))) |>  
  count(month)
```

###

Besides the use of `factor()` and `fct()` as described earlier, `col_factor()`, when used within `read_csv()` and similar import functions, is the most common way of creating factor variables.

## General Social Survey
### 

The `gss_cat` tibble is a data set in the **forcats** package. It’s a sample of data from the General Social Survey, a long-running US survey conducted by the independent research organization NORC at the University of Chicago. 

### Exercise 1

Type `gss_cat` and hit "Run Code."

```{r general-social-surve-1, exercise = TRUE}

```

```{r general-social-surve-1-hint-1, eval = FALSE}
gss_cat
```

```{r include = FALSE}
gss_cat
```

### 

There are 9 variables and more than 20,000 observations. Note how the `print()` method for tibbles, which is called whenever you just enter the name of a tibble, like `gss_cat`, gives the variable types across the top.

### Exercise 2

Look up the help page for `gss_cat` by typing `?gss_cat` at the Console. Copy/paste the Description.

```{r general-social-surve-2}
question_text(NULL,
	answer(NULL, correct = TRUE),
	allow_retry = TRUE,
	try_again_button = "Edit Answer",
	incorrect = NULL,
	rows = 2)
```

### 

When referring to a tibble (or other variable) which is part of a package, you can just use the variable name if you have already loaded the package. (Recall that running `library(tidyverse)` loads all the *Tidyverse* libraries, including **forcats**.) You can also refer to the variable directly using the double colon notation -- `::` -- i.e., `forcats::gss_cat`.


### Exercise 3

When factors are stored in a tibble, you can’t see their levels so easily. One way to view them is with `count()`. Pipe `gss_cat` to `count(race)`.

```{r general-social-surve-3, exercise = TRUE}

```

```{r general-social-surve-3-hint-1, eval = FALSE}
gss_cat |>
  ...(race)
```

```{r include = FALSE}
gss_cat |>
  count(race)
```

### 

The `<fct>` indicator above `race` indicates that it is a factor variable, not character.

<!-- This is where the chapter stops. We could add more examples working with this data. Or we could work through the exercises from 17.3.1 -->

## Modifying factor order
### 

When working with factors, one common operation is changing the order of the levels. Let's create this plot:

```{r make-plot1}
plot1 <- gss_cat |>
  summarize(n = n(),
            tvhours = mean(tvhours, na.rm = TRUE),
            .by = relig) |> 
  mutate(relig = fct_reorder(relig, tvhours)) |> 
  ggplot(aes(x = tvhours, y = relig)) +
  geom_point() +
  labs(title = "TV Watching and Religious Affiliation",
       subtitle = "Don't Knows watch a lot of TV",
       x = "TV Hours Watched Per Day",
       y = "Religious Affiliation")

plot1
```

### Exercise 1

Run `glimpse()` on `gss_cat`.

```{r modifying-factor-ord-1, exercise = TRUE}

```

```{r modifying-factor-ord-1-hint-1, eval = FALSE}
glimpse(...)
```

```{r include = FALSE}
glimpse(gss_cat)
```

### 

We will be working with two variables: `relig` and `tvhours`. `relig` is a factor variable reporting religious affiliation, if any. `tvhours` is hours per day spent watching TV, on average.

### Exercise 2

Pipe `gss_cat` to `summarize(n = n())`

```{r modifying-factor-ord-2, exercise = TRUE}

```

```{r modifying-factor-ord-2-hint-1, eval = FALSE}
gss_cat |>
  summarize(n = n())
```

### 

Note how the letter "n" is used in two ways. First, it is the name of a new variable `n`, created via `summarize()`. In statistics, it is common for the letter "n" to mean the number of observations. Second, `n()` is a function, hence the `()`, which calculates the number of observations. Since there is no `.by` argument, the result is a tibble with a single row.

### Exercise 3

Use the same pipe again, but add `.by = relig` as an argument/value pairing to `summarize()`.

```{r modifying-factor-ord-3, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r modifying-factor-ord-3-hint-1, eval = FALSE}
gss_cat |>
  summarize(n = n(),
            .by = relig)
```

### 

The result is a tibble with one row for each level of `relig`. (Older R code will often use the `group_by()` function when calculating statistics for each level of a factor. You should avoid this approach. Use the `.by` argument to `summarize()` and similar functions.)

### Exercise 4

Use the same code again, adding another variable creation step to `summarize()`: `tvhours = mean(tvhours)`.

```{r modifying-factor-ord-4, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r modifying-factor-ord-4-hint-1, eval = FALSE}
gss_cat |>
  summarize(n = n(),
            tvhours = mean(tvhours),
            .by = relig)
```

### 

Note that each argument (or variable creation step) in `summarize()` must be separated by a command. Alas, there are `NA` values present at least one person in every level of `relig`.


### Exercise 5

Modify the pipe by add `na.rm = TRUE` as an argument **within* the `mean()` function.

```{r modifying-factor-ord-5, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r modifying-factor-ord-5-hint-1, eval = FALSE}
gss_cat |>
  summarize(n = n(),
            tvhours = mean(tvhours, na.rm = TRUE),
            .by = relig)
```

### 

All statistical functions in R will produce a value of NA if even a single one of the input values is NA, consistent with the rules of mathematics. Most statistical functions have a `na.rm` --- short for NA remove --- which allows us to remove any NA values prior to the calculation.

### Exercise 6

Continue the pipe with a call to `ggplot()`, setting the `mapping` argument to `aes(x = tvhours, y = relig)`.

```{r modifying-factor-ord-6, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r modifying-factor-ord-6-hint-1, eval = FALSE}
gss_cat |>
  summarize(n = n(),
            tvhours = mean(tvhours, na.rm = TRUE),
            .by = relig) |> 
  ggplot(aes(x = tvhours, y = relig))
```

### 

Without a geom function, no data is plotted. But we still get the plotting area and the axis labels. Does the ordering of the religious affiliations on the y-axis seem reasonable? 

### Exercise 7

Add `geom_point()` to the pipe. Don't forget that calls to **ggplot** components are separated by plus signs, not pipes -- by `+` not `|>`.

```{r modifying-factor-ord-7, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r modifying-factor-ord-7-hint-1, eval = FALSE}
gss_cat |>
  summarize(n = n(),
            tvhours = mean(tvhours, na.rm = TRUE),
            .by = relig) |> 
  ggplot(aes(x = tvhours, y = relig)) +
  geom_point()
```

### 

It is hard to read this plot because there’s no overall pattern. We can improve it by reordering the levels of relig using `fct_reorder()`. `fct_reorder()` takes three arguments:

* `f`, the factor whose levels you want to modify.
* `x`, a numeric vector that you want to use to reorder the levels.
* Optionally, `fun`, a function that’s used if there are multiple values of `x` for each value of `f`. The default value is median.

### Exercise 8

Replace `y = relig` with `y = fct_reorder(relig, tvhours)` in your pipe.

```{r modifying-factor-ord-8, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r modifying-factor-ord-8-hint-1, eval = FALSE}
gss_cat |>
  summarize(n = n(),
            tvhours = mean(tvhours, na.rm = TRUE),
            .by = relig) |> 
  ggplot(aes(x = tvhours, y = fct_reorder(relig, tvhours))) +
  geom_point()
```

### 

Reordering religion makes it much easier to see that people in the “Don’t know” category watch much more TV, and Hinduism & Other Eastern religions watch much less.

### Exercise 9

As you start making more complicated transformations, we recommend moving them out of `aes()` and into a separate `mutate()` step. After the `summarize()` step, insert this line: `mutate(relig = fct_reorder(relig, tvhours)) |> `. Then, change `y = fct_reorder(relig, tvhours)` back to `y = relig`.

```{r modifying-factor-ord-9, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r modifying-factor-ord-9-hint-1, eval = FALSE}
gss_cat |>
  summarize(n = n(),
            tvhours = mean(tvhours, na.rm = TRUE),
            .by = relig) |> 
  mutate(... = fct_reorder(relig, ...)) |> 
  ggplot(aes(x = tvhours, ... = relig)) +
  geom_point()
```

### 

It is almost always better to complete your data transformations before starting your plot.

### Exercise 10

Finish the plot by adding a title, subtitle, and axis labels. Remember that the plot looks like this:

```{r modifying-factor-ord-10-example}
plot1
```

```{r modifying-factor-ord-10, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r modifying-factor-ord-10-hint-1, eval = FALSE}
... +
  labs(... = "TV Watching and Religious Affiliation",
       subtitle = ...,
       x = ...,
       ... = "Religious Affiliation")
```

### 

The subtitle of a plot should be the one sentence conclusion/summary/observation with which you most want viewers to come away.

### 

Let's create this graph now.

```{r make-plot2}
plot2 <- gss_cat |>
  filter(!is.na(age)) |> 
  count(age, marital) |>
  mutate(
    prop = n / sum(n),
    .by = age) |> 
  ggplot(aes(x = age, 
             y = prop, 
             color = fct_reorder2(marital, age, prop))) +
    geom_line(linewidth = 1) +
    scale_color_brewer(palette = "Set1") + 
    labs(color = "Marital", x = "Age", y = "Proportion")

plot2
```

### Exercise 11

Pipe gss_cat to `filter()`, with the argument set as `!is.na(age)`

```{r modifying-factor-ord-11, exercise = TRUE}

```

```{r modifying-factor-ord-11-hint-1, eval = FALSE}
gss_cat |>
  filter(...)
```

### 

This will remove all rows with a `NA` response to `age`, allowing for easier calculations in the future.

### Exercise 12

Continue the pipe to `count()`, with `age` and `marital` set as the arguments

```{r modifying-factor-ord-12, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r modifying-factor-ord-12-hint-1, eval = FALSE}
... |>
  count(age, ...)
```

### Exercise 13

Continue the pipe to `mutate().` Create the variable `prop` and set it `n/sum(n)`

```{r modifying-factor-ord-13, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r modifying-factor-ord-13-hint-1, eval = FALSE}
... |>
  mutate(... = n/sum(n))
```

### Exercise 14

Add the argument `.by` within the `mutate()` function. Set `.by` to `age`

```{r modifying-factor-ord-14, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r modifying-factor-ord-14-hint-1, eval = FALSE}
... |>
  mutate(prop = n/sum(n), .by = ...)
```

### 

Adding the `.by` argument allows us to sort the tibble by age for the mutate function.

### Exercise 15

Add `ggplot()` to the current pipe. Set `x` to `age`, `y` to `prop` and `color` to `marital`

```{r modifying-factor-ord-15, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r modifying-factor-ord-15-hint-1, eval = FALSE}
... |>
  ggplot(aes(x = ..., y = ..., color = ...))
```

### Exercise 16

Continue the pipe to `geom_line()`. Inside the function, set `linewidth` to 1.

```{r modifying-factor-ord-16, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r modifying-factor-ord-16-hint-1, eval = FALSE}
... |>
  geom_line(linewidth = ...)
```

### Exercise 17

Continue the pipe with `scale_color_brewer()`, with the argument `palette` set to "Set1".

```{r modifying-factor-ord-17, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r modifying-factor-ord-17-hint-1, eval = FALSE}
... |>
  scale_color_brewer(... = "Set1")
```

### Exercise 18

Finish the pipe with `labs()` giving an appropriate axes and legend titles.

```{r modifying-factor-ord-18, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r modifying-factor-ord-18-hint-1, eval = FALSE}
... |>
  labs(x = ..., y = ..., color = ...)
```

### 

This graph is confusing to read as the colors assigned to the lines don't match up well with the legend. We can use `fct_reorder2()` to solve this problem.

### Exercise 19

In the `ggplot()` function, change `color` from `marital` to `fct_reorder2(marital, age, prop)`

```{r modifying-factor-ord-19, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r modifying-factor-ord-19-hint-1, eval = FALSE}
... |>
  ggplot(aes(x = age, y = prop, color = fct_reorder2(marital, age, prop))) |>
  ...
```

### 

Rearranging the legend makes the plot easier to read because the legend colors now match the order of the lines on the far right of the plot. `fct_reorder2(f, x, y)` reorders the factor `f` by the `y` values associated with the largest `x` values.

## Modifying factor levels
### 

More powerful than changing the orders of the levels is changing their values. This allows you to clarify labels for publication, and collapse levels for high-level displays. The most general and powerful tool is `fct_recode()`. It allows you to recode, or change, the value of each level. 

### Exercise 1

Pipe `gss_cat` to `count(partyid)`


```{r modifying-factor-lev-1, exercise = TRUE}

```

```{r modifying-factor-lev-1-hint-1, eval = FALSE}
... |> 
  count(...)
```

```{r include = FALSE}
gss_cat |> 
  count(partyid)
```

### 

The levels of `partyid` are terse and inconsistent. Let’s tweak them to be longer and use a parallel construction.

### Exercise 2

Like most rename and recoding functions in the *Tidyverse*, the new values go on the left and the old values go on the right. Pipe `gss_cat` to `mutate()`. Within `mutate()`, use `partyid = fct_recode(partyid, "Republican, weak" = "Not str republican")` to change `partyid`.


```{r modifying-factor-lev-2, exercise = TRUE}

```

```{r modifying-factor-lev-2-hint-1, eval = FALSE}
gss_cat |>
  mutate(
    partyid = ...(partyid,
      "Republican, weak"      = ...
    )
  ) 
```

### 

Note how the second and seventh values for `partyid` have been changed from "Not str republican" to  "Republican, weak". `fct_recode()` is the easiest way to change the value for a given factor level. Sometimes, as here, we change the value "in place," that is, we replace `partyid` with `partyid`. Other times, we use `mutate()` to create a new variable.

### Exercise 3

Let's change all the values for `partyid`. Here is the mapping from new values to old values:

```{verbatim echo = TRUE}
      "Republican, strong"    = "Strong republican",
      "Republican, weak"      = "Not str republican",
      "Independent, near rep" = "Ind,near rep",
      "Independent, near dem" = "Ind,near dem",
      "Democrat, weak"        = "Not str democrat",
      "Democrat, strong"      = "Strong democrat"
```

Use these within the call to `fct_recode()`.


```{r modifying-factor-lev-3, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r modifying-factor-lev-3-hint-1, eval = FALSE}
gss_cat |>
  mutate(
    partyid = fct_recode(partyid,
      "Republican, strong"    = "Strong republican",
      "Republican, weak"      = "Not str republican",
      "Independent, near rep" = "Ind,near rep",
      "Independent, near dem" = ...,
      ...       = "Not str democrat",
      "Democrat, strong"      = "Strong democrat"
    )
  ) 
```

### 

`fct_recode()` will leave the levels that aren’t explicitly mentioned as they are, and will warn you if you accidentally refer to a level that doesn’t exist.

### Exercise 4

To combine groups, you can assign multiple old levels to the same new level. With the same pipe as above, use this mapping:

```{verbatim echo = TRUE}
      "Republican, strong"    = "Strong republican",
      "Republican, weak"      = "Not str republican",
      "Independent, near rep" = "Ind,near rep",
      "Independent, near dem" = "Ind,near dem",
      "Democrat, weak"        = "Not str democrat",
      "Democrat, strong"      = "Strong democrat",
      "Other"                 = "No answer",
      "Other"                 = "Don't know",
      "Other"                 = "Other party"
```

```{r modifying-factor-lev-4, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r modifying-factor-lev-4-hint-1, eval = FALSE}
gss_cat |>
  mutate(
    partyid = fct_recode(partyid,
      ...
    )
  )
```

### 

Use this technique with care: if you group together categories that are truly different, you will end up with misleading results.

### Exercise 5

Continue the pipe to `count(partyid)` to confirm that the recoding has worked.

```{r modifying-factor-lev-5, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r modifying-factor-lev-5-hint-1, eval = FALSE}
... |> 
  count(partyid)
```

```{r include = FALSE}
gss_cat |>
  mutate(
    partyid = fct_recode(partyid,
      "Republican, strong"    = "Strong republican",
      "Republican, weak"      = "Not str republican",
      "Independent, near rep" = "Ind,near rep",
      "Independent, near dem" = "Ind,near dem",
      "Democrat, weak"        = "Not str democrat",
      "Democrat, strong"      = "Strong democrat",
      "Other"                 = "No answer",
      "Other"                 = "Don't know",
      "Other"                 = "Other party")) |>
  count(partyid)
```

### 


Read the help page for [`fct_recode()`](https://forcats.tidyverse.org/reference/fct_recode.html) for more details.

### Exercise 6

If you want to collapse a lot of levels, `fct_collapse()` is a useful variant of `fct_recode()`. For each new variable, you can provide a vector of old levels. Replace the call to `fct_record()` in the previous pipe with this:

```{verbatim echo = TRUE}
fct_collapse(partyid,
      "other" = c("No answer", "Don't know", "Other party"),
      "rep" = c("Strong republican", "Not str republican"),
      "ind" = c("Ind,near rep", "Independent", "Ind,near dem"),
      "dem" = c("Not str democrat", "Strong democrat")
    )
```

```{r modifying-factor-lev-6, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r modifying-factor-lev-6-hint-1, eval = FALSE}
gss_cat |>
  ...(
    partyid = fct_collapse(...,
      "other" = c("No answer", "Don't know", "Other party"),
      "rep" = c("Strong republican", "Not str republican"),
      "ind" = c("Ind,near rep", "Independent", "Ind,near dem"),
      "dem" = c("Not str democrat", "Strong democrat")
    )
  ) |> 
  ...(partyid)
```

### 

Read the help page for [`fct_collapse()`](https://forcats.tidyverse.org/reference/fct_collapse.html) for more details. The `other_level` argument is sometimes useful.

### Exercise 7

Sometimes you just want to lump together the small groups to make a plot or table simpler. That’s the job of the `fct_lump_*()` family of functions.

Pipe `gss_cat` to `mutate()` with `relig = fct_lump_lowfreq(relig)` as its argument.

```{r modifying-factor-lev-7, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>


```{r modifying-factor-lev-7-hint-1, eval = FALSE}
gss_cat |>
  mutate(relig = ...)
```

### 

`fct_lump_lowfreq()` is a simple starting point that progressively lumps the smallest groups categories into “Other”, always keeping “Other” as the smallest category.

### Exercise 8

Continue the pipe to the function `count()` with `relig` as its argument.

```{r modifying-factor-lev-8, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r modifying-factor-lev-8-hint-1, eval = FALSE}
... |>
  count(...)
```

```{r include = FALSE}
gss_cat |>
  mutate(relig = fct_lump_lowfreq(relig)) |>
  count(relig)
```

### 

In this case it’s not very helpful: it is true that the majority of Americans in this survey are Protestant, but we’d probably like to see some more details! 

### Exercise 9

Instead, we can use the `fct_lump_n()` to specify that we want exactly 10 groups. Pipe `gss_cat` to `mutate()` with `fct_lump_n(relig, n = 10)` as its argument.

```{r modifying-factor-lev-9, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r modifying-factor-lev-9-hint-1, eval = FALSE}
gss_cat |>
  mutate(relig = ...)
```

### Exercise 10

Continue the pipe to `count()`. Add `relig` as an argument, as well as `sort = TRUE` to `count()`.

```{r modifying-factor-lev-10, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r modifying-factor-lev-10-hint, eval = FALSE}
... |>
  count(relig, ...)
```

```{r include = FALSE}
gss_cat |>
  mutate(relig = fct_lump_n(relig, n = 10)) |>
  count(relig, sort = TRUE)
```

### 

Read the documentation to learn about [`fct_lump_min()`](https://forcats.tidyverse.org/reference/fct_lump.html) and [`fct_lump_prop()`](https://forcats.tidyverse.org/reference/fct_lump.html) which are useful in other cases.

## Ordered factors
### 

Ordered factors, created with `ordered()`, imply a strict ordering and equal distance between levels: the first level is “less than” the second level by the same amount that the second level is “less than” the third level, and so on.

### Exercise 1

Run this code

```{r ordered-factors-1, exercise = TRUE}
ordered(c("a", "b", "c"))
```

```{r include = FALSE}
ordered(c("a", "b", "c"))
```

### 

You can recognize ordered factors when printing because they use `<` between the factor levels. We don't recommend using ordered factors unless you have a compelling reason for doing so.

## Summary
### 

This tutorial covered [Chapter 17: Factors](https://r4ds.hadley.nz/factors.html) from [*R for Data Science (2e)*](https://r4ds.hadley.nz/) by Hadley Wickham, Mine Çetinkaya-Rundel, and Garrett Grolemund. [**forcats**](https://forcats.tidyverse.org/) is the core *Tidyverse* package for working with categorical variables, called "factors" in R. Key commands include [`fct()`](https://forcats.tidyverse.org/reference/fct.html) for creating factors, [`fct_reorder()`](https://forcats.tidyverse.org/reference/fct_reorder.html) for changing the order of the levels, and [`fct_recode()`](https://forcats.tidyverse.org/reference/fct_recode.html) for recoding factors.

If you want to learn more about factors, read "[Wrangling categorical data in R](https://peerj.com/preprints/3163/))" by Amelia McNamara and Nicholas Horton.   

```{r download-answers, child = system.file("child_documents/download_answers.Rmd", package = "tutorial.helpers")}
```
