---
title: "Joins"
tutorial:
  id: "joins"
output: 
  learnr::tutorial:
    progressive: true
    allow_skip: true
runtime: shiny_prerendered
description: "Tutorial for Chapter 20: Joins"
---

```{r setup, include = FALSE}
library(learnr)
library(tutorial.helpers)
knitr::opts_chunk$set(echo = FALSE)
options(tutorial.exercise.timelimit = 60, 
        tutorial.storage = "local") 
```

```{r copy-code-chunk, child = system.file("child_documents/copy_button.Rmd", package = "tutorial.helpers")}
```

```{r info-section, child = system.file("child_documents/info_section.Rmd", package = "tutorial.helpers")}
```

<!-- CHECKLIST BEFORE STARTING: -->
<!-- * Edit yaml at the top of this file -->
<!-- * Save the file as "tutorial.Rmd" in a new directory under inst/tutorials/. -->
<!-- * Load any necessary libraries for the tutorial in the first code chunk -->
<!-- * Review: https://ppbds.github.io/tutorial.helpers/articles/instructions.html -->
<!-- * Delete this and the other commented instructions below. -->

## Introduction
###

This tutorial covers [Chapter 20: Strings](https://r4ds.hadley.nz/joins.html) from [*R for Data Science (2e)*](https://r4ds.hadley.nz/) by Hadley Wickham, Mine Çetinkaya-Rundel, and Garrett Grolemund. 
This chapter will introduce you to two important types of joins:
Mutating joins, which add new variables to one data frame from matching observations in another.

Filtering joins, which filter observations from one data frame based on whether or not they match an observation in another.

You will also learn about basic joins, keys, foreign keys and more which is basically like creating a relational database in SQL.


## Keys
###

## Basic Joins
### 

## How do joins work?
###

## Non-equi joins
###





## Summary
###

This tutorial covered [Chapter 20: Strings](https://r4ds.hadley.nz/joins.html) from [*R for Data Science (2e)*](https://r4ds.hadley.nz/) by Hadley Wickham, Mine Çetinkaya-Rundel, and Garrett Grolemund. 
This chapter introduced you to two important types of joins:
Mutating joins, which add new variables to one data frame from matching observations in another.

Filtering joins, which filter observations from one data frame based on whether or not they match an observation in another.

You also learned about basic joins, keys, foreign keys and more which is basically like creating a relational database in SQL.


```{r download-answers, child = system.file("child_documents/download_answers.Rmd", package = "tutorial.helpers")}
```
