---
title: Numbers
author: Srihith Garlapati
tutorial:
  id: numbers
output:
  learnr::tutorial:
    progressive: yes
    allow_skip: yes
runtime: shiny_prerendered
description: 'Tutorial for Chapter 14: Numbers'
---

```{r setup, include = FALSE}
library(learnr)
library(tutorial.helpers)
library(tidyverse)
library(primer.data)
library(readr)
knitr::opts_chunk$set(echo = FALSE)
options(tutorial.exercise.timelimit = 60, 
        tutorial.storage = "local") 


```

```{r copy-code-chunk, child = system.file("child_documents/copy_button.Rmd", package = "tutorial.helpers")}
```

```{r info-section, child = system.file("child_documents/info_section.Rmd", package = "tutorial.helpers")}
```

## Introduction
### 

This tutorial covers [Chapter 14: Numbers](https://r4ds.hadley.nz/numbers.html) from [*R for Data Science (2e)*](https://r4ds.hadley.nz/) by Hadley Wickham, Mine Çetinkaya-Rundel, and Garrett Grolemund. You will learn how to parse numbers, utilize numeric transformations on variables within datasets, and learn new ways to numerically summarize data. 

## Making Numbers
### 

Our ultimate goal in this section is to make use of the functions `parse_double()` and `parse_number()` to get specific values of a variable from a certain dataset.

### Exercise 1
This chapter mostly uses functions from base R, which are available without loading any packages. But we still need the tidyverse because we’ll use these base R functions inside of tidyverse functions like mutate() and filter(). 

###

`readr` provides two useful functions for parsing strings into numbers: `parse_double()` and `parse_number()`.

###

First, let's load in the tidyverse and the readr package to make use of these methods.
```{r making-numbers-1, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r making-numbers-1-hint, eval = FALSE}
library(tidyverse)
library(readr)
```
### Exercise 2

In most cases, you’ll get numbers already recorded in one of R’s numeric types: integer or double. In some cases, however, you’ll encounter them as strings.

###

Let's make a vector `x` that has the following values within it `"1.2"`, `"5.6"`, `"1e3"`. These values are strings for now.
```{r making-numbers-2, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r making-numbers-2-hint, eval = FALSE}
x <- c(..., ..., ...)
```

### Exercise 3
When you have numbers that have been written as strings we should use `parse_double()`. So let's use use `parse_double()` on the command we just created. 

```{r making-numbers-3, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r making-numbers-3-hint, eval = FALSE}
x <- c("...", "...", "...")
parse_double(x)
```

###

This should return an output of: `#> [1]    1.2    5.6 1000.0`.

### Exercise 4
Use `parse_number()` when the string contains non-numeric text that you want to ignore. This is particularly useful for currency data and percentages. Let's first make a vector y that has the string values: `"$1,234"`, `"USD 3,513"`, `59%`

```{r making-numbers-4, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r making-numbers-4-hint, eval = FALSE}
y <- c(..., ..., ...)
```

###

As you can see, in this vector `y` there are non-numeric text that we would normally want to ignore if we were to see this in a dataset. 

### Exercise 5

Let's use `parse_number()` to ignore this non-numeric text.

```{r making-numbers-5, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r making-numbers-5-hint, eval = FALSE}
y <- c(..., ..., ...)
parse_number(y)
```

###

The result you should've received is `#> [1] 1234 3513   59`

## Counts
### 

In this section, we will be reviewing how to use `count()` and other similar functions to make data analysis easier. 

### Exercise 1

Load in the package `tidyverse` and `dplyr`.
```{r counts-1, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r counts-1-hint, eval = FALSE}
library(tidyverse)
library(dplyr)
```

###

### Exercise 2

Now let's load in the dataset, `flights` that we will be making use of. This dataset contains information about all flights that departed from NYC (e.g. EWR, JFK and LGA) to destinations in the United States, Puerto Rico, and the American Virgin Islands) in 2013: 336,776 flights in total. To help understand what causes delays, it also includes a number of other useful datasets.

###


Let's count the number of planes of which departed from a certain destination. Make a pipeline from `flights` and use the function `count()` on the `dest`.
```{r counts-2, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r counts-2-hint, eval = FALSE}
flights
```

###

### Exercise 3


```{r counts-3, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r counts-3-hint, eval = FALSE}

```

###

### Exercise 4


```{r counts-4, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r counts-4-hint, eval = FALSE}

```

###

### Exercise 5


```{r counts-5, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r counts-5-hint, eval = FALSE}

```

###

### Exercise 6


```{r counts-6, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r counts-6-hint, eval = FALSE}

```

###

### Exercise 7


```{r counts-7, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r counts-7-hint, eval = FALSE}

```

###

### Exercise 8


```{r counts-8, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r counts-8-hint, eval = FALSE}

```

###

### Exercise 9


```{r counts-9, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r counts-9-hint, eval = FALSE}

```

###

### Exercise 10


```{r counts-10, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r counts-10-hint, eval = FALSE}

```

###

### Exercise 11


```{r counts-11, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r counts-11-hint, eval = FALSE}

```

###



## Numeric Transformations
### 

This section will go over transforming numeric vectors with the function `mutate()` and using various mathematic methods to make new columns in a certain dataset.

## More Transformations
### 

This section will allow you to make use of transformations on an actual dataset.

## Numeric Summaries
### 

This section will introduce more useful summary functions that will help to summarize your data much better.

## Summary
### 

This tutorial covered [Chapter 14: Numbers](https://r4ds.hadley.nz/numbers.html) from [*R for Data Science (2e)*](https://r4ds.hadley.nz/) by Hadley Wickham, Mine Çetinkaya-Rundel, and Garrett Grolemund. 

```{r download-answers, child = system.file("child_documents/download_answers.Rmd", package = "tutorial.helpers")}
```
