---
title: Numbers
author: David Kane
tutorial:
  id: numbers
output:
  learnr::tutorial:
    progressive: yes
    allow_skip: yes
runtime: shiny_prerendered
description: 'Tutorial for Chapter 14: Numbers'
---

```{r setup, include = FALSE}
library(learnr)
library(tutorial.helpers)
library(tidyverse)
knitr::opts_chunk$set(echo = FALSE)
options(tutorial.exercise.timelimit = 60, 
        tutorial.storage = "local") 


```

```{r copy-code-chunk, child = system.file("child_documents/copy_button.Rmd", package = "tutorial.helpers")}
```

```{r info-section, child = system.file("child_documents/info_section.Rmd", package = "tutorial.helpers")}
```

## Introduction
### 

This tutorial covers [Chapter 14: Numbers](https://r4ds.hadley.nz/numbers.html) from [*R for Data Science (2e)*](https://r4ds.hadley.nz/) by Hadley Wickham, Mine Çetinkaya-Rundel, and Garrett Grolemund. You will learn how to import data into your R project using [`read_csv()`](https://readr.tidyverse.org/reference/read_delim.html) and related functions from the [**readr**](https://readr.tidyverse.org/) package. You will also learn how to write out data to files with functions like [`write_csv()`](https://readr.tidyverse.org/reference/write_delim.html).

## Older material
### 

### Exercise 1

People write numbers differently in different parts of the world. For example, some countries use `.` in between the integer and fractional parts of a real number, while others use `,`.

### 

To begin, use the function  `parse_double()` with the argument "1.23".

```{r older-material-1, exercise = TRUE}

```

```{r older-material-1-hint-1, eval = FALSE}
parse_double("1.23")
```

### 

There are other `parse_*` functions such as `parse_logical()`, `parse_integer()`, and `parse_character()`.

### Exercise 2

What if the file you worked with used a comma to denote a decimal mark? Many functions in the [**readr**](https://readr.tidyverse.org) package provide a “locale” argument. You provide an object that specifies parsing options that differ from place to place. You can override the default value of `.` by creating a new locale and setting the `decimal_mark` argument.

### 

Use `parse_double()` with the argument "1,23". The second argument should set `locale` to `locale(decimal_mark = ",")`.

```{r older-material-2, exercise = TRUE}

```

```{r older-material-2-hint-1, eval = FALSE}
parse_double("1,23", locale = locale(..))
```

Note that `locale` is both an argument for `parse_*` functions and the name of a function itself, on which create a local object. See `?readr::locale` for [details](https://readr.tidyverse.org/articles/locales.html).

### Exercise 3

The function `parse_number()` ignores non-numeric characters before and after the number. 

### 

Use `parse_number()` with the argument "$100". Then, on the line below, use `parse_number()` again with the argument "20%".


```{r older-material-3, exercise = TRUE}

```

```{r older-material-3-hint-1, eval = FALSE}
parse_number("$100")
parse_number("20%")
```

### 

Unlike `parse_integer()` and `parse_double()`, `parse_number()` can parse both integers and doubles.

### Exercise 4

Use `parse_number()` with the argument "The cost of the package is $40.75" 

```{r older-material-4, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r older-material-4-hint-1, eval = FALSE}
parse_number("The cost of ...." )
```

<!-- DK: So many knowledge drop possibilities! -->


## Summary
### 

This tutorial covered [Chapter 14: Numbers](https://r4ds.hadley.nz/numbers.html) from [*R for Data Science (2e)*](https://r4ds.hadley.nz/) by Hadley Wickham, Mine Çetinkaya-Rundel, and Garrett Grolemund. 

```{r download-answers, child = system.file("child_documents/download_answers.Rmd", package = "tutorial.helpers")}
```
