---
title: Data tidying
author: Miriam Heiss
tutorial:
  id: data-tidying
output:
  learnr::tutorial:
    progressive: yes
    allow_skip: yes
runtime: shiny_prerendered
description: 'Tutorial for Chapter 6: Data tidying'
---

```{r setup, include = FALSE}
library(learnr)
library(tidyverse)
library(tutorial.helpers)
library(primer.data)
library(ggdist)
knitr::opts_chunk$set(echo = FALSE)
options(tutorial.exercise.timelimit = 60, 
        tutorial.storage = "local") 
```

```{r copy-code-chunk, child = system.file("child_documents/copy_button.Rmd", package = "tutorial.helpers")}
```

```{r info-section, child = system.file("child_documents/info_section.Rmd", package = "tutorial.helpers")}
```

## Introduction

### 

This tutorial covers [Chapter 6: Data tidying](https://r4ds.hadley.nz/data-tidy.html) from [*R for Data Science (2e)*](https://r4ds.hadley.nz/) by Hadley Wickham, Mine Çetinkaya-Rundel, and Garrett Grolemund.

Please restart your R session before proceeding. The short cut key for doing so is `Cmd-Shift-0`.

## Tidy Data

### 

You can represent the same underlying data in multiple ways by organizing values in a given dataset in different orders. But not all data is equally easy to use.

### Exercise 1

In the console, load in the *tidyverse* package. Copy and paste the loading message (the thing with the conflicts and check marks) into the box below.

```{r tidy-data-1}
question_text(NULL,
	answer(NULL, correct = TRUE),
	allow_retry = TRUE,
	try_again_button = "Edit Answer",
	incorrect = NULL,
	rows = 3)
```

### Exercise 2

Restart your R session with `Cmd-Shift-0`

### 

Load the `tidyverse` package again and copy and paste the first line of the section labeled `Attatching core tidyverse packages` into the box below.

```{r tidy-data-2}
question_text(NULL,
	answer(NULL, correct = TRUE),
	allow_retry = TRUE,
	try_again_button = "Edit Answer",
	incorrect = NULL,
	rows = 3)
```

### 

It should just show [dplyr](https://dplyr.tidyverse.org/) and [readr](https://readr.tidyverse.org/). 

###

`dplyr` has all your basic data transformation functions like `mutate()`, `filter()`, and `summarise()`. `readr` allows you to read data that is in the form of a csv, tsv, or spreadsheet.

### Exercise 3

Copy and paste the third line of the loading message into the box below.

```{r tidy-data-3}
question_text(NULL,
	answer(NULL, correct = TRUE),
	allow_retry = TRUE,
	try_again_button = "Edit Answer",
	incorrect = NULL,
	rows = 3)
```

It should have [ggplot2](https://ggplot2.tidyverse.org/) and [tibble](https://tibble.tidyverse.org/). 

###

`ggplot2` is how we make plots and maps in R. You provide the data, tell ggplot2 how to map variables to aesthetics, what graphical primitives to use, and it takes care of the details. `tibble` allows you to create data frames that are lazy and surly: they do less (i.e. they don’t change variable names or types, and don’t do partial matching) and complain more (e.g. when a variable does not exist).


### Exercise 4

Copy and paste the 4th line of the loading message into the box below. 

```{r tidy-data-4}
question_text(NULL,
	answer(NULL, correct = TRUE),
	allow_retry = TRUE,
	try_again_button = "Edit Answer",
	incorrect = NULL,
	rows = 3)
```

###

It should show [lubridate](https://lubridate.tidyverse.org/) and [**tidyr**](https://tidyr.tidyverse.org/)

###

`lubridate` is a more intuitive way to change and make dates in R, and allows for all of the quirks that our time has (such as leap days and daylight savings) that computers and programs don't understand. `tidyr` is the most important package for this tutorial. It makes sure that you have **Tidy Data**. The three main rules for having tidy data are: 

1. Each variable is a column; each column is a variable.
2. Each observation is a row; each row is an observation.
3. Each value is a cell; each cell is a single value.

###

### Exercise 5

Restart your R session again with `Cmd-Shift-0`.

###

Run the code provided in the boxes below.

```{r tidy-data-5-1, exercise = TRUE}
table1
```

```{r tidy-data-5-2, exercise = TRUE}
table2
```

```{r tidy-data-5-3, exercise = TRUE}
table3
```

###

These show Tuberculosis cases in three countries, China, Brazil, and Afghanistan.

###

They all show the same data (country, year, population, cases), but one of them is going to be way easier to work with.

### Exercise 6

Let's  take a look at `table2` first.

```{r tidy-data-6, exercise = TRUE}
table2
```

###

## Lengthening Data

### 

## Widening Data

### 

## Michigan voter experiment

### 

The `shaming` data set measures the impact of social pressure on voting. Nearly 350,000 people in Michigan were randomly assigned to 1 of 5 treatment groups before the 2006 Michigan primary. All 5 groups were sent mail before the primary: The "Civic Duty" group had an extra reminder that voting was a civic responsibility. The "Hawthorne" group was told that whether or not they voted, their decision would not be in the public record. The "Self" group was actually *sent* the public record of whether or not they voted in 2004. The "Neighbors" group was sent both their voting record and their neighbors' voting record from 2004. They then measured the impact of these "social pressures" on their choice to vote.

### 

Let's create the following plot

```{r make-shaming1}
shaming_p1 <- shaming |> 
  mutate(id = 1:length(primary_06)) |> 
  pivot_longer(cols = primary_00:general_04,
               names_to = "year",
               values_to = "voted") |> 
  mutate(voted = if_else(voted == "Yes", 1, 0)) |> 
  pivot_wider(names_from = year, 
              values_from = voted) |> 
  rowwise() |> 
  mutate(vote_sum = sum(primary_00, general_00, primary_02, general_02, 
                        primary_04, general_04)) |> 
  mutate(vote_sum = as.character(vote_sum),
         primary_06 = as.character(primary_06)) |> 
  ggplot(aes(age, fill = primary_06)) +
    geom_histogram(alpha = 0.5, bins = 30, position = "identity") +
    facet_wrap(~ vote_sum) + 
    coord_cartesian(xlim = c(20, 90)) +
    scale_fill_manual(values = c("#F8766D", "#00BFC4"), 
                      labels = c("Did not vote", "Voted")) +
    theme_classic() +
    labs(title = "Age Distributions by Voting History",
         subtitle = 
           "Those who have voted more are more likely to vote in the 2006 primary", 
         x = "Age", 
         y = "Number of Times Voted", 
         fill = "Voted in 2006 Primary",
         caption = "Gerber, Green, and Larimer (2008)") 

shaming_p1
```

### Exercise 1

`skim()` the `shaming` data set.

```{r michigan-voter-exper-1, exercise = TRUE}

```

```{r michigan-voter-exper-1-hint-1, eval = FALSE}
skim(...)
```

### 

Note that this data set has no missing values.

### Exercise 2

Let's convert the voting records for the election before the 2006 primary into a binary system (using only 0 and 1 as answers). We will use `pivot_longer()` to transform the six columns into one column. Make the change into binary. Then use `pivot_wider()` to transform the data back into the original format. *However*, since there is no column that uniquely identifies each row, `pivot_wider()` cannot match data with the data that belonged to its original row. We've run into a problem.

### 

Let's solve this problem. We must create a variable that uniquely identifies each row. Begin a pipe the `shaming` data set. Then use `mutate` to create the column `id`. Inside `mutate`, set `id` equal to the expression `1:length(primary_06)`(an expression for an integer series starting from 1 and ending with the number of rows in the column `primary_06`)

```{r michigan-voter-exper-2, exercise = TRUE}

```

```{r michigan-voter-exper-2-hint-1, eval = FALSE}
... |> 
  mutate(id = ...)
```

```{r michigan-voter-exper-2-hint-2, eval = FALSE}
... |> 
  mutate(id = 1:length(primary_06))
```

### 

By default, `pivot_wider()` attempts to use all columns to define unique rows. In many data sets, these columns provide uniquely identifying information on their own.

### Exercise 3

Now we can start making this a binary system. We first use `pivot_longer()` to combine the contents of the columns from `primary_00` to `general_04` into one column. Then, we set the column names to go to a new column called `"election"` and the values to go to a column called, `"voted"`. Remember that you can use the `:` to make consecutive column selection easier.

```{r michigan-voter-exper-3, exercise = TRUE}

```

<button onclick="transfer_code(this)">

Copy previous code

</button>

```{r michigan-voter-exper-3-hint-1, eval = FALSE}
... |> 
  pivot_longer(cols = ...:...,
               ...)
```

```{r michigan-voter-exper-3-hint-2, eval = FALSE}
pivot_longer(cols = primary_00:general_04,
             names_to = "...",
             values_to = "...")
```

### 

Because the columns we have selected are consecutive, we can use the `:` to more quickly name them as a group. Look at the `tidyr_tidy_select` help page for more operators which make it easier to select variables.

### Exercise 4

Now it's time to change this into a binary system. Use `mutate()` to change the `voted` column into a column of the same name, but which has `1` and `0` instead of `"Yes"` and `"No"`. Inside the `mutate()`, set `voted` equal to an `if_else()` function whose condition is `voted ==  "Yes"`. Set the results, for if true and false, to be `1` and `0`, respectively.

```{r michigan-voter-exper-4, exercise = TRUE}

```

<button onclick="transfer_code(this)">

Copy previous code

</button>

```{r michigan-voter-exper-4-hint-1, eval = FALSE}
... |> 
 mutate(voted = if_else(voted == "...", 1, ...))
```

### Exercise 5

Now onto the next part. Use `pivot_wider()` to transform the contents of `voted` back into six columns. Set the names of these new columns to come from `election` and their values to come from `voted`

```{r michigan-voter-exper-5, exercise = TRUE}

```

<button onclick="transfer_code(this)">

Copy previous code

</button>

```{r michigan-voter-exper-5-hint-1, eval = FALSE}
... |> 
pivot_wider(names_from = ..., 
              values_from = ...)
```

### 

Notice how quotes are not needed when referring to existing columns in `pivot_wider()` but are needed when creating new columns in `pivot_longer()`.

### Exercise 6

After that, add the function `rowwise()` to your pipe. Then use `mutate()` to create a new column called `vote_sum`. Inside `mutate()`, set `vote_sum` equal to the sum of the columns from `primary_00` to `general_04`.

```{r michigan-voter-exper-6, exercise = TRUE}

```

<button onclick="transfer_code(this)">

Copy previous code

</button>

```{r michigan-voter-exper-6-hint-1, eval = FALSE}
... |> 
  rowwise() |> 
  mutate(vote_sum = sum(primary_00, general_00, primary_02, general_02, 
                        primary_04, general_04))
```

### 

If we did not use `rowwise()` here, `mutate()` would interpret the sum as a summing together of all the rows in each of these six columns. The new column would have the same, very large value throughout all of its rows.

### Exercise 7

Use `mutate()` to change the `vote_sum` column into a column of the same name, whose contents are characters rather than numbers. Use the `as.character()` function.

```{r michigan-voter-exper-7, exercise = TRUE}

```

<button onclick="transfer_code(this)">

Copy previous code

</button>

```{r michigan-voter-exper-7-hint-1, eval = FALSE}
mutate(... = as.character(...))
```

```{r michigan-voter-exper-7-hint-2, eval = FALSE}
mutate(vote_sum = as.character(vote_sum))
```

### 

We are using these numbers which represent vote histories as different categories for our age distributions. Later when we put these on the y-axis, numeric form would confuse `ggplot()` because it would think we are putting a continuous variable on the y-axis.

### Exercise 8

Now use a new `mutate()` to change the `primary_06` column into a column of the same name, whose contents are characters rather than numbers. Use the `as.character()` function.

```{r michigan-voter-exper-8, exercise = TRUE}

```

<button onclick="transfer_code(this)">

Copy previous code

</button>

```{r michigan-voter-exper-8-hint-1, eval = FALSE}
... |> 
  mutate(... = as.character(...))
```

```{r michigan-voter-exper-8-hint-2, eval = FALSE}
... |> 
  mutate(primary_06 = as.character(primary_06))
```

### 

Likewise, when we use `primary_06`as a fill, `ggplot()` if it were a numeric variable.

### Exercise 9

`select()` the columns relevant for our plot: `age`, `vote_sum`, and `primary_06`.

```{r michigan-voter-exper-9, exercise = TRUE}

```

<button onclick="transfer_code(this)">

Copy previous code

</button>

```{r michigan-voter-exper-9-hint-1, eval = FALSE}
... |> 
select(..., ..., ...)
```

### Exercise 10

Now it's time to graph it. Continue your pipe with `ggplot()`. Map `age` to the x-axis, and `primary_06` to `fill`. Add the layer `geom_histogram()`.

```{r michigan-voter-exper-10, exercise = TRUE}

```

<button onclick="transfer_code(this)">

Copy previous code

</button>

```{r michigan-voter-exper-10-hint-1, eval = FALSE}
... |> 
  ggplot(mapping = aes(x = ..., 
                        fill = ...)) +
  geom_histogram()
```

### 

Remember that histogram has a default `y` which are the counts it tabulates for each bin from our data.

### Exercise 11

Within `geom_histogram()`, set the argument `position` equal to `"identity"`, set the argument `alpha` = `0.5`, and set the argument `bins` = 30.

```{r michigan-voter-exper-11, exercise = TRUE}

```

<button onclick="transfer_code(this)">

Copy previous code

</button>

```{r michigan-voter-exper-11-hint-1, eval = FALSE}
... |> 
  geom_histogram(position = "identity", alpha =  ..., 
                        bins = ...)
```

### 

For histogram and bar plots, `position` is by default `"stacked"`, which stacks different colors of the same x value on top of one another. `position` being equal to `"identity"`, makes it so that different colors overlap, and that both can be seen when the opacity of the colors is increased.

### Exercise 12

Add the layer `facet_wrap()` to facet the histogram by `vote_sum`.

```{r michigan-voter-exper-12, exercise = TRUE}

```

<button onclick="transfer_code(this)">

Copy previous code

</button>

```{r michigan-voter-exper-12-hint-1, eval = FALSE}
Within facet_wrap(), remember to put a ~ before the variable you want to facet by.
```

```{r michigan-voter-exper-12-hint-2, eval = FALSE}
... + 
  facet_wrap(~...) 
```

### Exercise 13

Add the layer `coord_cartesian` and use the argument to limit the x-axis domain to a minimum of 20 and a maximum of 90

```{r michigan-voter-exper-13, exercise = TRUE}

```

<button onclick="transfer_code(this)">

Copy previous code

</button>

```{r michigan-voter-exper-13-hint-1, eval = FALSE}
... +
  coord_cartesian(xlim = c(...,...))
```

### Exercise 14

Ok now what we're going to do is add the layer `scale_fill_manual` to change the legend options from `0` and `1` to `Did not vote` and `vote`. Add `scale_fill_manual` and set the first argument `values`, which determines the fill colors, equal to `c("#F8766D", "#00BFC4")`, the default red and blue in ggplot. Set the second argument `labels` equal to `c("Did not vote", "Voted")`

```{r michigan-voter-exper-14, exercise = TRUE}

```

<button onclick="transfer_code(this)">

Copy previous code

</button>

```{r michigan-voter-exper-14-hint-1, eval = FALSE}
... +
  scale_fill_manual(values = c("...", "..."), labels = c("...", "..."))
```

### Exercise 15

Adjust the feel of the graph by adding the theme `theme_classic()`. To finish your plot, use `labs()` to give the graph a title, subtitle, axes labels, legend heading, and caption.

```{r michigan-voter-exper-15, exercise = TRUE}

```

<button onclick="transfer_code(this)">

Copy previous code

</button>

```{r michigan-voter-exper-15-hint-1, eval = FALSE}
... +
  theme_classic()  +
  labs(... ,
     fill = "...",
     ...)
```

Reminder: This is what your plot should look like.

```{r shaming15-answer}
shaming_p1
```

### Exercise 16

Let's make the following plot, but this time, we're using density plots to visualize the same data.

```{r}
shaming_p2 <- shaming |> 
  mutate(id = 1:length(primary_06)) |> 
  pivot_longer(cols = primary_00:general_04,
               names_to = "year",
               values_to = "voted") |> 
  mutate(voted = if_else(voted == "Yes", 1, 0)) |> 
  pivot_wider(names_from = year, 
              values_from = voted) |> 
  rowwise() |> 
  mutate(vote_sum = sum(primary_00, general_00, primary_02, general_02, primary_04, general_04)) |> 
  mutate(vote_sum = as.character(vote_sum),
         primary_06 = as.character(primary_06)) |> 
  select(age, vote_sum, primary_06) |> 
  ggplot(aes(x = age, fill = primary_06)) +
  geom_density(alpha = 0.3) +
  facet_wrap(~ vote_sum) + 
  coord_cartesian(xlim = c(20, 90)) +
  scale_fill_manual(values = c("#F8766D", "#00BFC4"), labels = c("Did not vote", "Voted")) +
  theme_classic() +
  labs(title = "Age Distributions by Voting History",
       subtitle = "Among those of the same voting history, those who voted in the \n 2006 primary are of generally similar ages to those who did not", 
       x = "Age", 
       y = "Density", 
       fill = "2006 Primary",
       caption = "Gerber, Green, and Larimer (2008)") 
  
shaming_p2
```

### 

We have provided you with the beginning of code for the last plot (YOU'RE TOTALLY WELCOME). Continue your pipe with `ggplot()`. Map `age` to the x-axis, and `primary_06` to `fill`. Add the layer `geom_density()`. Within `geom_density()`, set `alpha` equal to 0.3.

```{r michigan-voter-exper-16, exercise = TRUE}
shaming |> 
  mutate(id = 1:length(primary_06)) |> 
  pivot_longer(cols = primary_00:general_04,
               names_to = "year",
               values_to = "voted") |> 
  mutate(voted = if_else(voted == "Yes", 1, 0)) |> 
  pivot_wider(names_from = year, 
              values_from = voted) |> 
  rowwise() |> 
  mutate(vote_sum = sum(primary_00, general_00, primary_02, general_02, primary_04, general_04)) |> 
  mutate(vote_sum = as.character(vote_sum),
         primary_06 = as.character(primary_06)) |> 
  select(age, vote_sum, primary_06) 
```

```{r michigan-voter-exper-16-hint-1, eval = FALSE}
... |> 
  ggplot(aes(..., fill = ...)) +
  geom_density(alpha = ...) 
```

### Exercise 17

Add the layer `facet_wrap()` to facet the histogram by `vote_sum`. Add the layer `coord_cartesian` and use the argument to limit the x-axis domain to a minimum of 20 and a maximum of 90.

```{r michigan-voter-exper-17, exercise = TRUE}

```

<button onclick="transfer_code(this)">

Copy previous code

</button>

```{r michigan-voter-exper-17-hint-1, eval = FALSE}
Within facet_wrap(), remember to put a ~ before the variable you want to facet by.
```

```{r michigan-voter-exper-17-hint-2, eval = FALSE}
... +
  coord_cartesian(xlim = c(...,...))
```

### Exercise 18

Ok now we're gonna add the layer `scale_fill_manual` to change the legend options from `0` and `1` to `Did not vote` and `vote`. Add the layer `scale_fill_manual` and set the first argument `values`, which determines the fill colors, equal to `c("#F8766D", "#00BFC4")`, the default red and blue in ggplot. Set the second argument `labels` equal to `c("Did not vote", "Voted")`.

```{r michigan-voter-exper-18, exercise = TRUE}

```

<button onclick="transfer_code(this)">

Copy previous code

</button>

```{r michigan-voter-exper-18-hint-1, eval = FALSE}
... +
  scale_fill_manual(values = c("...", "..."), labels = c("...", "..."))
```

### Exercise 19

Ok final part. Adjust the feel of your graph by adding `theme_classic()`. Finally, add labels to your graph. Pay particular attention to how the y-axis label and subtitle should change compared to the previous plot. If, as in our example, the subtitle is too long insert `\n` at the point in your subtitle where you would like a new line.

```{r michigan-voter-exper-19, exercise = TRUE}

```

<button onclick="transfer_code(this)">

Copy previous code

</button>

```{r michigan-voter-exper-19-hint-1, eval = FALSE}
... +
  theme_classic()  +
  labs(... ,
    subtitle = "...\n..."
     fill = "...",
     ...)
```

Reminder: This is what your plot should look like.

```{r shaming19-answer}
shaming_p2
```

### 

Plot outputs have a margin about the size of the R markdown margin. For overlong subtitles such as the one we used in the example, `ggplot()` will not wrap them automatically. To create a new line, insert `\n` meaning "new-line" at the appropriate point. `ggplot()` will print the following text on a new line and will not print `\n`.

### Exercise 20

Let's create the following plot, using another `ggdist` function to visualize the same data with point interval graphs.

```{r}
shaming_p3 <- shaming |> 
  mutate(id = 1:length(primary_06)) |> 
  pivot_longer(cols = primary_00:general_04,
               names_to = "year",
               values_to = "voted") |> 
  mutate(voted = if_else(voted == "Yes", 1, 0)) |> 
  pivot_wider(names_from = year, 
              values_from = voted) |> 
  rowwise() |> 
  mutate(vote_sum = sum(primary_00, general_00, primary_02, general_02, primary_04, general_04)) |> 
  mutate(vote_sum = as.character(vote_sum),
         primary_06 = as.character(primary_06)) |> 
  select(age, vote_sum, primary_06) |> 
  ggplot(aes(x = age, y = vote_sum, color = primary_06)) +
  stat_pointinterval(position = "dodge") +
  scale_color_manual(values = c("#F8766D", "#00BFC4"), labels = c("Did not vote", "Voted")) +
  theme_classic() +
  labs(title = "Age Distributions by Voting History",
       subtitle = "With the exception of the youngest, those who voted in the 2006 \n primary were always slightly older than those who did not", 
       x = "Age", 
       y = "Number of Times Voted",
       color = "2006 Primary",
       caption = "Gerber, Green, and Larimer (2008)") 

shaming_p3
```

### 

The middle point represents the mean of the distribution, the bolded line represents the 25th to 75th percentile, and the unbolded line represents the 5th to 95th percentiles.

### 

We have provided you with the beginning of code for the last plot (Again, you are so very welcome). Continue your pipe with `ggplot()`. Map `age` to the x-axis, `vote_sum` to the y-axis, and `primary_06` to `color`. Add the layer `stat_pointinterval()` and specify `position` to be `"dodge"`.

```{r michigan-voter-exper-20, exercise = TRUE}
shaming |> 
  mutate(id = 1:length(primary_06)) |> 
  pivot_longer(cols = primary_00:general_04,
               names_to = "year",
               values_to = "voted") |> 
  mutate(voted = if_else(voted == "Yes", 1, 0)) |> 
  pivot_wider(names_from = year, 
              values_from = voted) |> 
  rowwise() |> 
  mutate(vote_sum = sum(primary_00, general_00, primary_02, general_02, primary_04, general_04)) |> 
  mutate(vote_sum = as.character(vote_sum),
         primary_06 = as.character(primary_06)) |> 
  select(age, vote_sum, primary_06)
```

```{r michigan-voter-exper-20-hint-1, eval = FALSE}
... |> 
  ggplot(aes(..., color = ...)) +
  stat_pointinterval() 
```

### 

We use `color` instead of `fill` since we are dealing with lines rather than filled shapes.

### Exercise 21

Add the layer `coord_cartesian` and use the argument to limit the x-axis domain to a minimum of 20 and a maximum of 81, since none of our distributions have a 95th percentile greater than 81.

```{r michigan-voter-exper-21, exercise = TRUE}

```

<button onclick="transfer_code(this)">

Copy previous code

</button>

```{r michigan-voter-exper-21-hint-1, eval = FALSE}
... +
  coord_cartesian(xlim = c(...,...))
```

### Exercise 22

Ok now we're gonna add the layer `scale_color_manual` to change the legend options from `0` and `1` to `Did not vote` and `Voted`. Use `scale_color_manual` and set the first argument `values`, which determines the fill colors, equal to `c("#F8766D", "#00BFC4")`, the default red and blue in ggplot. Set the second argument `labels` equal to `c("Did not vote", "Voted")`

```{r michigan-voter-exper-22, exercise = TRUE}

```

<button onclick="transfer_code(this)">

Copy previous code

</button>

```{r michigan-voter-exper-22-hint-1, eval = FALSE}
... +
  scale_color_manual(values = c("...", "..."), labels = c("...", "..."))
```

### 

Just as you mapped to `color` instead of `fill`, you must use `scale_color_manual` in place of `scale_fill_manual`.

### Exercise 23

Now adjust the feel of your graph by adding `theme_classic()`. Finally, add labels to your graph. Use `color` rather than `fill` as we have done above. Pay particular attention to how the y-axis label and subtitle should change compared to the previous plot.

```{r michigan-voter-exper-23, exercise = TRUE}

```

<button onclick="transfer_code(this)">

Copy previous code

</button>

```{r michigan-voter-exper-23-hint-1, eval = FALSE}
... +
  theme_classic()  +
  labs(... ,
     fill = "...",
     ...)
```

Reminder: This is what your plot should look like.

```{r shaming22-answer}
shaming_p3
```

## Summary

### 

This tutorial covered [Chapter 6: Data tidying](https://r4ds.hadley.nz/data-tidy.html) from [*R for Data Science (2e)*](https://r4ds.hadley.nz/) by Hadley Wickham, Mine Çetinkaya-Rundel, and Garrett Grolemund.

```{r download-answers, child = system.file("child_documents/download_answers.Rmd", package = "tutorial.helpers")}
```
