---
title: Data tidying
author: Miriam Heiss
tutorial:
  id: data-tidying
output:
  learnr::tutorial:
    progressive: yes
    allow_skip: yes
runtime: shiny_prerendered
description: 'Tutorial for Chapter 6: Data tidying'
---

```{r setup, include = FALSE}
library(learnr)
library(tidyverse)
library(tutorial.helpers)
library(primer.data)
library(ggdist)
library(scales)
knitr::opts_chunk$set(echo = FALSE)
options(tutorial.exercise.timelimit = 60, 
        tutorial.storage = "local") 

# Needed for lengthening data plot

billboard_longer <- billboard |>
  pivot_longer(cols = starts_with("wk"),
               names_to = "week",
               values_to = "rank",
               values_drop_na = TRUE) |>
  mutate(week = parse_number(week))

# Needed for widening data plot
cms_pivoted <- cms_patient_experience |>
  pivot_wider(id_cols = starts_with("org"), 
            names_from = measure_cd,
            values_from = prf_rate)
```

```{r copy-code-chunk, child = system.file("child_documents/copy_button.Rmd", package = "tutorial.helpers")}
```

```{r info-section, child = system.file("child_documents/info_section.Rmd", package = "tutorial.helpers")}
```

<!-- DK: We need a lot of knowledge drops. And some questions split up into two or three steps. -->

## Introduction
### 

This tutorial covers [Chapter 6: Data tidying](https://r4ds.hadley.nz/data-tidy.html) from [*R for Data Science (2e)*](https://r4ds.hadley.nz/) by Hadley Wickham, Mine Çetinkaya-Rundel, and Garrett Grolemund.

Please restart your R session before proceeding. The short cut key for doing so is `Cmd-Shift-0`.

## Tidy Data
### 

You can represent the same underlying data in multiple ways by organizing values in a given dataset in different orders. But not all data is equally easy to use.

### Exercise 1

In the console, load in the *tidyverse* package. Copy and paste the loading message (the thing with the conflicts and check marks) into the box below.

```{r tidy-data-1}
question_text(NULL,
	answer(NULL, correct = TRUE),
	allow_retry = TRUE,
	try_again_button = "Edit Answer",
	incorrect = NULL,
	rows = 3)
```

### Exercise 2

Restart your R session with `Cmd-Shift-0`

### 

Load the `tidyverse` package again and copy and paste the first line of the section labeled `Attatching core tidyverse packages` into the box below.

```{r tidy-data-2}
question_text(NULL,
	answer(NULL, correct = TRUE),
	allow_retry = TRUE,
	try_again_button = "Edit Answer",
	incorrect = NULL,
	rows = 3)
```

### 

It should just show [dplyr](https://dplyr.tidyverse.org/) and [readr](https://readr.tidyverse.org/). 

### 

`dplyr` has all your basic data transformation functions like `mutate()`, `filter()`, and `summarise()`. `readr` allows you to read data that is in the form of a csv, tsv, or spreadsheet.

### Exercise 3

Copy and paste the third line of the loading message into the box below.

```{r tidy-data-3}
question_text(NULL,
	answer(NULL, correct = TRUE),
	allow_retry = TRUE,
	try_again_button = "Edit Answer",
	incorrect = NULL,
	rows = 3)
```

It should have [ggplot2](https://ggplot2.tidyverse.org/) and [tibble](https://tibble.tidyverse.org/). 

### 

`ggplot2` is how we make plots and maps in R. You provide the data, tell ggplot2 how to map variables to aesthetics, what graphical primitives to use, and it takes care of the details. `tibble` allows you to create data frames that are lazy and surly: they do less (i.e. they don’t change variable names or types, and don’t do partial matching) and complain more (e.g. when a variable does not exist).


### Exercise 4

Copy and paste the 4th line of the loading message into the box below. 

```{r tidy-data-4}
question_text(NULL,
	answer(NULL, correct = TRUE),
	allow_retry = TRUE,
	try_again_button = "Edit Answer",
	incorrect = NULL,
	rows = 3)
```

### 

It should show [lubridate](https://lubridate.tidyverse.org/) and [**tidyr**](https://tidyr.tidyverse.org/)

### 

`lubridate` is a more intuitive way to change and make dates in R, and allows for all of the quirks that our time has (such as leap days and daylight savings) that computers and programs don't understand. `tidyr` is the most important package for this tutorial. It makes sure that you have **Tidy Data**. The three main rules for having tidy data are: 

1. Each variable is a column; each column is a variable.
2. Each observation is a row; each row is an observation.
3. Each value is a cell; each cell is a single value.


### Exercise 5

Restart your R session again with `Cmd-Shift-0`.

### 

Run the code provided in the box below.

```{r tidy-data-5, exercise = TRUE}
table1
table2
table3
```

### 

These tables show Tuberculosis cases in three countries: China, Brazil, and Afghanistan.

### 

They all show the same data (country, year, population, cases), but one of them is going to be way easier to work with. Let's see if we can identify the tidy table.

### Exercise 6

Let's  take a look at `table2` first. Type and run `table2` in the box below to see it.

```{r tidy-data-6, exercise = TRUE}

```

```{r tidy-data-6-hint-1, eval = FALSE}
table2
```

### 

`table2` is a tibble with 4 columns and 12 rows. Let's see if the rules apply.

### Exercise 7

```{r tidy-data-7}
question_text(NULL,
	message = "Place correct answer here.",
	answer(NULL, correct = TRUE),
	allow_retry = FALSE,
	incorrect = NULL,
	rows = 6)
```

### 

### Exercise 8

These are the rules for tidy data: 

1. Each variable is a column; each column is a variable.
2. Each observation is a row; each row is an observation.
3. Each value is a cell; each cell is a single value.

### 

In the column `type`, it switches in between `cases` and `population`. In the column `count` it also switches in between the number of cases and the population. 

### 

Using the three rules, answer the questions below.


### Exercise 9

```{r tidy-data-9}
question_text("Are all the possible variables columns?",
	message = "No",
	answer(NULL, correct = TRUE),
	allow_retry = FALSE,
	incorrect = NULL,
	rows = 3)
```

### 

All the possible variables are `country`, `year`, `cases`, and `population`. The `type` column has two of the variables that are not observations, breaking the first two rules. The `count` column has observations that switch between cases and population, because there is no `cases` or `population` column.


### Exercise 10

Now let's look at `table3`. Type and run `table3` in the box below to see it.

```{r tidy-data-10, exercise = TRUE}

```

```{r tidy-data-10-hint-1, eval = FALSE}
table3
```

### 

Now, this looks like a very nice dataset with 3 columns and 6 rows, but it breaks one of the rules. These are the rules for tidy data: 

1. Each variable is a column; each column is a variable.
2. Each observation is a row; each row is an observation.
3. Each value is a cell; each cell is a single value.


### Exercise 11

Using the three rules, answer the question below.

```{r tidy-data-11}
question_text("Are all the cells a single value?",
	message = "No",
	answer(NULL, correct = TRUE),
	allow_retry = FALSE,
	incorrect = NULL,
	rows = 3)
```

### 

Each of the cells in the `rate` column are not a single observation. They are `population` divided by `cases`, so in order for the data to be considered tidy, we would need to separate those columns.

### Exercise 12

Let's look at `table1`. Type and run `table1` in the box below to see it.

```{r tidy-data-12, exercise = TRUE}

```

```{r tidy-data-12-hint-1, eval = FALSE}
table1
```

### 

`table1` is a tibble with 4 columns and 6 rows. Let's see if the rules apply.

### Exercise 13

As a refresher, these are the rules for tidy data: 

1. Each variable is a column; each column is a variable.
2. Each observation is a row; each row is an observation.
3. Each value is a cell; each cell is a single value.

### 

Using the three rules, answer the questions below.


```{r tidy-data-13}
question_text("Are all the rules true in this data?",
	message = "Yes",
	answer(NULL, correct = TRUE),
	allow_retry = FALSE,
	incorrect = NULL,
	rows = 3)
```

### 

This is a very good example of a tidy dataset. All the variables are columns. All the columns are variables. All the observations are rows, all the rows are observations, all the values are cells, and all the cells are single values.

### Exercise 14

Using the tidy data from `table1`, we are going to make this plot: 

```{r echo = FALSE}
plot1 <- ggplot(data = table1, mapping = aes(x = year, y = cases)) +
  geom_line(aes(group = country)) +
  geom_point(aes(color = country, shape = country)) +
  scale_x_continuous(breaks = c(1999, 2000)) +
  scale_y_continuous(labels = scales::comma) +
  labs(title = "Tuberculosis cases in three countries",
       subtitle = "Between 1999 and 2000",
       x = "Year",
       y = "Number of cases",
       color = "Country",
       shape = "Country")
```

```{r}
plot1
```

### 

Start making a plot with `ggplot`. Use `table1` as your data.

```{r tidy-data-14, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r tidy-data-14-hint-1, eval = FALSE}
ggplot(data = ...)
```

### Exercise 15

Add an `aes` with `year` for your x and `cases` for your y.

```{r tidy-data-15, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>


```{r tidy-data-15-hint-1, eval = FALSE}
ggplot(... = ..., mapping = aes(x = ..., y = ...))
```

### Exercise 16

Add `geom_line()` to your plot too.

```{r tidy-data-16, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>


```{r tidy-data-16-hint-1, eval = FALSE}
ggplot(...(...)) + 
  geom_...()
```

### Exercise 17

In `geom_line()`, we need to mess around with a couple things in order to get three separate lines instead of one big "N" on our plot. We need to add an `aes` inside of `geom_line(), with group equaling country (inside parenthesis).

```{r tidy-data-17, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>


```{r tidy-data-17-hint-1, eval = FALSE}
ggplot(...(...)) + 
  geom_line(aes(group = country))
```

### Exercise 18

Now, we need to add a second `geom`, `geom_point()`. Inside `geom_point()` we need another `aes`, this time with color and shape both equaling country. This will make three distinctive lines that you can clearly see which one is which.

```{r tidy-data-18, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>


```{r tidy-data-18-hint-1, eval = FALSE}
ggplot(...(...)) + 
  geom_...(...) +
  geom_point(aes(color = ..., shape = ...))
```

### 

Technically you could tell before, but this will make it easier to see which one is which. When a categorical variable is mapped to an aesthetic, `ggplot()` will automatically assign a unique value of the aesthetic to each unique level of the variable, a process known as scaling.

### Exercise 19

Let's fix the x and y axes

### 

For the x axis, we need make just two breaks, 1999 and 2000. To do this, we will need to add a layer (using `+`). The layer we are going to add is `scale_x_continuous`. We need to change what the breaks are, using `breaks = c(1999, 2000)`.

```{r tidy-data-19, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>


```{r tidy-data-19-hint-1, eval = FALSE}
ggplot(...(...)) + 
  geom_...(...) +
  geom_...(...) +
  scale_x_continuous(breaks = c(..., ....))
```

### Exercise 20

To add commas to the y axis, we will use `scale_y_continuous()` on another layer. Inside of the parenthesis, we will use `labels = scales::comma`. 

```{r tidy-data-20, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r tidy-data-20-hint-1, eval = FALSE}
ggplot(...(...)) + 
  geom_...(...) +
  geom_...(...) +
  scale_...(...) +
  scale_y_continuous(... = scales::comma)
```

### Exercise 21

The last step is to add labels to our plot.

### 

Reminder: This is what your plot should look like.

```{r}
plot1
```

### 

```{r tidy-data-21, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>


```{r tidy-data-21-hint-1, eval = FALSE}
... +
  labs(title = "...",
       subtitle = "...",
       x = "...",
       y = "..."
       color = "Country",
       shape = "Country")
```

### 

To add labels to the key on the side, you will need to label both the shape and the color "Country" (as that is what the key is telling us)

### 

Now that you understand what tidy data is and how to identify it using the three rules, we need to learn how to make data tidy.

## Lengthening Data
### 

The rules of tidy data might seem so obvious that you wonder if you’ll ever encounter data that isn’t tidy. Unfortunately, almost all real data is untidy, and you will need to clean it.

### Exercise 1

Restart your R session with `Cmd-Shift-0`. The tutorial will keep your place.

### 

For this section, we will need to see the `billboard` dataset. Type and run `billboard` in the box below

```{r lengthening-data-1, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r lengthening-data-1-hint-1, eval = FALSE}
billboard
```

### 

The `billboard` dataset consists of song rankings from the Billboard top 100 songs in the year 2000. 

### Exercise 2

This data is going to be extremely difficult to work with. It has *79* columns!

### 

Start a pipe with `billboard`.

```{r lengthening-data-2, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r lengthening-data-2-hint-1, eval = FALSE}
...|>
```

### 

How exactly can we tidy a dataset that is this wide? First, we need to figure out where the problem is coming from. The majority of the width is from the 76 columns that tell us the rank of the song on a given week in the year 2000.

### Exercise 3

Add `pivot_longer()` to the pipe.

```{r lengthening-data-3, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r lengthening-data-3-hint-1, eval = FALSE}
billboard |>
  pivot_...()
```

### 

We can use `pivot_longer()`, and condense the other 76 columns into just two columns: `week` and `rank`.

### Exercise 4

Inside `pivot_longer()`, we need to grab all the columns that start with `wk`, because that is something that all 76 have in common. We can do this by using `cols = starts_with("wk")`.

```{r lengthening-data-4, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r lengthening-data-4-hint-1, eval = FALSE}
billboard|>
  pivot_longer(cols = starts_with("..."))
```

### 

`cols` specifies which columns need to be pivoted, i.e. which columns aren’t variables. This argument uses the same syntax as `select()` so here we could use `!c(artist, track, date.entered)` or `starts_with("wk")`

### Exercise 5

Now, we need to specify what `pivot_longer()` needs to do with all these columns.

### 

We need to turn these 76 columns and all these observations into two columns: `rank` and `week`. We can do this by adding `names_to = "week"`, and `values_to = "rank` to `pivot_longer()`. (Make sure you have commas separating each line!) 

```{r lengthening-data-5, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r lengthening-data-5-hint-1, eval = FALSE}
billboard|>
  pivot_longer(cols = starts_with("wk"),
             names_to = "...",
             values_to = "...")
```

### 

`names_to` names the variable stored in the column names, we named that variable week, and `values_to` names the variable stored in the cell values, we named that variable rank.

### 

### Exercise 6

Now we need to get rid of all the NA's in the dataset. We can do this by using `values_drop_na = TRUE` inside `pivot_longer()`

```{r lengthening-data-6, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r lengthening-data-6-hint-1, eval = FALSE}
pivot_longer(...,
             ...,
             ...,
             values_drop_na = ...)
```

### 

These NAs don’t really represent unknown observations; they were forced to exist by the structure of the dataset2, so we can ask `pivot_longer()` to get rid of them by setting `values_drop_na = TRUE`.

### Exercise 7

Although this data is considered tidy, we can still clean it up to make it more understandable. We can get rid of all the extra `wk`'s in the `week` column to make everything look nicer.

### 

We do this by continuing our pipe with `mutate()`.

```{r lengthening-data-7, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r lengthening-data-7-hint-1, eval = FALSE}
(...)|>
  mutate()
```

### 

With `mutate()` we can convert character strings to numbers. This way of using `mutate()` doesn't create a new column, but changes it.

### Exercise 8

Inside of `mutate()`, we can set `week` equal to `parse_number(week)`. This will get rid of all the `wk`'s, replacing them with numbers.

```{r lengthening-data-8, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r lengthening-data-8-hint-1, eval = FALSE}
|>
  mutate(... = parse_number(...))
```

### 

`parse_number()` is a handy function that will extract the first number from a string, ignoring all other text.

### Exercise 9

Assign this mutated dataset to `billboard_longer`, using `<-`

```{r lengthening-data-9, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r lengthening-data-9-hint-1, eval = FALSE}
billboard_longer <- billboard |>
  pivot_longer(...)
```

### 

This will make it easier to make this dataset into a plot, that way we don't have to pipe the entire process for the improved dataset into `ggplot()`.

### 

We are going to be making this plot:

```{r echo = FALSE}
plot2 <- billboard_longer |> 
  ggplot(aes(x = week, y = rank, group = track)) + 
  geom_line(alpha = 0.25) + 
  scale_y_reverse()+
  labs(title = "Billboard top 100 rankings in the year 2000",
       subtitle = "Over the course of 76 Weeks",
       x = "Week",
       y = "Rank")
```

```{r}
plot2
```

### Exercise 10

Restart your R session again with `Cmd-Shift-0`.

### 

Start a pipe with `billboard_longer`.

```{r lengthening-data-10, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r lengthening-data-10-hint-1, eval = FALSE}
billboard_longer |>
```

### 

<!-- knowledge -->

### Exercise 11

Continue your pipe with `ggplot()`.

```{r lengthening-data-11, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r lengthening-data-11-hint-1, eval = FALSE}
billboard_longer |>
  ggplot()
```

### 

### Exercise 12

Inside of `ggplot()`, we need to add multiple things. First is `aes()`. Inside of `aes()`, we need x to equal `week`, y to equal `rank`, and group to equal `track`.

```{r lengthening-data-12, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r lengthening-data-12-hint-1, eval = FALSE}
ggplot(aes(x = ..., 
           y = ..., 
           group = ...))
```

### 

We didn't have to specify data because this plot is part of a pipe where the data is at the top. We could also make this plot without the pipe, with `billboard_longer` where data would go.

### Exercise 13

Next, we need to ad a `geom`. We will be using `geom_line()`. Inside `geom_line()` we need to change the opacity of the lines using `alpha`. We will set it to 0.25.

```{r lengthening-data-13, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r lengthening-data-13-hint-1, eval = FALSE}
ggplot(...) +
  geom_line(alpha = ...)
```

### 

<!-- knowledge -->

### Exercise 14

This plot is deceiving. This plot seems to tell us that so many songs get the top place, because so many lines are concentrated there. But rankings go from 100 (lowest) to 1 (highest), so we should probably flip this plot upside down.

### 

We do this with `scale_y_reverse()`. That will put 1 at the top and 100 at the bottom.

```{r lengthening-data-14, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r lengthening-data-14-hint-1, eval = FALSE}
geom_line(...) +
  scale_y_reverse()
```

### Exercise 15

Now, we need to add labels. Reminder: This is what your plot should look like

```{r}
plot2
```

### 

```{r lengthening-data-15, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r lengthening-data-15-hint-1, eval = FALSE}
labs(title = ...,
     subtitle = ...,
     x = ...,
     y = ...)
```

### 

Now you know how to pivot really wide datasets, and when to pivot. Now, we need to learn about what we need to do with the really long datasets.

### 

## Widening Data
### 

`pivot_wider()` makes datasets wider by increasing columns and reducing rows and helps when one observation is spread across multiple rows. This seems to arise less commonly in the wild, but it does seem to crop up a lot when dealing with governmental data, such as the Census.

### 

For this section, we'll be using the dataset `cms_patient_experience`, a patient experiences dataset from the Centers of Medicare and Medicaid Services. Please restart your R session before proceeding using `Cmd-Shift-0`.

<!-- DK: Why the need to restart? -->

### Exercise 1

Load in the `tidyverse` in the box below.

```{r widening-data-1, exercise = TRUE}

```

```{r widening-data-1-hint-1, eval = FALSE}
library(...)
```

### 

We are going to be focusing on the package `tidyr`, on the fourth line of the section in the loading message titled: "Attaching core tidyverse packages".

### Exercise 2

In the Console, look up the help page for `tidyr` using `?` before the package name. Copy and paste the Description from the help page into the box below.

```{r widening-data-2}
question_text(NULL,
	answer(NULL, correct = TRUE),
	allow_retry = TRUE,
	try_again_button = "Edit Answer",
	incorrect = NULL,
	rows = 3)
```

### 

Putting a question mark before a package name is a shortcut to looking up the help page. You could always go to the help pane in the bottom left corner, and search for `tidyr`, but it is a much faster way. 

### Exercise 3

Look up the help page for `cms_patient_experience` using two question marks in front of it. Copy and paste the the page down to the section titled "Format".

```{r widening-data-3}
question_text(NULL,
	answer(NULL, correct = TRUE),
	allow_retry = TRUE,
	try_again_button = "Edit Answer",
	incorrect = NULL,
	rows = 3)
```

### 

This gives you all the information for this data. What package it comes from, where the data is sourced from, and how to see and use the data.

###

This data comes from the package `tidyr`, so we won't need any other libraries other than the `tidyverse`.

### Exercise 4

Let's take a look at our data.

###

Run `cms_patient_experience` in the box below. 

```{r widening-data-4, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r widening-data-4-hint, eval = FALSE}
cms_patient_experience
```

###

`cms_patient_expercience` is a tibble with 5 columns and 500 observations. 

### Exercise 5

We will want to see 12 rows so we can see how this data is not tidy. You can do that with the function `print()`. Inside `print()` we need to specify `x` and `n`. `x` will be the data, and `n` will be the number of rows we see. Run `print()` with `x` and `n` specified in the box below


```{r widening-data-5, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r widening-data-5-hint, eval = FALSE}
print(x = cms_patient_experience, n = 12)
```

###

This data is not tidy! In the column `measure_cd`, we have 5 repeating rows that would work much better as separate columns.

### Exercise 6

Now, `pivot_wider()` essentially all the same functions as `pivot_longer()`. Remember how we used `names_to` and `values_to` to make more columns from all the extra rows? Well, for `pivot_wider()`, we just have to do the opposite. We will use `names_to` and `values_to` to add more columns to this dataset.

###

Start a pipe with `cms_patient_experience`.

```{r widening-data-6, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r widening-data-6-hint, eval = FALSE}
cms_patient_experience |>
```

### Exercise 7

Add `pivot_wider()` to the pipe.

```{r widening-data-7, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r widening-data-7-hint, eval = FALSE}
cms_patient_experience |>
  pivot_wider()
```

### Exercise 8

Inside `pivot_wider()`, we are going to take the names from the column `measure_cd`. We do this by using `names_from = measure_cd`. Add that inside `pivot_wider()`.

```{r widening-data-8, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r widening-data-8-hint, eval = FALSE}
pivot_wider(... = measure_cd)
```

### Exercise 9

Now, we need values for our new columns to have in them. We are going to take the values from the column `prf_rate` using `values_from = prf_rate`. Add that under `names_from`. (make sure you have a comma!)

```{r widening-data-9, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r widening-data-9-hint, eval = FALSE}
...(...,
      values_from = ...)
```

###

### Exercise 10

This data looks a little bit tidier, but there are still repeated rows in `org_nm`. We need to tell `pivot_wider` to uniquely identify the rows in every column (getting rid of duplicates). The columns we need to target are `org_pac_id` and `org_nm`. We can tell `pivot_wider()` to target these rows by adding `id_cols = starts_with("org")` to `pivot_wider()`.

```{r widening-data-10, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r widening-data-10-hint, eval = FALSE}
pivot_wider(id_cols = starts_with("org"), 
            names_from = ... ,
            values_from = ...)
```

###

This will get rid of any duplicate rows in the columns that start with "org'. And as we can see, there are many duplicates. `id_cols` cut the row count from 500 to 95.


## Michigan voter experiment
### 

The `shaming` data set measures the impact of social pressure on voting. Nearly 350,000 people in Michigan were randomly assigned to 1 of 5 treatment groups before the 2006 Michigan primary. All 5 groups were sent mail before the primary: The "Civic Duty" group had an extra reminder that voting was a civic responsibility. The "Hawthorne" group was told that whether or not they voted, their decision would not be in the public record. The "Self" group was actually *sent* the public record of whether or not they voted in 2004. The "Neighbors" group was sent both their voting record and their neighbors' voting record from 2004. They then measured the impact of these "social pressures" on their choice to vote.

### 

Let's create the following plot

```{r make-shaming1}
# This is bad. If you really study the plot, you can see that it makes little
# sense. The labels, especially for the 6 panels, aren't very good ay all. It is
# unclear what the plot is teaching us. There should be much better things to do.

# Also, I would prefer that we got rid of primer.data entirely from this
# package, since we want it to be as "clean" as possible when we try to get it
# on CRAN.

# So, for now, we leave it here. But, soon, we should replace it with a plot
# which still does some tidying but uses a built-in data set, or something from
# a package we use elsewhere in this package, like babynames.

shaming_p1 <- shaming |> 
  mutate(id = 1:length(primary_06)) |> 
  pivot_longer(cols = primary_00:general_04,
               names_to = "year",
               values_to = "voted") |> 
  mutate(voted = if_else(voted == "Yes", 1, 0)) |> 
  pivot_wider(names_from = year, 
              values_from = voted) |> 
  rowwise() |> 
  mutate(vote_sum = sum(primary_00, general_00, primary_02, general_02, 
                        primary_04, general_04)) |> 
  mutate(vote_sum = as.character(vote_sum),
         primary_06 = as.character(primary_06)) |> 
  ggplot(aes(age, fill = primary_06)) +
    geom_histogram(alpha = 0.5, bins = 30, position = "identity") +
    facet_wrap(~ vote_sum) + 
    coord_cartesian(xlim = c(20, 90)) +
    scale_fill_manual(values = c("#F8766D", "#00BFC4"), 
                      labels = c("Did not vote", "Voted")) +
    theme_classic() +
    labs(title = "Age Distributions by Voting History",
         subtitle = 
           "Those who have voted more are more likely to vote in the 2006 primary", 
         x = "Age", 
         y = "Number of Times Voted", 
         fill = "Voted in 2006 Primary",
         caption = "Gerber, Green, and Larimer (2008)") 

shaming_p1
```

### Exercise 1

`skim()` the `shaming` data set.

```{r michigan-voter-exper-1, exercise = TRUE}

```

```{r michigan-voter-exper-1-hint-1, eval = FALSE}
skim(...)
```

### 

Note that this data set has no missing values.

### Exercise 2

Let's convert the voting records for the election before the 2006 primary into a binary system (using only 0 and 1 as answers). We will use `pivot_longer()` to transform the six columns into one column. Make the change into binary. Then use `pivot_wider()` to transform the data back into the original format. *However*, since there is no column that uniquely identifies each row, `pivot_wider()` cannot match data with the data that belonged to its original row. We've run into a problem.

### 

Let's solve this problem. We must create a variable that uniquely identifies each row. Begin a pipe the `shaming` data set. Then use `mutate` to create the column `id`. Inside `mutate`, set `id` equal to the expression `1:length(primary_06)`(an expression for an integer series starting from 1 and ending with the number of rows in the column `primary_06`)

```{r michigan-voter-exper-2, exercise = TRUE}

```

```{r michigan-voter-exper-2-hint-1, eval = FALSE}
... |> 
  mutate(id = ...)
```

```{r michigan-voter-exper-2-hint-2, eval = FALSE}
... |> 
  mutate(id = 1:length(primary_06))
```

### 

By default, `pivot_wider()` attempts to use all columns to define unique rows. In many data sets, these columns provide uniquely identifying information on their own.

### Exercise 3

Now we can start making this a binary system. We first use `pivot_longer()` to combine the contents of the columns from `primary_00` to `general_04` into one column. Then, we set the column names to go to a new column called `"election"` and the values to go to a column called, `"voted"`. Remember that you can use the `:` to make consecutive column selection easier.

```{r michigan-voter-exper-3, exercise = TRUE}

```

<button onclick="transfer_code(this)">

Copy previous code

</button>

```{r michigan-voter-exper-3-hint-1, eval = FALSE}
... |> 
  pivot_longer(cols = ...:...,
               ...)
```

```{r michigan-voter-exper-3-hint-2, eval = FALSE}
pivot_longer(cols = primary_00:general_04,
             names_to = "...",
             values_to = "...")
```

### 

Because the columns we have selected are consecutive, we can use the `:` to more quickly name them as a group. Look at the `tidyr_tidy_select` help page for more operators which make it easier to select variables.

### Exercise 4

Now it's time to change this into a binary system. Use `mutate()` to change the `voted` column into a column of the same name, but which has `1` and `0` instead of `"Yes"` and `"No"`. Inside the `mutate()`, set `voted` equal to an `if_else()` function whose condition is `voted ==  "Yes"`. Set the results, for if true and false, to be `1` and `0`, respectively.

```{r michigan-voter-exper-4, exercise = TRUE}

```

<button onclick="transfer_code(this)">

Copy previous code

</button>

```{r michigan-voter-exper-4-hint-1, eval = FALSE}
... |> 
 mutate(voted = if_else(voted == "...", 1, ...))
```

### Exercise 5

Now onto the next part. Use `pivot_wider()` to transform the contents of `voted` back into six columns. Set the names of these new columns to come from `election` and their values to come from `voted`

```{r michigan-voter-exper-5, exercise = TRUE}

```

<button onclick="transfer_code(this)">

Copy previous code

</button>

```{r michigan-voter-exper-5-hint-1, eval = FALSE}
... |> 
pivot_wider(names_from = ..., 
              values_from = ...)
```

### 

Notice how quotes are not needed when referring to existing columns in `pivot_wider()` but are needed when creating new columns in `pivot_longer()`.

### Exercise 6

After that, add the function `rowwise()` to your pipe. Then use `mutate()` to create a new column called `vote_sum`. Inside `mutate()`, set `vote_sum` equal to the sum of the columns from `primary_00` to `general_04`.

```{r michigan-voter-exper-6, exercise = TRUE}

```

<button onclick="transfer_code(this)">

Copy previous code

</button>

```{r michigan-voter-exper-6-hint-1, eval = FALSE}
... |> 
  rowwise() |> 
  mutate(vote_sum = sum(primary_00, general_00, primary_02, general_02, 
                        primary_04, general_04))
```

### 

If we did not use `rowwise()` here, `mutate()` would interpret the sum as a summing together of all the rows in each of these six columns. The new column would have the same, very large value throughout all of its rows.

### Exercise 7

Use `mutate()` to change the `vote_sum` column into a column of the same name, whose contents are characters rather than numbers. Use the `as.character()` function.

```{r michigan-voter-exper-7, exercise = TRUE}

```

<button onclick="transfer_code(this)">

Copy previous code

</button>

```{r michigan-voter-exper-7-hint-1, eval = FALSE}
mutate(... = as.character(...))
```

```{r michigan-voter-exper-7-hint-2, eval = FALSE}
mutate(vote_sum = as.character(vote_sum))
```

### 

We are using these numbers which represent vote histories as different categories for our age distributions. Later when we put these on the y-axis, numeric form would confuse `ggplot()` because it would think we are putting a continuous variable on the y-axis.

### Exercise 8

Now use a new `mutate()` to change the `primary_06` column into a column of the same name, whose contents are characters rather than numbers. Use the `as.character()` function.

```{r michigan-voter-exper-8, exercise = TRUE}

```

<button onclick="transfer_code(this)">

Copy previous code

</button>

```{r michigan-voter-exper-8-hint-1, eval = FALSE}
... |> 
  mutate(... = as.character(...))
```

```{r michigan-voter-exper-8-hint-2, eval = FALSE}
... |> 
  mutate(primary_06 = as.character(primary_06))
```

### 

Likewise, when we use `primary_06`as a fill, `ggplot()` if it were a numeric variable.

### Exercise 9

`select()` the columns relevant for our plot: `age`, `vote_sum`, and `primary_06`.

```{r michigan-voter-exper-9, exercise = TRUE}

```

<button onclick="transfer_code(this)">

Copy previous code

</button>

```{r michigan-voter-exper-9-hint-1, eval = FALSE}
... |> 
select(..., ..., ...)
```

### Exercise 10

Now it's time to graph it. Continue your pipe with `ggplot()`. Map `age` to the x-axis, and `primary_06` to `fill`. Add the layer `geom_histogram()`.

```{r michigan-voter-exper-10, exercise = TRUE}

```

<button onclick="transfer_code(this)">

Copy previous code

</button>

```{r michigan-voter-exper-10-hint-1, eval = FALSE}
... |> 
  ggplot(mapping = aes(x = ..., 
                        fill = ...)) +
  geom_histogram()
```

### 

Remember that histogram has a default `y` which are the counts it tabulates for each bin from our data.

### Exercise 11

Within `geom_histogram()`, set the argument `position` equal to `"identity"`, set the argument `alpha` = `0.5`, and set the argument `bins` = 30.

```{r michigan-voter-exper-11, exercise = TRUE}

```

<button onclick="transfer_code(this)">

Copy previous code

</button>

```{r michigan-voter-exper-11-hint-1, eval = FALSE}
... |> 
  geom_histogram(position = "identity", alpha =  ..., 
                        bins = ...)
```

### 

For histogram and bar plots, `position` is by default `"stacked"`, which stacks different colors of the same x value on top of one another. `position` being equal to `"identity"`, makes it so that different colors overlap, and that both can be seen when the opacity of the colors is increased.

### Exercise 12

Add the layer `facet_wrap()` to facet the histogram by `vote_sum`.

```{r michigan-voter-exper-12, exercise = TRUE}

```

<button onclick="transfer_code(this)">

Copy previous code

</button>

```{r michigan-voter-exper-12-hint-1, eval = FALSE}
Within facet_wrap(), remember to put a ~ before the variable you want to facet by.
```

```{r michigan-voter-exper-12-hint-2, eval = FALSE}
... + 
  facet_wrap(~...) 
```

### Exercise 13

Add the layer `coord_cartesian` and use the argument to limit the x-axis domain to a minimum of 20 and a maximum of 90

```{r michigan-voter-exper-13, exercise = TRUE}

```

<button onclick="transfer_code(this)">

Copy previous code

</button>

```{r michigan-voter-exper-13-hint-1, eval = FALSE}
... +
  coord_cartesian(xlim = c(...,...))
```

### Exercise 14

Ok now what we're going to do is add the layer `scale_fill_manual` to change the legend options from `0` and `1` to `Did not vote` and `vote`. Add `scale_fill_manual` and set the first argument `values`, which determines the fill colors, equal to `c("#F8766D", "#00BFC4")`, the default red and blue in ggplot. Set the second argument `labels` equal to `c("Did not vote", "Voted")`

```{r michigan-voter-exper-14, exercise = TRUE}

```

<button onclick="transfer_code(this)">

Copy previous code

</button>

```{r michigan-voter-exper-14-hint-1, eval = FALSE}
... +
  scale_fill_manual(values = c("...", "..."), labels = c("...", "..."))
```

### Exercise 15

Adjust the feel of the graph by adding the theme `theme_classic()`. To finish your plot, use `labs()` to give the graph a title, subtitle, axes labels, legend heading, and caption.

```{r michigan-voter-exper-15, exercise = TRUE}

```

<button onclick="transfer_code(this)">

Copy previous code

</button>

```{r michigan-voter-exper-15-hint-1, eval = FALSE}
... +
  theme_classic()  +
  labs(... ,
     fill = "...",
     ...)
```

Reminder: This is what your plot should look like.

```{r shaming15-answer}
shaming_p1
```

### Exercise 16

Let's make the following plot, but this time, we're using density plots to visualize the same data.

```{r}
shaming_p2 <- shaming |> 
  mutate(id = 1:length(primary_06)) |> 
  pivot_longer(cols = primary_00:general_04,
               names_to = "year",
               values_to = "voted") |> 
  mutate(voted = if_else(voted == "Yes", 1, 0)) |> 
  pivot_wider(names_from = year, 
              values_from = voted) |> 
  rowwise() |> 
  mutate(vote_sum = sum(primary_00, general_00, primary_02, general_02, primary_04, general_04)) |> 
  mutate(vote_sum = as.character(vote_sum),
         primary_06 = as.character(primary_06)) |> 
  select(age, vote_sum, primary_06) |> 
  ggplot(aes(x = age, fill = primary_06)) +
  geom_density(alpha = 0.3) +
  facet_wrap(~ vote_sum) + 
  coord_cartesian(xlim = c(20, 90)) +
  scale_fill_manual(values = c("#F8766D", "#00BFC4"), labels = c("Did not vote", "Voted")) +
  theme_classic() +
  labs(title = "Age Distributions by Voting History",
       subtitle = "Among those of the same voting history, those who voted in the \n 2006 primary are of generally similar ages to those who did not", 
       x = "Age", 
       y = "Density", 
       fill = "2006 Primary",
       caption = "Gerber, Green, and Larimer (2008)") 
  
shaming_p2
```

### 

We have provided you with the beginning of code for the last plot (YOU'RE TOTALLY WELCOME). Continue your pipe with `ggplot()`. Map `age` to the x-axis, and `primary_06` to `fill`. Add the layer `geom_density()`. Within `geom_density()`, set `alpha` equal to 0.3.

```{r michigan-voter-exper-16, exercise = TRUE}
shaming |> 
  mutate(id = 1:length(primary_06)) |> 
  pivot_longer(cols = primary_00:general_04,
               names_to = "year",
               values_to = "voted") |> 
  mutate(voted = if_else(voted == "Yes", 1, 0)) |> 
  pivot_wider(names_from = year, 
              values_from = voted) |> 
  rowwise() |> 
  mutate(vote_sum = sum(primary_00, general_00, primary_02, general_02, primary_04, general_04)) |> 
  mutate(vote_sum = as.character(vote_sum),
         primary_06 = as.character(primary_06)) |> 
  select(age, vote_sum, primary_06) 
```

```{r michigan-voter-exper-16-hint-1, eval = FALSE}
... |> 
  ggplot(aes(..., fill = ...)) +
  geom_density(alpha = ...) 
```

### Exercise 17

Add the layer `facet_wrap()` to facet the histogram by `vote_sum`. Add the layer `coord_cartesian` and use the argument to limit the x-axis domain to a minimum of 20 and a maximum of 90.

```{r michigan-voter-exper-17, exercise = TRUE}

```

<button onclick="transfer_code(this)">

Copy previous code

</button>

```{r michigan-voter-exper-17-hint-1, eval = FALSE}
Within facet_wrap(), remember to put a ~ before the variable you want to facet by.
```

```{r michigan-voter-exper-17-hint-2, eval = FALSE}
... +
  coord_cartesian(xlim = c(...,...))
```

### Exercise 18

Ok now we're gonna add the layer `scale_fill_manual` to change the legend options from `0` and `1` to `Did not vote` and `vote`. Add the layer `scale_fill_manual` and set the first argument `values`, which determines the fill colors, equal to `c("#F8766D", "#00BFC4")`, the default red and blue in ggplot. Set the second argument `labels` equal to `c("Did not vote", "Voted")`.

```{r michigan-voter-exper-18, exercise = TRUE}

```

<button onclick="transfer_code(this)">

Copy previous code

</button>

```{r michigan-voter-exper-18-hint-1, eval = FALSE}
... +
  scale_fill_manual(values = c("...", "..."), labels = c("...", "..."))
```

### Exercise 19

Ok final part. Adjust the feel of your graph by adding `theme_classic()`. Finally, add labels to your graph. Pay particular attention to how the y-axis label and subtitle should change compared to the previous plot. If, as in our example, the subtitle is too long insert `\n` at the point in your subtitle where you would like a new line.

```{r michigan-voter-exper-19, exercise = TRUE}

```

<button onclick="transfer_code(this)">

Copy previous code

</button>

```{r michigan-voter-exper-19-hint-1, eval = FALSE}
... +
  theme_classic()  +
  labs(... ,
    subtitle = "...\n..."
     fill = "...",
     ...)
```

Reminder: This is what your plot should look like.

```{r shaming19-answer}
shaming_p2
```

### 

Plot outputs have a margin about the size of the R markdown margin. For overlong subtitles such as the one we used in the example, `ggplot()` will not wrap them automatically. To create a new line, insert `\n` meaning "new-line" at the appropriate point. `ggplot()` will print the following text on a new line and will not print `\n`.

### Exercise 20

Let's create the following plot, using another `ggdist` function to visualize the same data with point interval graphs.

```{r}
shaming_p3 <- shaming |> 
  mutate(id = 1:length(primary_06)) |> 
  pivot_longer(cols = primary_00:general_04,
               names_to = "year",
               values_to = "voted") |> 
  mutate(voted = if_else(voted == "Yes", 1, 0)) |> 
  pivot_wider(names_from = year, 
              values_from = voted) |> 
  rowwise() |> 
  mutate(vote_sum = sum(primary_00, general_00, primary_02, general_02, primary_04, general_04)) |> 
  mutate(vote_sum = as.character(vote_sum),
         primary_06 = as.character(primary_06)) |> 
  select(age, vote_sum, primary_06) |> 
  ggplot(aes(x = age, y = vote_sum, color = primary_06)) +
  stat_pointinterval(position = "dodge") +
  scale_color_manual(values = c("#F8766D", "#00BFC4"), labels = c("Did not vote", "Voted")) +
  theme_classic() +
  labs(title = "Age Distributions by Voting History",
       subtitle = "With the exception of the youngest, those who voted in the 2006 \n primary were always slightly older than those who did not", 
       x = "Age", 
       y = "Number of Times Voted",
       color = "2006 Primary",
       caption = "Gerber, Green, and Larimer (2008)") 

shaming_p3
```

### 

The middle point represents the mean of the distribution, the bolded line represents the 25th to 75th percentile, and the unbolded line represents the 5th to 95th percentiles.

### 

We have provided you with the beginning of code for the last plot (Again, you are so very welcome). Continue your pipe with `ggplot()`. Map `age` to the x-axis, `vote_sum` to the y-axis, and `primary_06` to `color`. Add the layer `stat_pointinterval()` and specify `position` to be `"dodge"`.

```{r michigan-voter-exper-20, exercise = TRUE}
shaming |> 
  mutate(id = 1:length(primary_06)) |> 
  pivot_longer(cols = primary_00:general_04,
               names_to = "year",
               values_to = "voted") |> 
  mutate(voted = if_else(voted == "Yes", 1, 0)) |> 
  pivot_wider(names_from = year, 
              values_from = voted) |> 
  rowwise() |> 
  mutate(vote_sum = sum(primary_00, general_00, primary_02, general_02, primary_04, general_04)) |> 
  mutate(vote_sum = as.character(vote_sum),
         primary_06 = as.character(primary_06)) |> 
  select(age, vote_sum, primary_06)
```

```{r michigan-voter-exper-20-hint-1, eval = FALSE}
... |> 
  ggplot(aes(..., color = ...)) +
  stat_pointinterval() 
```

### 

We use `color` instead of `fill` since we are dealing with lines rather than filled shapes.

### Exercise 21

Add the layer `coord_cartesian` and use the argument to limit the x-axis domain to a minimum of 20 and a maximum of 81, since none of our distributions have a 95th percentile greater than 81.

```{r michigan-voter-exper-21, exercise = TRUE}

```

<button onclick="transfer_code(this)">

Copy previous code

</button>

```{r michigan-voter-exper-21-hint-1, eval = FALSE}
... +
  coord_cartesian(xlim = c(...,...))
```

### Exercise 22

Ok now we're gonna add the layer `scale_color_manual` to change the legend options from `0` and `1` to `Did not vote` and `Voted`. Use `scale_color_manual` and set the first argument `values`, which determines the fill colors, equal to `c("#F8766D", "#00BFC4")`, the default red and blue in ggplot. Set the second argument `labels` equal to `c("Did not vote", "Voted")`

```{r michigan-voter-exper-22, exercise = TRUE}

```

<button onclick="transfer_code(this)">

Copy previous code

</button>

```{r michigan-voter-exper-22-hint-1, eval = FALSE}
... +
  scale_color_manual(values = c("...", "..."), labels = c("...", "..."))
```

### 

Just as you mapped to `color` instead of `fill`, you must use `scale_color_manual` in place of `scale_fill_manual`.

### Exercise 23

Now adjust the feel of your graph by adding `theme_classic()`. Finally, add labels to your graph. Use `color` rather than `fill` as we have done above. Pay particular attention to how the y-axis label and subtitle should change compared to the previous plot.

```{r michigan-voter-exper-23, exercise = TRUE}

```

<button onclick="transfer_code(this)">

Copy previous code

</button>

```{r michigan-voter-exper-23-hint-1, eval = FALSE}
... +
  theme_classic()  +
  labs(... ,
     fill = "...",
     ...)
```

Reminder: This is what your plot should look like.

```{r shaming22-answer}
shaming_p3
```

## Summary
### 

This tutorial covered [Chapter 6: Data tidying](https://r4ds.hadley.nz/data-tidy.html) from [*R for Data Science (2e)*](https://r4ds.hadley.nz/) by Hadley Wickham, Mine Çetinkaya-Rundel, and Garrett Grolemund.

```{r download-answers, child = system.file("child_documents/download_answers.Rmd", package = "tutorial.helpers")}
```
