---
title: Data tidying
author: Miriam Heiss
tutorial:
  id: data-tidying
output:
  learnr::tutorial:
    progressive: yes
    allow_skip: yes
runtime: shiny_prerendered
description: 'Tutorial for Chapter 6: Data tidying'
---

```{r setup, include = FALSE}
library(learnr)
library(tidyverse)
library(datasets)
library(tutorial.helpers)
knitr::opts_chunk$set(echo = FALSE)
options(tutorial.exercise.timelimit = 60, 
        tutorial.storage = "local") 
```

```{r copy-code-chunk, child = system.file("child_documents/copy_button.Rmd", package = "tutorial.helpers")}
```

```{r info-section, child = system.file("child_documents/info_section.Rmd", package = "tutorial.helpers")}
```

## Introduction
###

This tutorial covers [Chapter 6: Data tidying](https://r4ds.hadley.nz/data-tidy.html) from [*R for Data Science (2e)*](https://r4ds.hadley.nz/) by Hadley Wickham, Mine Çetinkaya-Rundel, and Garrett Grolemund.

## Tidy Data
###

You can represent the same underlying data in multiple ways by organizing values in a given dataset in different orders. But not all datasets are equally easy to use.

### Exercise 1

If you take a close look at the tidyverse loading message below, you should see a package called `tidyr`. 

<!-- Load the library tidyverse in the console, and copy and paste the loading message in the box below. -->

<!-- MH: Not sure how to get warning and loading messages to show up -->

```{r echo = TRUE, warning = TRUE, output = TRUE, message = TRUE}
library(tidyverse)
```

```{r tidy-data-1, exercise = TRUE}

```

### Exercise 2

Run `library(tidyverse)` in the chunk below:

```{r tidy-data-2, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r tidy-data-2-hint, eval = FALSE}
library(...)
```





## Lengthening Data
###

## Widening Data
###

## Summary
###

This tutorial covered [Chapter 6: Data tidying](https://r4ds.hadley.nz/data-tidy.html) from [*R for Data Science (2e)*](https://r4ds.hadley.nz/) by Hadley Wickham, Mine Çetinkaya-Rundel, and Garrett Grolemund.

```{r download-answers, child = system.file("child_documents/download_answers.Rmd", package = "tutorial.helpers")}
```