---
title: Terminal
author: David Kane
tutorial:
  id: terminal
output:
  learnr::tutorial:
    progressive: yes
    'allow_skip:': yes
runtime: shiny_prerendered
description: Terminal Tutorial
---

```{r setup, include = FALSE}
library(learnr)
library(tutorial.helpers)
library(tidyverse)
library(knitr)
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(out.width = '90%')
options(tutorial.exercise.timelimit = 60, 
        tutorial.storage = "local") 
```

```{r copy-code-chunk, child = system.file("child_documents/copy_button.Rmd", package = "tutorial.helpers")}
```

```{r info-section, child = system.file("child_documents/info_section.Rmd", package = "tutorial.helpers")}
```

<!-- Make sure that everything now works on Windows. Note all the DK comments which mention Windows. -->

<!-- Add an exercise which downloads a file from the web, and then moves it to a new directory. This is good practice for the class session doing the same. Put it after Options. Call it "Files". -->

<!-- Paths topic needs work. -->

<!-- Paths and Options topics are both (too?) long. -->

<!-- Change Terminal and Paths to not use any symbols? Save all those for the Symbols topic. Not sure if this is possible. How do you move one directory up without `..`? -->

<!-- Lots of opportunities for knowledge drops. See Chapter 7 in R4DS. -->

<!-- Chapter 7 includes some discussion about relative and absolute paths. Include some of that stuff here? Certainly we should emphasize that you should only ever use relative paths, not absolute paths. -->

<!-- Consider using and teaching the **here** package. -->

<!-- Instead of the bs about the defaults being Mac/Linux mentioned in the introduction, incorporate Windows nomenclature throughout. -->

## Introduction
### 

This tutorial covers material which is not actually in [*R for Data Science (2e)*](https://r4ds.hadley.nz/quarto.html) by Hadley Wickham, Mine Çetinkaya-Rundel, and Garrett Grolemund. But the material is in keeping with the spirit of that book. Some of the material from [Chapter 6 Workflow: scripts and projects](https://r4ds.hadley.nz/workflow-scripts) appears here.

This tutorial introduces the structure and uses of the Terminal from within RStudio. The older, more general term for the terminal is the "command line." All modern operating systems provide a command line, from which sophisticated users can modify files and directories, using commands like `pwd`, `ls`, `touch`, `mkdir`, `mv`, `rm`, and `cd`. We also discuss regular expressions as well as some of the metacharacters (like, `*`, `^` and `$`) from which they are constructed.

### 

Note that a default Windows installation is different from Mac or Linux. If you are using Windows, you must first install RTools. Download the latest version from [here](https://cran.r-project.org/bin/windows/Rtools/). Currently, the Rtools44 installer is what you want. The link is in the seventh paragraph of [this page](https://cran.r-project.org/bin/windows/Rtools/rtools44/rtools.html).

### 

For the most part, Windows works like Mac/Unix. There are two main exceptions which are important for this tutorial:

* Environment variables look like `$R_HOME` on Mac/Unix and `%R_HOME%` on Windows. That is, in both cases, we are using the `R_HOME` environment variable. We just access this variable differently depending on the operating system.

* Directory separators are forward slashes -- `/` -- on Max/Unix and backward slashes -- `\` -- on Windows.

The tutorial defaults use Mac/Unix nomenclature. If you are using Windows, you need to translate, for example:

```
ls $R_HOME/etc
```

into

```
ls %R_HOME%\etc
```


## Terminal
### 

The Terminal tab is next to the Console tab within the Console pane in the lower left portion of RStudio. (Yes, it is a bit confusing that we call the entire structure the Console **pane** even though only one part of it is the Console **tab**.) We use the Console tab to talk to R. We use the Terminal tab to talk directly with the computer. Sadly, the Console and the Terminal speak different languages.

### Exercise 1

Hit the "return" key (the "enter" key on Windows) two times in the Terminal to see what happens. The terminal has a string of characters ending in a \$ before your cursor, called the **prompt**. After a command has been executed, a prompt will be generated on a new line to let you know that Terminal is ready for a new command. Copy and paste the three (blank) lines from the Terminal as your answer below. We do a lot of **c**opy/**p**asting of **c**ommands/**r**esponses, so we abbreviate those instructions as CP/CR.

```{r terminal-1}
question_text(NULL,
    answer(NULL, correct = TRUE),
    allow_retry = TRUE,
    try_again_button = "Edit Answer",
    incorrect = NULL,
    rows = 3)
```

### 

Your answer should look something like:

````
Davids-MacBook-Pro-2:averages dkane$ 
Davids-MacBook-Pro-2:averages dkane$ 
Davids-MacBook-Pro-2:averages dkane$ 
````

Your answer will look different from mine. Default prompts vary from computer to computer. In my case, the first part of the prompt is the name of my computer: `Davids-MacBook-Pro-2`. The second part, separated with a colon, is the name of the directory, `averages`, in which I am running the tutorial. The third part, separated by a space, is `dkane`, my username.

### 

The prompt acts as a quick way to tell you where in the computer you are. The Terminal is sensitive to which folder location you are in, just like the R Console is.

### Exercise 2

Let's figure out the location of the folder in which we currently are. (Note that the terms "folder" and "directory" mean the same thing.) To see your current location within your computer, type the command `pwd` (**p**resent **w**orking **d**irectory) in the Terminal. Hit return/enter key to run the command. 

Going forward, we will instruct you to "run" a given command. You do this by typing the command at the prompt and then hitting the return/enter key.

CP/CR.

```{r terminal-2}
question_text(NULL,
    answer(NULL, correct = TRUE),
    allow_retry = TRUE,
    try_again_button = "Edit Answer",
    incorrect = NULL,
    rows = 3)
```

### 

Your answer should look something like mine:

````
Davids-MacBook-Pro-2:averages dkane$ pwd
/Users/dkane/Desktop/projects/averages
Davids-MacBook-Pro-2:averages dkane$ 
````

Notice how the last directory (folder) in the **path** is included in the prompt.


### Exercise 3

Let's see a **l**i**s**t of what is where we are, i.e. in our working directory. Run `ls`. CP/CR.

```{r terminal-3}
question_text(NULL,
    answer(NULL, correct = TRUE),
    allow_retry = TRUE,
    try_again_button = "Edit Answer",
    incorrect = NULL,
    rows = 3)
```

### 

Note how, if you do not give any more information, `ls` assumes that you want a list of what is in your current working directory.

### Exercise 4

Let's **m**a**k**e a **dir**ectory called `example`. Run `mkdir example`. Run `ls` to confirm that the new directory exists inside of the current working directory. CP/CR.

```{r terminal-4}
question_text(NULL,
    answer(NULL, correct = TRUE),
    allow_retry = TRUE,
    try_again_button = "Edit Answer",
    incorrect = NULL,
    rows = 3)
```

### 

`mkdir` creates an empty directory. We --- and most other people --- use the terms "folder" and "directory" interchangeably.

### Exercise 5

Move into the `example` directory by running `cd example`. Run `pwd`. CP/CR.


```{r terminal-5}
question_text(NULL,
	answer(NULL, correct = TRUE),
	allow_retry = TRUE,
	try_again_button = "Edit Answer",
	incorrect = NULL,
	rows = 3)
```

### 

Your answer should look like:

````
Davids-MacBook-Pro-2:averages dkane$ cd example
Davids-MacBook-Pro-2:example dkane$ pwd
/Users/dkane/Desktop/projects/averages/example
Davids-MacBook-Pro-2:example dkane$ 
````

The `example` directory is one level below the directory in which I started.


### Exercise 6

Let's create a file called `my.txt`. Run `touch my.txt`. Run `ls` to confirm that the new file exists inside the current directory. CP/CR.

```{r terminal-6}
question_text(NULL,
    answer(NULL, correct = TRUE),
    allow_retry = TRUE,
    try_again_button = "Edit Answer",
    incorrect = NULL,
    rows = 3)
```

### 

````
Davids-MacBook-Pro-2:example dkane$ touch my.txt
Davids-MacBook-Pro-2:example dkane$ ls
my.txt
Davids-MacBook-Pro-2:example dkane$ 
````

`touch` creates an empty file.


### Exercise 7

Make a copy of `my.txt` called `my_2.txt` by using the `cp` command. That is, run `cp my.txt my_2.txt`. Confirm that this worked by running `ls`. CP/CR.

```{r terminal-7}
question_text(NULL,
	answer(NULL, correct = TRUE),
	allow_retry = TRUE,
	try_again_button = "Edit Answer",
	incorrect = NULL,
	rows = 3)
```

### 

Your answer should look like:

````
Davids-MacBook-Pro-2:example dkane$ cp my.txt my_2.text
Davids-MacBook-Pro-2:example dkane$ ls
my.txt          my_2.text
Davids-MacBook-Pro-2:example dkane$ 
````

### Exercise 8

We rename files with the `mv` command, which is derived from **m**o**v**e. The first argument is the file we want to rename, the second argument is the new name

Let's rename `my.txt` to `fake.txt`. Run `mv my.txt fake.txt`. Run `ls` to confirm. CP/CR.

```{r terminal-8}
question_text(NULL,
    answer(NULL, correct = TRUE),
    allow_retry = TRUE,
    try_again_button = "Edit Answer",
    incorrect = NULL,
    rows = 3)
```

### 

Your answer should look like:

````
Davids-MacBook-Pro-2:example dkane$ mv my.txt fake.txt
Davids-MacBook-Pro-2:example dkane$ ls
fake.txt          my_2.text
Davids-MacBook-Pro-2:example dkane$ 
````

### 

From the computer's point of view, **renaming** a file is the same thing as **moving** a file. In fact, moving is more general since it allows to change both the name **and** the location of a file.


### Exercise 9

Let's **r**e**m**ove the file `fake.txt`. Run `rm fake.txt`. Run `ls` to confirm. CP/CR.

```{r terminal-9}
question_text(NULL,
    answer(NULL, correct = TRUE),
    allow_retry = TRUE,
    try_again_button = "Edit Answer",
    incorrect = NULL,
    rows = 3)
```

### 

**Be careful** with the `rm` command in the Terminal. Unlike moving files to Trash on your computer, it is (usually) **irreversible**.

### Exercise 10

We want to remove the `example` directory. But we can't do that while we are "located" in that directory. That is, the computer is aware that our Terminal process --- the entity which is executing our commands like `ls` and `cp` --- is currently working from the `example` directory. 

Run `cd ..`. Run `pwd` to confirm. CP/CR.

```{r terminal-10}
question_text(NULL,
	answer(NULL, correct = TRUE),
	allow_retry = TRUE,
	try_again_button = "Edit Answer",
	incorrect = NULL,
	rows = 3)
```

### 

Your answer should look like:

````
Davids-MacBook-Pro-2:example dkane$ cd ..
Davids-MacBook-Pro-2:averages dkane$ pwd
/Users/dkane/Desktop/projects/averages
Davids-MacBook-Pro-2:averages dkane$ 
````

Again, note how the prompt changed because we moved directories. The `..` symbol --- two periods --- always indicates the directory one level up, i.e., the directory in which `example` is located.


### Exercise 11

Run `rm -r example` to emove the directory `example`. CP/CR.

```{r terminal-11}
question_text(NULL,
    answer(NULL, correct = TRUE),
    allow_retry = TRUE,
    try_again_button = "Edit Answer",
    incorrect = NULL,
    rows = 3)
```

`-r` is an **option** that allows us to delete a directory. We will learn more about options later. The `r` stands for **r**ecursive because, when you delete a directory, you are also (recursively) deleting everything within that directory.

It is good to clean up. Don't leave junk files lying around.

## Paths
### 

You will practice using paths both inside and outside the current working directory. You will learn a few shortcuts to make this easier.

### Exercise 1

By default, the Terminal in RStudio starts with a working directory which is the same as the folder in which the current R project resides, i.e., the folder with the `.Rproj` file in it. Run `pwd` to show your current working directory. CP/CR.

```{r paths-1}
question_text(NULL,
    answer(NULL, correct = TRUE),
    allow_retry = TRUE,
    try_again_button = "Edit Answer",
    incorrect = NULL,
    rows = 3)
```

### 

The Output pane (bottom right corner) in RStudio includes several tabs, the most important of which is the Files tab. Click on it. At the top of the Files tab, you will see a series of directories, separated by `>`, on your computer, beginning with the Home icon. This should match the path generated by `pwd`.

### Exercise 2

Use `mkdir` to make a directory named `paths` inside your current working directory. CP/CR.

```{r paths-2}
question_text(NULL,
    answer(NULL, correct = TRUE),
    allow_retry = TRUE,
    try_again_button = "Edit Answer",
    incorrect = NULL,
    rows = 3)
```

### 

The new directory is created directly inside the working directory.

### Exercise 3

Let's **c**hange the working **d**irectory to `paths`. Run `cd paths`. Run `pwd` to confirm that you have changed the working directory. CP/CR.

```{r paths-3}
question_text(NULL,
    answer(NULL, correct = TRUE),
    allow_retry = TRUE,
    try_again_button = "Edit Answer",
    incorrect = NULL,
    rows = 3)
```

### 

Your answer should look something like:

````
Davids-MacBook-Pro-2:averages dkane$ cd paths
Davids-MacBook-Pro-2:paths dkane$ pwd
/Users/dkane/Desktop/projects/averages/paths
Davids-MacBook-Pro-2:paths dkane$ 
````

Note how the prompt changed after I ran the `cd paths` command. Before, the prompt included `averages` because that was the name of the current working directory. Using `cd` changed the current working directory to `paths`, causing the prompt to change as well.

### Exercise 4

Use `mkdir` to make a directory called `lessons` inside `paths` by running `mkdir lessons`. Confirm that this worked by running `ls`. CP/CR.

```{r paths-4}
question_text(NULL,
    answer(NULL, correct = TRUE),
    allow_retry = TRUE,
    try_again_button = "Edit Answer",
    incorrect = NULL,
    rows = 3)
```

### 

Your answer should look like this:

````
Davids-MacBook-Pro-2:paths dkane$ mkdir lessons
Davids-MacBook-Pro-2:paths dkane$ ls
lessons
Davids-MacBook-Pro-2:paths dkane$ 
````

Because we have changed working directories, the `lessons` directory is created inside the `paths` directory. It is the only object in the `paths` directory.

### Exercise 5

Make a directory within `lessons` called `fruits` by running `mkdir lessons/fruits`. In Windows, this would be `mkdir lessons\fruits`. The only difference between the Mac/Linux command and the Windows command is that the former uses a forward slash, `/`, to separate out directories while the latter uses a back slash, `\`.

Run `ls lessons` to check to see if `fruits` exists inside the `lessons` directory. CP/CR.

```{r paths-5}
question_text(NULL,
    answer(NULL, correct = TRUE),
    allow_retry = TRUE,
    try_again_button = "Edit Answer",
    incorrect = NULL,
    rows = 3)
```

### 

Your answer should look like this:

````
Davids-MacBook-Pro-2:paths dkane$ mkdir lessons/fruits
Davids-MacBook-Pro-2:paths dkane$ ls lessons/
fruits
Davids-MacBook-Pro-2:paths dkane$ 
````

### 

We could not, as before, simply give `fruits` as an argument to `mkdir` because then the directory would be created directly inside `paths`, the current working directory, rather than within `lessons`. To refer to a location other than the working directory, we need to use a **path**, which describes the location directory by directory. The path above is called a **relative path** because it assumes the working directory as its starting point.

### Exercise 6

Let's change our working directory to `fruits`. Run `cd lessons/fruits`. In Windows, `cd lessons\fruits`.

As you are typing this command, type only the `f` in `fruits` and then press the tab key (twice may be necessary on some computers). Pressing tab autocompletes the name of a file or directory. 

CP/CR.

```{r paths-6}
question_text(NULL,
    answer(NULL, correct = TRUE),
    allow_retry = TRUE,
    try_again_button = "Edit Answer",
    incorrect = NULL,
    rows = 3)
```

### 

Get in the practice of using the tab key. Avoid typing whenever possible. Be lazy!

### Exercise 7

Use `touch` to make a text file inside `fruits`, named `pineapple.txt`. Recall that your current location --- i.e., your working directory --- is now `fruits`. So, to make a new file within `fruits`, you just need `touch pineapple.txt`. Run `ls` to confirm. CP/CR.

### 

```{r paths-7}
question_text(NULL,
    answer(NULL, correct = TRUE),
    allow_retry = TRUE,
    try_again_button = "Edit Answer",
    incorrect = NULL,
    rows = 3)
```

Your answer should look like this:

````
Davids-MacBook-Pro-2:fruits dkane$ touch pineapple.txt
Davids-MacBook-Pro-2:fruits dkane$ ls
pineapple.txt
Davids-MacBook-Pro-2:fruits dkane$ 
````

### 

`pineapple.txt` is still a relative path, but since the location is directly inside the assumed starting point, i.e. the working directory, the name of the file is all we need. If we had used `lessons/fruits/apple.txt`, as we would have had to do if our working directory were `paths`, we would have gotten an error because those directories would not be recognized within our actual starting point of `fruits`.

### Exercise 8

Use `touch` to make two more text files inside `fruits`, named `pear.txt` and `does-not-belong`. That is, run `touch pear.txt does-not-belong`. Confirm by running `ls`. CP/CR.

### 

```{r paths-8}
question_text(NULL,
    answer(NULL, correct = TRUE),
    allow_retry = TRUE,
    try_again_button = "Edit Answer",
    incorrect = NULL,
    rows = 3)
```

Your answer should look like this:

````
Davids-MacBook-Pro-2:fruits dkane$ touch pear.txt does-not-belong
Davids-MacBook-Pro-2:fruits dkane$ ls
does-not-belong pear.txt        pineapple.txt
Davids-MacBook-Pro-2:fruits dkane$ 
````

As always, note how the prompt reminds us that we are in the `fruits` directory.


### Exercise 9

Let's change our working directory to `lessons`. Run `cd ..` to go to the folder immediately above the working directory. Use `pwd` to confirm that you are in the right directory. CP/CR.

```{r paths-9}
question_text(NULL,
    answer(NULL, correct = TRUE),
    allow_retry = TRUE,
    try_again_button = "Edit Answer",
    incorrect = NULL,
    rows = 3)
```

Your answer should look like this:

````
Davids-MacBook-Pro-2:fruits dkane$ cd ..
Davids-MacBook-Pro-2:lessons dkane$ pwd
/Users/dkane/Desktop/projects/averages/paths/lessons
Davids-MacBook-Pro-2:lessons dkane$ 
````

### 

`..` is shorthand for the directory immediately above the current working directory. The phrase "current working directory" is a bit redundant. You don't need the word "current."

### Exercise 10

Use `mkdir` to make a directory within `lessons`, named `tbd`. Run `ls` to confirm. CP/CR.


```{r paths-10}
question_text(NULL,
    answer(NULL, correct = TRUE),
    allow_retry = TRUE,
    try_again_button = "Edit Answer",
    incorrect = NULL,
    rows = 3)
```

### 

Your answer should look like this:

````
Davids-MacBook-Pro-2:lessons dkane$ mkdir tbd
Davids-MacBook-Pro-2:lessons dkane$ ls
fruits  tbd
Davids-MacBook-Pro-2:lessons dkane$ 
````



### Exercise 11

Use `mv` to **m**o**v**e the file `does-not-belong` from the `fruits` directory to the `tbd` directory by running `mv fruits/does-not-belong tbd`. Run `ls tbd` to confirm. (Again, change forward slashes to backward slashes if you are using Windows.)

CP/CR.

```{r paths-11}
question_text(NULL,
	answer(NULL, correct = TRUE),
	allow_retry = TRUE,
	try_again_button = "Edit Answer",
	incorrect = NULL,
	rows = 3)
```

### 

Your answer should look like this:

````
Davids-MacBook-Pro-2:lessons dkane$ mv fruits/does-not-belong tbd
Davids-MacBook-Pro-2:lessons dkane$ ls tbd
does-not-belong
Davids-MacBook-Pro-2:lessons dkane$ 
````

### 

We can act on files which are not in the working directory as long as we provide a path --- either relative, as here, or absolute --- to those files. In this case, neither `mv` nor `ls` are acting on the current directory. Instead, they are acting on lower directories: `fruits` and `tbd`.


### Exercise 12

Use `cd` to change the working directory up to `paths` by running `cd ..`. Confirm with `pwd`. Use `mv` and the `.` shorthand to move the directory `tbd` directly inside of `paths` by running `mv lessons/tbd .`. Confirm with `ls`. CP/CR.

```{r paths-12}
question_text(NULL,
    answer(NULL, correct = TRUE),
    allow_retry = TRUE,
    try_again_button = "Edit Answer",
    incorrect = NULL,
    rows = 3)
```

### 

Your answer should look like:

````
Davids-MacBook-Pro-2:lessons dkane$ cd ..
Davids-MacBook-Pro-2:paths dkane$ pwd
/Users/dkane/Desktop/projects/averages/paths
Davids-MacBook-Pro-2:paths dkane$ mv lessons/tbd .
Davids-MacBook-Pro-2:paths dkane$ ls
lessons tbd
Davids-MacBook-Pro-2:paths dkane$ 
````

When using `mv`, there is not difference between moving a directory and moving a file.

### Exercise 13

Move up one more directory, back to the original working directory, by running `cd ..`. Confirm with `pwd`. Run `rm -r paths` to remove all the directories and files with which we have been working. CP/CR.


```{r paths-13}
question_text(NULL,
    answer(NULL, correct = TRUE),
    allow_retry = TRUE,
    try_again_button = "Edit Answer",
    incorrect = NULL,
    rows = 3)
```

Your answer should look like:

````
Davids-MacBook-Pro-2:averages dkane$ pwd
/Users/dkane/Desktop/projects/averages
Davids-MacBook-Pro-2:averages dkane$ rm -r paths
Davids-MacBook-Pro-2:averages dkane$ 
````

### 

In olden times, professional programmers would spend a lot of time learning the various options to commands like `rm`. Now, we just ask ChatGPT or a similar tool.

## Important symbols
### 

We have seen one important symbol --- `..` --- already. This section will explore it along with `.` and `~`.

It would be tedious to have to use `..` every time we wanted to refer to anywhere outside of our working directory. To refer to a specific folder outside of the working directory, we can instead use an **absolute path**. An absolute path, as opposed to the relative paths we have been using, describes a place on your computer without respect to a working directory. An absolute path begins with the root directory, which is represented by a lone `/`, at least in Mac/Linux. The command `pwd` returns an absolute path.

<!-- DK: Edit above to explain Windows. -->

### Exercise 1

Run `pwd`. This returns the current working directory, which should be the same location from which you began this tutorial. Note thatthe return value is an **absolute path**. It tells you exactly where you --- meaning your Terminal instance --- are located on the computer.

```{r important-symbols-1}
question_text(NULL,
	answer(NULL, correct = TRUE),
	allow_retry = TRUE,
	try_again_button = "Edit Answer",
	incorrect = NULL,
	rows = 3)
```

### 

### Exercise 2

<!-- DK: Could do more here. Perhaps discuss environment variables, including GITHUB_PAT and where it lives. -->

Run `cd` alone. Now, use `pwd` to check where you are. This is called the **home directory**, within which all an user's directories and files are contained. Run `ls` to see what is inside your home directory. CP/CR.

```{r important-symbols-2}
question_text(NULL,
    answer(NULL, correct = TRUE),
    allow_retry = TRUE,
    try_again_button = "Edit Answer",
    incorrect = NULL,
    rows = 3)
```

### Exercise 3

Use `cd` to change the the working directory back to your default working directory. You can again use the absolute path from the `pwd` in Exercise 1. CP/CR.

```{r important-symbols-3}
question_text(NULL,
    answer(NULL, correct = TRUE),
    allow_retry = TRUE,
    try_again_button = "Edit Answer",
    incorrect = NULL,
    rows = 3)
```

### 

Since the default working directory is within the home directory, the full absolute path was not necessary. We could have omitted the user-name and everything before it.

### Exercise 4

Run `ls ~`. The list should be identical to the list you generated above in the home directory. Because the home directory is used in the absolute path of so many files, its path (`/Users/your-user-name-here`) can be written with the shorthand `~`, which is a tilde, pronounced TIL-D. CP/CR.


```{r important-symbols-4}
question_text(NULL,
    answer(NULL, correct = TRUE),
    allow_retry = TRUE,
    try_again_button = "Edit Answer",
    incorrect = NULL,
    rows = 3)
```

### 

The shorthands we have learned so far, `.`, `..`, `~`, stand for the text of the paths for the corresponding directory. They are not specific to any function and can be used in any situation where the text of the relevant path could be used.

### Exercise 5

<!-- DK: Awkward. Maybe mention tab completion. -->

Let's change the working directory to a directory of your choice within the home directory using the shorthand `~`. Run, for example, `cd ~/Desktop`.


```{r important-symbols-5}
question_text(NULL,
    answer(NULL, correct = TRUE),
    allow_retry = TRUE,
    try_again_button = "Edit Answer",
    incorrect = NULL,
    rows = 3)
```

### 

Here, unlike above, an absolute path was useful because `Desktop`, for example, is not within our working directory.

### Exercise 6

Use `cd` to change back to the default working directory using the shorthand for the home directory to write its absolute path. CP/CR.

<!-- DK: huh? -->

```{r important-symbols-6}
question_text(NULL,
    answer(NULL, correct = TRUE),
    allow_retry = TRUE,
    try_again_button = "Edit Answer",
    incorrect = NULL,
    rows = 3)
```

### 

Depending on whether the folder you chose contains the default working directory, an absolute path may or may not have been necessary.



## Options
### 

Options modify the behavior of command line functions like `ls`.

### Exercise 1

Use `pwd` to confirm that you are in your default working directory. Make a directory directly inside the working directory, named `options`, by running `mkdir options`. `cd` into `options`. Run `ls` to confirm that it is empty. CP/CR.


```{r options-1}
question_text(NULL,
    answer(NULL, correct = TRUE),
    allow_retry = TRUE,
    try_again_button = "Edit Answer",
    incorrect = NULL,
    rows = 3)
```

### 

My answer looks like:

````
Davids-MacBook-Pro-2:averages dkane$ pwd
/Users/dkane/Desktop/projects/averages
Davids-MacBook-Pro-2:averages dkane$ mkdir options
Davids-MacBook-Pro-2:averages dkane$ cd options/
Davids-MacBook-Pro-2:options dkane$ ls
Davids-MacBook-Pro-2:options dkane$ 
````

Note how easy it is to string together several commands in order to accomplish our goals.


### Exercise 2

Add a text file to the `options` directory, named `.my-hidden.txt` --- make sure to include the `.` at the front of the file name --- by running `touch .my-hidden.txt`.  CP/CR.


```{r options-2}
question_text(NULL,
    answer(NULL, correct = TRUE),
    allow_retry = TRUE,
    try_again_button = "Edit Answer",
    incorrect = NULL,
    rows = 3)
```

### 

Other than `.`, avoid using special characters or spaces anywhere in file names. These are difficult to work with in the Terminal.

### Exercise 3

Run `ls` to look for the file you created. CP/CR.

```{r options-3}
question_text(NULL,
	answer(NULL, correct = TRUE),
	allow_retry = TRUE,
	try_again_button = "Edit Answer",
	incorrect = NULL,
	rows = 3)
```

### 

Your answer should be something like:

````
Davids-MacBook-Pro-2:options dkane$ ls
Davids-MacBook-Pro-2:options dkane$ 
````

If you did the previous exercise correctly, you should **not** see the new file. This is because we prefixed the name with `.`, which hid the file from normal view. Files whose name begins with a `.` are called "hidden" files for this reason.


### Exercise 4

Run `ls -a`. Using the **option** `-a`, you should be able to see **a**ll the files in the directory.  CP/CR.


```{r options-4}
question_text(NULL,
    answer(NULL, correct = TRUE),
    allow_retry = TRUE,
    try_again_button = "Edit Answer",
    incorrect = NULL,
    rows = 3)
```

### 

Options are rpreceded by a `-` and come before the argument. Your answer should look like:

````
Davids-MacBook-Pro-2:options dkane$ ls -a
.               ..              .my-hidden.txt
Davids-MacBook-Pro-2:options dkane$ 
````

The `.` file refers to the options directory itself. The `..` refers to the directory in which the options directy is located. Every directory has both a `.` and a `..` file inside of it, but we only see those files if we use the `-a` option.

### 

Hidden files live all over your computer. You can turn on your ability to see hidden files in the R Studio Files pane (bottom right corner) by clicking the "More" button towards the top.

### Exercise 5

The **environmental variable** `R_HOME` represents the path to a folder of set-up files that R has installed on your computer. Type and enter `echo $R_HOME` alone to make sure we are telling you the truth. 

Recall that, on Windows, this would be `echo %R_HOME%` and `cd %R_HOME%`.

CP/CR.

```{r options-5}
question_text(NULL,
    answer(NULL, correct = TRUE),
    allow_retry = TRUE,
    try_again_button = "Edit Answer",
    incorrect = NULL,
    rows = 3)
```

### 

Your answer should look like:

````
Davids-MacBook-Pro-2:options dkane$ echo $R_HOME
/Library/Frameworks/R.framework/Resources
Davids-MacBook-Pro-2:options dkane$ 
````

Note how the prompt has changed because I am now in the `options` directory.


### Exercise 6

Run `ls $HOME`. This should generate a list of the items in your home directory. `$HOME` is an environmental variable that represents the home directory.  

If you are on Windows, you can try `ls %HOME%`, but, sometimes, `%HOME%` is not defined on Windows. In that case, `ls %USERPROFILE%` should work. 

CP/CR.


```{r options-6}
question_text(NULL,
    answer(NULL, correct = TRUE),
    allow_retry = TRUE,
    try_again_button = "Edit Answer",
    incorrect = NULL,
    rows = 3)
```

### 

There are many different environmental variables for places throughout your computer. You can also make your own.

### Exercise 7

Use `pwd` to confirm that you are still in `options`. Use `ls` to see what is inside `R_HOME` by running `ls $R_HOME`. CP/CR.

```{r options-7}
question_text(NULL,
	answer(NULL, correct = TRUE),
	allow_retry = TRUE,
	try_again_button = "Edit Answer",
	incorrect = NULL,
	rows = 3)
```

### 

Don't worry about the details of these files. Our main point is that environmental variables are very handy because they make it easy to refer to a specific location on a computer, like where R is installed, even though the absolute path to this location is different on every person's computer.

### Exercise 8

Run `ls -l $R_HOME`, so that we can see more information in a **l**ong list.  CP/CR.


```{r options-8}
question_text(NULL,
    answer(NULL, correct = TRUE),
    allow_retry = TRUE,
    try_again_button = "Edit Answer",
    incorrect = NULL,
    rows = 3)
```

### 

My answer looks like this:

````
Davids-MacBook-Pro-2:options dkane$ ls -l $R_HOME
lrwxrwxr-x  1 root  admin  26 Nov 27 11:20 /Library/Frameworks/R.framework/Resources -> Versions/Current/Resources
Davids-MacBook-Pro-2:options dkane$
````

This is an example of a situation in which `$R_HOME` and `$R_HOME/` produce different results, as we will see in the next exercise. When the location pointed to be an environmental variable is a link, as in this case, using `ls -l $R_HOME` just gives us back the link.   


### Exercise 9

<!-- DK: Is the behavior for the above and this question different on Windows? -->

Run `ls -l $R_HOME/`. Just adding the trailing backward slash (or forward slash in Windows?) changes the output.  CP/CR.


```{r options-9}
question_text(NULL,
    answer(NULL, correct = TRUE),
    allow_retry = TRUE,
    try_again_button = "Edit Answer",
    incorrect = NULL,
    rows = 3)
```

### 

My answer looks like:

````
Davids-MacBook-Pro-2:options dkane$ ls -l $R_HOME/
total 200
-rw-rw-r--    1 root  admin  18011 Oct 31 21:02 COPYING
-rw-rw-r--    1 root  admin    488 Oct 31 21:02 Info.plist
lrwxrwxr-x    1 root  admin      5 Nov 27 11:20 R -> bin/R
-rwxrwxr-x    1 root  admin  70944 Oct 31 21:04 Rscript
-rw-rw-r--    1 root  admin     46 Oct 31 21:02 SVN-REVISION
drwxrwxr-x   28 root  admin    896 Nov 27 11:20 bin
drwxrwxr-x   23 root  admin    736 Nov 27 11:20 doc
...
````

In a long list, names are in the last column and file type is in the first: if the string in the first column begins with `d` it is a **d**irectory; if it begins with `-` it is a file; if it begins with `l` it is a **l**ink to another file or directory on the computer. The middle columns give, from left to right, the number of files, author, permission, file size (in bytes), and the date/time last modified.

### Exercise 10

Check what is inside the `doc` directory by running `ls -l $R_HOME/doc`. CP/CR.


```{r options-10}
question_text(NULL,
    answer(NULL, correct = TRUE),
    allow_retry = TRUE,
    try_again_button = "Edit Answer",
    incorrect = NULL,
    rows = 3)
```

### 

As you could guess from the name, the directory contains mostly files, including `AUTHORS`.

### Exercise 11

Let's make a copy of `$R_HOME/doc/AUTHORS` in `options`. Use the command `cp`, whose syntax is identical to `mv`, and the `.` shorthand.  That is, run `cp $R_HOME/doc/AUTHORS .`. CP/CR.

On Windows, we would refer to this file as `%R_HOME%\doc\AUTHORS`. Recall that, in Windows, environment variables are surrounded by percentage signs -- `%` -- and directories are separated by backward slashes: `\`.


```{r options-11}
question_text(NULL,
    answer(NULL, correct = TRUE),
    allow_retry = TRUE,
    try_again_button = "Edit Answer",
    incorrect = NULL,
    rows = 3)
```

### 

Like `mv`, `cp` can change a file name. Like `mv`, if the destination file name is omitted, `cp` will retain the same name by default.

### Exercise 12

Try to use `cp` and `.` to copy the `doc` directory into `options`. You will get an error informing you that the thing you tried to copy was a directory. Run the same command using the option `-r`.  CP/CR.


```{r options-12}
question_text(NULL,
    answer(NULL, correct = TRUE),
    allow_retry = TRUE,
    try_again_button = "Edit Answer",
    incorrect = NULL,
    rows = 3)
```

### 

As we have seen for `rm`, `-r` allows us to use `cp` with directories. This is different from `mv`, which does not need `-r`.

### Exercise 13

Let's view just the initial lines of the `AUTHORS`. Run `head AUTHORS`.  CP/CR.


```{r options-13}
question_text(NULL,
    answer(NULL, correct = TRUE),
    allow_retry = TRUE,
    try_again_button = "Edit Answer",
    incorrect = NULL,
    rows = 3)
```

### 

Now you know how R got its name.

### Exercise 14

Let's view more than just the initial 10 lines. The option `-n` allows you to specify the number of initial lines you want to see. Run `head -n 18 AUTHORS`.  CP/CR.


```{r options-14}
question_text(NULL,
    answer(NULL, correct = TRUE),
    allow_retry = TRUE,
    try_again_button = "Edit Answer",
    incorrect = NULL,
    rows = 3)
```

### 

The option `-n` takes an argument which allows head to display any number of initial lines.

### Exercise 15

Use the command `tail` and the same `-n` option to view the last 15 lines of `AUTHORS`.  CP/CR.


```{r options-15}
question_text(NULL,
    answer(NULL, correct = TRUE),
    allow_retry = TRUE,
    try_again_button = "Edit Answer",
    incorrect = NULL,
    rows = 3)
```

### 

If the argument of `-n`is greater than the number of lines in the file, `head` or `tail` will just print the whole file.

### Exercise 16

Let's view the whole file at once. Run `cat AUTHORS`.  CP/CR.

```{r options-16}
question_text(NULL,
    answer(NULL, correct = TRUE),
    allow_retry = TRUE,
    try_again_button = "Edit Answer",
    incorrect = NULL,
    rows = 3)
```

### 

`cat` stands for con**cat**enate, which means "to chain together." While it is often used to a single text file, the command can take two arguments and print the files together.

### Exercise 17

Run `cat -n AUTHORS`. Notice that `-n` has no argument.  CP/CR.


```{r options-17}
question_text(NULL,
    answer(NULL, correct = TRUE),
    allow_retry = TRUE,
    try_again_button = "Edit Answer",
    incorrect = NULL,
    rows = 3)
```

### 

`cat -n` prints the text file with line numbers. Note that the names end at line 28.

### Exercise 18

Let's make our own text file. Start by running `echo` which takes a quoted string, for example, "Your First and Last Name Here" as its argument. This "echoes" the argument by printing it in the next line of the terminal.  CP/CR.


```{r options-18}
question_text(NULL,
    answer(NULL, correct = TRUE),
    allow_retry = TRUE,
    try_again_button = "Edit Answer",
    incorrect = NULL,
    rows = 3)
```

### 

`echo` can also take a string that is not in quotes as its argument.

### Exercise 19

We can change the place that we `echo` to, from the Terminal to a text file, using `>` to redirect the output Run `echo "Your First and Last Name Here" > test.txt`. Then use `ls` to see if the file `test.txt` exists within your working directory. Then use `cat` to read this file. After you have confirmed that the file contains your name, use `rm` to remove `test.txt`.  CP/CR.


```{r options-19}
question_text(NULL,
    answer(NULL, correct = TRUE),
    allow_retry = TRUE,
    try_again_button = "Edit Answer",
    incorrect = NULL,
    rows = 3)
```

### 

<!-- DK? -->


### Exercise 20

`>` can be used to redirect the output of any command, not just `echo`. Let's try to make a new text file in `options` with just the first 28 lines of `AUTHORS` called `AUTHORS-fake`. Run `head -n 28 AUTHORS > AUTHORS-fake`. CP/CR.


```{r options-20}
question_text(NULL,
    answer(NULL, correct = TRUE),
    allow_retry = TRUE,
    try_again_button = "Edit Answer",
    incorrect = NULL,
    rows = 3)
```

### 

<!-- DK? -->


### Exercise 21

Using `>>`, whose syntax is the same as `>`, we can append output to an existing document. Use `echo` and `>>` to append your name to the list of `AUTHORS-fake`. CP/CR.


```{r options-21}
question_text(NULL,
    answer(NULL, correct = TRUE),
    allow_retry = TRUE,
    try_again_button = "Edit Answer",
    incorrect = NULL,
    rows = 3)
```

### 

Like `>`, `>>` is not limited to `echo`, but can be used to append the output of any command.

### Exercise 22

Use `tail` and `>>` to append the last 11 lines of `AUTHORS` to `AUTHORS-fake`. Refer to the syntax in Exercise 18 if needed. CP/CR.


```{r options-22}
question_text(NULL,
    answer(NULL, correct = TRUE),
    allow_retry = TRUE,
    try_again_button = "Edit Answer",
    incorrect = NULL,
    rows = 3)
```

### 

Your fake is not great because your name is likely not placed in correct alphabetical order. 

### Exercise 23

Use `cd` and `..` to change your directory to the default working directory. Use `rm -r` to remove the `options` directory. CP/CR.

```{r options-23}
question_text(NULL,
    answer(NULL, correct = TRUE),
    allow_retry = TRUE,
    try_again_button = "Edit Answer",
    incorrect = NULL,
    rows = 3)
```

### 

<!-- DK? -->

## Wildcards
### 

You will learn how to search for files by using **wildcards** to form **regular expressions**.

### Exercise 1

Use `pwd` to confirm that you are in your default working directory. Use `mkdir` to create a directory named `wildcards`. CP/CR.

```{r wildcards-1}
question_text(NULL,
    answer(NULL, correct = TRUE),
    allow_retry = TRUE,
    try_again_button = "Edit Answer",
    incorrect = NULL,
    rows = 3)
```

### 

The `mkdir` command does not produce any output. To confirm that it had the desired effect, you could run `ls`. 

### Exercise 2

Change your working directory to `wildcards` using `cd`. Confirm the change with `pwd`. CP/CR.

```{r wildcards-2}
question_text(NULL,
	answer(NULL, correct = TRUE),
	allow_retry = TRUE,
	try_again_button = "Edit Answer",
	incorrect = NULL,
	rows = 3)
```

### 

You answer should look something like:

````
Davids-MacBook-Pro-2:averages dkane$ cd wildcards/
Davids-MacBook-Pro-2:wildcards dkane$ pwd
/Users/dkane/Desktop/projects/averages/wildcards
Davids-MacBook-Pro-2:wildcards dkane$ 
````

Note how the prompt changed after we switche directories.

Keep track of your current location. There is a difference between being in the directory above `wildcards` and being inside of `wildcards`. Right now, you are inside `wildcard`.

### Exercise 3

Using `touch`, Make three text files named `hat.txt`, `cat.txt`, and `ate.txt`. Confirm that you have done so by running `ls`. (No worries if you make a mistake. Just fix it by removing the misnamed file.) CP/CR.

```{r wildcards-3}
question_text(NULL,
    answer(NULL, correct = TRUE),
    allow_retry = TRUE,
    try_again_button = "Edit Answer",
    incorrect = NULL,
    rows = 3)
```

### 

`touch` can take multiple arguments, making a file for each one. `touch hat.txt cat.txt ate.txt` is equivalent to running `touch` three times.

### Exercise 4

Terminal commands allow the use of **wildcards**. A wildcard `*` represents zero or more of any character, and it is used to work with similarly named files as a grouping.

### 

Let's generate a list of all our files whose names contain the letter `a`. First simply try to run `ls a`. You will get an error message --- "No such file or directory" --- because there is no file named `a` in the working directory. CP/CR.

```{r wildcards-4}
question_text(NULL,
    answer(NULL, correct = TRUE),
    allow_retry = TRUE,
    try_again_button = "Edit Answer",
    incorrect = NULL,
    rows = 3)
```

### 

Though we have been using `ls` to return all the content of a directory, `ls` can also take a name as an argument, as above.

### Exercise 5

Let's try again. Run `ls a*`. The only return will be `ate.txt`. Because you put `*` after `a`, `ls` returns only files that began with `a`. The "star" wildcard --- `*` --- matches anything. CP/CR.

```{r wildcards-5}
question_text(NULL,
    answer(NULL, correct = TRUE),
    allow_retry = TRUE,
    try_again_button = "Edit Answer",
    incorrect = NULL,
    rows = 3)
```

### 

Though `*` will match strings of any length, its positioning matters. `a*` is not the same thing as `*a`. The first matches any string (i.e., any file name) which starts with `a`. The second matches any string which ends with `a`.

### Exercise 6

Now run `ls *a*`. This should return `hat.txt cat.txt ate.txt`. CP/CR.

```{r wildcards-6}
question_text(NULL,
    answer(NULL, correct = TRUE),
    allow_retry = TRUE,
    try_again_button = "Edit Answer",
    incorrect = NULL,
    rows = 3)
```

Note that these **regular expressions** are case-sensitive. A regular expression is a "pattern" which you can use for many purposes, including as a search criteria.

### Exercise 7

<!-- DK: Too big a jump to this question? -->

Make a list of all the files with names, before the suffix, which end in `t` by running `ls *t.*` CP/CR.

```{r wildcards-7}
question_text(NULL,
    answer(NULL, correct = TRUE),
    allow_retry = TRUE,
    try_again_button = "Edit Answer",
    incorrect = NULL,
    rows = 3)
```

### 

This should return `hat.txt` and `cat.txt`. Read this regular expression --- `*t.*` --- as matching any string which includes a `t` followed directly by a period, regardless of any characters which are also included, before or after. `ate.txt` is not matched because the `t` and the `.` are separated by `e`.

### Exercise 8

Save the output of our list of all the file names that contained `a` to a new file names `afiles.txt`. Keep pressing the **up** arrow key to navigate through your previous commands until you get to the command you used above: `ls *a*`. Then add `> afiles.txt` afterwards and run the command. Run `ls` to confirm that it worked. CP/CR.

```{r wildcards-8}
question_text(NULL,
    answer(NULL, correct = TRUE),
    allow_retry = TRUE,
    try_again_button = "Edit Answer",
    incorrect = NULL,
    rows = 3)
```

### 

You can use your `up` and `down` arrow keys to navigate command history. The `>` symbol outputs the result of the left-hand side command into the right-hand side location.

### Exercise 9

Wildcards can be used with more than `ls`. Let's move `hat.txt` and `cat.txt` together into one directory. 

First, make a new directory called `seuss` within the `wildcards` directory in which you are currently located. 

Run `ls` to confirm. CP/CR

```{r wildcards-9}
question_text(NULL,
    answer(NULL, correct = TRUE),
    allow_retry = TRUE,
    try_again_button = "Edit Answer",
    incorrect = NULL,
    rows = 3)
```

### 

You answer should look like:

````
Davids-MBP:wildcards dkane$ mkdir seuss
Davids-MBP:wildcards dkane$ ls
afiles.txt      ate.txt         cat.txt         hat.txt         seuss
Davids-MBP:wildcards dkane$ 
````


The `wildcards` directory, in which you are currently located, now includes 5 items, one of which is the `seuss` directory.

### Exercise 10

Run `ls -l`. CP/CR

```{r wildcards-10}
question_text(NULL,
	answer(NULL, correct = TRUE),
	allow_retry = TRUE,
	try_again_button = "Edit Answer",
	incorrect = NULL,
	rows = 3)
```

### 

The `-l` option provides a **l**ong-format listing of the contents of the directory, including information about the type of content (file versus directory) and creation date/times. The leading `d` in the line for `seuss` indicates that it is a directory. 

<!-- DK: Give a tour of what the different columns mean?  -->

### Exercise 11

Run `ls *t.*`. CP/CR.

```{r wildcards-11}
question_text(NULL,
	answer(NULL, correct = TRUE),
	allow_retry = TRUE,
	try_again_button = "Edit Answer",
	incorrect = NULL,
	rows = 3)
```

### 

Before moving or otherwise acting on a group of files, we often want to ensure that we have the correct grouping. Running `ls` with the appropriate regular expression is an easy way to test.

### Exercise 12

Use the previous command (since we have confirmed that it generates the list of files we want to move) but replace `ls` with `mv` and then add `seuss`, which indicates the location to which we want to move the files. In other words, run `mv *t.* seuss`. CP/CR.

```{r wildcards-12}
question_text(NULL,
	answer(NULL, correct = TRUE),
	allow_retry = TRUE,
	try_again_button = "Edit Answer",
	incorrect = NULL,
	rows = 3)
```

### 

The `mv` command, like `mkdir`, does not produce any output. You need to check by hand to ensure it has done what you wanted.

### Exercise 13

Run `ls -R`. CP/CR.

```{r wildcards-13}
question_text(NULL,
	answer(NULL, correct = TRUE),
	allow_retry = TRUE,
	try_again_button = "Edit Answer",
	incorrect = NULL,
	rows = 3)
```

### 

Your answer should look like:

````
Davids-MBP:wildcards dkane$ ls -R
afiles.txt      ate.txt         seuss

./seuss:
cat.txt hat.txt
Davids-MBP:wildcards dkane$ 
````

The `-R` option to the `ls` command stands for **r**ecursive. We see, not simply the contents of the current directory (two files and the `seuss` directory) but also the contents of the `seuss` directory itself.

### Exercise 14

Use `cd ..` to move out of the `wildcards` directory, one level up. Run `pwd`. CP/CR.

```{r wildcards-14}
question_text(NULL,
	answer(NULL, correct = TRUE),
	allow_retry = TRUE,
	try_again_button = "Edit Answer",
	incorrect = NULL,
	rows = 3)
```

### 

In order to delete a directory, you can not be located within that directory. We want to delete (that is, "remove") the `wildcards` directory since we always want to clean up any temporary files/directories which we have created.

### Exercise 15

Run `rm wildcards`. That command will fail because "wildcards: is a directory". You can't just `rm` directories in the same way that you `rm` files. Instead, you need to use `rm -r`, where the `-r` indicates **r**ecursive. Run `rm -r wildcards`. CP/CR.

```{r wildcards-15}
question_text(NULL,
	answer(NULL, correct = TRUE),
	allow_retry = TRUE,
	try_again_button = "Edit Answer",
	incorrect = NULL,
	rows = 3)
```

### 

Note that both `wildcards` and `wildcards/` refer to the same entity, the `wildcards` directory. The Terminal often "adds" a forward slash (`/`) --- or, on Windows, a backward slash (`\`) --- to the end of directory names to indicate that the object is a directory, not simply a file.

## grep
### 

In the previous section, we used **wildcards** to create **regular expressions** in order to generate lists of file names which matched a specific pattern. In this section, we will use `grep` to look for words inside of files. Fortunately, wildcards and regular expressions work the same with `grep`.

### Exercise 1

Make a directory called `grepping` using `mkdir`. Run `ls`. CP/CR.

```{r grep-1}
question_text(NULL,
    answer(NULL, correct = TRUE),
    allow_retry = TRUE,
    try_again_button = "Edit Answer",
    incorrect = NULL,
    rows = 3)
```

### 

We call this folder `grepping` rather than `grep` because you should not use a command, like `grep`, as the name of a file or directory. It is possible to do this, just like it is possible to jump off a bridge. But, in both cases, you will probably regret it.

### Exercise 2

`cd` into the `grepping` directory. Run this command:

<!-- DK: Does this work on Windows? I think so. -->

```
printf "hat.txt\ncat.txt\nate.txt\n\n" > files.txt
```

The `\n` symbol will insert a new line between `hat.txt`, `cat.txt`, and `ate.txt`. 

Run `cat files.txt`. CP/CR.

```{r grep-2}
question_text(NULL,
	answer(NULL, correct = TRUE),
	allow_retry = TRUE,
	try_again_button = "Edit Answer",
	incorrect = NULL,
	rows = 3)
```

### 

The extra `\n` at the end ensures that there is a blank line at the end of the file, which is always a good idea when dealing with text files. 

Don't worry about the `printf` command. It is not used very often. We didn't use `echo` because it is annoying to use in the case of multiple lines of input when you are writing code which needs to work on both Mac/Unix and Windows.

### Exercise 3

Let's search the text file for lines containing the letter "a". Run `grep a files.txt`. CP/CR.

```{r grep-3}
question_text(NULL,
    answer(NULL, correct = TRUE),
    allow_retry = TRUE,
    try_again_button = "Edit Answer",
    incorrect = NULL,
    rows = 3)
```

### 

Remember to use the *tab* key to take advantage of auto-completion. Don't type! Tab complete!

This should return all three lines from the file, which are the names of the three files we initially created. The blank line is not returned because it does not include an "a".

`grep` is used to search for a string of characters within a text file. When it finds a match, it prints the line. You can put the regular expression, in this case `a`, in quotes but that is not required.

### Exercise 4

Let's search through `files.txt` for lines that begin with the letter "a". Run `grep ^a files.txt`. CP/CR.

```{r grep-4}
question_text(NULL,
    answer(NULL, correct = TRUE),
    allow_retry = TRUE,
    try_again_button = "Edit Answer",
    incorrect = NULL,
    rows = 3)
```

### 

This should return `ate.txt`.

`^` is put at the beginning of a regular expression to indicate that the expression must occur at the beginning of a line. A dollar sign --- `$` --- at the end of a regular expression indicates that the expression must occur at the end of the line.

<!-- DK: What does the below mean? -->

We will not attempt the previous search we made of these files with wildcards because it is difficult, though possible, to search for strings with special characters like `.` with `grep`.

### Exercise 5

Run `ls $R_HOME` --- `ls %R_HOME%` on Windows --- to make a list of the contents of the R home directory. CP/CR.

```{r grep-5}
question_text(NULL,
	answer(NULL, correct = TRUE),
	allow_retry = TRUE,
	try_again_button = "Edit Answer",
	incorrect = NULL,
	rows = 3)
```

### 

`R_HOME` is an environment variable which should behave similarly regardless of your computer. There should be several directories and files in this listing, including `COPYING` and `tests`.

### Exercise 6

Add `>` to `ls $R_HOME` to save this output to a new file named `RHOME.txt` located withing the `grepping` directory. That is, run

```
ls $R_HOME > RHOME.txt
```

or, on Windows, 

```
ls %R_HOME% > RHOME.txt
```

Run `ls`. CP/CR.

```{r grep-6}
question_text(NULL,
    answer(NULL, correct = TRUE),
    allow_retry = TRUE,
    try_again_button = "Edit Answer",
    incorrect = NULL,
    rows = 3)
```

### 

The `>` symbol is known as the "output redirection operator." Because all the output has been sent (or redirected) into a new file, `RHOME.txt`, this command produces no output.

There are now two files in `grepping`: `files.txt` and `R_HOME.txt`.

### Exercise 7

Run `wc RHOME.txt`. The `wc` command stands for **w**ord **c**ount. CP/CR.

```{r grep-7}
question_text(NULL,
	answer(NULL, correct = TRUE),
	allow_retry = TRUE,
	try_again_button = "Edit Answer",
	incorrect = NULL,
	rows = 3)
```

### 

The first number returned by `wc` is the number of lines in the file. The second is the total number of words. The third is the number of characters. In this case, the number of lines is the same as the number of words.

### Exercise 8

Use `grep` to search `RHOME.txt` for lines that contain `R`. That is, run:

`grep R RHOME.txt`

CP/CR.

```{r grep-8}
question_text(NULL,
    answer(NULL, correct = TRUE),
    allow_retry = TRUE,
    try_again_button = "Edit Answer",
    incorrect = NULL,
    rows = 3)
```

### 

<!-- DK: Correct? -->

A mistyped `grep` command can cause the Terminal to get stuck and not generate a new prompt. If this happens, navigate to the top of the Terminal page and click the dropdown menu where Terminal says "busy." Select "Interrupt Terminal Command" from this menu in order to generate a new prompt and start again.

### Exercise 9

Use `grep` and `^` to search `RHOME.txt` for lines that **begin** with `R`. That is, run:

`grep ^R RHOME.txt`

CP/CR.

```{r grep-9}
question_text(NULL,
    answer(NULL, correct = TRUE),
    allow_retry = TRUE,
    try_again_button = "Edit Answer",
    incorrect = NULL,
    rows = 3)
```

### 

`^` and `$` are called metacharacters. It is often useful to put quotes around a regular expression which includes metacharacters.

### Exercise 10

Use `grep` and `$` to search `RHOME.txt` for lines that end with `c`. That is, run:

`grep c$ RHOME.txt`

CP/CR.

```{r grep-10}
question_text(NULL,
    answer(NULL, correct = TRUE),
    allow_retry = TRUE,
    try_again_button = "Edit Answer",
    incorrect = NULL,
    rows = 3)
```

### 

This should return two lines: `doc` and `etc`. (Your computer set up might be different, so don't worry if your answer is not the same.)

### Exercise 11

Use `grep` to search `RHOME.txt` for lines than contain the string `ib`. That is, run:

`grep ib RHOME.txt`

CP/CR

```{r grep-11}
question_text(NULL,
    answer(NULL, correct = TRUE),
    allow_retry = TRUE,
    try_again_button = "Edit Answer",
    incorrect = NULL,
    rows = 3)
```

### 

`egrep` is an even more powerful version of `grep`.  Learning more about `egrep` is beyond the scope of this tutorial.

### Exercise 12

Use `cd ..` to change the working directory to the directory which is one level above your current location. This should be the default working directory from which we started. Run `pwd`. CP/CR.

```{r grep-12}
question_text(NULL,
    answer(NULL, correct = TRUE),
    allow_retry = TRUE,
    try_again_button = "Edit Answer",
    incorrect = NULL,
    rows = 3)
```

### 

If you keep your R projects in the same place on your computer, only the last directory in the path for the default working directory will change between projects.

### Exercise 13

Use `rm -r` to remove `grepping`. CP/CR.

```{r grep-13}
question_text(NULL,
    answer(NULL, correct = TRUE),
    allow_retry = TRUE,
    try_again_button = "Edit Answer",
    incorrect = NULL,
    rows = 3)
```

### 

It is always a good idea to clean up after yourself. As the signs say when taking a hike in the forest: "Take only photos. Leave only footprints."

## Summary
### 

This tutorial covered material which is not actually in [*R for Data Science (2e)*](https://r4ds.hadley.nz/quarto.html) by Hadley Wickham, Mine Çetinkaya-Rundel, and Garrett Grolemund. But the material is in keeping with the spirit of that book. Some of the material from [Chapter 6 Workflow: scripts and projects](https://r4ds.hadley.nz/workflow-scripts) appeared here.

This tutorial introduced the structure and uses of the Terminal from within RStudio. The older, more general term for the terminal is the "command line." All modern operating systems provide a command line, from which sophisticated users can modify files and directories, using commands like `pwd`, `ls`, `touch`, `mkdir`, `mv`, `rm`, and `cd`. We also discussed regular expressions as well as some of the metacharacters (like, `*`, `^` and `$`) from which they are constructed.

Want to learn more about the command line and related topics? Read [*The Unix Workbench*](https://seankross.com/the-unix-workbench/) by Sean Kross.

```{r download-answers, child = system.file("child_documents/download_answers.Rmd", package = "tutorial.helpers")}
```
