---
title: Layers
author: David Kane and Mann Talati
tutorial:
  id: layers
output:
  learnr::tutorial:
    progressive: yes
    allow_skip: yes
runtime: shiny_prerendered
description: 'Tutorial for Chapter 10: Layers'
---

```{r setup, include = FALSE}
library(learnr)
library(tutorial.helpers)
library(tidyverse)
library(maps)
knitr::opts_chunk$set(echo = FALSE)
options(tutorial.exercise.timelimit = 60, 
        tutorial.storage = "local") 

nz <- map_data("nz")

car_names <- c(`f` = "Front Wheel Drive",
               `4` = "4 Wheel Drive",
               `r` = "Rear Wheel Drive")

polling_data <- read_csv("data/vis4_polling_data.csv",
              col_types = cols(createddate = col_date(),
                               approval_type = col_character(),
                               rate = col_double()))
```

```{r copy-code-chunk, child = system.file("child_documents/copy_button.Rmd", package = "tutorial.helpers")}
```

```{r info-section, child = system.file("child_documents/info_section.Rmd", package = "tutorial.helpers")}
```

<!-- Have students load up all libraries, wherever you like. -->

<!-- Note that the later material sucks! There are not enough knowledge drops.  -->

<!-- This was a bit too long. Also, much if it is sloppy! Need to revisit. -->

<!-- Change elements of theme.. -->

<!-- Each section should begin with two questions looking at the data, even if they are the same questions each time. print, glimpse. -->

<!-- Make the plots cooler! Add links to  https://ggplot2-book.org/. And https://socviz.co/. Or maybe put these in case studies. Love this one: https://socviz.co/groupfacettx.html#histograms. -->

<!--  Delete entire sections of you don't like them! Put in more ### as continue buttons. Drop some knowledge at the end of each exercise.  Always start by looking at the data with just print the data name (trains) or glimpse() or summary() or skim(). Always use select() for each question to keep just the variables you need. Nothing wrong with an arbitrary filter() in the pipe. Nothing wrong with using a simple mutate() to create a new variable, like change height to inches.  
-->

## Introduction
### 

This tutorial covers [Chapter 10: Layers](https://r4ds.hadley.nz/layers.html) from [*R for Data Science (2e)*](https://r4ds.hadley.nz/) by Hadley Wickham, Mine Çetinkaya-Rundel, and Garrett Grolemund. You will learn how to use [**ggplot2**](https://ggplot2.tidyverse.org/) and related packages to create persuasive graphics. 


## Aesthetic mappings: Simple `geom_point()`
### 

### Exercise 1

Run `glimpse()` on `mpg`.

```{r aesthetic-mappings-simple-1, exercise = TRUE}

```

```{r aesthetic-mappings-simple-1-hint-1, eval = FALSE}
glimpse(...)
```

### 

The mpg data frame bundled with the ggplot2 package contains 234 observations on 38 car models

###

### Exercise 2

Load the **tidyverse** library because that will be used with the mpg data

```{r aesthetic-mappings-simple-2, exercise = TRUE}

```


```{r aesthetic-mappings-2-simple-hint-1, eval = FALSE}
library(...)
```

### 

“The greatest value of a picture is when it forces us to notice what we never expected to see.” — John Tukey

### Exercise 3

Run `ggplot()` on `mpg`.

```{r aesthetic-mappings-simple-3, exercise = TRUE}

```


```{r aesthetic-mappings-3-simple-hint-1, eval = FALSE}
ggplot(...)
```

Note that this just creates an empty area in which a plot can be placed.

###

**ggplot2** is an R package for creating customizable, high-quality data visualizations using a layered grammar of graphics approach. Often used to see the data through different types of graphs.

### Exercise 4

Continue the code by adding the `aes()` function within the call to `ggplot()`, setting the `x` argument to `displ` and the `y` argument to equal `hwy`.  

```{r aesthetic-mappings-simple-4, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r aesthetic-mappings-4-simple-hint-1, eval = FALSE}
ggplot(mpg, aes(x = ..., 
                ... = hwy))
```

### 

We want to visualize the relationship between the car's engine size and how it correlates to the car's fuel efficiency on the highway

###

### Exercise 5

Add `geom_point()` --- don't forget the `+` after the class to `ggplot()` --- to see the actual data.

```{r aesthetic-mappings-simple-5, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r aesthetic-mappings-simple-5-hint-1, eval = FALSE}
ggplot(mpg, aes(x = displ, 
                y = hwy)) +
  ...()
```

### 

### Exercise 6

Let's add some color to this graph. Set the `color` argument within `aes()` to `class`. This will separate each color based on the `class` of the car.

```{r aesthetic-mappings-simple-6, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r aesthetic-mappings-6-simple-hint-1, eval = FALSE}
ggplot(mpg, aes(x = ..., 
                y = ..., 
                color = ...)) +
  geom_point()
```

### 

Learn more about the `ggplot()` function by checking out its [help page](https://ggplot2.tidyverse.org/reference/ggplot.html).

###

## Aesthetic mappings: Advanced `geom_point()`
### 

When you are working with data sets that have over a thousand data points, you will have [overplotting](https://r-graphics.org/recipe-scatter-overplot). In this section, you will learn to minimize the overplotting by changing the size of each point.

### 

Let's make this scatterplot using `geom_point()`.

```{r aesthetic-mappings-advanced-0}
point_p2 <- ggplot(data = diamonds,
                   mapping = aes(x = price,
                                 y = carat,
                                 color = clarity)) +
  geom_point(size = .25, alpha = .5) +
  scale_x_log10() +
  labs(title = "Diamond Carat and Price",
       subtitle = "3+ carat diamonds are rare and of low quality",
       x = "Price",
       y =  "Carat",
       color = "Clarity",
       caption = "Source: R datasets")

point_p2
```

### Exercise 1

Type `?diamonds` to look at the help page for the `diamonds` tibble. If you cannot run it in your code chunk below, run it in your Console. Remember to load `library(tidyverse)` beforehand.

```{r aesthetic-mappings-advanced-1, exercise = TRUE}

```

```{r aesthetic-mappings-advanced-1-hint-1, eval = FALSE}
?diamonds
```

###
The `diamonds` dataset is a collection of data that contains information about diamonds, including their carat weight, cut, color, clarity, and price for analysis and modeling purposes.

### Exercise 2

Explore the `diamonds` dataset. Since we are looking at the "carats" and "clarity" categories, copy and paste the definitions for "carats" and "clarity" below.

```{r aesthetic-mappings-advanced-2}
  question_text(NULL,
    answer(NULL, correct = TRUE),
    allow_retry = TRUE,
    try_again_button = "Edit Answer",
    incorrect = NULL,
    rows = 3)
```

### Exercise 3

Start a new pipe with `diamonds`. Select the columns `carat`, `price`, and `clarity`.

```{r aesthetic-mappings-advanced-3, exercise = TRUE}

```

```{r aesthetic-mappings-advanced-3-hint-1, eval = FALSE}
diamonds |>
  select(..., ..., ...)
```

### Exercise 4

Continue your pipe and add `ggplot()`. Map `price` to the x-axis and `carat` to the y-axis.

```{r aesthetic-mappings-advanced-4, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r aesthetic-mappings-advanced-4-hint-1, eval = FALSE}
... |>
  ggplot(mapping = aes(x = ...,
                       y = ...))
```

###

If you want to add any themes to your plots check out: [**ggplot2**](https://ggplot2.tidyverse.org/articles/ggplot2-specs.html) contains all aesthetic specifications that are possible for the data. In addition, [**ggplot2: elegant graphics**](https://ggplot2-book.org/) provides all possible aesthetics that any user will use.

### Exercise 5

Add the `geom_point()` layer. Within `aes()` set `clarity` to `color`.

```{r aesthetic-mappings-advanced-5, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r aesthetic-mappings-advanced-5-hint-1, eval = FALSE}
geom_point() is a layer so use + to add it
```

```{r aesthetic-mappings-advanced-5-hint-2, eval = FALSE}
... |>
  ggplot(mapping = aes(x = ...,
                       y = ...,
                       color = ...)) +
  geom_point()
```

### 

`ggplot()` contains two specific uses: color and shape. We often use color when we are dealing with a vast amount of data. When compared to color, shape can only use up to 6 discrete values, meaning that we cannot utilize shape to its fullest with more data.

### Exercise 6

There are several ways to minimize overplotting. One such way is through the use of `alpha`. In `geom_point()` set `alpha` to 0.5.

```{r aesthetic-mappings-advanced-6, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r aesthetic-mappings-advanced-6-hint-1, eval = FALSE}
... +
  geom_point(alpha = ...)
```

### 

With the use of alpha we can control the transparency of the dots.

### Exercise 7

Now, within `geom_point()` set `size` to .25.

```{r aesthetic-mappings-advanced-7, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r aesthetic-mappings-advanced-7-hint-1, eval = FALSE}
... +
  geom_point(alpha = ...,
             size = ...)
```

### 

Size controls the diameter of the points. The default diameter size is 1.

### Exercise 8

Use the layer `scale_x_log10()` function to make the scale of the plot logarithmic.

```{r aesthetic-mappings-advanced-8, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r aesthetic-mappings-advanced-8-hint-1, eval = FALSE}
Remembder you are adding a layers here so use +
```

```{r aesthetic-mappings-advanced-8-hint-2, eval = FALSE}
... +
  scale_x_log10()
```

### 

Remember: a logarithmic scale means that each increment increases by a factor of the base of the logarithm. In our case, that's base 10. We want to implement the use of a logarithmic scale here because we want to be able to visualize all of our data without having any of it too close to each other.

### Exercise 9

To finish your plot, use `labs()` to give the graph a title, subtitle, axis labels, legend label, and caption of your choosing. To set the legend label, use the argument `color`.

Reminder: This is what your plot should look like.

```{r}
point_p2
```

```{r aesthetic-mappings-advanced-9, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r aesthetic-mappings-advanced-9-hint-1, eval = FALSE}
... +
  labs(...,
       color = "Clarity",
       ...)
```

### 

Now we know how to minimize the over plotting, it will come in handy in future graphs. Another way to minimize the over plotting is through `geom_jitter()`.

## Geometric Objects with `geom_smooth()`
### 

In this exercise, we will map Joe Biden's approval rating until June 30th, the first ~6 months of his term. We will explore different types of `geom_smooth()` in this section.

### 

We will be attempting to recreate this plot below.

```{r}
biden_p <- ggplot(data = polling_data, 
       mapping = aes(x = createddate,
                     y = rate,
                     color = approval_type)) +
  geom_point(alpha = 0.5) + 
  geom_smooth(method = "gam",
              formula = y ~ s(x, bs = "cs"),
              se = TRUE) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
  theme(legend.position="bottom") +
  labs(x = "Month",
       y = "Percent",
       title = "Biden's Approval Rating From Jan 22 to Jun 30",
       subtitle = "While Biden's approval rating has stayed relatively static, his disapproval rating \nhas been rising.",
       color = "",
       caption = "Source: Five Thirty Eight Data")

biden_p
```

### Exercise 1

We will first explore the file that contains this data. Run `summary()` on `polling_data` in the console below.

```{r geometric-objects-with-smooth-1, exercise = TRUE}

```

```{r geometric-objects-with-smooth-1-hint-1, eval = FALSE}
summary(...)
```

### 

This data was taken from the fivethirtyeight website. 

### Exercise 2

Write down the name of the three variables.

```{r geometric-objects-with-smooth-2}
question_text(NULL,
	message = "createddate, rate, approval_type",
	answer(NULL,
	correct = TRUE),
	allow_retry = FALSE,
	incorrect = NULL,
	rows = 6)
```

### 

`createddate` stands for the day the poll was finished. `rate` is the % of approval or disapproval, and `approval_type` indicates whether or not the rate is approval or disapproval.

### Exercise 3

We will get started building the plot. Start the function `ggplot()`, and set the argument `data` to `polling_data`.

```{r geometric-objects-with-smooth-3, exercise = TRUE}

```

```{r geometric-objects-with-smooth-3-hint-1, eval = FALSE}
ggplot(data = ...)
```

### 

Remember: you should not see a plot here as you have not mapped anything, nor do you have a geom layer. 

### Exercise 4

Now, map `createddate` to the x-axis, and map `rate` to the y-axis.

```{r geometric-objects-with-smooth-4, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r geometric-objects-with-smooth-4-hint-1, eval = FALSE}
ggplot(data = polling_data,
       mapping = aes(x = ...,
                     y = ...))
```

### 

`createddate` is a date column, and `rate` is a double column.

### Exercise 5

We want to be able to see both Joe Biden's approval and disapproval rate. Therefore, we'll change the color of the dots depending on whether the dot represents an approval/disapproval rating. Set `color` to `approval_type`.

```{r geometric-objects-with-smooth-5, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r geometric-objects-with-smooth-5-hint-1, eval = FALSE}
ggplot(data = polling_data,
       mapping = aes(x = createddate,
                     y = rate,
                     color = ...))
```

### 

Every geom function in ggplot2 takes a mapping argument, either defined locally in the geom layer or globally in the `ggplot()` layer. However, not every aesthetic works with every geom. Now we're ready to start building the plot.

### Exercise 6

Next, we will add the points to the graph. Add the layer `geom_point()`.

```{r geometric-objects-with-smooth-6, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r geometric-objects-with-smooth-6-hint-1, eval = FALSE}
ggplot(data = polling_data,
       mapping = aes(x = createddate,
                     y = rate,
                     color = approval_type)) +
  ...()
```

### 

Currently, we have a bit of overplotting. We will rectify this in the next code chunk.

### Exercise 7

In the `geom_point()` layer set the argument `alpha` to 0.5.

```{r geometric-objects-with-smooth-7, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r geometric-objects-with-smooth-7-hint-1, eval = FALSE}
ggplot(data = ...,
       mapping = aes(x = ...,
                     y = ...,
                     color = ...)) +
  geom_point(alpha = ...)
```

### 

Good! Now you should be able to see more distinct points.

### Exercise 8

We want to see the trend line for Biden's approval/disapproval ratings throughout his term. Add the layer of `geom_smooth().`

```{r geometric-objects-with-smooth-8, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r geometric-objects-with-smooth-8-hint-1, eval = FALSE}
... + 
  ...()
```

### 

If you just add this layer, you will get a few warning messages. We will fix this in the following exercises. If you place mappings in a geom function, `ggplot2` will treat them as local mappings for the layer. It will use these mappings to extend or overwrite the global mappings for that layer only. This makes it possible to display different aesthetics in different layers.

### Exercise 9

Within `geom_smooth()`, set the `method` argument to "gam". We use "gam" since there is a large quantity of data points.

```{r geometric-objects-with-smooth-9, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r geometric-objects-with-smooth-9-hint-1, eval = FALSE}
... + 
  geom_smooth(method = ...)
```

### 

"gam" stands for **generalized additive model**. It is a modified form of a linear model. If you are curious to learn more, [GAMs in R](https://noamross.github.io/gams-in-r-course/chapter1) is a good resource.

### Exercise 10

Within `geom_smooth()` set `formula` to `y ~ s(x, bs = "cs")`.

```{r geometric-objects-with-smooth-10, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r geometric-objects-with-smooth-10-hint-1, eval = FALSE}
... + 
  geom_smooth(method = "gam",
              formula = ...)
```

### 

Just like `y ~ x` indicates a relationship between x and y, `y ~ s(x, bs = "cs")` indicates a "generalized additive model" relationship. With a generalized additive model we are able to handle large data.

### Exercise 11

Set `se = TRUE`.

```{r geometric-objects-with-smooth-11, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r geometric-objects-with-smooth-11-hint-1, eval = FALSE}
... + 
  geom_smooth(method = "gam",
              formula = y ~ s(x, bs = "cs"),
              se = ...)
```

### 

With `se` we are able to create a margin of error that makes our data more accurate.

### Exercise 12

Next, we want to change the labels on the graph so they are in percentage format. Use scale_y_continuous to do so. Set the `labels` argument to `scales::percent_format(accuracy = 1)`.

```{r geometric-objects-with-smooth-12, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r geometric-objects-with-smooth-12-hint-1, eval = FALSE}
...+
  scale_y_continuous(labels = ...)
```

### 

We use scale_y_continuous when we want to modify the y-axis for continuous data. [This documentation here](https://ggplot2.tidyverse.org/reference/scale_continuous.html) will tell you more, and all the possible arguments you can use with `scale_y_continuous`

### Exercise 13

Currently our legend is to the right of the graph. Let's change the way that we see our legend. Add `theme()` to the pipe and set `legend.position` to "bottom".

```{r geometric-objects-with-smooth-13, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r geometric-objects-with-smooth-13-hint, eval = FALSE}
+ 
  theme(...="bottom")
```

###

With the help of `legend.position`. We have the ability to set it to the `left`, `right`, `top`, and `bottom` based on how we would like to see it within our graph.

### Exercise 14

Finally, use `labs()` to add an appropriate heading, subtitle, caption, etc. 

Reminder - your graph should look like this:

```{r}
biden_p
```

```{r geometric-objects-with-smooth-14, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r geometric-objects-with-smooth-14-hint-1, eval = FALSE}
... +
  labs(x = ...,
       y = ...,
       title = ...,
       subtitle = ...,
       color = ...,
       caption = ...)
```

### 

Good! Now you have a plot of Biden's approval rating for the first six months of his term.

## Geometric Objects with `geom_histogram()` and `geom_boxplot()`
###

Let's create the following histogram.

```{r}
hist_p <- mpg |> 
  mutate(year = as.character(year)) |> 
  ggplot(aes(cty, fill = year)) +
  geom_histogram(bins = 6,
                 color = "black") +
  labs(title = "Histogram of City Fuel Efficiency in 1999 and 2008 Models", 
       subtitle = "1999 had more models of high city fuel efficiency", 
       x = "City MPG", 
       y = "Count",
       fill = "Year",
       caption = "Source: EPA") 

hist_p
```

### Exercise 1

Run `?mpg` in the code chunk below. 

```{r geometric-objects-wi-1, exercise = TRUE}

```

This dataset is about fuel economy of cars from the Environmental Protection Agency.

### Exercise 2

Run `glimpse()` on the mpg data set. Take a look at the different variables available.

```{r geometric-objects-wi-2, exercise = TRUE}

```

```{r geometric-objects-wi-2-hint-1, eval = FALSE}
glimpse(...)
```

### 

The variables that we care about are `year` and `cty`. Year indicates the year of manufacture, and `cty` indicates the city miles per gallon.

### Exercise 3

Use the `$` to extract the column `year` as a vector.

```{r geometric-objects-wi-3, exercise = TRUE}

```

```{r geometric-objects-wi-3-hint-1, eval = FALSE}
mpg$...
```

### 

Notice that there are only two different years, 1999 and 2008.

### Exercise 4

Start a new pipe with `mpg`. Use `mutate()` and the function `as.character()` to change the `year` column from a numeric variable to a character variable.

```{r geometric-objects-wi-4, exercise = TRUE}

```

```{r geometric-objects-wi-4-hint-1, eval = FALSE}
...  |> 
  mutate(... = as.character(...))
```

### 

We make this change so that we can use year as a fill color when we plot. We use the `as.character()` function to convert an object into its character representation for textual manipulation or analysis.

### Exercise 5

Select the columns `year` and `cty`.

```{r geometric-objects-wi-5, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r geometric-objects-wi-5-hint-1, eval = FALSE}
...  |> 
  select(..., ...)
```

### Exercise 6

Continue your pipe and add `ggplot()`. Map `cty` to the x-axis and `year` to `fill`.

```{r geometric-objects-wi-6, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r geometric-objects-wi-6-hint-1, eval = FALSE}
... |> 
  ggplot(aes(x = ..., fill = ...))
```

### 

We don't have to use `mapping = aes()` because it is in the variable spot for mapping, meaning we don't have to specify `mapping = aes()`. Again, you should not see anything. The `geom` layer is what displays objects on the plot.

### Exercise 7

Now, we will add a geom layer. Add the layer `geom_histogram()`. 

```{r geometric-objects-wi-7, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r geometric-objects-wi-7-hint-1, eval = FALSE}
geom_histogram() is a layer so use + to add it
```

```{r geometric-objects-wi-7-hint-2, eval = FALSE}
...|> 
  ... +
  ...()
```

### 

You may notice that there are too many "bins" for our liking. There can be too many bins if the number chosen for a histogram or other visualizations is too high compared to the size of data, which results in loss of clarity. We'll change this now.

### Exercise 8

Within `geom_histogram()` set the number of `bins` to 6.

```{r geometric-objects-wi-8, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r geometric-objects-wi-8-hint-1, eval = FALSE}
... +
  geom_histogram(bins = ...)
```

You can change the width of each bin with the `binwidth` argument. We won't be doing that now, but you can do that in future plots. Read more about [**binwidth**](https://r-graphics.org/recipe-quick-histogram) and how it can be used.

### Exercise 9

Within `geom_histogram()` set  `color` to `"black"`. 

```{r geometric-objects-wi-9, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r geometric-objects-wi-9-hint-1, eval = FALSE}
... +
  geom_histogram(bins = 6, 
                 color = ...)
```

Remember, `color` only changes the outline of the bars. If you want to change the fill, use `fill`. 

### Exercise 10

To finish your plot, use `labs()` to give the graph a title, subtitle, axis labels, legend labels, and caption of your choosing. To change the legend labels use the argument `fill`. 

```{r geometric-objects-wi-10, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r geometric-objects-wi-10-hint-1, eval = FALSE}
... +
  labs(..., 
       fill = "Year",
       ...)
```

Reminder: This is what your graph should look like. 

```{r show-hist-p}
hist_p
```

### 

In the previous graph we used `color` to change the legend labels, while here we used `fill`. The argument within labels matches the argument you use within `ggplot()`. `color` is used for objects like lines and points. `fill` is used for larger shapes. 

Are you satisfied with the way we have made the bin sizes? Do the axes labels make this easy to understand? We will do better later. 

[**Histograms**](https://socviz.co/groupfacettx.html#histograms) contain many more vast resources with items that contain comparing distributions in a new way.

### Exercise 11

Let's try the same data for before but with a `geom_boxplot()` instead of a `geom_histogram()`. Box plots compute the five-number summary of the distribution and then display that summary as a specially formatted box.

```{r geometric-objects-wi-11, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r geometric-objects-wi-11-hint-1, eval = FALSE}
... +
  ...()
```

### 

ggplot2 provides a vast amount of opportunities, however, there are a countless number of plots that can be made through different geoms and extensions. Check out [**possible options**](https://exts.ggplot2.tidyverse.org/gallery/) for any future implementations or possibilities. 

## Facets and Statistical Transformations with `geom_col()`, `facet_wrap()`, and `facet_grid()`
### 

Let's make the same plot, using a related function, `geom_col()`. Unlike `geom_bar()`, you have to specify the value for the y-axis when using this function. 

```{r make-col-p}
col_p <- mpg |> 
  select(drv, fl, class) |>
  summarize(total = n(),
            .by = c(drv, fl, class)) |> 
  ggplot(aes(x = fl, 
             y = total, 
             fill = class)) +
    geom_col() +
    facet_wrap(~ drv,
               labeller = as_labeller(car_names)) +
    labs(title = "The Fuel Type Of Cars Based On Their Class Type",
         subtitle = "Higher Proportion of 4 Wheel and Front Wheel Drive Participants",
         x = "Fuel Type",
         y = "Number of Participants",
         fill = "Class")
col_p
```

### Exercise 1

Start a pipe with `mpg`. Select the columns `drv`, `fl` and `class`. 

```{r facets-statistical-transformations-1, exercise = TRUE}

```

```{r facets-statistical-transformations-1-hint-1, eval = FALSE}
mpg |> 
  select(..., ..., ...)
```

### Exercise 2

Copy/paste the previous code. Use `summarize()` to create a new table with the column `total` which is equal to the function `n()`. 

```{r facets-statistical-transformations-2, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r facets-statistical-transformations-2-hint-1, eval = FALSE}
... |> 
  summarize(total = ...)
```

```{r geomcol-2-hint-2, eval = FALSE}
... |> 
  summarize(total = n())
```

### Exercise 3

Add a `.by` argument to `summarize()` with the values set to `drv`, `fl`, and `class`. You need to place these three values with a vector using `c()`. 

```{r facets-statistical-transformations-3, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r facets-statistical-transformations-3-hint-1, eval = FALSE}
... |> 
  summarize(total = n(),
            .by = c(..., ..., ...))  
```

### Exercise 4

Continue your pipe with `ggplot()`. Map the variable `fl` to the x-axis. Map `total` to the `y` axis. Add the layer `geom_col()`.

```{r facets-statistical-transformations-4, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r facets-statistical-transformations-4-hint-1, eval = FALSE}
ggplot(aes(x = ..., 
           y = ...) +
  ...()
```

### 

`geom_col()` will automatically take the three to four values for each `fl` in the `total` column and stack them to form one bar. This behavior is from the fact that the `position` argument for `geom_col()` is set by default to "stack". 

### Exercise 5

Now, map `class` to `fill`.

```{r facets-statistical-transformations-5, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r facets-statistical-transformations-5-hint-1, eval = FALSE}
ggplot(data = ... , 
          aes(x = ..., 
              y = ...,
              fill = ...) +
  geom_col()
```

### Exercise 6

Add the layer `facet_wrap()` and facet by the variable `drv`. 

```{r facets-statistical-transformations-6, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r facets-statistical-transformations-6-hint-1, eval = FALSE}
... + 
  facet_wrap(~ ...)
```

###

The `facet_wrap()` function arranges the panels in a grid, wrapping them onto multiple rows and columns. It is useful for visualizing and comparing subsets of data simultaneously.

### Exercise 7

We want to change the labels of the facet_wrap that we just created. We can do this by setting `labeller` to `as_labeller(car_names)`. Note: `car_names` has already been created for you.

```{r facets-and-statistic-7, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r facets-and-statistic-7-hint, eval = FALSE}
facet_wrap(~...,
               ... = as_labeller(car_names)
```

###

Through the `as_labeller()` function we are able to change the labels of `facet_wrap()`.

### Exercise 8

To finish your plot, use `labs()` to give the graph a title, subtitle, axes labels, legend labels, and caption of your choosing.

Reminder: This is what your plot should look like.

```{r}
col_p
```

```{r facets-statistical-transformations-8, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r facets-statistical-transformations-8-hint-1, eval = FALSE}
... +
  labs(..., 
       fill = "...",
       ...)
```

###

Through this we are able to view our data with a `facet_wrap` which splits a plot into subplots that each display one subset of the data based on a categorical variable.

## Statistical Transformations and Position Adjustments with `geom_bar()`
### 

Let's create this graph using `geom_bar()`.

```{r}
bar_p <- ggplot(data = midwest, 
                   mapping = aes(x = state, 
                                 fill = category)) + 
  geom_bar(alpha = 1/5, 
           position = "dodge") + 
  labs(title = "Categories found in each state",
       subtitle = "The 5 states in the midwest with each category present.",
       x = "State",
       y = "Count",
       fill = "Category")

bar_p
```

### Exercise 1


Run `glimpse()` on the `midwest` data set. 

```{r statistical-transfor-1, exercise = TRUE}

```

```{r statistical-transfor-1-hint-1, eval = FALSE}
...(midwest)
```

###


### Exercise 2

Add `ggplot()` and map `state` to the x axis of the `midwest` data and set `fill` to `category`.

```{r statistical-transfor-2, exercise = TRUE}

```

```{r statistical-transfor-2-hint-1, eval = FALSE}
ggplot(data = ..., 
       mapping = aes(x =..., 
                     fill =...))
```

### 

Here we have used `state` data on the x axis and the count on the y axis is automatically implemented. Within a bar graph, we only input the x variable due to the y variable automatically being created usually as count. Through this we can see count within each `state` in the `midwest` data.

### Exercise 3

Add the layer `geom_bar()`.

```{r statistical-transfor-3, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r statistical-transfor-3-hint, eval = FALSE}
... +
  geom_bar()
```

###

`geom_bar()` displays the frequency or count of a categorical variable as bars of equal width. By default, `geom_bar()` counts the number of occurrences of each category and maps them to the height of the bars.

### Exercise 4

Within `geom_bar()`, set `alpha` to 1/5, `position` to `"dodge"`.

```{r statistical-transfor-4, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r statistical-transfor-4-hint-1, eval = FALSE}
... + geom_bar(alpha = ..., 
               position = ...)
```

### Exercise 5

Finally, use `labs()` to give the graph a title, subtitle, axis labels, legend labels, and caption of your choosing.

```{r statistical-transfor-5, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r statistical-transfor-5-hint-1, eval = FALSE}
... +
  labs(..., 
       fill = "...",
       ...)
```

Reminder: This is what your plot should look like.

```{r make-bar-p}
bar_p
```

### 

We like using bar charts here because they bin your data and then plot bin counts, the number of points that fall in each bin.

###

For more information on other stat_summary plots besides bar graphs check out [Statistical transformations](https://r4ds.hadley.nz/layers.html#statistical-transformations) from [Chapter 10: Layers](https://r4ds.hadley.nz/layers.html) in [*R for Data Science (2e)*](https://r4ds.hadley.nz/).

## Coordinate Systems
###

Coordinate systems are probably the most complicated part of ggplot2. The default coordinate system is the Cartesian coordinate system where the x and y positions act independently to determine the location of each point. We are going to create one of these systems with the US.

### Exercise 1

Load the `maps` package with `library()`.

```{r coordinate-systems-1, exercise = TRUE}

```

```{r coordinate-systems-1-hint, eval = FALSE}
library(...)
```

###

The `maps` package includes the area of each matching region in the map is computed, and regions which match the same element of regions have their areas combined. Each region is assumed planar, with vertices specified by the x and y components of the map object.

### Exercise 2

Run `summary()` on `nz` to view its contents.

```{r coordinate-systems-2, exercise = TRUE}

```

```{r coordinate-systems-2-hint, eval = FALSE}
nz <- map_data("nz")
```

### Exercise 3

We now want to plot our `nz` data with `ggplot()`

```{r coordinate-systems-3, exercise = TRUE}

```

```{r coordinate-systems-3-hint, eval = FALSE}
ggplot(...)
```

###

We don't have to use `data = nz` because it is in the variable spot for data, meaning we don't have to specify `data = nz`.

### Exercise 4

Set x-axis to `long` and y-axis to `lat` and group to `group`

```{r coordinate-systems-4, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r coordinate-systems-4-hint, eval = FALSE}
ggplot(nz, aes(x = ..., 
               y = ...,
               group = ...))
```

### Exercise 5

Now we want to add the layer of `geom_polygon()`.

```{r coordinate-systems-5, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r coordinate-systems-5-hint, eval = FALSE}
... +
  ...()
```

###

`geom_polygon()` is used to create filled polygons or regions on a plot. Commonly used to visualize geographic data, such as maps, choropleth plots, or any other data that can be represented as polygons.

### Exercise 6

Now we want to add color the item that we have plotted. Set `fill` to white and `color` to black

```{r coordinate-systems-6, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r coordinate-systems-6-hint, eval = FALSE}
... +
  geom_polygon(fill = ..., 
               color = ...)
```

### Exercise 7

Finally let us use the `coord_quickmap()` which sets the aspect ratio correctly for geographic maps. This is very important because we are plotting spatial data with `ggplot2`

```{r coordinate-systems-7, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r coordinate-systems-7-hint, eval = FALSE}
... +
  coord_quickmap()
```

###

Now you should see the New Zealand data plotted out based on the spatial data. We can visualize more data through `coord_polar()`. Take a look at the use of [`coord_polar()`](https://r4ds.hadley.nz/layers.html#coordinate-systems)

## Summary
### 

This tutorial covered [Chapter 10: Layers](https://r4ds.hadley.nz/layers.html) from [*R for Data Science (2e)*](https://r4ds.hadley.nz/) by Hadley Wickham, Mine Çetinkaya-Rundel, and Garrett Grolemund. You learned how to use [**ggplot2**](https://ggplot2.tidyverse.org/) and related packages to create persuasive graphics.

```{r download-answers, child = system.file("child_documents/download_answers.Rmd", package = "tutorial.helpers")}
```
