---
title: Data visualization
author: Anish Bellamkonda
tutorial:
  id: data-visualization
output:
  learnr::tutorial:
    progressive: yes
    allow_skip: yes
runtime: shiny_prerendered
description: 'Tutorial for Chapter 2: Data Visualization'
---

```{r setup, include = FALSE}
library(learnr)
library(tutorial.helpers)
library(ggplot2)
library(tidyverse)
library(palmerpenguins)
library(ggthemes)
knitr::opts_chunk$set(echo = FALSE)
options(tutorial.exercise.timelimit = 60, 
        tutorial.storage = "local") 
```

```{r copy-code-chunk, child = system.file("child_documents/copy_button.Rmd", package = "tutorial.helpers")}
```

```{r info-section, child = system.file("child_documents/info_section.Rmd", package = "tutorial.helpers")}
```

## Introduction
### 

This tutorial covers [Chapter 2: Data Visualization](https://r4ds.hadley.nz/data-visualize.html) from [*R for Data Science (2e)*](https://r4ds.hadley.nz/) by Hadley Wickham, Mine Çetinkaya-Rundel, and Garrett Grolemund. 

## Creating first ggplot
### 

<!-- AB: Need to chnage the code so eveyrhting has a pipe -->

Our Ultimate Goal in this section is to create a plot displaying the relationship between flipper lengths and body masses of these penguins, taking into consideration the species of the penguin.Before doing so, lets load the palmerpenguins package using `library()`.

### Exercise 1

```{r creating-first-ggplo-1, exercise = TRUE}

```

```{r creating-first-ggplo-1-hint-1, eval = FALSE}
library(palmerpenguins)
```

### 

A `variable`: a quanitty, quality, or property that you can measure, `value`: state of a variable when you measure it, `observation`: a set of measurements made under similiar conditions.

### Exercise 2

Run `?palmerpenguins`. After doing so, copy and paste the description here.

```{r creating-first-ggplo-2}
question_text(NULL,
	answer(NULL, correct = TRUE),
	allow_retry = TRUE,
	try_again_button = "Edit Answer",
	incorrect = NULL,
	rows = 3)
```

### 

The data include size measurements, clutch observations, and blood isotope ratios for adult foraging Adélie, Chinstrap, and Gentoo penguins observed on islands in the Palmer Archipelago near Palmer Station, Antarctica. 

### Exercise 3

Run the `library()` command with `ggthemes` as the argument.

```{r creating-first-ggplo-3, exercise = TRUE}

```

```{r creating-first-ggplo-3-hint-1, eval = FALSE}
library(ggthemes)
```

### 

We can infer from the name of the package itself that `ggthemes` is used for aesthetic looks of a graph and make it fancy.



### Exercise 4

Type `penguins` and hit Run Code.


```{r creating-first-ggplo-4, exercise = TRUE}

```

```{r creating-first-ggplo-4-hint-1, eval = FALSE}
penguins
```

### 

Tabular data is a set of values, each associated with a variable and an observation. Tabular data is tidy if each value is placed in its own “cell”, each variable in its own column, and each observation in its own row. `penguins` is tabular data.

### Exercise 5

Run `glimpse()` with `penguins` as its argument.

```{r creating-first-ggplo-5, exercise = TRUE}

```

```{r creating-first-ggplo-5-hint-1, eval = FALSE}
glimpse(...)
```

### 

In RStudio, you can click on the name of the data frame in the Environment pane or run `View(penguins)` to open an interactive data viewer.

### Exercise 6

<!-- AB: teach what a pipe is  -->

The key command for making plots is `ggplot()` from the **ggplot2** package. Run `ggplot(data = penguins)`.

```{r creating-first-ggplo-6, exercise = TRUE}

```

```{r creating-first-ggplo-6-hint-1, eval = FALSE}
ggplot(data = penguins)
```

### 

There is another way to create a plot without using `|>` which is to use the data arguemnt within the `ggplot()` like `ggplot(data = penguins)`.


### Exercise 7

Run `sqrt(16)`

```{r creating-first-ggplo-7, exercise = TRUE}

```

```{r creating-first-ggplo-7-hint-1, eval = FALSE}

```

### 

### Exercise 8

Run `16 |> sqrt()`
```{r creating-first-ggplo-8, exercise = TRUE}

```

```{r creating-first-ggplo-8-hint-1, eval = FALSE}

```

### 

The difference between the previous code was it that we are using the pipe operator `|>` which takes anything from the left side and plugs it into the function in the right side.

### Exercise 9

Run `penguins` with the pipe operator `|>` and plug it into `ggplot()` like this
`penguins |>
  ggplot()`
```{r creating-first-ggplo-9, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r creating-first-ggplo-9-hint-1, eval = FALSE}
penguins |>
    ggplot()
```

### 

penguins which is the input is being plugged into the function `ggplot()` which gives the same output as `ggplot(data = penguins)`. While Indentation may not affect code, it is a good habit to do because it helps go through the code much faster.




### Exercise 10

Copy the previous code and within `ggplot()`, run `mapping = aes(x = flipper_length_mm)` and see what you get.
```{r creating-first-ggplo-10, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r creating-first-ggplo-10-hint-1, eval = FALSE}
penguins |>
  ggplot(... = aes(x = ...))
```

### 

`mapping = aes(x = flipper_length_mm)` specifies that the variable `flipper_length_mm` is mapped to the x-axis of a plot. This allows the length of the penguin's flipper to be plotted on the x-axis.

### Exercise 11

Copy the previous code, within the `aes()` set `y` equal to `body_mass_g` and run it.
```{r creating-first-ggplo-11, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r creating-first-ggplo-11-hint-1, eval = FALSE}
penguins |>
  ggplot(mapping = aes(x = ..., y = ...))
```

To display the data points we need to add a geometric object(geom). We can define a `geom`: the geometrical object that a plot uses to represent data. These geometric objects are made available in ggplot2 with functions that start with `geom_`. 

### 

### Exercise 12

The function `geom_point()` adds a layer of points to your plot, which creates a scatterplot so copy the previous code and run `+ geom_point()`

```{r creating-first-ggplo-12, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r creating-first-ggplo-12-hint-1, eval = FALSE}
penguins |>
  ggplot(
    mapping = aes(x = flipper_length_mm, y = ...)
    ) +
  geom_point()
```

### 

`geom_point()` is not the only geometric objects, there are more than this single one and we will learn about them in other sections in this tutorial.

### Exercise 13

Scatter plot are useful for displaying relationship between x and y but it is also helpful to ask yourself if there are other variable which might contribute to the relationship. For example, does the relationship between x and y differ for different species of penguins. To find the relationship, lets add more the aesthetic `color` within `aes()` to `species`.
```{r creating-first-ggplo-13, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r creating-first-ggplo-13-hint-1, eval = FALSE}
penguins |>
  ggplot(
  mapping = aes(x = flipper_length_mm, y = ..., ... = species)
)+
  geom_point()
```

### 

When you ran it, you see that the data points differ by the color of the species which just makes the scatter plot more insightful. Color inst just the only aesthetics. There are more, to find out more go check out [aesthetic mappings](https://ggplot2.tidyverse.org/reference/aes.html)


### Exercise 14

Now let's add one more layer: a smooth curve which will dispaly the reltionship between body mass and flipper length. To do so, lets add a new geometric object:`geom_smooth()` and within that we will specity it to be a best of fit line of a linear model.
```{r creating-first-ggplo-14, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r creating-first-ggplo-14-hint-1, eval = FALSE}
penguins |>
  ggplot(
  mapping = aes(x = flipper_length_mm, y = body_mass_g, color = ...)
) +
  geom_point() +
  geom_smooth(... = "lm")
```

### 

If you don't specify the model for geom_smooth(), RStudio will automatically set the method to a linear model, adding on you will get a message saying that it was set to `lm`.

### Exercise 15

We have added new lines to the plot, but it does not match the plot in beginning, which displayed only one line for the entire dataset instead of separate lines for each penguin species.

In **ggplot2**, when aesthetic mappings are defined at the global level, they are passed down to all subsequent geom layers in the plot. However, each geom function in ggplot2 can also accept a mapping argument, allowing for local-level aesthetic mappings that are added to those inherited from the global level. To achieve our desired outcome of coloring the points based on the species without separating the lines for each species, we should specify `color = species` for `geom_point()` only.

```{r creating-first-ggplo-15, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r creating-first-ggplo-15-hint-1, eval = FALSE}
penguins |>
  ggplot(
  mapping = ...(x = flipper_length_mm, y = body_mass_g)
) +
  geom_point(mapping = ...(color = species)) +
  ...(method = "lm")
```

### 

Great progress! The plot now resembles our ultimate goal, although it's not perfect yet. We still need to differentiate the species of penguins by using different shapes for each species, and also improve the labels.

### Exercise 16

Representing information using only colors on a plot may not be a good idea as individuals perceive colors differently due to color blindness or other color vision differences. Therefore, it's advisable to map the `species` to the `shape` aesthetic in addition to `color` to ensure everyone can differentiate the species.

```{r creating-first-ggplo-16, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r creating-first-ggplo-16-hint-1, eval = FALSE}
penguins |>
  ggplot(
  mapping = aes(x = flipper_length_mm, y = body_mass_g)
) +
  geom_point(mapping = aes(color = ..., shape = ...)) +
  geom_smooth(method = "lm")
```

### 

### Exercise 17

Now that we have the data points, legend, and lines set, lets change the labels for the graph and make it more insightful and visually appealing. Copy the previous code and add a new layear using `+` and type `labs()`. Within labs, type `title` and set it equal to "Body mass and flipper length" in quotes.Let's also add the subtitle by typing a comma and then 'subtitle'. Set  `Subtitle` to "Dimensions for Adelie, Chinstrap, and Gentoo Penguins". Run the code and see what you get.

```{r creating-first-ggplo-17, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r creating-first-ggplo-17-hint-1, eval = FALSE}
penguins |>
  ggplot(
  mapping = aes(x = flipper_length_mm, y = body_mass_g)
) +
  geom_point(mapping = aes(color = species, shape = species)) +
  geom_smooth(method = "lm")+
  labs( ... = "Body mass and flipper length"+
          subtitle = "Dimensions for Adelie, Chinstrap, and Gentoo Penguins")
```

### 

### Exercise 18

We see that the title and subtitle have been added, now lets modify the x and y axises, type a comma and then type `x`. Then set `x` to "Flipper Length (mm)" in quotes. Type another comma and a `y`, then set it equal to "Body mass (g)" in quotes.

```{r creating-first-ggplo-18, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r creating-first-ggplo-18-hint-1, eval = FALSE}
penguins |>
  ggplot(
  mapping = aes(x = flipper_length_mm, y = body_mass_g)
) +
  geom_point(mapping = aes(color = species, shape = species)) +
  geom_smooth(method = "lm")+
  labs( ... = "Body mass and flipper length"+
          subtitle = "Dimensions for Adelie, Chinstrap, and Gentoo Penguins",
        x = ..., y = ...)
```

### 

### Exercise 19

You might see that it looks the same as the one in the beginning but we forgot one minor thing which is capitalizing the legend, we can do so by typing a comma and setting both `color` and `shape` equal to "Species" in quotes.
```{r creating-first-ggplo-19, exercise = TRUE}
penguins |>
  ggplot(
  mapping = aes(x = flipper_length_mm, y = body_mass_g)
) +
  geom_point(mapping = aes(color = species, shape = species)) +
  geom_smooth(method = "lm")+
  labs( ... = "Body mass and flipper length"+
          subtitle = "Dimensions for Adelie, Chinstrap, and Gentoo Penguins",
        x = ..., y = ...,
         color = "Species", shape = "Species")
```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r creating-first-ggplo-19-hint-1, eval = FALSE}

```

### 

## Exercise
### 

## Summary
### 

This tutorial covered [Chapter 2: Data Visualization](https://r4ds.hadley.nz/data-visualize.html) from [*R for Data Science (2e)*](https://r4ds.hadley.nz/) by Hadley Wickham, Mine Çetinkaya-Rundel, and Garrett Grolemund. 


```{r download-answers, child = system.file("child_documents/download_answers.Rmd", package = "tutorial.helpers")}
```
