---
title: "Data visualization"
tutorial:
  id: "data-visualization"
output: 
  learnr::tutorial:
    progressive: TRUE
    allow_skip: TRUE
runtime: shiny_prerendered
description: "Tutorial for Chapter 2: Data Visualization"
---

```{r setup, include = FALSE}
library(learnr)
library(ggplot2)
library(tidyverse)
library(palmerpenguins)
library(ggthemes)
knitr::opts_chunk$set(echo = FALSE)
options(tutorial.exercise.timelimit = 60, 
        tutorial.storage = "local") 
```

```{r copy-code-chunk, child = system.file("child_documents/copy_button.Rmd", package = "tutorial.helpers")}
```

```{r info-section, child = system.file("child_documents/info_section.Rmd", package = "tutorial.helpers")}
```

## Introduction
###

This tutorial covers [Chapter 2: Data Visualization](https://r4ds.hadley.nz/data-visualize.html) from [*R for Data Science (2e)*](https://r4ds.hadley.nz/) by Hadley Wickham, Mine Çetinkaya-Rundel, and Garrett Grolemund. 

## Creating first ggplot
###

<!-- AB: Need to chnage the code so eveyrhting has a pipe -->

Our Ultimate Goal in this section is to create a plot displaying the relationship between flipper lengths and body masses of these penguins, taking into consideration the species of the penguin.

### Exercise 1

Before diving into making our first plot, let's have a quick look at the dataset which we would be using to make a plot. Use the `glimpse()` function and within it, type in the name of the data frame which is `penguins` and see what kind of variables there are.
```{r creating-first-ggplo-1, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r creating-first-ggplo-1-hint, eval = FALSE}
glimpse(penguins)
```
What `glimpse()` does is 


###

### Exercise 2

To create our 
```{r creating-first-ggplo-2, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r creating-first-ggplo-2-hint, eval = FALSE}
ggplot(data = penguins)
```

###

### Exercise 3


```{r creating-first-ggplo-3, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r creating-first-ggplo-3-hint, eval = FALSE}
...(
  data = penguins,
  mapping = aes(x = flipper_length_mm, y = ...)
)
```

###


### Exercise 4


```{r creating-first-ggplo-4, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r creating-first-ggplo-4-hint, eval = FALSE}
ggplot(
  data = penguins,
 ... = aes(x = ..., y = body_mass_g)
) +
  geom_point()
```

###

### Exercise 5


```{r creating-first-ggplo-5, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r creating-first-ggplo-5-hint, eval = FALSE}
ggplot(
  data = penguins,
  mapping = ...(x = flipper_length_mm, y = body_mass_g, ... = species)
) +
  geom_point()
```

###


### Exercise 6


```{r creating-first-ggplo-6, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r creating-first-ggplo-6-hint, eval = FALSE}
ggplot(
  ... = penguins,
  mapping = aes(x = flipper_length_mm, y = body_mass_g, color = ...)
) +
  geom_point() +
  geom_smooth(... = "lm")
```

###

### Exercise 7


```{r creating-first-ggplo-7, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r creating-first-ggplo-7-hint, eval = FALSE}
ggplot(
  data = penguins,
  mapping = ...(x = flipper_length_mm, y = body_mass_g)
) +
  geom_point(mapping = ...(color = species)) +
  ...(method = "lm")
```

###

### Exercise 8


```{r creating-first-ggplo-8, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r creating-first-ggplo-8-hint, eval = FALSE}
ggplot(
  data = ...,
  mapping = aes(x = flipper_length_mm, y = body_mass_g)
) +
  geom_point(mapping = aes(color = ..., shape = ...)) +
  geom_smooth(method = "lm")
```

###

## Exericse
###








## Summary
###

This tutorial covered [Chapter 2: Data Visualization](https://r4ds.hadley.nz/data-visualize.html) from [*R for Data Science (2e)*](https://r4ds.hadley.nz/) by Hadley Wickham, Mine Çetinkaya-Rundel, and Garrett Grolemund. 


```{r download-answers, child = system.file("child_documents/download_answers.Rmd", package = "tutorial.helpers")}
```
