---
title: Data visualization
tutorial:
  id: data-visualization
output:
  learnr::tutorial:
    progressive: yes
    allow_skip: yes
runtime: shiny_prerendered
description: 'Tutorial for Chapter 2: Data Visualization'
---

```{r setup, include = FALSE}
library(learnr)
library(ggplot2)
library(tidyverse)
library(palmerpenguins)
library(ggthemes)
knitr::opts_chunk$set(echo = FALSE)
options(tutorial.exercise.timelimit = 60, 
        tutorial.storage = "local") 
```

```{r copy-code-chunk, child = system.file("child_documents/copy_button.Rmd", package = "tutorial.helpers")}
```

```{r info-section, child = system.file("child_documents/info_section.Rmd", package = "tutorial.helpers")}
```

## Introduction
### 

This tutorial covers [Chapter 2: Data Visualization](https://r4ds.hadley.nz/data-visualize.html) from [*R for Data Science (2e)*](https://r4ds.hadley.nz/) by Hadley Wickham, Mine Çetinkaya-Rundel, and Garrett Grolemund. 

## Creating first ggplot
### 

<!-- AB: Need to chnage the code so eveyrhting has a pipe -->

Our Ultimate Goal in this section is to create a plot displaying the relationship between flipper lengths and body masses of these penguins, taking into consideration the species of the penguin.Before doing so, lets load the palmerpenguins package using `library()`.

### Exercise 1

```{r creating-first-ggplo-1, exercise = TRUE}

```

```{r creating-first-ggplo-1-hint-1, eval = FALSE}
library(palmerpenguins)
```

### 

To make this turtorial easier for you to comprehend, let's define a `variable`: a quanitty, quality, or property that you can measure, `value`: state of a variable when you measure it, `observation`: a set of measurements made under similiar conditions.

### Exercise 2

Before creating our first plot, we should learn a little more about the data set we will be using so going to help page of the palmerpenguins by typing `?palmerpenguins`. After doing so, copy and paste the description here.
```{r creating-first-ggplo-2}
question_text(NULL,
	answer(NULL, correct = TRUE),
	allow_retry = TRUE,
	try_again_button = "Edit Answer",
	incorrect = NULL,
	rows = 3)
```

### 

In this pacakge, we have the `penguins` dataset which contains data about different species(Adelie, Chinstrap, or Gentoo), flipper_length, and mass of a penguin.

### Exercise 3

Let's load the next library for this plot which is the `ggthemes` with `library()`
```{r creating-first-ggplo-3, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r creating-first-ggplo-3-hint-1, eval = FALSE}
library(ggthemes)
```

### 

We can infer from the name of the package itself that `ggthemes` is used for aesthetic looks of a graph and make it fancy.



### Exercise 4

Before diving into making our first plot, let's have a quick look at the dataset which we would be using to make a plot. Just type in the name of the data set whihc is palmerpenguins and see what you get.
```{r creating-first-ggplo-4, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r creating-first-ggplo-4-hint-1, eval = FALSE}
penguins
```

### 

To view the data set in alternative way is to use `glimpse(penguins)` which shows all the variables compared to just printing the data set.

### Exercise 5

To make a plot, we need to follow these steps:

Type in the name of the dataset we want to use, in this case penguins.

Type a space and then |> (this is called a "pipe" symbol), which lets us chain multiple functions together.

Press enter.

Indent the next line and type ggplot(), which creates a plot.

By doing this, we're essentially telling the computer to use the penguins data set as the starting point, and then use the ggplot() function to create a plot. The |> symbol lets us pass the penguins dataset to other functions in a chain, which can help us do more complex things with the data.
```{r creating-first-ggplo-5, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r creating-first-ggplo-5-hint-1, eval = FALSE}
penguins |>
  ggplot()
```

### 

There is another way to create a plot without using `|>` which is to use the data arguemnt within the `ggplot()` like `ggplot(data = penguins)`.

### Exercise 6

In the previous exercise, we saw an empty plot with only the outline. This happened because we didn't tell ggplot() how to map the variables from our dataset to the aesthetics of our plot.

To fix this, we need to use the mapping argument and set it to aes. Inside aes, we need to specify which variable to use for the x-axis and which one for the y-axis. For this specific example, we want to use flipper_length_mm for the x-axis and body_mass_g for the y-axis.

So, in summary, we need to add the following code after the ggplot() function:
`mapping = aes(x = flipper_length_mm, y = body_mass_g)`
```{r creating-first-ggplo-6, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r creating-first-ggplo-6-hint-1, eval = FALSE}
penguins |>
  ggplot(
  mapping = aes(x = flipper_length_mm, y = ...)
)
```

### 

Without mapping, ggplot() will still create a plot, but it may not be what we want. It will use default aesthetics for the plot, such as the x and y axes labels, colors, shapes, etc., which may not be the best choice for our data. We can customize these aesthetics later, but it requires more code.

### Exercise 7

After running the code in previous code, we see that the plot has more structure but the data point themselves are not yet on the plot. To represent the obeservarions from out data set on out plot, we need to use a geometric object  and for this we will use point object. To do, add a new layer using `+` after the ggplot function and add `geom_pont()` wich will add the points.
```{r creating-first-ggplo-7, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r creating-first-ggplo-7-hint-1, eval = FALSE}
penguins |>
  ggplot(
  mapping = aes(x = flipper_length_mm, y = ...)
)+
  geom_point()
```

### 

`geom_point()` is not the only geometric objects, there are more than this single one and we will learn about them in other sections in this tutorial.

### Exercise 8

Scatter plot are useful for displaying relationship between x and y but it is also helpful to ask yourself if there are other variable which might contribute to the relationship. For example, does the relationship between x and y differ for different species of penguins. To find the relationship, lets add more the aesthetic `color` within `aes()` to `species`.
```{r creating-first-ggplo-8, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r creating-first-ggplo-8-hint-1, eval = FALSE}
penguins |>
  ggplot(
  mapping = aes(x = flipper_length_mm, y = ..., ... = species)
)+
  geom_point()
```

### 

When you ran it, you see that the data points differ by the color of the species which just makes the scatter plot more insightful. Color inst just the only aesthetics. There are more, to find out more go check out [aesthetic mappings](https://ggplot2.tidyverse.org/reference/aes.html)


### Exercise 9

Now let's add one more layer: a smooth curve which will dispaly the reltionship between body mass and flipper length. To do so, lets add a new geometric object:`geom_smooth()` and within that we will specity it to be a best of fit line of a linear model.
```{r creating-first-ggplo-9, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r creating-first-ggplo-9-hint-1, eval = FALSE}
penguins |>
  ggplot(
  mapping = aes(x = flipper_length_mm, y = body_mass_g, color = ...)
) +
  geom_point() +
  geom_smooth(... = "lm")
```

### 

If you don't specify the model for geom_smooth(), RStudio will automatically set the method to a linear model, adding on you will get a message saying that it was set to `lm`.

### Exercise 10

We have added new lines to the plot, but it does not match the plot in beginning, which displayed only one line for the entire dataset instead of separate lines for each penguin species.

In ggplot2, when aesthetic mappings are defined at the global level, they are passed down to all subsequent geom layers in the plot. However, each geom function in ggplot2 can also accept a mapping argument, allowing for local-level aesthetic mappings that are added to those inherited from the global level. To achieve our desired outcome of coloring the points based on the species without separating the lines for each species, we should specify color = species for geom_point() only.
```{r creating-first-ggplo-10, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r creating-first-ggplo-10-hint-1, eval = FALSE}
penguins |>
  ggplot(
  mapping = ...(x = flipper_length_mm, y = body_mass_g)
) +
  geom_point(mapping = ...(color = species)) +
  ...(method = "lm")
```

### 

Great progress! The plot now resembles our ultimate goal, although it's not perfect yet. We still need to differentiate the species of penguins by using different shapes for each species, and also improve the labels.

### Exercise 11

Representing information using only colors on a plot may not be a good idea as individuals perceive colors differently due to color blindness or other color vision differences. Therefore, it's advisable to map the `species` to the `shape` aesthetic in addition to `color` to ensure everyone can differentiate the species.
```{r creating-first-ggplo-11, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r creating-first-ggplo-11-hint-1, eval = FALSE}
penguins |>
  ggplot(
  mapping = aes(x = flipper_length_mm, y = body_mass_g)
) +
  geom_point(mapping = aes(color = ..., shape = ...)) +
  geom_smooth(method = "lm")
```

### 



### Exercise 12
Now that we have the data points, legend, and lines set, lets change the labels for the graph and make it more insightful and visually appealing. Copy the previous code and add a new layear using `+` and type `labs()`. Within labs, type `title` and set it equal to "Body mass and flipper length" in quotes.Let's also add the subtitle by typing a comma and then 'subtitle'. Set  `Subtitle` to "Dimensions for Adelie, Chinstrap, and Gentoo Penguins". Run the code and see what you get.

```{r creating-first-ggplo-12, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r creating-first-ggplo-12-hint, eval = FALSE}
penguins |>
  ggplot(
  mapping = aes(x = flipper_length_mm, y = body_mass_g)
) +
  geom_point(mapping = aes(color = species, shape = species)) +
  geom_smooth(method = "lm")+
  labs( ... = "Body mass and flipper length"+
          subtitle = "Dimensions for Adelie, Chinstrap, and Gentoo Penguins")
```

###
### Exercise 13

We see that the title and subtitle have been added, now lets modify the x and y axises, type a comma and then type `x`. Then set `x` to "Flipper Length (mm)" in quotes. Type another comma and a `y`, then set it equal to "Body mass (g)" in quotes.
```{r creating-first-ggplo-13, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r creating-first-ggplo-13-hint, eval = FALSE}
penguins |>
  ggplot(
  mapping = aes(x = flipper_length_mm, y = body_mass_g)
) +
  geom_point(mapping = aes(color = species, shape = species)) +
  geom_smooth(method = "lm")+
  labs( ... = "Body mass and flipper length"+
          subtitle = "Dimensions for Adelie, Chinstrap, and Gentoo Penguins",
        x = ..., y = ...)
```

###
### Exercise 14

You might see that it looks the same as the one in the beginning but we forgot one minor thing which is capitalizing the legend, we can do so by typing a comma and setting both `color` and `shape` equal to "Species" in quotes.
```{r creating-first-ggplo-14, exercise = TRUE}
penguins |>
  ggplot(
  mapping = aes(x = flipper_length_mm, y = body_mass_g)
) +
  geom_point(mapping = aes(color = species, shape = species)) +
  geom_smooth(method = "lm")+
  labs( ... = "Body mass and flipper length"+
          subtitle = "Dimensions for Adelie, Chinstrap, and Gentoo Penguins",
        x = ..., y = ...,
         color = "Species", shape = "Species")
```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r creating-first-ggplo-14-hint, eval = FALSE}

```

###


## Exercise
###



## Summary
### 

This tutorial covered [Chapter 2: Data Visualization](https://r4ds.hadley.nz/data-visualize.html) from [*R for Data Science (2e)*](https://r4ds.hadley.nz/) by Hadley Wickham, Mine Çetinkaya-Rundel, and Garrett Grolemund. 


```{r download-answers, child = system.file("child_documents/download_answers.Rmd", package = "tutorial.helpers")}
```
