---
title: Data visualization
author: Anish Bellamkonda
tutorial:
  id: data-visualization
output:
  learnr::tutorial:
    progressive: yes
    allow_skip: yes
runtime: shiny_prerendered
description: 'Tutorial for Chapter 2: Data Visualization'
---

```{r setup, include = FALSE}
library(learnr)
library(tutorial.helpers)
library(ggplot2)
library(tidyverse)
library(palmerpenguins)
library(ggthemes)
knitr::opts_chunk$set(echo = FALSE)
options(tutorial.exercise.timelimit = 60, 
        tutorial.storage = "local") 
```

```{r copy-code-chunk, child = system.file("child_documents/copy_button.Rmd", package = "tutorial.helpers")}
```

```{r info-section, child = system.file("child_documents/info_section.Rmd", package = "tutorial.helpers")}
```

## Introduction
### 

This tutorial covers [Chapter 2: Data Visualization](https://r4ds.hadley.nz/data-visualize.html) from [*R for Data Science (2e)*](https://r4ds.hadley.nz/) by Hadley Wickham, Mine Çetinkaya-Rundel, and Garrett Grolemund. 

## Creating first ggplot
### 

<!-- AB: Need to chnage the code so eveyrhting has a pipe -->

Our Ultimate Goal in this section is to create a plot displaying the relationship between flipper lengths and body masses of these penguins, taking into consideration the species of the penguin.Before doing so, lets load the palmerpenguins package using `library()`.

### Exercise 1

```{r creating-first-ggplo-1, exercise = TRUE}

```

```{r creating-first-ggplo-1-hint-1, eval = FALSE}
library(palmerpenguins)
```

### 

A `variable`: a quanitty, quality, or property that you can measure, `value`: state of a variable when you measure it, `observation`: a set of measurements made under similiar conditions.

### Exercise 2

Run `?palmerpenguins`. After doing so, copy and paste the description here.

```{r creating-first-ggplo-2}
question_text(NULL,
	answer(NULL, correct = TRUE),
	allow_retry = TRUE,
	try_again_button = "Edit Answer",
	incorrect = NULL,
	rows = 3)
```

### 

The data include size measurements, clutch observations, and blood isotope ratios for adult foraging Adélie, Chinstrap, and Gentoo penguins observed on islands in the Palmer Archipelago near Palmer Station, Antarctica. 

### Exercise 3

Run the `library()` command with `ggthemes` as the argument.

```{r creating-first-ggplo-3, exercise = TRUE}

```

```{r creating-first-ggplo-3-hint-1, eval = FALSE}
library(ggthemes)
```

### 

We can infer from the name of the package itself that `ggthemes` is used for aesthetic looks of a graph and make it fancy.



### Exercise 4

Type `penguins` and hit Run Code.


```{r creating-first-ggplo-4, exercise = TRUE}

```

```{r creating-first-ggplo-4-hint-1, eval = FALSE}
penguins
```

### 

Tabular data is a set of values, each associated with a variable and an observation. Tabular data is tidy if each value is placed in its own “cell”, each variable in its own column, and each observation in its own row. `penguins` is tabular data.

### Exercise 5

Run `glimpse()` with `penguins` as its argument.

```{r creating-first-ggplo-5, exercise = TRUE}

```

```{r creating-first-ggplo-5-hint-1, eval = FALSE}
glimpse(...)
```

### 

In RStudio, you can click on the name of the data frame in the Environment pane or run `View(penguins)` to open an interactive data viewer.

### Exercise 6

<!-- AB: teach what a pipe is  -->

The key command for making plots is `ggplot()` from the **ggplot2** package. Run `ggplot(data = penguins)`.

```{r creating-first-ggplo-6, exercise = TRUE}

```

```{r creating-first-ggplo-6-hint-1, eval = FALSE}
ggplot(data = penguins)
```

### 

There is another way to create a plot without using `|>` which is to use the data arguemnt within the `ggplot()` like `ggplot(data = penguins)`.


### Exercise 7

Run `sqrt(16)`

```{r creating-first-ggplo-7, exercise = TRUE}

```

```{r creating-first-ggplo-7-hint-1, eval = FALSE}

```

### 

### Exercise 8

Run `16 |> sqrt()`
```{r creating-first-ggplo-8, exercise = TRUE}

```

```{r creating-first-ggplo-8-hint-1, eval = FALSE}

```

### 

The difference between the previous code was it that we are using the pipe operator `|>` which takes anything from the left side and plugs it into the function in the right side.

### Exercise 9

Run `penguins` with the pipe operator `|>` and plug it into `ggplot()` like this
`penguins |> ggplot()`
```{r creating-first-ggplo-9, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r creating-first-ggplo-9-hint-1, eval = FALSE}
penguins |>
  ggplot()
```

### 

penguins which is the input is being plugged into the function `ggplot()` which gives the same output as `ggplot(data = penguins)`. While Indentation may not affect code, it is a good habit to do because it helps go through the code much faster.




### Exercise 10

Copy the previous code and within `ggplot()`, run `mapping = aes(x = flipper_length_mm)` and see what you get.
```{r creating-first-ggplo-10, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r creating-first-ggplo-10-hint-1, eval = FALSE}
penguins |>
  ggplot(... = aes(x = ...))
```

### 

`mapping = aes(x = flipper_length_mm)` specifies that the variable `flipper_length_mm` is mapped to the x-axis of a plot. This allows the length of the penguin's flipper to be plotted on the x-axis.

### Exercise 11

Copy the previous code, within the `aes()` set `y` equal to `body_mass_g` and run it.
```{r creating-first-ggplo-11, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r creating-first-ggplo-11-hint-1, eval = FALSE}
penguins |>
  ggplot(mapping = aes(x = ..., y = ...))
```

To display the data points we need to add a geometric object(geom). We can define a `geom`: the geometrical object that a plot uses to represent data. These geometric objects are made available in ggplot2 with functions that start with `geom_`. 

### 

### Exercise 12

The function `geom_point()` adds a layer of points to your plot, which creates a scatterplot so copy the previous code and run `+ geom_point()`

```{r creating-first-ggplo-12, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r creating-first-ggplo-12-hint-1, eval = FALSE}
penguins |>
  ...(
    mapping = aes(x = flipper_length_mm, y = ...)
    ) +
  geom_point()
```

### 

`geom_point()` is not the only geometric objects, there are more than this single one and we will learn about them in other sections in this tutorial.

### Exercise 13

Within `aes()`, run `color = species`.
```{r creating-first-ggplo-13, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r creating-first-ggplo-13-hint-1, eval = FALSE}
penguins |>
  ggplot(
  mapping = aes(x = flipper_length_mm, y = ..., ... = species)
)+
  geom_point()
```

### 

Scatter plot are useful for displaying relationship between x and y but it is also helpful to ask yourself if there are other variable which might contribute to the relationship. For example, does the relationship between x and y differ for different species of penguins.

When you ran it, you see that the data points differ by the color of the species which just makes the scatter plot more insightful. Color isnt just the only aesthetics. There are more, to find out more go check out [aesthetic mappings](https://ggplot2.tidyverse.org/reference/aes.html)


### Exercise 14

Copy the previous code, run `geom_smooth()` and within it, set `method` to `lm`. 
```{r creating-first-ggplo-14, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r creating-first-ggplo-14-hint-1, eval = FALSE}
penguins |>
  ggplot(
  mapping = aes(x = flipper_length_mm, y = body_mass_g, color = ...)
) +
  geom_point() +
  geom_smooth(... = "lm")
```

### 

`geom_smooth()`: used to add a smooth curve to a scatterplot, which can help identify trends and patterns in data by fitting a line or smoothing function. It offers different smoothing methods like linear or polynomial regression and loess smoothing


### Exercise 15

We have added new lines to the plot, but it does not match the plot in beginning, which displayed only one line for the entire dataset instead of separate lines for each penguin species. Remove the `color` aesthetic within `ggplot()` and add it to `geom_point` by running `geom_point(mapping = aes(color = species))`


```{r creating-first-ggplo-15, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r creating-first-ggplo-15-hint-1, eval = FALSE}
penguins |>
  ggplot(
  mapping = ...(x = flipper_length_mm, y = body_mass_g)
) +
  geom_point(mapping = ...(color = species)) +
  ...(method = "lm")
```

### 

In **ggplot2**, when aesthetic mappings are defined at the global level, they are passed down to all subsequent geom layers in the plot. However, each geom function in ggplot2 can also accept a mapping argument, allowing for local-level aesthetic mappings that are added to those inherited from the global level. 

### Exercise 16

Not all individuals perceive colors differently due to color blindness or other color vision differences. Therefore,  map the `species` to the `shape` aesthetic in addition to `color` within `geom_point()`.
`
```{r creating-first-ggplo-16, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r creating-first-ggplo-16-hint-1, eval = FALSE}
penguins |>
  ggplot(
  mapping = aes(x = flipper_length_mm, y = body_mass_g)
) +
  geom_point(mapping = aes(color = ..., shape = ...)) +
  geom_smooth(method = "lm")
```

### 

### Exercise 17

Now that we have the data points, lets add the title, substitle, labels and etc. Copy the previous code,   add a new layer using `+` and run `labs()`. Within labs, set `title` equal to "Body mass and flipper length" in quotes.

```{r creating-first-ggplo-17, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r creating-first-ggplo-17-hint-1, eval = FALSE}
penguins |>
  ggplot(
  mapping = aes(x = flipper_length_mm, y = body_mass_g)
) +
  geom_point(mapping = aes(color = species, shape = species)) +
  geom_smooth(method = "lm")+
  labs( ... = "Body mass and flipper length")
```

### 

### Exercise 18

Copy the previous code, lets add the subtitle by running `subtitle = "Dimensions for Adelie, Chinstrap, and Gentoo Penguins"` after the title.

```{r creating-first-ggplo-18, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r creating-first-ggplo-18-hint-1, eval = FALSE}
penguins |>
  ggplot(
  mapping = aes(x = flipper_length_mm, y = body_mass_g)
) +
  geom_point(mapping = aes(color = species, shape = species)) +
  geom_smooth(method = "lm")+
  labs( ... = "Body mass and flipper length",
        subtitle = "...")
```

### 

### Exercise 19

Lets modify the x and y axises, type a comma and then run `x` and set it to `"Flipper Length (mm)"`  Type another comma and a `y`, then set it equal to "Body mass (g)" in quotes.

```{r creating-first-ggplo-19, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r creating-first-ggplo-19-hint-1, eval = FALSE}
penguins |>
  ggplot(
  mapping = aes(x = flipper_length_mm, y = body_mass_g)
) +
  geom_point(mapping = aes(color = species, shape = species)) +
  geom_smooth(method = "lm")+
  labs( ... = "Body mass and flipper length"+
          subtitle = "Dimensions for Adelie, Chinstrap, and Gentoo Penguins",
        x = ...)
```

### 

### Exercise 20

Next to `x` axis label, type another comma and a `y`, then set it equal to `"Body mass (g)"` in quotes.

```{r creating-first-ggplo-20, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r creating-first-ggplo-20-hint-1, eval = FALSE}
penguins |>
  ggplot(
  mapping = aes(x = flipper_length_mm, y = body_mass_g)
) +
  geom_point(mapping = aes(color = species, shape = species)) +
  geom_smooth(method = "lm")+
  labs( ... = "Body mass and flipper length"+
          subtitle = "Dimensions for Adelie, Chinstrap, and Gentoo Penguins",
        x = ..., y = ...)
```

### 

### Exercise 21

You might see that it looks the same as the one in the beginning but we forgot one minor thing which is capitalizing the legend, we can do so by typing a comma and setting both `color` and `shape` equal to "Species" in quotes.
```{r creating-first-ggplo-21, exercise = TRUE}
penguins |>
  ggplot(
  mapping = aes(x = flipper_length_mm, y = body_mass_g)
) +
  geom_point(mapping = aes(color = species, shape = species)) +
  geom_smooth(method = "lm")+
  labs( ... = "Body mass and flipper length"+
          subtitle = "Dimensions for Adelie, Chinstrap, and Gentoo Penguins",
        x = ..., y = ...,
         color = "Species", shape = "Species")
```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r creating-first-ggplo-21-hint-1, eval = FALSE}

```

### 

## Exercise
### 

To practice the new skills we learned from the plot we made in the previous exercises, we will make another plot which displays the relationship between `bill_depth_mm` and `bill_lenght_mm`.

### Exercise 1


Run `penguins` with the pipe operator `|>` and plug it into `ggplot()` 
```{r exercise-1, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r exercise-1-hint-1, eval = FALSE}
penguins |>
  ggplot()
```

### 






### Exercise 2

Copy the previous code and within `ggplot()`, set `x` to `bill_length_mm`.
```{r exercise-2, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r exercise-2-hint-1, eval = FALSE}
penguins |>
  ggplot(... = aes(x = ...))
```

### 



### Exercise 3

Copy the previous code, within the `aes()`, set `y` to `bill_depth_mm` and run it.
```{r exercise-3, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r exercise-3-hint-1, eval = FALSE}
penguins |>
  ggplot(mapping = aes(x = ..., y = ...))
```



### 

### Exercise 4

Copy the previous code and run `+ geom_point()`

```{r exercise-4, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r exercise-4-hint-1, eval = FALSE}
penguins |>
  ...(
    mapping = aes(x = bill_length_mm, y = ...)
    ) +
  geom_point()
```

### 



### Exercise 5

Within `aes()`, run `color = species`.
```{r exercise-5, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r exercise-5-hint-1, eval = FALSE}
penguins |>
  ggplot(
  mapping = aes(x = bill_length_mm, y = ..., ... = species)
)+
  geom_point()
```

### 




### Exercise 6

Copy the previous code, run `geom_smooth()` and within it, set `method` to `lm`. 
```{r exercise-6, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r exercise-6-hint-1, eval = FALSE}
penguins |>
  ggplot(
  mapping = aes(x = bill_length_mm, y = bill_depth_mm, color = ...)
) +
  geom_point() +
  geom_smooth(... = "lm")
```

### 




### Exercise 7

Remove the `color` aesthetic within `ggplot()` and add it to `geom_point` by running `geom_point(mapping = aes(color = species))` to get three different lines for each species.


```{r exercise-7, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r exercise-7-hint-1, eval = FALSE}
penguins |>
  ggplot(
  mapping = ...(x = bill_length_mm, y = bill_depth_mm)
) +
  geom_point(mapping = ...(color = species)) +
  ...(method = "lm")
```

### 


### Exercise 8

Map the `species` variable to the `shape` aesthetic in addition to `color` within `geom_point()`.
`
```{r exercise-8, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r exercise-8-hint-1, eval = FALSE}
penguins |>
  ggplot(
  mapping = aes(x = bill_length_mm, y = bill_depth_mm)
) +
  geom_point(mapping = aes(color = ..., shape = ...)) +
  geom_smooth(method = "lm")
```

### 

### Exercise 9

Now that we have the data points, lets add the title, subtitle, labels and etc. Copy the previous code,   add a new layer using `+` and run `labs()`. Within labs, set `title` equal to "Penguins Beak's Length and Depth" , `subtitle` to "Dimensions for Adelie, Chinstrap, and Gentoo Penguins" , `x` to  "Beak Length (mm)", `y` = "Beak Depth (mm)", `color` to "Species" and `shape` to "Species".

```{r exercise-9, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r exercise-9-hint-1, eval = FALSE}
penguins |>
  ggplot(mapping = aes(x = ..., y = bill_depth_mm, color = ... , shape = species) )+
  geom_point() +
  geom_smooth(method = "lm")+
  labs( title = "Penguins Beak's Length and Depth", subtitle = ...,
        x = ... , ... = "Beak Depth (mm)",color = "Species", ... = "Species")
```

Reminder: This is what our graph should look like

```{r}
#| message: false
#| warning: false
plot1 <- penguins |>
  ggplot(mapping = aes(x = bill_length_mm, y = bill_depth_mm, color = species, shape = species) )+
  geom_point() +
  geom_smooth(method = "lm")+
  labs( title = "Penguins Beak's Length and Depth", subtitle = "Dimensions for Adelie, Chinstrap, and Gentoo Penguins",
        x = "Beak Length (mm)", y = "Beak Depth (mm)",color = "Species", shape = "Species")
plot1
```


### 



## geom_bar
### 

## geom_histogram
### 

## geom_density
### 

## geom_boxplot
### 

## Summary
### 

This tutorial covered [Chapter 2: Data Visualization](https://r4ds.hadley.nz/data-visualize.html) from [*R for Data Science (2e)*](https://r4ds.hadley.nz/) by Hadley Wickham, Mine Çetinkaya-Rundel, and Garrett Grolemund. 


```{r download-answers, child = system.file("child_documents/download_answers.Rmd", package = "tutorial.helpers")}
```
