---
title: Data visualization
tutorial:
  id: data-visualization
output:
  learnr::tutorial:
    progressive: yes
    allow_skip: yes
runtime: shiny_prerendered
description: 'Tutorial for Chapter 2: Data Visualization'
---

```{r setup, include = FALSE}
library(learnr)
library(ggplot2)
library(tidyverse)
library(palmerpenguins)
library(ggthemes)
knitr::opts_chunk$set(echo = FALSE)
options(tutorial.exercise.timelimit = 60, 
        tutorial.storage = "local") 
```

```{r copy-code-chunk, child = system.file("child_documents/copy_button.Rmd", package = "tutorial.helpers")}
```

```{r info-section, child = system.file("child_documents/info_section.Rmd", package = "tutorial.helpers")}
```

## Introduction
### 

This tutorial covers [Chapter 2: Data Visualization](https://r4ds.hadley.nz/data-visualize.html) from [*R for Data Science (2e)*](https://r4ds.hadley.nz/) by Hadley Wickham, Mine Çetinkaya-Rundel, and Garrett Grolemund. 

## Creating first ggplot
### 

<!-- AB: Need to chnage the code so eveyrhting has a pipe -->

Our Ultimate Goal in this section is to create a plot displaying the relationship between flipper lengths and body masses of these penguins, taking into consideration the species of the penguin.Before doing so, lets load the palmerpenguins package using `library()`.
### Exercise 1

```{r creating-first-ggplo-1, exercise = TRUE}

```

```{r creating-first-ggplo-1-hint-1, eval = FALSE}
library(palmerpenguins)
```


### 

To make this turtorial easier for you to comprehend, let's define a `variable`: a quanitty, quality, or property that you can measure, `value`: state of a variable when you measure it, `observation`: a set of measurements made under similiar conditions.

### Exercise 2

Before creating our first plot, we should learn a little more about the data set we will be using so going to help page of the palmerpenguins by typing `?palmerpenguins`. After doing so, copy and paste the description here.
```{r creating-first-ggplo-2}
question_text(NULL,
	answer(NULL, correct = TRUE),
	allow_retry = TRUE,
	try_again_button = "Edit Answer",
	incorrect = NULL,
	rows = 3)
```


### 
In this pacakge, we have the `penguins` dataset which contains data about different species(Adelie, Chinstrap, or Gentoo), flipper_length, and mass of a penguin.

### Exercise 3

Let's load the next library for this plot which is the `ggthemes` with `library()`
```{r creating-first-ggplo-3, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r creating-first-ggplo-3-hint-1, eval = FALSE}
library(ggthemes)
```

### 

We can infer from the name of the package itself that `ggthemes` is used for aesthetic looks of a graph and make it fancy.



### Exercise 4

Before diving into making our first plot, let's have a quick look at the dataset which we would be using to make a plot. Just type in the name of the data set whihc is palmerpenguins and see what you get.
```{r creating-first-ggplo-4, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r creating-first-ggplo-4-hint-1, eval = FALSE}
penguins
```

### 

To view the data set in alternative way is to use `glimpse(penguins)` which shows all the variables compared to just printing the data set.

### Exercise 5

To create our first ever plot, we need to type out of `penguins`, then give a space and type in `|>` which is called creating a pipe and what it does is chaining multiple fucntions together. Then click enter, now the starting poitn must be indented, then type `ggplot()` which will create a plot.
```{r creating-first-ggplo-5, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r creating-first-ggplo-5-hint-1, eval = FALSE}
penguins |>
  ggplot()
```

There is another way to create a plot without using `|>` which is to use the data arguemnt within the `ggplot()` like `ggplot(data = penguins)`.

### 

### Exercise 6

When we ran the code in the last exercise, we see the outline of the plot but no axis or data point. So lets tell `ggplot()` how the varibales from out dataset are map to aesthetics of our plot. So set `mapping` to  `aes` and within it set "x" to the flipper_length_mm, and y to body_mass_g. 
```{r creating-first-ggplo-6, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r creating-first-ggplo-6-hint-1, eval = FALSE}
penguins |>
  ggplot(
  mapping = aes(x = flipper_length_mm, y = ...)
)
```

### 

### Exercise 7

After running the code in previous code, we see that the plot has more structure but the data point yhemselves are not yet on the plot. To represent the obeservarions from out data set on out plot, we need to use a geometric object  and for this we will use point object. To do, add a new layer using `+` after the ggplot function and add `geom_pont()` wich will add the points.
```{r creating-first-ggplo-7, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r creating-first-ggplo-7-hint-1, eval = FALSE}
penguins |>
  ggplot(
  mapping = aes(x = flipper_length_mm, y = ...)
)+
  geom_point()
```

`geom_point()` is not the only geometric objects, there are more than this single one and we will learn about them in other sections in this tutorial.

### 

### Exercise 8

Scatter plot are useful for displaying relationship between x and y but it is also helpful to ask yourself if there are other variable which might contribute to the relationship. For example, does the relationship between x and y differ for different species of penguins. To find the relationship, lets add more the aesthetic `color` within `aes()` to `species`.
```{r creating-first-ggplo-8, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r creating-first-ggplo-8-hint-1, eval = FALSE}
penguins |>
  ggplot(
  mapping = aes(x = flipper_length_mm, y = ..., ... = species)
)+
  geom_point()
```

When you ran it, you see that the data points differ by the color of the species which just makes the scatter plot more insightful. Color inst just the only aesthetics. There are more, to find out more go check out [aesthetic mappings](https://ggplot2.tidyverse.org/reference/aes.html)


### 

### Exercise 9

Now let's add one more layer: a smooth curve which will dispaly the reltionship between body mass and flipper length. To do so, lets add a new geometric object:`geom_smooth()` and within that we will specity it to be a best of fit line of a linear model.
```{r creating-first-ggplo-9, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r creating-first-ggplo-9-hint-1, eval = FALSE}
penguins |>
  ggplot(
  mapping = aes(x = flipper_length_mm, y = body_mass_g, color = ...)
) +
  geom_point() +
  geom_smooth(... = "lm")
```

### 

### Exercise 10

```{r creating-first-ggplo-10, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r creating-first-ggplo-10-hint-1, eval = FALSE}
penguins |>
  ggplot(
  mapping = ...(x = flipper_length_mm, y = body_mass_g)
) +
  geom_point(mapping = ...(color = species)) +
  ...(method = "lm")
```

### 

### Exercise 11

```{r creating-first-ggplo-11, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r creating-first-ggplo-11-hint-1, eval = FALSE}
penguins |>
  ggplot(
  mapping = aes(x = flipper_length_mm, y = body_mass_g)
) +
  geom_point(mapping = aes(color = ..., shape = ...)) +
  geom_smooth(method = "lm")
```

### 

## Exericse
### 

## Summary
### 

This tutorial covered [Chapter 2: Data Visualization](https://r4ds.hadley.nz/data-visualize.html) from [*R for Data Science (2e)*](https://r4ds.hadley.nz/) by Hadley Wickham, Mine Çetinkaya-Rundel, and Garrett Grolemund. 


```{r download-answers, child = system.file("child_documents/download_answers.Rmd", package = "tutorial.helpers")}
```
