---
title: A field guide to base R
author: David Kane and Aryan Kancherla
tutorial:
  id: a-field-guide-to-base-R
output:
  learnr::tutorial:
    progressive: yes
    allow_skip: yes
runtime: shiny_prerendered
description: 'Tutorial for Chapter 28: A field guide to base R'
---

```{r setup, include = FALSE}
library(learnr)
library(tutorial.helpers)
library(tidyverse)

knitr::opts_chunk$set(echo = FALSE)
options(tutorial.exercise.timelimit = 60, 
        tutorial.storage = "local") 


x <- c("one", "two", "three", "four", "five")

x2 <- c(10, 3, NA, 5, 8, 1, NA)

x3 <- c(abc = 1, def = 2, xyz = 5)

df <- tibble(
  x = 1:3, 
  y = c("a", "e", "f"), 
  z = runif(3)
)

df0 <- tibble(
  x = c(2, 3, 1, 1, NA), 
  y = letters[1:5], 
  z = runif(5)
) 

df2 <- data.frame(x1 = 1)

tb <- tibble(
  x = 1:4,
  y = c(10, 4, 1, 21)
)

tb2 <- tibble(x1 = 1)


```

```{r copy-code-chunk, child = system.file("child_documents/copy_button.Rmd", package = "tutorial.helpers")}
```

```{r info-section, child = system.file("child_documents/info_section.Rmd", package = "tutorial.helpers")}
```

## Introduction
### 

This tutorial covers [Chapter 28: A field guide to base R](https://r4ds.hadley.nz/base-r) from [*R for Data Science (2e)*](https://r4ds.hadley.nz/) by Hadley Wickham, Mine Çetinkaya-Rundel, and Garrett Grolemund. 

In this chapter, we’ll focus on four big topics: subsetting with [, subsetting with [[ and $, the apply family of functions, and for loops. To finish off, we’ll briefly discuss two essential plotting functions. These topics are particularly useful as you do more programming and will help you read code you’ll encounter in the wild.

## Subsetting with the single bracket operator
### 

The `[` bracket can be used to extract sub-components from data frames and vectors, with the syntax `x[i]`. `x` represents the vector and `i` represents the position of the value inside of `x` (1st element is position 1, second element is position 2, and so forth). 

### Exercise 1

Load the tidyverse library

```{r subsetting-with-the--1, exercise = TRUE}

```

### 

There are five main types of things that you can subset a vector with, which will be covered in the following exercises:

1) A vector of positive integers
2) A vector of negative integers
3) A logical vector
4) A character vector
5) Nothing

### Exercise 2

Press "Run Code".

```{r subsetting-with-the--2, exercise = TRUE}
x <- c("one", "two", "three", "four", "five")
```

### 

This code sets the variable `x` to the vector `c("one", "two", "three", "four", "five")`. 

### Exercise 3

Extract the first element of `x` by typing `x[]` with the number `1` inside the brackets

```{r subsetting-with-the--3, exercise = TRUE}

```

```{r subsetting-with-the--3-hint-1, eval = FALSE}
x[...]
```

### 

As you can see, the code extracts the first value of the `x` vector, printing out the value "one". 

### Exercise 4

We can also pass in a vector to `[]`, containing various positions to extract from `x`. In the code chunk below, extract the 2nd, 3rd, and 4th values of `x`, using `c()`

```{r subsetting-with-the--4, exercise = TRUE}

```

```{r subsetting-with-the--4-hint-1, eval = FALSE}
x[c(..., ..., ...)]
```

### 

When you run the code above, you should see that it extracts "two", "three", and "four". By using vectors inside `[]`, you can extract multiple elements at once.

### Exercise 5

You can also pass in a vector of negative values. In the code chunk below, type `x[]`, placing the vector `c(-1, -2)` inside the brackets. Oberseve what happens.

```{r subsetting-with-the--5, exercise = TRUE}

```

```{r subsetting-with-the--5-hint-1, eval = FALSE}
x[c(..., ...)]
```

### 

Negative values drop the elements at the specified positions; the code above drops the first and second elements, returning "three", "four", and "five".

### Exercise 6

Subsetting with logical vectors are another type of thing that you can subset a vector with. Create a vector with the values `10`, `3`, `NA`, `5`, `8`, `1`, and `NA`. Save this to a variable named `x2`.

```{r subsetting-with-the--6, exercise = TRUE}

```

```{r subsetting-with-the--6-hint-1, eval = FALSE}
... <- c(10, 3, NA, 5, 8, 1, NA)
```

### 

Subsetting with a logical vector keeps all values corresponding to a TRUE value. This is most often useful in conjunction with the comparison functions.

### Exercise 7

`is.na()` is a function that identifies missing values in vectors, data frames, etc. In the code chunk below, type `x2[]`, placing `is.na(x2)` inside of the brackets.


```{r subsetting-with-the--7, exercise = TRUE}

```

```{r subsetting-with-the--7-hint-1, eval = FALSE}
x2[is.na(x2)]
```

### 

This code prints out the missing values (the `NA`'s) stored inside `x2`. Unlike `filter()`, NA indicies will be included in the output as `NA`'s.

### Exercise 8

Let's extract all the even values of `x2`. In the code chunk below, type `x2[]` and extract the even numbers using the `%%` operator.


```{r subsetting-with-the--8, exercise = TRUE}

```

```{r subsetting-with-the--8-hint-1, eval = FALSE}
x2[x2 %% 2 == 0]
```

### 

The `%%` operator is used to calculate the remainder of the division of two numbers. So, by placing `x2 %% 2 == 0` inside the brackets, the code will search through each element of `x2` and return all of the numbers with a remainder of 0 when divided by 2 (thus being an even number). And as mentioned previously, all `NA` indicies will be included in the output as `NA`

### Exercise 9

The last kind of vector that you can subset is a character vector. Using the following vector below, lets extract the `xyz` element. On a new line, type `x3` followed with a pair of brackets, passing in the string `"xyz"`

```{r subsetting-with-the--9, exercise = TRUE}
x3 <- c(abc = 1, def = 2, xyz = 5)
```

```{r subsetting-with-the--9-hint-1, eval = FALSE}
x3["..."]
```

### 

Along with sub-setting logical, character, and integer vectors, you can also subset nothing. For example, by running `x[]` on the code above, it'll just return the vector `x`.


### Exercise 10

Sub-setting can not only be used on vectors; it works on data sets too. In the code chunk below, create a tibble called `df`. The first argument, `x`, should contain a range of numbers from 1 to 3. The second argument, `y`, should be set to the vector `c("a", "e", "f")`. The third argument, `z`, should be set to `runif(3)`. After completing this, run `df` on a new line.


```{r subsetting-with-the--10, exercise = TRUE}

```

```{r subsetting-with-the--10-hint-1, eval = FALSE}
... <- tibble(x = ... : ...,
              y = c("...", "...", "..."),
              z = runif(...)
              )
```

### 

There are many ways to use `[` with data sets, but the most common way is to subset by selecting rows and columns, with the syntax `df[row, col]`.


### Exercise 11

Using `df`,  let's extract the letter "a". On a new line, type `df` followed by a pair of brackets. In the brackets, type `1,2`. 

```{r subsetting-with-the--11, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r subsetting-with-the--11-hint-1, eval = FALSE}
df[... , ...]
```

### 

This extracts the element in the 1st row, 2nd column, which is the letter "a". 

### Exercise 12

You can also leave the row/column input blank when extracting data from a data set. For example, `df[rows, ]` returns the specified row(s) and all columns in the data set, while `df[, cols]` returns all rows and the specified column(s) in the data set. 

In the code chunk below, extract all the rows in `df`, as well as the columns `x` and `z`.

```{r subsetting-with-the--12, exercise = TRUE}

```

```{r subsetting-with-the--12-hint-1, eval = FALSE}
df[, c("...", "...")]
```

### 

As you can see, by leaving the first part of the subset blank (the part before the comma), the code returns all of the rows in `df`, but only returns columns `x` and `z` due to the vector inputted after the comma.

### Exercise 13

There’s an important difference between tibbles and data frames when it comes to `[` . In the tutorials, we’ve mainly used tibbles, which are data frames, but tibbles tweak some behaviors to make your life a little easier. In most places, you can use “tibble” and “data frame” interchangeably, so when we want to draw particular attention to R’s built-in data frame, we’ll write data.frame. 

Press "Run Code".


```{r subsetting-with-the--13, exercise = TRUE}
df1 <- data.frame(x = 1:3)
df2 <- tibble(x = 1:3)

df1
df2
```

### 

If `df` is a data.frame, then `df[, cols]` will return a vector if `col` selects a single column and a data frame if it selects more than one column. If `df` is a tibble, then `[` will always return a `tibble.

### Exercise 14

Press "Run code".

```{r subsetting-with-the--14, exercise = TRUE}
df0 <- tibble(
  x = c(2, 3, 1, 1, NA), 
  y = letters[1:5], 
  z = runif(5)
)

df0
```

### 

### Exercise 15

The `dplyr` package contains many equivalents of subsetting, such as the `filter()`, `arrange()`, and`select()` functions.


Let's observe the `filter()` equivalent. Pipe `df0` (with the `|>`) to the filter function. Pass `x > 1` into `filter()`.


```{r subsetting-with-the--15, exercise = TRUE}

```

```{r subsetting-with-the--15-hint-1, eval = FALSE}
... |> 
  ...(x > ...)
```

### 

`filter()` is equivalent to subsetting the rows with a logical vector, taking care to exclude any missing values. For this scenario, the equivalent subset code would be `df0[!is.na(df0$x) & df0$x > 1, ]`. Running this code produces the same result.

### 

The `$` symbol is used to pull out columns from data frames.

### Exercise 16

`arrange()` is equivalent to subsetting the rows with an integer vector, usually created with order(). In the code chunk below, pipe `df0` to `arrange()`, passing in `x, y`. Then, on a new line, paste its equivalent: `df0[order(df0$x, df0$y),]`


```{r subsetting-with-the--16, exercise = TRUE}

```

```{r subsetting-with-the--16-hint-1, eval = FALSE}
... |> 
  ...(x, y)

df0[order(df0$x, df0$y), ]
```

### 

You can use `order(decreasing = TRUE)` to sort all columns in descending order or `-rank(col)` to sort columns in decreasing order individually.

### Exercise 17

`select()` is equivalent to subsetting columns with character vector. In the code chunk below, pipe `df0` to `select()`, passing in `x, z` to `select()`. Then, on a new line, paste its equivalent: `df0[, c("x", "z")]`


```{r subsetting-with-the--17, exercise = TRUE}

```

```{r subsetting-with-the--17-hint-1, eval = FALSE}
... |>
  ...(x, z)

df0[, c("x", "z")]
```

### 

`filter()`, `arrange()`, and `select()` are very useful functions that help organize data. You will use these functions quite often when analyzing data. 


## Subsetting with double brackes and a dollar sign
### 

The single bracket operator, `[`, which selects many elements, is paired with the double bracket operator, `[[`, and `$`, which extract a single element. In this section, we’ll show you how to use `[[` and `$` to pull columns out of data frames, discuss a couple more differences between data.frames and tibbles, and emphasize some important differences between `[` and `[[` when used with lists.


### Exercise 1

In the code chunk below, create a tibble. The first value should be `x`, which is equal to a range of integers from 1 to 4. The second value should be `y`, which is a vector containing 10, 4, 1, and 21. Save this tibble to the name `tb` and print it on a new line.

```{r subsetting-with-doub-1, exercise = TRUE}

```

```{r subsetting-with-doub-1-hint-1, eval = FALSE}
... <- tibble(x = .. : .., y = c(..., ..., ..., ...))
```

### 

### Exercise 2

The double bracket operator, `[[`, and dollar sign, `$`, can be used to extract columns out of a data frame. `[[` can access by position or by name, and `$` is specialized for access by name.

In the code chunk below, lets extract the elements of column `x`. In the code chunk below, type in `tb` followed by `[[]]`. Inside the inner bracket, type in `1`.



```{r subsetting-with-doub-2, exercise = TRUE}

```

```{r subsetting-with-doub-2-hint-1, eval = FALSE}
tb[[...]]
```

### 

In this scenario, the `[[]]` are being used to return the values by position. By placing the number `1` inside the brackets, the code returns the values in the first position, which are the range of integers stored in column `x`. 


### Exercise 3

Now, lets extract the same elements by name. Copy the code above, replacing the code inside the inner bracket with `"x"`


```{r subsetting-with-doub-3, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

### 

By entering either the position or the name of the column, you can extract the values it contains. 

### Exercise 4

As mentioned previously, the `$` is specialized for accessing columns by name. In the code chunk below, lets extract the elements inside `y`. Type in the name of the tibble, `tb`, followed by `$` and the name of the column, `y`. 

```{r subsetting-with-doub-4, exercise = TRUE}

```

```{r subsetting-with-doub-4-hint-1, eval = FALSE}
tb...y
```

### 

As you can see, the `$` extracted the values inside column `y`, which are 10, 4, 1, and 21.

### Exercise 5

The `$` can also be used to create new columns, which is the Base R equivalent of the `mutate()` method. In the code chunk below, type in `tb$z`, setting it equal to `tb$x + tb$y` (use `<-` to do this). On a new line, type in `tb`.

```{r subsetting-with-doub-5, exercise = TRUE}

```

```{r subsetting-with-doub-5-hint-1, eval = FALSE}
tb$z <- ... + ...
```

### 

There are several other base R approaches to creating new columns including with `transform()`, `with()`, and `within()`. Hadley Wickham, one of the authors of [*R for Data Science*](https://r4ds.hadley.nz/), collected a [few examples](https://gist.github.com/hadley/1986a273e384fb2d4d752c18ed71bedf).

### Exercise 6

Press "Run code".

```{r subsetting-with-doub-6, exercise = TRUE}
diamonds
```

### 

Using `$` directly is convenient when performing quick summaries. Lets use this to calculate some values of the **diamonds** package. 

### Exercise 7

Lets find out the maximum carat in the **diamonds** package, using the `max()` function. Extract the `carat` column using `$` and place the code inside the `max()` function. 

```{r subsetting-with-doub-7, exercise = TRUE}

```

```{r subsetting-with-doub-7-hint-1, eval = FALSE}
max(...$...)
```

### 

In this scenario, the code above would be the Base R equivalent of the `summarize()` function. 

### Exercise 8

The **dplyr** package also provides a base R equivalent for the double bracket operator, `[[`, and `$` called `pull()`. `pull()` takes either a variable name or variable position and returns just that column.

Let's replicate the code above, using `pull()`. Pipe the **diamonds** package to `pull()`, passing in `carat`. Then, create another pipe to the `max()` function.

```{r subsetting-with-doub-8, exercise = TRUE}

```

```{r subsetting-with-doub-8-hint-1, eval = FALSE}
diamonds |> ...(carat) |> ...()
```

### 

Just like the previous exercise, the code returns the maximum carat, which is `5.01`. 

### Exercise 9

An important difference between tibbles and data frames is that tibbles are much more strict when extracting columns with `$`. 

Press "Run code".

```{r subsetting-with-doub-9, exercise = TRUE}
df2 <- data.frame(x1 = 1)
df2$x
```

### 

Although there is no column named `x` in `df2`, the code is still able to output the values in column `x1`. This is because data frames are able to match the prefix of any variable's name (so-called **partial matching**) without returning an error if the column doesn't exist. 

### Exercise 10

However, tibbles are much stricter: they only ever match variable names exactly and they will generate a warning if the column you are trying to access doesn’t exist.

Press "Run code".

```{r subsetting-with-doub-10, exercise = TRUE}
tb2 <- tibble(x1 = 1)
tb2$x
```

### 

Since there is no column named exactly `x1` in the tibble `tb2`, the code will print warning message and `NULL`. For this reason we sometimes joke that tibbles are lazy and surly: they do less and complain more.







## Apply family
## For loops
## Plots
## Summary
### 

This tutorial covered [Chapter 28: A field guide to base R](https://r4ds.hadley.nz/base-r) from [*R for Data Science (2e)*](https://r4ds.hadley.nz/) by Hadley Wickham, Mine Çetinkaya-Rundel, and Garrett Grolemund. 
`


```{r download-answers, child = system.file("child_documents/download_answers.Rmd", package = "tutorial.helpers")}
```
