---
title: Communication
author: David Kane and Srihith Garlapati
tutorial:
  id: communication
output:
  learnr::tutorial:
    progressive: yes
    allow_skip: yes
runtime: shiny_prerendered
description: 'Tutorial for Chapter 12: Communication'
---

```{r setup, include = FALSE}
library(learnr)
library(tutorial.helpers)
library(tidyverse)
library(babynames)
library(ggrepel)
library(patchwork)

knitr::opts_chunk$set(echo = FALSE)
options(tutorial.exercise.timelimit = 60, 
        tutorial.storage = "local") 

labels_tib <- tibble(
  x = 1:10,
  y = cumsum(x^2)
)

labels_graphic = ggplot(mpg, aes(x = displ, y = hwy)) +
  geom_point(aes(color = class)) +
  geom_smooth(se = FALSE) +
  labs(
    x = "Engine displacement (L)",
    y = "Highway fuel economy (mpg)",
    color = "Car type",
    title = "Fuel efficiency generally decreases with engine size",
    subtitle = "Two seaters (sports cars) are an exception because of their light weight",
    caption = "Data from fueleconomy.gov"
  )

label_info <- mpg |>
  arrange(desc(displ)) |>
  slice_head(n = 1, by = drv) |>
  mutate(
    drive_type = case_when(
      drv == "f" ~ "front-wheel drive",
      drv == "r" ~ "rear-wheel drive",
      drv == "4" ~ "4-wheel drive"
    )
  ) |>
  select(displ, hwy, drv, drive_type)

potential_outliers <- mpg |>
  filter(hwy > 40 | (hwy > 20 & displ > 5))

trend_text <- "Larger engine sizes tend to have lower fuel economy." |>
  str_wrap(width = 30)

base <- ggplot(mpg, aes(x = displ, y = hwy)) +
  geom_point(aes(color = class))

random_vals <- tibble(
  x = rnorm(10000),
  y = rnorm(10000)
)
```

```{r copy-code-chunk, child = system.file("child_documents/copy_button.Rmd", package = "tutorial.helpers")}
```

```{r info-section, child = system.file("child_documents/info_section.Rmd", package = "tutorial.helpers")}
```

<!-- DK: Don't paste all the lines in the pipe in the hint. Just give the new code: -->

<!-- ... |>  -->
<!--   new_code(this) -->

<!-- DK: Delete most uses of "Use your previous code". Raises the question, what is the CP/CR equivalent of "Continue the pipe"? -->

## Introduction
### 

This tutorial covers [Chapter 12: Communication](https://r4ds.hadley.nz/communication.html)  from [*R for Data Science (2e)*](https://r4ds.hadley.nz/) by Hadley Wickham, Mine Çetinkaya-Rundel, and Garrett Grolemund. In this tutorial we will be making use of three core packages from *ggplot2*: [**scales**](https://scales.r-lib.org/) to override the default breaks, [**ggrepel**](https://ggrepel.slowkow.com), and [**patchwork**](https://patchwork.data-imaginist.com). Some key commands that you will be making use of are [`quote()`](https://rdrr.io/r/base/substitute.html) which simply returns its argument, [`geom_label_repel()`](https://rdrr.io/cran/ggrepel/man/geom_text_repel.html) which adds text directly to the plot, and [`label_dollar()`](https://scales.r-lib.org/reference/label_dollar.html) which formats numbers as currency. 

## Labels
### 

In this section, you will learn that good labels makes for a great graphic. By the end of this section, we will have learnt how to make an exploratory graphic into an expository graphic. 

### 

By the end of this section we should have this great expository plot:

```{r include = TRUE}
labels_graphic
```

### Exercise 1

Let's first load in the package **tidyverse** using the `library()` function. 

```{r labels-1, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r labels-1-hint-1, eval = FALSE}
library(...)
```

### 

Let's start with the basics and make a plot with well-formatted labels.

### Exercise 2

Now, let's explore the dataset that we want to create a plot on. Type in `mpg` and hit "Run Code". 

```{r labels-2, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r labels-2-hint-1, eval = FALSE}
mpg
```

### 

The `mpg` dataset provides fuel economy data from 1999 and 2008 for 38 popular models of cars. The dataset is shipped with [**ggplot2**](https://ggplot2.tidyverse.org/) package.

### Exercise 3

Create a new pipeline. Pipe `mpg` to the `ggplot()` function. Within the `ggplot()` function, map `x` to the `displ` variable and `y` to the hwy variable using the `aes()` function.

```{r labels-3, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r labels-3-hint-1, eval = FALSE}
mpg |> 
  ggplot(aes(x = ..., y = ...))
```

### 

### Exercise 4

Using your previous code, add the `geom_point()` function to the pipeline.

```{r labels-4, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r labels-4-hint-1, eval = FALSE}
mpg |> 
  ggplot(aes(x = ..., y = ...)) +
  geom_point()
```

### Exercise 5

Using your previous code, within the `geom_point()` function, map the argument `color` to the variable `class` using the `aes()` function. 

```{r labels-5, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r labels-5-hint-1, eval = FALSE}
mpg |> 
  ggplot(aes(x = ..., y = ...)) +
  geom_point(aes(color = ...))
```

### 

Now that we have a basic dot plot. We need to add a trend line that can allow us to see the trend of the points very clearly.

### Exercise 6

Using your previous code, add the `geom_smooth()` function to the pipeline.

```{r labels-6, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r labels-6-hint-1, eval = FALSE}
mpg |> 
  ggplot(aes(x = ..., y = ...)) +
  geom_point(aes(color = ...)) +
  geom_smooth()
```

### Exercise 7

Using your previous code, within your call to the function `geom_smooth()` add the argument called `se` and set it to `FALSE`. This removes the gray background of the line.

```{r labels-7, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r labels-7-hint-1, eval = FALSE}
mpg |> 
  ggplot(aes(x = ..., y = ...)) +
  geom_point(aes(color = ...)) +
  geom_smooth(se = ...)
```

### 

Our graphic is currently exploratory, let's turn it into an expository graphic with some good labels.

### Exercise 8

Using your previous code, add a `labs()` function to the pipeline and add a `title`, `subtitle`, `x` axis title, `y` axis title, `color` title, and a `caption`. 

(Hint the `caption` should be the following: `"Data from fueleconomy.gov"`)

```{r labels-8, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r labels-8-hint-1, eval = FALSE}
mpg |> 
  ggplot(aes(x = ..., y = ...)) +
  geom_point(aes(color = ...)) +
  geom_smooth(se = ...) +
  labs(
    title = ...,
    subtitle = ...,
    x = ...,
    y = ...,
    color = ...,
    caption = ...
  )
```

### 

The purpose of a plot title is to summarize the main finding. Avoid titles that just describe what the plot is, e.g., “A scatterplot of engine displacement vs. fuel economy”. The argument, `subtitle`, adds additional detail in a smaller font beneath the `title` and the argument, `caption`, adds text at the bottom right of the plot, often used to describe the source of the data.

### 

Reminder: The graphic should look something like this

```{r include = TRUE}
labels_graphic
```

### Exercise 9

It’s also possible to use mathematical equations instead of text strings. Let's take a look at a tibble called `labels_tib`. Type in `labels_tib` and hit "Run Code". 

```{r labels-9, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r labels-9-hint-1, eval = FALSE}
labels_tib
```

### 

We can use mathematical equations instead of textstrings for the labels if we switch `""` for the function `quote()`

### Exercise 10

Create a new pipeline. Pipe `labels_tib` to the function `ggplot()`. Within this function, map `x` to `x` and `y` to `y` using the `aes()` function.

```{r labels-10, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r labels-10-hint-1, eval = FALSE}
labels_tib |> 
  ggplot(aes(x = ..., y = ...))
```

### 

### Exercise 11

Using your previous code, add the `geom_point()` function to the pipeline.

```{r labels-11, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r labels-11-hint-1, eval = FALSE}
labels_tib |> 
  ggplot(aes(x = ..., y = ...)) +
  geom_point()
```

### Exercise 12

Using your previous code, add the `labs()` function to the pipeline. Within the `labs()` function, set the `x` axis title to `quote(x[i])`.

```{r labels-12, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r labels-12-hint-1, eval = FALSE}
labels_tib |> 
  ggplot(aes(x = ..., y = ...)) +
  geom_point() +
  labs(
    x = ...
  )
```

### 

### Exercise 13

Using your previous code, within your call to the `labs()` function add the `y` axis title and set it to `quote(sum(x[i] ^ 2, i == 1, n))`. 

```{r labels-13, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r labels-13-hint-1, eval = FALSE}
labels_tib |> 
  ggplot(aes(x = ..., y = ...)) +
  geom_point() +
  labs(
    x = ...,
    y = ...
  )
```

### Exercise 14

Type in `?plotmath` in the Console and learn more about what kind of syntax you can input into the `quote()` function. Copy/paste an interesting syntax and a meaning to the textbox below and hit "Submit".

```{r labels-14}
question_text(NULL,
	message = "Place correct answer here.",
	answer(NULL, correct = TRUE),
	allow_retry = FALSE,
	incorrect = NULL,
	rows = 6)
```

## Annotations
### 

<!-- DK: Build the whole plot as a pipe first so that results are vomitted out with each exercise. Then, if necessary at the end, you can assign it to an object.  -->

In this section we will learn how to label an individual observation or a group of observations instead of just labelling major components of a plot. 

### Exercise 1

Load in the `ggrepel` package using the `library()` function.

```{r annotations-1, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r annotations-1-hint-1, eval = FALSE}
library(...)
```

### Exercise 2

Let's make a new dataframe called `label_info` and set it to equal the dataset `mpg`.

```{r annotations-2, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r annotations-2-hint-1, eval = FALSE}
label_info <- mpg
```

### Exercise 3

Using your previous code, create a new pipeline. Pipe `mpg` to the `arrange()` function. Within this function, add the argument `desc(displ)`.

```{r annotations-3, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r annotations-3-hint-1, eval = FALSE}
label_info <- mpg |> 
  arrange(...)
```

### Exercise 4

Using your previous code, add the `slice_head()` function. Within this function set the argument `n` to `1`.

```{r annotations-4, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r annotations-4-hint-1, eval = FALSE}
label_info <- mpg |> 
  arrange(...) |> 
  slice_head(n = ...)
```

### 

### Exercise 5

Using your previous code, within your call to `slice_head()`, add a `by` argument and set it to `drv` (drivetype). 

```{r annotations-5, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r annotations-5-hint-1, eval = FALSE}
label_info <- mpg |> 
  arrange(...) |> 
  slice_head(n = ..., by = ...)
```

### 

The by argument will group by whatever is inputted into the argument. 

### Exercise 6

Using your previous code, add the `mutate()` function to the pipeline. Within this function, create a new variable called `drive_type` and set it to: 
`case_when(drv == "f" ~ "front-wheel drive", drv == "r" ~ "rear-wheel drive", drv == "4" ~ "4-wheel drive")`


```{r annotations-6, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r annotations-6-hint-1, eval = FALSE}
label_info <- mpg |> 
  arrange(...) |> 
  slice_head(n = ..., by = ...) |> 
  mutate(drive_type = case_when(...))
```

### Exercise 7

Using your previous code, add the `select()` function to the pipeline. Select the following columns: `displ`, `hwy`, `drv`, `drive_type`.

```{r annotations-7, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r annotations-7-hint-1, eval = FALSE}
label_info <- mpg |> 
  arrange(...) |> 
  slice_head(n = ..., by = ...) |> 
  mutate(drive_type = case_when(...)) |> 
  select(...)
```

### 

So, far we have pulled out the cars with the highest engine size in each drive type and saved their information as a new data frame called `label_info`.

### Exercise 8

Display `label_info` by typing `label_info` below and hitting "Run Code".

```{r annotations-8, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r annotations-8-hint-1, eval = FALSE}
label_info
```

### 

We will now use this new data frame to directly label the three groups and replace the legend by placing labels directly on the plot. 

### Exercise 9

Create a new pipeline. Pipe `mpg` to the `ggplot()` function. Within this function, map `x` to `displ`, `y` to `hwy`, and `color` to `drv` using `aes()`.

```{r annotations-9, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r annotations-9-hint-1, eval = FALSE}
mpg |> 
  ggplot(aes(x = ..., y = ..., color = ...))
```

### Exercise 10

Using your previous code, add the `geom_point()` function to the pipeline. Within this function set the argument `alpha` to `0.3`.

```{r annotations-10, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r annotations-10-hint-1, eval = FALSE}
mpg |> 
  ggplot(aes(x = ..., y = ..., color = ...)) |> 
  geom_point(alpha = ...)
```

### Exercise 11

Using your previous code, add the `geom_smooth()` function to the pipeline. Within this function set the `se` argument to `FALSE`. 

```{r annotations-11, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r annotations-11-hint-1, eval = FALSE}
mpg |> 
  ggplot(aes(x = ..., y = ..., color = ...)) |> 
  geom_point(alpha = ...) |> 
  geom_smooth(se = ...)
```

### Exercise 12

Using your previous code, add the `geom_text()` function to the pipeline. Within this function set the argument `data` to the dataframe `label_info`. 

```{r annotations-12, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r annotations-12-hint-1, eval = FALSE}
mpg |> 
  ggplot(aes(x = ..., y = ..., color = ...)) |> 
  geom_point(alpha = ...) |> 
  geom_smooth(se = ...) |> 
  geom_text(
    data = ...
  )
```

### 

This code chunk should have returned an error stating that it *requires the following missing aesthetics: label*. It returns an error because we haven't yet mapped the label argument to a variable in the dataframe.


### Exercise 13

Using your previous code, within your call to `geom_text()`, map `x` to `displ`, `y` to `hwy`, and `label` to `drive_type` using `aes()`.

```{r annotations-13, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r annotations-13-hint-1, eval = FALSE}
mpg |> 
  ggplot(aes(x = ..., y = ..., color = ...)) |> 
  geom_point(alpha = ...) |> 
  geom_smooth(se = ...) |> 
  geom_text(
    data = ..., 
    aes(...)
  )
```

### Exercise 14

Using your previous code, within your call to `geom_text()`, add the argument `fontface` and set it to `"bold"`. 

```{r annotations-14, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r annotations-14-hint-1, eval = FALSE}
mpg |> 
  ggplot(aes(x = ..., y = ..., color = ...)) |> 
  geom_point(alpha = ...) |> 
  geom_smooth(se = ...) |> 
  geom_text(
    data = ..., 
    aes(...), 
    fontface = ...
  )
```

### Exercise 15

Using your previous code, within your call to `geom_text()`, add the argument `size` and set it to `5`.

```{r annotations-15, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r annotations-15-hint-1, eval = FALSE}
mpg |> 
  ggplot(aes(x = ..., y = ..., color = ...)) |> 
  geom_point(alpha = ...) |> 
  geom_smooth(se = ...) |> 
  geom_text(
    data = ..., 
    aes(...), 
    fontface = ..., 
    size = ...
  )
```

### 

The `fontface` and `size` arguments we can customize the look of the text labels. In this case, these arguments make the labels larger than the rest of the text on the plot and bolded.

### Exercise 16

Using your previous code, within your call to `geom_text()` add the `hjust` and `vjust` arguments. Set the `hjust` argument to `"right"` and the `vjust` argument to `"bottom"`. 

```{r annotations-16, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r annotations-16-hint-1, eval = FALSE}
mpg |> 
  ggplot(aes(x = ..., y = ..., color = ...)) |> 
  geom_point(alpha = ...) |> 
  geom_smooth(se = ...) |> 
  geom_text(
    data = ..., 
    aes(...), 
    fontface = ..., 
    size = ...,
    hjust = ...,
    vjust = ...
  )
```

### 

Note the use of `hjust` (horizontal justification) and `vjust` (vertical justification) to control the alignment of the label.

### Exercise 17

Using your previous code, add the `theme()` function to the pipeline. Within this function, add the argument `legend.position` and set it to `"none"`. 

```{r annotations-17, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r annotations-17-hint-1, eval = FALSE}
mpg |> 
  ggplot(aes(x = ..., y = ..., color = ...)) |> 
  geom_point(alpha = ...) |> 
  geom_smooth(se = ...) |> 
  geom_text(
    data = ..., 
    aes(...), 
    fontface = ..., 
    size = ...
  ) |> 
  theme(legend.position = ...)
```

### 

The function `theme()` with the argument `legend.position = "none"` turns all the legends off — we’ll talk about it more shortly.)

### Exercise 18

Using your previous code, remove the `hjust` and `vjust` arguments from within the `geom_text()` function. In addition, change the `geom_text()` function to the `geom_label_repel()` function. 

```{r annotations-18, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r annotations-18-hint-1, eval = FALSE}
mpg |> 
  ggplot(aes(x = ..., y = ..., color = ...)) |> 
  geom_point(alpha = ...) |> 
  geom_smooth(se = ...) |> 
  geom_label_repel(
    data = ..., 
    aes(...), 
    fontface = ..., 
    size = ...
  ) + 
  theme(legend.position = ...)
```

### 

The annotated plot we made before is hard to read because the labels overlap with each other and with the points. We can use the `geom_label_repel()` function from the **ggrepel** package to address both of these issues. This useful package will automatically adjust labels so that they don’t overlap as you can see above.

## More Annotations
### 

In this section, we will discuss and learn more ways to add annotations to your plots.

### Exercise 1

Create a new variable called `potential_outliers` and set it equal to the dataset `mpg`. 

```{r more-annotations-1, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r more-annotations-1-hint-1, eval = FALSE}
potential_outliers <- mpg
```

### Exercise 2

Using your previous code, create a new pipeline and pipe `mpg` to the `filter()` function. Within this function add the following code: `hwy > 40 | (hwy > 20 & displ > 5)`. 

```{r more-annotations-2, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r more-annotations-2-hint-1, eval = FALSE}
potential_outliers <- mpg |> 
  filter(...)
```

### Exercise 3

Create a new pipeline. Pipe `mpg` to the `ggplot()` function. Within this function map `x` to `displ` and `y` to `hwy` using `aes()`. 

```{r more-annotations-3, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r more-annotations-3-hint-1, eval = FALSE}
mpg |> 
  ggplot(aes(x = ..., y = ...))
```

### Exercise 4

Using your previous code, add the `geom_point()` function to the pipeline. 

```{r more-annotations-4, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r more-annotations-4-hint-1, eval = FALSE}
mpg |> 
  ggplot(aes(x = ..., y = ...)) +
  geom_point()
```

### Exercise 5

Using your previous code, add the `geom_text_repel()` function. Within this function add the argument `data` and set it to `potential_outliers`.

```{r more-annotations-5, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r more-annotations-5-hint-1, eval = FALSE}
mpg |> 
  ggplot(aes(x = ..., y = ...)) +
  geom_point() +
  geom_text_repel(data = ...)
```

### 

This code chunk should have returned an error stating that it *requires the following missing aesthetics: label*. It returns an error because we haven't yet mapped the label argument to a variable in the dataframe. 

### 

In addition to `geom_text()` and `geom_label()`, you have many other geoms in **ggplot2** available to help annotate your plot. For example, you can use `geom_hline()` and `geom_vline()` to add reference lines. We often make them thick (`linewidth = 2`) and white (`color = white`), and draw them underneath the primary data layer. That makes them easy to see, without drawing attention away from the data.

### Exercise 6

Using your previous code, within your call to `geom_text_repel()`, map `label` to `model` using `aes()`.

```{r more-annotations-6, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r more-annotations-6-hint-1, eval = FALSE}
mpg |> 
  ggplot(aes(x = ..., y = ...)) +
  geom_point() +
  geom_text_repel(data = ..., 
                  aes(label = ...))
```

### 

Another geom that can help annotate your plot is `geom_rect()`. You can use `geom_rect()` to draw a rectangle around points of interest. The boundaries of the rectangle are defined by aesthetics `xmin`, `xmax`, `ymin`, `ymax`. Alternatively, look into the [**ggforce**](https://ggforce.data-imaginist.com/) package, specifically [`geom_mark_hull()`](https://ggforce.data-imaginist.com/reference/geom_mark_hull.html), which allows you to annotate subsets of points with hulls.

### Exercise 7

Using your previous code, add the `geom_point()` function to the pipeline. Within this function, add the argument `data` and set it to `potential_outliers`. 

```{r more-annotations-7, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r more-annotations-7-hint-1, eval = FALSE}
mpg |> 
  ggplot(aes(x = ..., y = ...)) +
  geom_point() +
  geom_text_repel(data = ..., 
                  aes(label = ...)) +
  geom_point(data = ...)
```

### 

You can use `geom_segment()` with the arrow argument to draw attention to a point with an arrow. Use aesthetics `x` and `y` to define the starting location, and `xend` and `yend` to define the end location. 

### Exercise 8

Using your previous code, within your most recent call to `geom_point()`, add the argument `color` and set it to `"red"`.

```{r more-annotations-8, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r more-annotations-8-hint-1, eval = FALSE}
mpg |> 
  ggplot(aes(x = ..., y = ...)) +
  geom_point() +
  geom_text_repel(data = ..., 
                  aes(label = ...)) +
  geom_point(data = ..., color = ...) 
```

### 

### Exercise 9

Using your previous code, add another `geom_point()` function to the pipeline. Within this function, add an argument called `data` and set it to `potential_outliers`. 

```{r more-annotations-9, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r more-annotations-9-hint-1, eval = FALSE}
mpg |> 
  ggplot(aes(x = ..., y = ...)) +
  geom_point() +
  geom_text_repel(data = ..., 
                  aes(label = ...)) +
  geom_point(data = ..., color = ...) +
  geom_point(data = ...)
```

### Exercise 10

Using your previous code, within your most previous call to geom_point(), add an argument called `color` and set it to `"red"`.

```{r more-annotations-10, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r more-annotations-10-hint-1, eval = FALSE}
mpg |> 
  ggplot(aes(x = ..., y = ...)) +
  geom_point() +
  geom_text_repel(data = ..., 
                  aes(label = ...)) +
  geom_point(data = ..., color = ...) +
  geom_point(data = ..., 
             color = ...)
```

### Exercise 11

Using your previous code, within your most previous call to the `geom_point()` function, add an argument called `size` and set it to `3`. 

```{r more-annotations-11, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r more-annotations-11-hint-1, eval = FALSE}
mpg |> 
  ggplot(aes(x = ..., y = ...)) +
  geom_point() +
  geom_text_repel(data = ..., 
                  aes(label = ...)) +
  geom_point(data = ..., color = ...) +
  geom_point(data = ..., 
             color = ...,
             size = ...)
```

### Exercise 12

Using your previous code, within your most previous call to the `geom_point()` function, add the `shape` argument and set it to `"circle open"`. 

```{r more-annotations-12, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r more-annotations-12-hint-1, eval = FALSE}
mpg |> 
  ggplot(aes(x = ..., y = ...)) +
  geom_point() +
  geom_text_repel(data = ..., 
                  aes(label = ...)) +
  geom_point(data = ..., color = ...) +
  geom_point(data = ..., 
             color = ...,
             size = ..., 
             shape = ...)
```

### 

Note another handy technique used here: we added a second layer of large, hollow points to further highlight the labelled points.

### Exercise 13

Another handy function for adding annotations to plots is `annotate()`. As a rule of thumb, geoms are generally useful for highlighting a subset of the data while `annotate()` is useful for adding one or few annotation elements to a plot.

### 

Create a new variable called `trend_text` and set it to `"Larger engine sizes tend to have lower fuel economy."`

```{r more-annotations-13, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r more-annotations-13-hint-1, eval = FALSE}
trend_text <- ...
```

### 

To demonstrate using `annotate()`, let’s create some text to add to our plot. The text is a bit long, so we’ll use `stringr::str_wrap()` to automatically add line breaks to it given the number of characters you want per line.

### Exercise 14

Using your previous code, create a new pipeline and pipe `trend_text` with the `str_wrap()` function. Within this function add an argument called `width` and set it to `30`. 

```{r more-annotations-14, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r more-annotations-14-hint-1, eval = FALSE}
trend_text <- ... |> 
  str_wrap(width = ...)
```

### Exercise 15

Now let's make use of this annotation within a plot. Create a new pipeline and pipe `mpg` to the `ggplot()` function. Within this function map `x` to `displ` and `y` to `hwy` using `aes()`.

```{r more-annotations-15, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r more-annotations-15-hint-1, eval = FALSE}
mpg |> 
  ggplot(aes(x = ..., y = ...))
```

### Exercise 16

Using your previous code, add the `geom_point()` function to the pipeline.

```{r more-annotations-16, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r more-annotations-16-hint-1, eval = FALSE}
mpg |> 
  ggplot(aes(x = ..., y = ...)) +
  geom_point()
```

### 

### Exercise 17

Using your previous code, add the `annotate()` function to the pipeline. Within this function set the argument `geom` to `"label"`, `x` to `3.5`, and `y` to `38`. 
 
```{r more-annotations-17, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r more-annotations-17-hint-1, eval = FALSE}
mpg |> 
  ggplot(aes(x = ..., y = ...)) +
  geom_point() +
  annotate(
    geom = ...,
    x = ..., 
    y = ...
  )
```

### 

This code chunk should have returned an error stating that it *requires the following missing aesthetics: label*. It returns an error because we haven't yet mapped the label argument to a variable in the dataframe. Now let's end the annotation about the trend into the plot while mapping the label argument.

### Exercise 18

Using your previous code, within your most recent call to the `annotate()` function, add the argument `label` and set it to `trend_text`. 

```{r more-annotations-18, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r more-annotations-18-hint-1, eval = FALSE}
mpg |> 
  ggplot(aes(x = ..., y = ...)) +
  geom_point() +
  annotate(
    geom = ...,
    x = ..., 
    y = ...,
    label = trend_text
  )
```

### 

Now, let's alighn it horizontally using the argument `hjust`. 

### Exercise 19

Using your previous code, within your most recent call to the `annotate()` function, add the arguments `hjust` and set it to `"left"` and `color` and set it to `"red"`.

```{r more-annotations-19, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r more-annotations-19-hint-1, eval = FALSE}
mpg |> 
  ggplot(aes(x = ..., y = ...)) +
  geom_point() +
  annotate(
    geom = ...,
    x = ..., 
    y = ...,
    label = ..., 
    hjust = ...,
    color = ...
  )
```

### 

Annotation is a powerful tool for communicating main takeaways and interesting features of your visualizations. The only limit is your imagination (and your patience with positioning annotations to be aesthetically pleasing)!

## Scales
### 

In this section, we will learn about how to display plots more appropriately by using scales.

### Exercise 1

First load in the package **scales** using the `library()` function.

```{r scales-1, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r scales-1-hint-1, eval = FALSE}
library(...)
```

### 

The most common use of the scales package is to customise to control the appearance of axis and legend labels. Use a break_ function to control how breaks are generated from the limits, and a label_ function to control how breaks are turned in to labels.

### Exercise 2

When you make a regular plot, **ggplot2* automatically adds scales for you. Let's make a simple plot to demonstrate this.

### 

Create a new pipeline and pipe `mpg` to the `ggplot()` function. Within this function map `x` to `displ` and `y` to `hwy` using `aes()`. 

```{r scales-2, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r scales-2-hint-1, eval = FALSE}
mpg |> 
  ggplot(aes(x = ..., y = ...))
```

### Exercise 3

Using your previous code, add the `geom_point()` function to the pipeline. Within this function map the argument `color` to `class` using `aes()`.

```{r scales-3, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r scales-3-hint-1, eval = FALSE}
mpg |> 
  ggplot(aes(x = ..., y = ...)) + 
  geom_point(aes(color = ...))
```

### 

**ggplot2** automatically adds default scales behind the scenes such as `scale_x_continuous()`, `scale_y_continuous()`, and `scale_color_discrete()`.

### Exercise 4

We can test this previous statement, by adding those functions at the end of our previous pipeline. 

### 

Using your previous code, add the `scale_x_continuous()`, `scale_y_continuous()`, and the `scale_color_discrete()` functions to the pipeline. 

```{r scales-4, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r scales-4-hint-1, eval = FALSE}
mpg |> 
  ggplot(aes(x = ..., y = ...)) + 
  geom_point(aes(color = ...)) +
  scale_x_continuous() +
  ... +
  scale_color_discrete()
```

### 

We can see that the two data frames are the same. Note the naming scheme for scales: scale_ followed by the name of the aesthetic, then _, then the name of the scale. The default scales are named according to the type of variable they align with: continuous, discrete, datetime, or date.

### Exercise 5

Create a new pipeline and pipe `mpg` to the `ggplot()` function. Within this function, map `x` to `displ`, `y` to `hwy`, and `color` to `drv` using `aes()`. 

```{r scales-5, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r scales-5-hint-1, eval = FALSE}
mpg |> 
  ggplot(aes(x = ..., y = ..., color = ...))
```

### 

There are two primary arguments that affect the appearance of the ticks on the axes and the keys on the legend: `breaks` and `labels`. Breaks controls the position of the ticks, or the values associated with the keys.

### Exercise 6

Using your previous code, add the `scale_y_continuous()` function to the pipeline. Within this function add the `breaks` argument and set it to `seq(15, 40, by = 5)`.

```{r scales-6, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r scales-6-hint-1, eval = FALSE}
mpg |> 
  ggplot(aes(x = ..., y = ..., color = ...)) +
  scale_y_continuous(breaks = ...)
```

### 

Labels controls the text label associated with each tick/key. The most common use of `breaks` is to override the default choice.

### Exercise 7

Create a new pipeline and pipe `mpg` to the `ggplot()` function. Within this function map `x` to `displ`, `y` to `hwy`, and `color` to `drv` using `aes()`

```{r scales-7, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r scales-7-hint-1, eval = FALSE}
mpg |> 
  ggplot(aes(x = ..., y = ..., color = ...))
```

### 

You can use labels in the same way (a character vector the same length as breaks), but you can also set it to NULL to suppress the labels altogether. This can be useful for maps, or for publishing plots where you can’t share the absolute numbers.

### Exercise 8

Using your previous code, add the `geom_point()` function to the pipeline.

```{r scales-8, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r scales-8-hint-1, eval = FALSE}
mpg |> 
  ggplot(aes(x = ..., y = ..., color = ...)) +
  geom_point()
```

### Exercise 9

Using your previous code, add the `scale_x_continuous()` function to the pipeline. Within this function, add the `labels` argument and set it to `NULL`. 

```{r scales-9, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r scales-9-hint-1, eval = FALSE}
mpg |> 
  ggplot(aes(x = ..., y = ..., color = ...)) +
  geom_point() +
  scale_x_continuous(labels = ...)
```

### Exercise 10

Using your previous code, add the `scale_y_continuous()` function to the pipeline. Within this function, add the `labels` argument and set it to `NULL`. 

```{r scales-10, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r scales-10-hint-1, eval = FALSE}
mpg |> 
  ggplot(aes(x = ..., y = ..., color = ...)) +
  geom_point() +
  scale_x_continuous(labels = ...) +
  scale_y_continuous(labels = ...)
```

### Exercise 11

Using your previous code, add the `scale_color_discrete()` function to the pipeline. Within this function, add the `labels` argument and set it to `c("4" = "4-wheel", "f" = "front", "r" = "rear")`. 

```{r scales-11, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r scales-11-hint-1, eval = FALSE}
mpg |> 
  ggplot(aes(x = ..., y = ..., color = ...)) +
  geom_point() +
  scale_x_continuous(labels = ...) +
  scale_y_continuous(labels = ...) +
  scale_color_discrete(labels = ...)
```

### 

You can also use breaks and labels to control the appearance of legends. For discrete scales for categorical variables, labels can be a named list of the existing levels names and the desired labels for them.

### Exercise 12

This time we will be using the **scales** package to create more efficient axis labels. Create a new pipeline and pipe `diamonds` to the `ggplot()` function. Within this function, pipe map `x` to `price` and `y` to `cut`.

```{r scales-12, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r scales-12-hint-1, eval = FALSE}
diamonds |> 
  ggplot(aes(x = ..., y = ...))
```

### Exercise 13

Using your previous code, add the `geom_boxplot()` function to the pipeline. Within this function add the `alpha` argument and set it to `0.05`. 

```{r scales-13, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r scales-13-hint-1, eval = FALSE}
diamonds |> 
  ggplot(aes(x = ..., y = ...)) +
  geom_boxplot(alpha = ...)
```

### Exercise 14

Using your previous code, add the `scale_x_continuous()` function. Within this function add the `labels` argument and set it equal to the `label_dollar()` function.

```{r scales-14, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r scales-14-hint-1, eval = FALSE}
diamonds |> 
  ggplot(aes(x = ..., y = ...)) +
  geom_boxplot(alpha = ...) +
  scale_x_continuous(labels = ...)
```

### 

The `labels` argument coupled with labelling functions from the **scales** package is also useful for formatting numbers as currency, percent, etc. The plot shows default labelling with `label_dollar()`, which adds a dollar sign as well as a thousand separator comma.

### Exercise 15

Another handy label function is `label_percent()`. Create a new pipeline and pipe `diamonds` to the `ggplot()` function. Within this function, map `x` to `cut` and `fill` to `clarity`.

```{r scales-15, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r scales-15-hint-1, eval = FALSE}
diamonds |> 
  ggplot(aes(x = ..., fill = ...))
```

### Exercise 16

Using your previous code, add the `geom_bar()` function. Within this function add the `position` argument and set it equal to `"fill"`. 

```{r scales-16, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r scales-16-hint-1, eval = FALSE}
diamonds |> 
  ggplot(aes(x = ..., fill = ...)) +
  geom_bar(position = ...)
```

### Exercise 17

Using your previous code, add the `scale_y_continuous()` function to the pipeline. Within this function, add the `name` argument and set equal to `"Percentage"` and add the `labels` argument and set it equal to `label_percent()`. 

```{r scales-17, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r scales-17-hint-1, eval = FALSE}
diamonds |> 
  ggplot(aes(x = ..., fill = ...)) +
  geom_bar(position = ...) +
  scale_y_continuous(name = ..., labels = ...)
```

### Exercise 18

Type in `presidential` and hit "Run Code".

```{r scales-18, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r scales-18-hint-1, eval = FALSE}
presedential
```

### 

The dataset, `presidential` gives the names of each president, the start and end date of their term, and their party of 12 US presidents from Eisenhower to Trump. This data is in the public domain.

### Exercise 19

Create a new pipeline and pipe `presidential` to the `mutate()` function. Within this function, create a new variable `id` and set it equal to `33` plus `row_number()`.

```{r scales-19, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r scales-19-hint-1, eval = FALSE}
presidential |> 
  mutate(id = ...)
```

### Exercise 20

Using your previous code, add the `ggplot()` function to the pipeline. Within this function, map `x` to `start` and `y` to `id` using `aes()`. 

```{r scales-20, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r scales-20-hint-1, eval = FALSE}
presidential |> 
  mutate(id = ...) |> 
  ggplot(aes(x = ..., y = ...))
```

### Exercise 21

Using your previous code, add the `geom_point()` function to the pipeline. 

```{r scales-21, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r scales-21-hint-1, eval = FALSE}
presidential |> 
  mutate(id = ...) |> 
  ggplot(aes(x = ..., y = ...)) +
  geom_point()
```

### Exercise 22

Using your previous code, add the `geom_segment()` function. Within this function, map `xend` to `end` and `yend` to `id`.

```{r scales-22, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r scales-22-hint-1, eval = FALSE}
presidential |> 
  mutate(id = ...) |> 
  ggplot(aes(x = ..., y = ...)) +
  geom_point() +
  geom_segment(aes(xend = ..., yend = ...))
```

### 

Another use of breaks is when you have relatively few data points and want to highlight exactly where the observations occur.

### Exercise 23

Using your previous code, add the `scale_x_date()` function. Within this function add the argument `name` and set it equal to `NULL`.

```{r scales-23, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r scales-23-hint-1, eval = FALSE}
presidential |> 
  mutate(id = ...) |> 
  ggplot(aes(x = ..., y = ...)) +
  geom_point() +
  geom_segment(aes(xend = ..., yend = ...)) +
  scale_x_date(name = ...)
```

### Exercise 24

Using your previous code, within your call to the function `scale_x_date()`, add the argument `breaks` and set it equal to `presidential$start`.

```{r scales-24, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r scales-24-hint-1, eval = FALSE}
presidential |> 
  mutate(id = ...) |> 
  ggplot(aes(x = ..., y = ...)) +
  geom_point() +
  geom_segment(aes(xend = ..., yend = ...)) +
  scale_x_date(name = ..., breaks = ...)
```

### 

Note that for the breaks argument we pulled out the start variable as a vector with presidential$start because we can’t do an aesthetic mapping for this argument. Also note that the specification of breaks and labels for date and datetime scales is a little different:

### Exercise 25

Using your previous code, within your call to the function `scale_x_date()`, add the argument `date_labels` and set it equal to `"'%y"`.

```{r scales-25, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r scales-25-hint-1, eval = FALSE}

```

### 

The argument `date_labels` takes a format specification, in the same form as `parse_datetime()`. This plot shown here shows when each US president started and ended their term.

## Legend Layout
### 

In this section we will learn how to adjust the placement of the legend and how to customize the legend of your plot.

### Exercise 1

First create a new variable called `base`. Set `base` equal to the dataset `mpg`. (Add your code before `base` displayed below)

```{r legend-layout-1, exercise = TRUE}

base
```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r legend-layout-1-hint-1, eval = FALSE}
base <- mpg
```

### Exercise 2

Using your previous code, create a new pipeline and pipe `mpg` to the `ggplot()` function. Within this function map `x` to `displ` and `y` to `hwy`. (Add your code before `base` displayed below)

```{r legend-layout-2, exercise = TRUE}

base
```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r legend-layout-2-hint-1, eval = FALSE}
base <- mpg |> 
  ggplot(aes(x = ..., y = ...))
```

### Exercise 3

Using your previous code, add the `geom_point()` function to the pipeline. Within this newly added function map `color` to `class`. (Add your code before `base` displayed below)

```{r legend-layout-3, exercise = TRUE}

base
```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r legend-layout-3-hint-1, eval = FALSE}
base <- mpg |> 
  ggplot(aes(x = ..., y = ...)) +
  geom_point(aes(color = ...))
```

### 

To control the overall position of the legend, you need to use a theme() setting. The theme setting legend.position controls where the legend is drawn. The default is `legend.position = "right"`. 

### Exercise 4

Using your previous code, add the `theme()` function to the pipeline. Within this function, add the `legend.position` argument and set it equal to `"left".` (Add your code before `base` displayed below)

```{r legend-layout-4, exercise = TRUE}

base
```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r legend-layout-4-hint-1, eval = FALSE}
base <- mpg |> 
  ggplot(aes(x = ..., y = ...)) +
  geom_point(aes(color = ...)) +
  theme(legend.position = "left")
```

### 

As you can see, the legend's position is now to the left of the plot.

### Exercise 5

Using your previous code, within your call to the function `theme()` change the argument's definition from `"left"` to `"top"`. (Add your code before `base` displayed below)

```{r legend-layout-5, exercise = TRUE}

base
```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r legend-layout-5-hint-1, eval = FALSE}
base <- mpg |> 
  ggplot(aes(x = ..., y = ...)) +
  geom_point(aes(color = ...)) +
  theme(legend.position = "top")
base
```

### 

We can see that the legend is taking too much space at the top of the plot. So let's adjust the way the legend is displayed.

### Exercise 6

Using your previous code, add the `guides()` function to the pipeline. Within this function, add the `color` argument and set it equal to `guide_legend(nrow = 3)`. (Add your code before `base` displayed below)

```{r legend-layout-6, exercise = TRUE}

base
```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r legend-layout-6-hint-1, eval = FALSE}
base <- mpg |> 
  ggplot(aes(x = ..., y = ...)) +
  geom_point(aes(color = ...)) +
  theme(legend.position = "top") +
  guides 
```

### 

This fixes our previous problem. If your plot is short and wide, place the legend at the legend at the top or bottom, and if it’s tall and narrow, place the legend at the left or right. You can also use legend.position = "none" to suppress the display of the legend altogether.

## Replacing a Scale
### 

In this section, we will learn about switching out a scale for a new one. The scales that you would wanna switch out are the continuous position scales and the color scales.

### Exercise 1

We will be working with a dataset called `diamonds`. Type in `diamonds` and hit "Run Code".

```{r replacing-a-scale-1, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r replacing-a-scale-1-hint-1, eval = FALSE}

```

### 

The dataset `diamonds` us a dataset containing the prices and other attributes of almost 54,000 diamonds. The variables are as follows:

### Exercise 2

Create a new pipeline. Pipe `diamonds` to the `ggplot()` function. Within this function map `x` to `carat` and `y` to `price`.

```{r replacing-a-scale-2, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r replacing-a-scale-2-hint-1, eval = FALSE}
diamonds |> 
  ggplot(aes(x = ..., y = ...))
```

### Exercise 3

Using your previous code, add the `geom_bin2d()` function to the pipeline.

```{r replacing-a-scale-3, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r replacing-a-scale-3-hint-1, eval = FALSE}
... +
  geom_bin2d()
```

### 

It’s very useful to plot transformations of your variable. For example, it’s easier to see the precise relationship between carat and price if we log transform them.

### Exercise 4

Using your previous code, within the `ggplot()` function, add the `log10()` function to `carat` and `price`. 

```{r replacing-a-scale-4, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r replacing-a-scale-4-hint-1, eval = FALSE}
diamonds |> 
  ggplot(aes(x = log10(...), y = ...)) +
  geom_bin2d()
```

### 

However, the disadvantage of this transformation is that the axes are now labelled with the transformed values, making it hard to interpret the plot. Instead of doing the transformation in the aesthetic mapping, we can instead do it with the scale. 

### Exercise 5

Using your previous code, within the `ggplot()` function remove the `log10()` around `carat` and `price`. Then add the function `scale_x_log10()` to the pipeline. 

```{r replacing-a-scale-5, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r replacing-a-scale-5-hint-1, eval = FALSE}
diamonds |> 
  ggplot(aes(x = ..., y = ...)) +
  geom_bin2d() +
  scale_x_log10()
```

### Exercise 6

Using your previous code, add the `scale_y_log10()` function to the pipeline.

```{r replacing-a-scale-6, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r replacing-a-scale-6-hint-1, eval = FALSE}
... + 
  scale_y_log10()
```

### 

This plot is much easier on the eyes due to the axes being much more organized than the previous plot.

### Exercise 7

Create a new pipeline. Pipe `mpg` to the `ggplot()` function. Within this function map `x` to `displ` and `y` to `hwy`.

```{r replacing-a-scale-7, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r replacing-a-scale-7-hint-1, eval = FALSE}
mpg |> 
  ggplot(aes(x = ..., y = ...))
```

### Exercise 8

Using your previous code, add the `geom_point()` function to the pipeline. Within this function, map `color` to `drv`.

```{r replacing-a-scale-8, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r replacing-a-scale-8-hint-1, eval = FALSE}
,,, +
  geom_point(aes(color = ...))
```

### 

Another scale that is frequently customized is color. The default categorical scale picks colors that are evenly spaced around the color wheel. Useful alternatives are the ColorBrewer scales which have been hand tuned to work better for people with common types of color blindness.

### Exercise 9

Using your previous code, add a scale function, `scale_color_brewer()` to the pipeline. Within this function, add the argument `palette` and set it equal to `"Set1"`. 

```{r replacing-a-scale-9, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r replacing-a-scale-9-hint-1, eval = FALSE}
... +
  scale_color_brewer(palette = ...)
```

### 

Don’t forget simpler techniques for improving accessibility. If there are just a few colors, you can add a redundant shape mapping. This will also help ensure your plot is interpretable in black and white.

### Exercise 10

Using your previous code, within your call to the `geom_point()` function, add to your map, by mapping `shape` to `drv`. 

```{r replacing-a-scale-10, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r replacing-a-scale-10-hint-1, eval = FALSE}
... +
  geom_point(aes(color = ..., shape = ...)) +
  ....
```

### 

The ColorBrewer scales are documented online at https://colorbrewer2.org/ and made available in R via the RColorBrewer package, by Erich Neuwirth.

### Exercise 11

Create a new pipeline and pipe the dataset `presidential` to the `mutate()` function. Within the `mutate()` function, create a new variable called `id` and set it equal to `33` plus `row_number()`.

```{r replacing-a-scale-11, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r replacing-a-scale-11-hint-1, eval = FALSE}
presidential |> 
  mutate(id = ...)
```

### 

### Exercise 12

Using your previous code, add the `ggplot()` function to the pipeline and map `x` to `start`, `y` to `id`, and `color` to `party`.

```{r replacing-a-scale-12, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r replacing-a-scale-12-hint-1, eval = FALSE}
... + 
  ggplot(aes(x = ..., y = ..., color = ...))
```

### Exercise 13

Using your previous code, add the `geom_point()` function to the pipeline. 

```{r replacing-a-scale-13, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r replacing-a-scale-13-hint-1, eval = FALSE}
... +
  geom_point()
```

### Exercise 14

Using your previous code, add the `geom_segment()` function. Within this function, map `xend` to `end` and `yend` to `id`.

```{r replacing-a-scale-14, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r replacing-a-scale-14-hint-1, eval = FALSE}
... +
  geom_segment(aes(xend = ..., yend = ...))
```

### Exercise 15

Using your previous code, add the `scale_color_manual()` function to the pipeline. Within this function add the argument `values` and set it to equal `c(Republican = "#E81B23", Democratic = "#00AEF3")`. 

```{r replacing-a-scale-15, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r replacing-a-scale-15-hint-1, eval = FALSE}
... +
  scale_color_manual(values = ...)
```

### 

If we wanted to map presidential party to color, we want to use the standard mapping of red for Republicans and blue for Democrats. One approach for assigning these colors is using hex color codes as shown above. 

### Exercise 16

For our next plot, we will create a plot with the dataset `random_vals`. Type in `random_vals` and hit "Run Code".

```{r replacing-a-scale-16, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r replacing-a-scale-16-hint-1, eval = FALSE}
random_vals
```

### Exercise 17

Create a new pipeline. Pipe `random_vals` to the `ggplot()` function. Within this function, map `x` to `x` and `y` to `y`. 

```{r replacing-a-scale-17, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r replacing-a-scale-17-hint-1, eval = FALSE}
random_vals |> 
  ggplot(aes(x, y))
```

### 

For continuous color, you can use the built-in scale_color_gradient() or scale_fill_gradient(). If you have a diverging scale, you can use scale_color_gradient2(). That allows you to give, for example, positive and negative values different colors.

### Exercise 18

Using your previous code, add the `geom_hex()` function to the pipeline.

```{r replacing-a-scale-18, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r replacing-a-scale-18-hint-1, eval = FALSE}
... +
  geom_hex()
```

### Exercise 19

Using your previous code, add the `coord_fixed()` function to the pipeline.

```{r replacing-a-scale-19, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r replacing-a-scale-19-hint-1, eval = FALSE}
... +
  coord_fixed()
```

### 

Another option is to use the viridis color scales. The designers, Nathaniel Smith and Stéfan van der Walt, carefully tailored continuous color schemes that are perceptible to people with various forms of color blindness as well as perceptually uniform in both color and black and white. These scales are available as continuous (c), discrete (d), and binned (b) palettes in **ggplot2**.

### Exercise 20

Using your previous code, add the `scale_color_viridis_c()` function to the pipeline. 

```{r replacing-a-scale-20, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r replacing-a-scale-20-hint-1, eval = FALSE}
... +
  scale_fill_viridis_c()
```

### 

This is the plot that shows a continuous plot in the viridis color scale.

### Exercise 21

Using your previous code, change the `scale_fill_viridis_c()` to `scale_fill_viridis_b()`.

```{r replacing-a-scale-21, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r replacing-a-scale-21-hint-1, eval = FALSE}
... +
  scale_fill_viridis_b()
```

### 

As you can see the color gradient looks chunky and unlike the continuous plot before. This is because it is a binned plot on the viridis color scale. 

### Exercise 22

Using your previous code, add the `labs()` function to the pipeline. Within this function, add a title, an x axis title, and a y axis title to the plot.

```{r replacing-a-scale-22, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r replacing-a-scale-22-hint-1, eval = FALSE}
... +
  labs(title = ...,
       x = ..., 
       y = ...)
```

### 

Note that all color scales come in two varieties: scale_color_*() and scale_fill_*() for the color and fill aesthetics respectively (the color scales are available in both UK and US spellings).

## Babynames
### 

The `babynames` data set comes from the SSA (Social Security Administration). After 1986, all baby's were required to be given a social security number. This data contains all the names, year born, and sex of the babies. The number of names is counted and is converted into a proportion. We will be using this data to analyze the popularity of top 10 names each year.

### 

Let's create this graph:

<!-- DK: Change sex to Sex in legend title; change F/M to Female and Male in legend body; place Male on top. A couple of steps should be separated out into separate questions. Deal with "## Selecting by prop" message. -->

```{r message = FALSE}
baby_p <- babynames |>
  group_by(year, sex) |> 
  top_n(n = 10) |>
  ungroup() |> 
  summarise(total = sum(prop), .by = c(sex, year)) |>
  ggplot(mapping = aes(x = year, y = total, color = sex)) +
    geom_point() +
    scale_y_continuous(labels = scales::percent) +
    labs(title = "Total Popularity of Top 10 Names",
         subtitle = "The most popular names are beccoming much less popular",
         x = "Year",
         y = "Top 10 Names as Percentage of All Names") 

baby_p
```

### Exercise 1

Use `library()` to load the **babynames** package.

```{r babynames-1, exercise = TRUE}

```

```{r babynames-1-hint-1, eval = FALSE}
library(...)
```

### 

Learn more about the package from its [website](http://hadley.github.io/babynames/).

### Exercise 2

Recall that your Console and the exercise code chunks for this tutorial use different R sessions. The fact that you loaded **babynames** in an exercise code chunk does not mean that it is loaded in your Console.

In your Console, after loading the **babynames** package, look up the help page for the `babynames` tibble with `?babynames`. Copy/paste the "Format" paragraph.

```{r babynames-2}
question_text(NULL,
	answer(NULL, correct = TRUE),
	allow_retry = TRUE,
	try_again_button = "Edit Answer",
	incorrect = NULL,
	rows = 3)
```

### 

The variable `prop` is more subtle than it might, at first, appear. It is the proportion, **within a given `year` and `sex`**, of the total number of babies who have that given name. 


### Exercise 3

Run `summary()` on the `babynames` tibble.

```{r babynames-3, exercise = TRUE}

```

```{r babynames-3-hint-1, eval = FALSE}
summary(...)
```

### 

Note that we have no missing values.

### Exercise 4

Print out `babynames`

```{r babynames-4, exercise = TRUE}

```

```{r babynames-4-hint-1, eval = FALSE}
babynames
```

### 

There are almost 2 million observations.

### Exercise 5

Pipe `babynames` to `group_by(year, sex)`. This will lead future commands in the pipe to do all calculations on a per year/sex basis, at least until we `ungroup()` the data.

```{r babynames-5, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r babynames-5-hint-1, eval = FALSE}
babynames |> 
  group_by(year, sex) 
```

### 

Modern R code rarely uses `group_by()` anymore, prefering to `.by`/`by` arguments to *Tidyverse* functions like `summarise()`. 


### Exercise 6

Continue the pipe with `top_n(n = 10)`.

```{r babynames-6, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r babynames-6-hint-1, eval = FALSE}
... |> 
  top_n(n = 10)
```

### 

This pulls out the 10 most popular names, for each year and, within each year, for each sex.

Note that there should be 2,760 rows in the resulting tibble: 138 years times 10 names times 2 sexes. Instead we have 2,761 rows. Can you figure out why? Does it mess up our plot?

### Exercise 7

Continue the pipe with `ungroup()`.

```{r babynames-7, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r babynames-7-hint-1, eval = FALSE}
... |> 
  ungroup()
```

### 

One reason that `group_by()` is being phased out is that, unless you remember to `ungroup()` in the pipe, you will get weird results and/or errors. `ungroup()` removes the grouping variables which you had placed on the tibble previously.


### Exercise 8

Continue the pipe with `summarise()`, using the argument `total = sum(prop)`.

```{r babynames-8, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r babynames-8-hint-1, eval = FALSE}
... |>
  summarise(total = sum(prop))
```

### 

This produces a somewhat nonsense result. The sum of all the values of `prop` for all these years and both sexes does not really mean anything.

### Exercise 9

Edit the call to the `summarise()` by adding `.by` argument with value `c(sex`, `year`). 

```{r babynames-9, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r babynames-9-hint-1, eval = FALSE}
... |> 
  summarise(total = sum(prop), .by = c(...))
```

### 

This generates the data which we want to plot.

### Exercise 10

Continue your pipe with `ggplot()`. Within `aes()`, map `year` to the x-axis, `total` to the y-axis, and `sex` to color. Also add `geom_point()`.

```{r babynames-10, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r babynames-10-hint-1, eval = FALSE}
...|>
  ggplot(mapping = aes(x = ..., y = ..., color = ...)) +
    geom_point()
```

### Exercise 11

To finish your plot, use `labs()` to give the graph a title, subtitle, and axis labels. 

```{r babynames-11, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r babynames-11-hint-1, eval = FALSE}
... +
  labs(...)
```

Reminder: This is what your plot should look like.

```{r show-baby}
baby_p
```

### 

Can you see where that extra row we ignored earlier in the analysis messes things up for one point in the male set in the 1880s?


## Summary
### 

This tutorial covered [Chapter 12: Communication](https://r4ds.hadley.nz/communication.html) from [*R for Data Science (2e)*](https://r4ds.hadley.nz/) by Hadley Wickham, Mine Çetinkaya-Rundel, and Garrett Grolemund. You learned how to use [**ggplot2**](https://ggplot2.tidyverse.org/) and related packages to create persuasive graphics.


```{r download-answers, child = system.file("child_documents/download_answers.Rmd", package = "tutorial.helpers")}
```
