---
title: Missing Values
author: Matthew Ru
tutorial:
  id: missing-values
output:
  learnr::tutorial:
    progressive: yes
    allow_skip: yes
runtime: shiny_prerendered
description: 'Tutorial for Chapter 19: Missing Values'
---

```{r setup, include = FALSE}
library(learnr)
library(tutorial.helpers)
library(tidyverse)
knitr::opts_chunk$set(echo = FALSE)
options(tutorial.exercise.timelimit = 60, 
        tutorial.storage = "local") 

treatment <- tribble(
  ~person,           ~treatment, ~response,
  "Derrick Whitmore", 1,         7,
  NA,                 2,         10,
  NA,                 3,         NA,
  "Katherine Burke",  1,         4
)
```

```{r copy-code-chunk, child = system.file("child_documents/copy_button.Rmd", package = "tutorial.helpers")}
```

```{r info-section, child = system.file("child_documents/info_section.Rmd", package = "tutorial.helpers")}
```

## Introduction
###

This tutorial covers [Chapter 19: Missing Values](https://r4ds.hadley.nz/missing-values.html) from [*R for Data Science (2e)*](https://r4ds.hadley.nz/) by Hadley Wickham, Mine Çetinkaya-Rundel, and Garrett Grolemund. The primary focus of this tutorial will be to teach you how to use commands like [`fill()`] to fill in missing values, [`coalesce()`] to replace missing values with another value, and [`na_if()`] to replace certain values with a missing value, [`NA`]. Additionally we will look at functions like [`complete()`] which lets you generate missing values from a set of variables, and how to use [`anti_join()`] for missing values when joining data sets.


## Explicit missing values
###

### Exercise 1

Load the **tidyverse** package with the `library()` command.

```{r explicit-missing-val-1, exercise = TRUE}

```

```{r explicit-missing-val-1-hint, eval = FALSE}
library(...)
```

###

The functions for working with missing data mostly come from dplyr and tidyr, which are core members of the tidyverse.

### Exercise 2

Run this code in order to create a tibble called `treatment`.

```{r explicit-missing-val-2, exercise = TRUE}
treatment <- tribble(
  ~person,           ~treatment, ~response,
  "Derrick Whitmore", 1,         7,
  NA,                 2,         10,
  NA,                 3,         NA,
  "Katherine Burke",  1,         4
)
```

###

A common use for missing values is as a data entry convenience. When data is entered by hand, missing values sometimes indicate that the value in the previous row has been repeated (or carried forward).

### Exercise 3

Write `treatment` and hit "Run Code".

```{r explicit-missing-val-3, exercise = TRUE}

```

```{r explicit-missing-val-3-hint, eval = FALSE}
treatment
```

###

<!-- DK: Drop some knowledge. -->

### Exercise 4

Pipe `treatment` to `fill()` with the argument `response`.

```{r explicit-missing-val-4, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r explicit-missing-val-4-hint, eval = FALSE}
treatment |>
  ...(response)
```

###

This treatment is sometimes called “last observation carried forward”, or locf for short.

### Exercise 5

Use the same code, but replace `response` with `everything()`. Recall that `everything()` is a function which returns all the variables in a tibble.

```{r explicit-missing-val-5, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r explicit-missing-val-5-hint, eval = FALSE}
treatment |>
  fill(...)
```

###

You can use the `.direction` argument to fill in missing values that have been generated in more exotic ways.



```{r download-answers, child = system.file("child_documents/download_answers.Rmd", package = "tutorial.helpers")}
```

