---
title: Missing Values
author: Matthew Ru
tutorial:
  id: missing-values
output:
  learnr::tutorial:
    progressive: yes
    allow_skip: yes
runtime: shiny_prerendered
description: 'Tutorial for Chapter 19: Missing Values'
---


```{r setup, include = FALSE}
library(learnr)
library(tutorial.helpers)
library(tidyverse)
library(nycflights13)
library(primer.data)
knitr::opts_chunk$set(echo = FALSE)
options(tutorial.exercise.timelimit = 60, 
        tutorial.storage = "local") 
```

```{r copy-code-chunk, child = system.file("child_documents/copy_button.Rmd", package = "tutorial.helpers")}
```

```{r info-section, child = system.file("child_documents/info_section.Rmd", package = "tutorial.helpers")}
```

## Introduction
###

This tutorial covers [Chapter 19: Missing Values](https://r4ds.hadley.nz/missing-values.html) from [*R for Data Science (2e)*](https://r4ds.hadley.nz/) by Hadley Wickham, Mine Çetinkaya-Rundel, and Garrett Grolemund. The primary focus of this tutorial will be to teach you how to use commands like [`fill()`] to fill in missing values, [`coalesce()`] to replace missing values with another value, and [`na_if()`] to replace certain values with a missing value, [`NA`]. Additionally we will look at functions like [`complete()`] which lets you generate missing values from a set of variables, and how to use [`anti_join()`] for missing values when joining data sets.


Ultimately, this chapter will teach you several versatile ways to deal with missing values, identify types of missing values, and show you what to do when encountering an empty group.

## Explicit Missing Values
###
