---
title: Iteration
author: David Kane and Melissa Ban
tutorial:
  id: iteration
output:
  learnr::tutorial:
    progressive: yes
    allow_skip: yes
runtime: shiny_prerendered
description: 'Tutorial for Chapter 27: Iteration'
---

```{r setup, include = FALSE}
library(learnr)
library(tutorial.helpers)
library(tidyverse)

knitr::opts_chunk$set(echo = FALSE)
options(tutorial.exercise.timelimit = 60, 
        tutorial.storage = "local") 

df <- tibble(
  a = rnorm(10),
  b = rnorm(10),
  c = rnorm(10),
  d = rnorm(10)
)

df2 <- tibble(
  grp = sample(2, 10, replace = TRUE),
  a = rnorm(10),
  b = rnorm(10),
  c = rnorm(10),
  d = rnorm(10)
)

rnorm_na <- function(n, n_na, mean = 0, sd = 1) {
  sample(c(rnorm(n - n_na, mean = mean, sd = sd), rep(NA, n_na)))
}

df_miss <- tibble(
  a = rnorm_na(5, 1),
  b = rnorm_na(5, 1),
  c = rnorm_na(5, 2),
  d = rnorm(5)
)

expand_dates <- function(df) {
  df |> 
    mutate(
      across(where(is.Date), list(year = year, month = month, day = mday))
    )
}

df_date <- tibble(
  name = c("Amy", "Bob"),
  date = ymd(c("2009-08-03", "2010-01-16"))
)

summarize_means <- function(df, summary_vars = where(is.numeric)) {
  df |> 
    summarize(
      across({{ summary_vars }}, \(x) mean(x, na.rm = TRUE)),
      n = n()
    )
}

long <- df |> 
  pivot_longer(a:d) |> 
  group_by(name) |> 
  summarize(
    median = median(value),
    mean = mean(value))

df_paired <- tibble(
  a_val = rnorm(10),
  a_wts = runif(10),
  b_val = rnorm(10),
  b_wts = runif(10),
  c_val = rnorm(10),
  c_wts = runif(10),
  d_val = rnorm(10),
  d_wts = runif(10)
)

gapminder_1952 <- readxl::read_excel("data/gapminder/1952.xlsx")
gapminder_1957 <- readxl::read_excel("data/gapminder/1957.xlsx")
gapminder_1962 <- readxl::read_excel("data/gapminder/1962.xlsx")
gapminder_1952 <- readxl::read_excel("data/gapminder/1967.xlsx")
gapminder_1957 <- readxl::read_excel("data/gapminder/1972.xlsx")
gapminder_1962 <- readxl::read_excel("data/gapminder/1977.xlsx")
gapminder_1952 <- readxl::read_excel("data/gapminder/1982.xlsx")
gapminder_1957 <- readxl::read_excel("data/gapminder/1987.xlsx")
gapminder_1962 <- readxl::read_excel("data/gapminder/1992.xlsx")
gapminder_1952 <- readxl::read_excel("data/gapminder/1997.xlsx")
gapminder_1957 <- readxl::read_excel("data/gapminder/2002.xlsx")
gapminder_2007 <- readxl::read_excel("data/gapminder/2007.xlsx")

paths <- list.files("data/gapminder", pattern = "[.]xlsx$", full.names = TRUE)

# DK: Had to comment out the creation of this object since it was causing a
# weird error during Github checks. Let's discuss.

# gapminder <- paths |> 
#   set_names(basename) |> 
#   map(readxl::read_excel) |> 
#   list_rbind(names_to = "year") |> 
#   mutate(year = parse_number(year))

process_file <- function(path) {
  df <- read_csv(path)
  
  df |> 
    filter(!is.na(id)) |> 
    mutate(id = tolower(id)) |> 
    pivot_longer(jan:dec, names_to = "month")
}
```

```{r copy-code-chunk, child = system.file("child_documents/copy_button.Rmd", package = "tutorial.helpers")}
```

```{r info-section, child = system.file("child_documents/info_section.Rmd", package = "tutorial.helpers")}
```

<!-- DK: Only use the code button where it makes sense to do so. -->

<!-- Avoid group_by() if possible. .by/by in the next function should work. -->

<!-- Add `paths` creation to setup code chunk. -->

## Introduction
### 

This tutorial covers [Chapter 27: Iteration](https://r4ds.hadley.nz/iteration.html) from [*R for Data Science (2e)*](https://r4ds.hadley.nz/) by Hadley Wickham, Mine Çetinkaya-Rundel, and Garrett Grolemund.

<!-- finish this part later. -->

## Modifying multiple columns
### 

<!-- DK: Say something! -->

### Exercise 1

Load the **tidyverse** package with the `library()` command.

```{r modifying-multiple-c-1, exercise = TRUE}

```

```{r modifying-multiple-c-1-hint-1, eval = FALSE}
library(...)
```

```{r, include=FALSE}
library(tidyverse)
```


### 

Recall that we have already learnt a number of tools that perform the same action for multiple "things":

`facet_wrap()` and `facet_grid()` draws a plot for each subset.

`group_by()` plus `summarize()` computes summary statistics for each subset.

`unnest_wider()` and `unnest_longer()` create new rows and columns for each element of a list-column.

### Exercise 2

Hit "Run Code" to create the `df` tibble.

```{r modifying-multiple-c-2, exercise = TRUE}
df <- tibble(
  a = rnorm(10),
  b = rnorm(10),
  c = rnorm(10),
  d = rnorm(10)
)
```

### 

This is a simple tibble, now you wish to count the number of observations and compute the median of every column.

<!-- Add smth abt functional programming -->

### Exercise 3

Type `df` and hit "Run Code".

```{r modifying-multiple-c-3, exercise = TRUE}

```


```{r modifying-multiple-c-3-hint-1, eval = FALSE}

```

```{r, include = FALSE}
df
```


### Exercise 4

Pipe `df` to `summarise()`.

```{r modifying-multiple-c-4, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r modifying-multiple-c-4-hint-1, eval = FALSE}
df |> 
  ...()
```

```{r, include=FALSE}
df |> 
  summarise()
```

Note that this creates a tibble with one row and zero columns.

### 

The result above is due to absence of arguments and values in the function `summarise()`. 

### Exercise 5

Add `n = n()` to the call to `summarise()`.

```{r modifying-multiple-c-5, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r modifying-multiple-c-5-hint-1, eval = FALSE}
df |> 
  summarise(n = ...)
```

```{r, include=FALSE}
df |> 
  summarise(n = n())
```


### 

Recall that this create a new column named `n` which records the number of observations in `df`.

### Exercise 6

Add the median of `a` and median of `b`, `med_a = median (a)` and `med_b = median (b)`, to the call to `summarise()`. Remember the arguments to summarize must be separated by commas.

```{r modifying-multiple-c-6, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r modifying-multiple-c-6-hint-1, eval = FALSE}
df |> 
  summarise(n = n(),
            ... = median(a),
            med_b = ...(b))
```

```{r, include=FALSE}
df |> 
  summarise(n = n(),
            med_a = median(a),
            med_b = median(b))
```


### 

Now you have a table of 3 columns: `n`, `med_a`, and `med_b`. 

### Exercise 7

Copy and paste the previous code, then add `med_c = median(c)` and `med_d = median(d)` to the call to `summarise ()`.

```{r modifying-multiple-c-7, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r modifying-multiple-c-7-hint-1, eval = FALSE}
df |> 
  summarise(n = n(),
            med_a = median(a),
            med_b = median(b),
            ... = ...,
            med_d = median(...))
```

```{r, include=FALSE}
df |> 
  summarise(n = n(),
            med_a = median(a),
            med_b = median(b),
            med_c = median(c),
            med_d = median(d))
```


### 

You should see a tibble with five columns: `n`, `med_a`, `med_b`, `med_c`, and `med_d`.

### 

You may find the above process repetitive -- we were simply copy-pasted `med_a = median(a)` 4 times --- which, violates our rule of thumb to never copy and paste more than twice.

This would get even more tedious if you have tens or even hundreds of columns!

### 

Solution? `across()` is here to make our lifes easier.

### Exercise 8

Run the code below.

```{r modifying-multiple-c-8, exercise = TRUE}
df |> 
  summarise(n = n())
```

<!-- DK: drop some knowledg. tells us about across. link to its help page. -->

### Exercise 9

Copy the previous code. Now, add `across(a:d, median)` to the call to `summarise()`. Hit "Run Code," see what happens.

```{r modifying-multiple-c-9, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r modifying-multiple-c-9-hint-1, eval = FALSE}
df |> 
  summarise(n = n(),
            ...(a:d, ...))
```

```{r, include=FALSE}
df |> 
  summarise(n = n(),
            across(a:d, median))
```

<!-- DK: Do not hard code exercise numbers. Use "as above" or "as previously." -->

You should get the exact same table as in Exercise 7.

### 

We will discuss 3 important specifications of `across()` in the following sections.

### 

The first specification is `.cols`, used to select columns.

### Exercise 10

Let's start by running the typed-in code. Hit "Run Code".

```{r modifying-multiple-c-10, exercise = TRUE}
df2 <- tibble(
  grp = sample(2, 10, replace = TRUE),
  a = rnorm(10),
  b = rnorm(10),
  c = rnorm(10),
  d = rnorm(10)
)
```

### Exercise 11

Pipe `df2` into `summarise()`.

```{r modifying-multiple-c-11, exercise = TRUE}

```


```{r modifying-multiple-c-11-hint-1, eval = FALSE}
df2 |> 
  summarise()
```

### 

Remember that when it comes to grouping, we always prefer `.by` instead of `group_by` where possible.

### Exercise 12

Add `.by = grp` to the call to `summarise()`

```{r modifying-multiple-c-12, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r modifying-multiple-c-12-hint-1, eval = FALSE}
df2 |> 
  summarise (.by = grp)
```

### 

A reminder that `.cols` function uses the same specifications as `select()`, so you can use functions like `starts_with()` and `ends_with()` to select columns based on their name.

### Exercise 13

Copy the previous code. Add `across (everything(), median)` to the call to `summarise ()`. You should have a comma between arguments to summarize.

```{r modifying-multiple-c-13, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r modifying-multiple-c-13-hint-1, eval = FALSE}
df2 |> 
  summarise (.by = grp,
             across(everything(), median))
```

### 

You should get a 2 x 5 table.

Note grouping columns (`grp` here) are not included in `across()`, because they're automatically preserved by `summarise()`.

### 

There are 2 additional selection techniques that are particularly useful for `across()`: `everything()` and `where()`.

### 

We have explored `everything()` in practice 12, it selects every (non-grouping) column.

### 

`where()` is a bit more complicated. It allows you to select columns based on their type:

`where(is.numeric)` selects all numeric columns.

`where(is.character)` selects all string columns.

`where(is.Date)` selects all date columns.

`where(is.POSIXct)` selects all date-time columns.

`where(is.logical)` selects all logical columns.

### 

You can combine these with Boolean algebra. For example:

`!where(is.numeric)` selects all non-numeric columns

`starts_with ("a") & where (is.logical)` selects all logical columns whose name starts with "a".

### 

We will now explore `.fns`, the second specification of `across()`.

This argument defines how each column will be transformed.

### 

What happens if we have some missing values in our data? We will use the exercises below to explore.

### Exercise 14

Hit "Run Code" to run the following code.

```{r modifying-multiple-c-14, exercise = TRUE}
df_miss <- tibble(
  a = rnorm_na(5, 1),
  b = rnorm_na(5, 1),
  c = rnorm_na(5, 2),
  d = rnorm(5)
)
```

### 

It's important to note that we're passing the function `.fns` to `across()`, so `across()` can call it; we're not calling it ourselves. That means the function name should never be followed by `()`.

ie. never write `median ()`.

If you forget, you'll get an error.

### Exercise 15

Pipe `df_miss` into `summarize()`

```{r modifying-multiple-c-15, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r modifying-multiple-c-15-hint-1, eval = FALSE}
df_miss |> 
  summarize ()
```

### Exercise 16

Copy the previous code. Add `across(a:d, median)` and `n = n()` to the call to `summarise ()`.

```{r modifying-multiple-c-16, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r modifying-multiple-c-16-hint-1, eval = FALSE}
df_miss |> 
  summarize (across(a:d, median), 
             n = n())
```

### 

You should get a 1 x 5 table with N/A as median for variables `a`, `b`, and `c`.

This is because `median()` propagates those missing values, giving us a suboptimal output (ie. get N/A as median).

### 

Solution?

Recall that `na. rm = TRUE` can remove the NA values. Therefore, it would be nice if we could pass along `na.rm = TRUE` to `median()`.

### Exercise 17

Pipe `df_miss` into `summarize()`

```{r modifying-multiple-c-17, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r modifying-multiple-c-17-hint-1, eval = FALSE}
df_miss |> 
  summarize()
```

### Exercise 18

Copy the previous code. Add `across()` and `n = n()` to the call to `summarise ()`.

```{r modifying-multiple-c-18, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r modifying-multiple-c-18-hint-1, eval = FALSE}
df_miss |> 
  summarise(across(),
            n = n())
```

### Exercise 19

Continue with the previous code. add argument `a:d` and `function(x) median(x, na.rm = TRUE)` to the function `across`.

```{r modifying-multiple-c-19, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r modifying-multiple-c-19-hint-1, eval = FALSE}
df_miss |> 
  summarise(across(a:d, 
                   function(x) median(x, na.rm = TRUE)),
            n = n())
```

### 

Now you should get a 1 x 5 table with all medians as numerical numbers.

Notice here, instead of calling `median()` directly, we created a new function, `function(x)` that calls `median()` with the desired arguments. This is how you can avoid error.

### Exercise 20

Use the previous code, but replace `function` with `\`

```{r modifying-multiple-c-20, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r modifying-multiple-c-20-hint-1, eval = FALSE}
df_miss |> 
  ...
    across(..., \(x) median(x, na.rm = TRUE)),
    ...
  )
```

### 

The above showcases a less verbose way by replacing `function` with `\`.

### 

To know how many values were removed from the `median()`, 2 other functions can be supplied to across(): one to compute the median, the other to count the missing values.

### Exercise 21

Copy and paste your code from exercise 18.

```{r modifying-multiple-c-21, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r modifying-multiple-c-21-hint-1, eval = FALSE}
df_miss |> 
  summarise(across(),
            n = n())
```

### Exercise 22

Add a new argument, `a:d` and `list()`, to the function `across`.

```{r modifying-multiple-c-22, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r modifying-multiple-c-22-hint-1, eval = FALSE}
df_miss |> 
  summarise(across(a:d, 
                   list()),
            n = n())
```

### 

You supply multiple functions by using a named list to `.fns`.

### Exercise 23

Add arguments `median = \(x) median(x, na.rm = TRUE)` and `n_miss = \(x) sum(is.na(x))` to the call to `list()`.

```{r modifying-multiple-c-23, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r modifying-multiple-c-23-hint-1, eval = FALSE}
df_miss |> 
  summarize( across (a:d, 
                     list (median = \(x) median(x, na.rm = TRUE),
                           n_miss = \(x) sum(is.na(x)))),
             n = n())

```

### 

Recall that `sum(is.na(x))` gives you the total number of missing values in x. Here, we also use the new function `\(x)` that calls `sum()` with the desired arguments.

### 

The last specification of `across()` is `.names` argument, which we will explore next.

### Exercise 24

This code is from Exercise 23. Hit `Run Code`.

```{r modifying-multiple-c-24, exercise = TRUE}
df_miss |> 
  summarize( across (a:d, 
                     list (median = \(x) median(x, na.rm = TRUE),
                           n_miss = \(x) sum(is.na(x)))),
             n = n())
```

### 

The result of across() is named according to the specification provided in the `.names` argument. We could specify our own if we wanted the name of the function to come first.

### Exercise 25

Add `.names = "{.fn}_{.col}"` to the call to `across()`.

```{r modifying-multiple-c-25, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r modifying-multiple-c-25-hint-1, eval = FALSE}
df_miss |> 
  summarize( across (a:d, 
                     list (median = \(x) median(x, na.rm = TRUE),
                           n_miss = \(x) sum(is.na(x))),
                     .names = "{.fn}_{.col}"),
             n = n())
```

### 

You should see a 1 x 9 table. Now, compare your table to the table produced in Exercise 24. What did you notice?

```         
Table from Exercise 24:
# > # A tibble: 1 × 9
# >   a_median a_n_miss b_median b_n_miss c_median c_n_miss d_median d_n_miss
# >      <dbl>    <int>    <dbl>    <int>    <dbl>    <int>    <dbl>    <int>
# > 1    0.139        1    -1.11        1   -0.387        2     1.15        0
```

```         
Tabble from Exercise 25:
# > # A tibble: 1 × 9
# >   median_a n_miss_a median_b n_miss_b median_c n_miss_c median_d n_miss_d
# >      <dbl>    <int>    <dbl>    <int>    <dbl>    <int>    <dbl>    <int>
# > 1    0.139        1    -1.11        1   -0.387        2     1.15        0
```

### 

You should notice that the only difference is the name of columns. ie. a_median vs. median_a.

By default, columns are named using a glue specification like {.col}\_{.fn}. `.col` is the name of the original column (ie. `a`, `b`) and `.fn` is the name of the function (ie. `median`). Hence the name of column in Exercise 24 is `a_median`.

You can use `.names` argument to supply your own glue specification. In Exercise 25, we define `.names` as "{.fn}\_{.col}", therefore the column produced has name `median_a`.

### Exercise 26

For this practice, pipe `df_miss` into `mutate()`.

```{r modifying-multiple-c-26, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r modifying-multiple-c-26-hint, eval = FALSE}
df_miss |> 
  mutate()
```

### 

The `.names` argument is particularly important when you use `across()` with `mutate()`.

### Exercise 27

Add `across(a:d, \(x) coalesce(x, 0))` to the call to `mutate()`.

```{r modifying-multiple-c-27, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r modifying-multiple-c-27-hint, eval = FALSE}
df_miss |> 
  mutate(across(a:d, 
                \(x) coalesce(x, 0)))
```

### 

Recall that argument `coalesce()` replace NAs with 0.

Note that by default, the output of `across()` is given the same names as the inputs. This means that `across()` inside of `mutate()` will replace existing columns, like in Exercise 27.

To create new columns, we need to specify `.names`.

### Exercise 28

Hit `Run Code`

```{r modifying-multiple-c-28, exercise = TRUE}
df_miss |> 
  mutate(across())
```

### Exercise 29

Add arguments `a:d`, `\(x) abs(x)` and `.names = "{.col}_abs"` to the call to `across()`. Make sure to separate each argument with comma.

```{r modifying-multiple-c-29, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r modifying-multiple-c-29-hint, eval = FALSE}
df_miss |> 
  mutate(
    across(a:d, \(x) abs(x), .names = "{.col}_abs"))
```

### 

Now, you should get a 5 x 8 table, with 4 new columns `a_abs`, `b_abs`, `c_abs` and `d_abs`.

### Exercise 30

Pipe `df_miss` into `filter()`.

```{r modifying-multiple-c-30, exercise = TRUE}

```

```{r modifying-multiple-c-30-hint, eval = FALSE}
df_miss |> 
  filter()
```

### 

`across()` is a great match for `summarize()` and `mutate()`, but it's more awkward to use with `filter()`.

We use two variants of `across()` called `if_any()` and `if_all()` with `filter()`.

### Exercise 31

Add `if_any()` to the call to `filter()`

```{r modifying-multiple-c-31, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r modifying-multiple-c-31-hint, eval = FALSE}
df_miss |> 
  filter(if_any())
```

### 

`if_any()` is the same as using `|` in-between the arguments.

### Exercise 32

Add `a:d` and `is.na` to the function `if_any()`.

```{r modifying-multiple-c-32, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r modifying-multiple-c-32-hint, eval = FALSE}
df_miss |> 
  filter(if_any(a:d, is.na))
```

### 

The code above is the same as `df_miss |> filter(is.na(a) | is.na(b) | is.na(c) | is.na(d))`.

### 

The other alternative, `if_all`, is the same as using `&`.

### Exercise 33

Copy and past the previous code. Change `if_any` to `if_all`, see what happens.

```{r modifying-multiple-c-33, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r modifying-multiple-c-33-hint, eval = FALSE}
df_miss |> 
  filter(if_any(a:d, is.na))
```

You will get a 0 x 4 table!

### Exercise 34

Hit `Run Code`.

```{r modifying-multiple-c-34, exercise = TRUE}
expand_dates <- function(df) {
  df |> 
    mutate(
      across(where(is.Date), list(year = year, month = month, day = mday))
    )
}

df_date <- tibble(
  name = c("Amy", "Bob"),
  date = ymd(c("2009-08-03", "2010-01-16"))
)
```

### 

`across()` is particularly useful to program with because it allows you to operate on multiple columns. This is an example of using lubridate functions to expand all date columns into year, month, and day columns.

### Exercise 35

Pipe `df_date` into `expand_dates()`.

```{r modifying-multiple-c-35, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r modifying-multiple-c-35-hint, eval = FALSE}
df_date |> 
  expand_dates()
```

### 

This should give you a 2 x 5 table with `date_year`, `date_month`, and `date_day` broken down.

### Exercise 36

Hit `Run Code` to run the following function.

```{r modifying-multiple-c-36, exercise = TRUE}
summarize_means <- function(df, summary_vars = where(is.numeric)) {
  df |> 
    summarize(
      across({{ summary_vars }}, \(x) mean(x, na.rm = TRUE)),
      n = n()
    )
}
```

### 

`across()` makes it easy to supply multiple columns in a single argument.

### Exercise 37

Pipe `diamonds` into `group_by(cut)`, then pipe into `summarize_means()`.

```{r modifying-multiple-c-37, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r modifying-multiple-c-37-hint, eval = FALSE}
diamonds |> 
  group_by(cut) |> 
  summarize_means()
```

### 

We mentioned that where possible, `.by()` should be sued instead of `group_by()`. However here 3we encounter a situation where it is not possible to use `.by()`.

Notice that `summarize_means()` is a self-defined function, as in Exercise 36, which was not built to incorporate `.by()`. Therefore, we stick with `group_by()` in this practice.

### Exercise 38

Now, add argument `c(carat, x:z)` to the function `summarize_means()`.

```{r modifying-multiple-c-38, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r modifying-multiple-c-38-hint, eval = FALSE}
diamonds |> 
  group_by(cut) |> 
  summarize_means(c(carat, x:z))
```

### 

Another interesting function associated with `across()` is `pivot_longer()`, which makes data sets longer by increasing the number of rows and decreasing the number of columns.

### Exercise 39

Hit `Run Code`. Pay attention to the table generated

```{r modifying-multiple-c-39, exercise = TRUE}
df |> 
  summarize(across(a:d, 
                   list(median = median, mean = mean)))
```

### 

What did you notice about this table? It's too wide and hence hard to compare values.

There's a way to compute the same values by pivoting longer and then summarizing.

### Exercise 40

Hit `Run Code` to create the following tibble.

```{r modifying-multiple-c-40, exercise = TRUE}
df_paired <- tibble(
  a_val = rnorm(10),
  a_wts = runif(10),
  b_val = rnorm(10),
  b_wts = runif(10),
  c_val = rnorm(10),
  c_wts = runif(10),
  d_val = rnorm(10),
  d_wts = runif(10)
)
```

### Exercise 41

Type in `df_paired` and hit `Run Code`. 

```{r modifying-multiple-c-41, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r modifying-multiple-c-41-hint, eval = FALSE}
df_paired
```

### 

Now you should see the tibble you have created!


### Exercise 42

Pipe `df_paired` into function `pivot_longer()`.

```{r modifying-multiple-c-42, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r modifying-multiple-c-42-hint, eval = FALSE}
df_paired |> 
  pivot_longer()
```

You will receive an error message, because we haven't put in any arguments. `cols` must select at least one column.

### 

`pivot_longer()` makes datasets longer by increasing the number of rows and decreasing the number of columns.

### Exercise 43

Add arguments `everything()` to the function `pivot_longer()`. 

```{r modifying-multiple-c-43, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r modifying-multiple-c-43-hint, eval = FALSE}
df_paired |> 
  pivot_longer(everything())
```

### 

Remember that the `cols` argument specifies which columns to pivot and `everything()` specifies that we want to pivot every column.

### Exercise 44

Add a second argument, `names_to = c("group", ".value")` to the call to `pivot_longer()`.

```{r modifying-multiple-c-44, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r modifying-multiple-c-44-hint, eval = FALSE}
df_paired |> 
  pivot_longer(everything(),
               names_to = c("group", ".value"))
```

### 

`names_to()` is used to create variables from the column names. Here, we creates two new columns: `group` and `.value`.

### Exercise 45

Finally let's add argument `names_sep = "_"` to the call to `pivot_longer()`.

```{r modifying-multiple-c-45, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r modifying-multiple-c-45-hint, eval = FALSE}
df_paired |> 
  pivot_longer(everything(),
               names_to = c("group", ".value"),
               names_sep = "_")
```

### 

The `names_sep` argument tell `pivot_longer()` what separates a `group` value from a  `.value` in each of the column names. Here, that's an underscore `_`.

### 

<!-- Quetion: how to break this part down... -->

Here's a quick summary of the above mentioned functions:

`.cols` specifies which columns you want to iterate over

`.fns` specifies what to do with each column

`.names` is used when you need additional control over the names of output columns, which is particularly important when you use `across()` with `mutate()`.

`if_any()` and `if_all()` are two important variations which work with `filter()`.

## Reading multiple files
### 

In this section, you'll learn how to use `purrr::map()` to do something to every file in a directory.

<!-- Will add some intro later - MB-->

### Exercise 1

Hit `Run Code` to open the following excel data file.

```{r reading-multiple-fil-1, exercise = TRUE}
data2019 <- readxl::read_excel("data/y2019.xlsx")
```

### 

There are many arguments to `read_excel()`, we can directly call its document name (remember to put in `""`).

Here, we are assigning this excel to new variable `data2019`.

### Exercise 2

Now, I want to look at data from year 2020, 2021 and 2022. 
Copy and past the previous code 4 times, change `"data/y2019.xlsx"` to `"data/y2020.xlsx"`, `"data/y2021.xlsx"` and `"data/y2022.xlsx"`.

```{r reading-multiple-fil-2, exercise = TRUE}

```

<button onclick="transfer_code(this)">Copy previous code</button>

```{r reading-multiple-fil-2-hint-1, eval = FALSE}
data2019 <- readxl::read_excel("...")
data2020 <- readxl::read_excel("...")
data2021 <- readxl::read_excel("...")
data2022 <- readxl::read_excel("...")
```

### 

Now, you can connect the excels together using `dplyr::bind_rows()`.

### Exercise 3

Now, you can connect the excels together using `dplyr::bind_rows()`. Type `bind_rows()` in the code chunck below, then hit `Run Code`.

```{r reading-multiple-fil-3, exercise = TRUE}

```

<button onclick="transfer_code(this)">Copy previous code</button>

```{r reading-multiple-fil-3-hint-1, eval = FALSE}
bind_rows()
```

### Exercise 4

Add `data2019`, `data2020`, `data2021` and `data2022` to the call to `bind_rows()`

```{r reading-multiple-fil-4, exercise = TRUE}

```

<button onclick="transfer_code(this)">Copy previous code</button>

```{r reading-multiple-fil-4-hint-1, eval = FALSE}
bind_rows(data2019, data2020, data2021, data2022)
```

### 

What we did above is a doable- when there's only 4 data sets. It also violates our rule for not copy-paste more than 3 times.

In the rest of this section, you will learn a better and quicker method of doing so.

### Exercise 5

Hit `Run Code`

```{r reading-multiple-fil-5, exercise = TRUE}
paths <- list.files("data/gapminder", pattern = "[.]xlsx$", full.names = TRUE)

paths
```

### 

Notice that we used function `list.files()` here, which basicallt lists the files in a directory.

Just like `across()` in the previous section, `list.files()` also have three arguments.

### 

This is the code from last exercise: `paths <- list.files("data/gapminder", pattern = "[.]xlsx$", full.names = TRUE)`

The first argument, `path`, is the directory to look in. In this case, it is `"data/gapminder"`.

`pattern` is a regular expression used to filter the file names. The most common pattern is something like `[.]xlsx$` or `[.]csv$` to find all files with a specified extension.

`full.names` determines whether or not the directory name should be included in the output. You almost always want this to be `TRUE`.

### 

Based on the 12 paths created from the last exercise, call `read_excel()` 12 times to get 12 data frames. 

Your code should look something like this:

```
gapminder_1952 <- readxl::read_excel("data/gapminder/1952.xlsx")
gapminder_1957 <- readxl::read_excel("data/gapminder/1957.xlsx")
gapminder_1962 <- readxl::read_excel("data/gapminder/1962.xlsx")
gapminder_1952 <- readxl::read_excel("data/gapminder/1967.xlsx")
gapminder_1957 <- readxl::read_excel("data/gapminder/1972.xlsx")
gapminder_1962 <- readxl::read_excel("data/gapminder/1977.xlsx")
gapminder_1952 <- readxl::read_excel("data/gapminder/1982.xlsx")
gapminder_1957 <- readxl::read_excel("data/gapminder/1987.xlsx")
gapminder_1962 <- readxl::read_excel("data/gapminder/1992.xlsx")
gapminder_1952 <- readxl::read_excel("data/gapminder/1997.xlsx")
gapminder_1957 <- readxl::read_excel("data/gapminder/2002.xlsx")
gapminder_2007 <- readxl::read_excel("data/gapminder/2007.xlsx")
```

### 

Separate excel sheets are always easier to work with once we put them into a single object. You can use `list()` to combine them together, then assign the list bto a new variable, `files`.

your code should look like this:

```
files <- list(
  readxl::read_excel("data/gapminder/1952.xlsx"),
  readxl::read_excel("data/gapminder/1957.xlsx"),
  readxl::read_excel("data/gapminder/1962.xlsx"),
  ...,
  readxl::read_excel("data/gapminder/2007.xlsx")
)

```

### Exercise 6

type `files[]` and hit `Run Code`, see what happens. Note that we are using `[]` here instead of `()`.

```{r reading-multiple-fil-6, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r reading-multiple-fil-6-hint-1, eval = FALSE}
files[]
```

### 

Remember that after you have data frames in a list, you can use `files[[i]]` to extract the ith element in the list.


### Exercise 7

Type `map()` in the code chucnk, then hit `Run Code`.

```{r reading-multiple-fil-7, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r reading-multiple-fil-7-hint-1, eval = FALSE}
map()
```

### 

The `map()` functions transform their input by applying a function to each element of a list or atomic vector and returning an object of the same length as the input.

### Exercise 8

Add argument `paths` and `readxl::read_excel` to the call to `map()`.

```{r reading-multiple-fil-8, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r reading-multiple-fil-8-hint-1, eval = FALSE}
map(paths, readxl::read_excel)
```

### 

The `map()` vector takes the format fo `map(x, f)`, it is a way to make good use of the `path` vector. 

### Exercise 9

Type `length()`, then hit `Run Code`.

```{r reading-multiple-fil-9, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r reading-multiple-fil-9-hint-1, eval = FALSE}
length()
```

This will give you an error because we haven't added any argumenyts yet.

### 

Function `length()` can get or set the length of vectors (including lists) and factors, and of any other R object for which a method has been defined.

### Exercise 10

Add `file`, the collection of tibbles we just created, to thew function `lenght()`.

```{r reading-multiple-fil-10, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r reading-multiple-fil-10-hint-1, eval = FALSE}
length(files)
```

### 

Now you know how many files are in `files`.

### Exercise 11

Run the code below to see what happens.

```{r reading-multiple-fil-11, exercise = TRUE}
files[[1]]
```

### 

This is another data structure that doesn’t display particularly compactly with str() so you might want to load it into RStudio and inspect it with View().

### Exercise 12

Type `list_rbind()` in the space below. Hit `Run Code`.

```{r reading-multiple-fil-12, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r reading-multiple-fil-12-hint-1, eval = FALSE}
list_rbind() 
```

### 

`list_rbind()` combines elements into a data frame by row-binding them together with `vctrs::vec_rbind()`.

### Exercise 13

Add `files` to the call to `list_rbind()`.

```{r reading-multiple-fil-13, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r reading-multiple-fil-13-hint-1, eval = FALSE}
list_rbind(files)
```

### 

There are two more functions that combine list elements into a single data structure than just `list_rbind()`.

1. `list_c()` combines elements into a vector by **concatenating them together** with vctrs::vec_c().

2. `list_rbind()` combines elements into a data frame by **row-binding them together** with vctrs::vec_rbind().

3. `list_cbind()` combines elements into a data frame by **column-binding them together** with vctrs::vec_cbind().

### Exercise 14

Pipe `paths` into fucntion `map()`.

```{r reading-multiple-fil-14, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r reading-multiple-fil-14-hint-1, eval = FALSE}
paths |> 
  map
```

### Exercise 15

Add `\(path) readxl::read_excel(path, n_max = 1)` into the function `map()`.

```{r reading-multiple-fil-15, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r reading-multiple-fil-15-hint-1, eval = FALSE}
paths |> 
  map(\(path) readxl::read_excel(path, n_max = 1))
```

### 

`n_max = 1` allows us to peak at the first few rows of the data.

### Exercise 16

Pipe the code above into function `list_rbind()`.

```{r reading-multiple-fil-16, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r reading-multiple-fil-16-hint-1, eval = FALSE}
paths |> 
  map(\(path) readxl::read_excel(path, n_max = 1)) |> 
  list_rbind()
```

### 

This combination of `map()` and `list_rbind()` allows all sporeasheet to be viewed at once and combined into one. Like what we did above with `n_max = 1`, you can also customize how many rows you want to see.

### 

Another important observation is that there's no "year" column indicating which row of data belong to which spreadsheet. 

### Exercise 17

Pipe `paths` into `set_names(basename)`

```{r reading-multiple-fil-17, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r reading-multiple-fil-17-hint-1, eval = FALSE}
paths |> set_names(basename) 
```

### 

Using `basename()` in the `set_names()` function, we can extract just the file name from the full path.

### Exercise 18

Pipe the previous code into `map(readxl::read_excel)`

```{r reading-multiple-fil-18, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r reading-multiple-fil-18-hint-1, eval = FALSE}
paths |> 
  set_names(basename) |> 
  map(readxl::read_excel)
```

### 

The names created in the last exercise automatically carried along by all the map functions, so the list of data frames will have those same names.

### Exercise 19

Assign a new variable `file` to the previous code.

```{r reading-multiple-fil-19, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r reading-multiple-fil-19-hint-1, eval = FALSE}
file <- paths |> 
  set_names(basename) |> 
  map(readxl::read_excel)
```

### 

What we've created above is actually shorthand for:

```
files <- list(
  "1952.xlsx" = readxl::read_excel("data/gapminder/1952.xlsx"),
  "1957.xlsx" = readxl::read_excel("data/gapminder/1957.xlsx"),
  "1962.xlsx" = readxl::read_excel("data/gapminder/1962.xlsx"),
  ...,
  "2007.xlsx" = readxl::read_excel("data/gapminder/2007.xlsx"))
```

Imagine how much time we would save!

### Exercise 20

Type `files[[""]]`, with any `year.xlsx` between the `""`

```{r reading-multiple-fil-20, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r reading-multiple-fil-20-hint-1, eval = FALSE}
files[["1962.xlsx"]]
```

### 

You should get the file of the respective year you chose! 

### Exercise 21

Copy and paste code from Exercise 18 to the space below. Hit `Run Code`.

```{r reading-multiple-fil-21, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r reading-multiple-fil-21-hint, eval = FALSE}
paths |> 
  set_names(basename) |> 
  map(readxl::read_excel)
```


### Exercise 22

Pipe the code above into `list_rbind() `, then add argument `names_to = "year"` to this function.

```{r reading-multiple-fil-22, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r reading-multiple-fil-22-hint, eval = FALSE}
paths |> 
  set_names(basename) |> 
  map(readxl::read_excel) |> 
  ...(names_to = "year")
```

###

The argument `names_to` gives the name of the variable that will be created from the data stored in the column names

### Exercise 23

Pipe the previous code into `mutate()` to create a new column, with the argument `year = parse_number(year)`.

```{r reading-multiple-fil-23, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r reading-multiple-fil-23-hint, eval = FALSE}
paths |> 
  set_names(basename) |> 
  map(readxl::read_excel) |> 
  list_rbind(names_to = "year") |> 
  mutate(year = parse_number(year))
```

###

`parse_number` parses the first number it finds, dropping any non-numeric characters before the first number and all characters after the first number. The grouping mark specified by the locale is ignored inside the number.

###

Overall, we used the  `names_to` argument to `list_rbind()` to tell it to save the names into a new column called year then use `readr::parse_number()` to extract the number from the string.


### Exercise 24

Assign the previous code to the variable `gapminder`

```{r reading-multiple-fil-24, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r reading-multiple-fil-24-hint, eval = FALSE}
gapminder <- paths |> 
  set_names(basename) |> 
  map(readxl::read_excel) |> 
  list_rbind(n...) |> 
  mutate(...)
```


### Exercise 25

Type `write_csv()` in the empty code chunk below, hit `Run Code`.

```{r reading-multiple-fil-25, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r reading-multiple-fil-25-hint, eval = FALSE}
write_csv()
```

###

In brief, `write_csv()` is used to write a data frame to a delimited file.

### Exercise 26

Add arguments `gapminder` and `"gapminder.csv"` to the call to `write_csv()`

```{r reading-multiple-fil-26, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r reading-multiple-fil-26-hint, eval = FALSE}
write_csv(gapminder, "gapminder.csv")
```

###

This is how we save the work we have previously done, a single csv file is now created and ready for future use.

### Exercise 27

Run the following code.

```{r reading-multiple-fil-27, exercise = TRUE}
df <- read_csv(path)
```

### Exercise 28

Pipe the newly created variable `df` into `filter()`, with argument `!is.na(id)`.

```{r reading-multiple-fil-28, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r reading-multiple-fil-28-hint, eval = FALSE}
df |> 
    filter(!is.na(id))
```

###

`is.na()` is used to deal with missing values in the dataset or data frame.

### Exercise 29

Pipe the code above into `mutate()` with argument `id = tolower(id)`.
```{r reading-multiple-fil-29, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r reading-multiple-fil-29-hint, eval = FALSE}
  df |> 
    filter(!is.na(id)) |> 
    mutate(...)
```


### Exercise 30

Then use `pivot_longer()` and add arguments `jan:dec` and `names_to = "month"` to the call. Make sure to add a `,` between the two arguments.

```{r reading-multiple-fil-30, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r reading-multiple-fil-30-hint, eval = FALSE}
  df |> 
    filter(!is.na(id)) |> 
    mutate(id = tolower(id)) |> 
    pivot_longer(...)
```

###

Just a recap, `pivot_longer()` makes datasets longer by increasing the number of rows and decreasing the number of columns.

### Exercise 31

Assigned variable name `process_file` to <- `function(path)`,followed by a `{}` with the previous code. 

```{r reading-multiple-fil-31, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r reading-multiple-fil-31-hint, eval = FALSE}
process_file <- function(path) {
  df <- read_csv(path)
  
  df |> 
    filter(!is.na(id)) |> 
    mutate(id = tolower(id)) |> 
    pivot_longer(jan:dec, names_to = "month")
}
```

###

In this step, we write a function that takes a file and does all the modifications (in this case, we read a bunch of files, filter out missing values, pivot, and then combine).

### Exercise 32

Pipe `paths` into `map()`.

```{r reading-multiple-fil-32, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r reading-multiple-fil-32-hint, eval = FALSE}
paths |> 
  map()
```


### Exercise 33

Add `process_file` as argument to `map()`.

```{r reading-multiple-fil-33, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r reading-multiple-fil-33-hint, eval = FALSE}
paths |> 
  map(process_file)
```

### Exercise 34

Pipe the above into `list_rbind()`.

```{r reading-multiple-fil-34, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r reading-multiple-fil-34-hint, eval = FALSE}
paths |> 
  map(process_file) |> 
  list_rbind()
```

###

This is an example of doing one round of iteration with a complex function.



###

<!-- End of Section knowledge drop -->

Just a recap, there are three basic steps:

use `list.files()` to list all the files in a directory, then use `purrr::map()` to read each of them into a list, then use `purrr::list_rbind()` to combine them into a single data frame. We'll then discuss how you can handle situations of increasing heterogeneity, where you can't do exactly the same thing to every file.

## Summary
### 

This tutorial covered [Chapter 27: Iteration](https://r4ds.hadley.nz/iteration.html) from [*R for Data Science (2e)*](https://r4ds.hadley.nz/) by Hadley Wickham, Mine Çetinkaya-Rundel, and Garrett Grolemund.

```{r download-answers, child = system.file("child_documents/download_answers.Rmd", package = "tutorial.helpers")}
```
