---
title: Iteration
author: David Kane and Melissa Ban
tutorial:
  id: iteration
output:
  learnr::tutorial:
    progressive: yes
    allow_skip: yes
runtime: shiny_prerendered
description: 'Tutorial for Chapter 27: Iteration'
---

```{r setup, include = FALSE}
library(learnr)
library(tutorial.helpers)
library(tidyverse)

knitr::opts_chunk$set(echo = FALSE)
options(tutorial.exercise.timelimit = 60, 
        tutorial.storage = "local") 

df <- tibble(
  a = rnorm(10),
  b = rnorm(10),
  c = rnorm(10),
  d = rnorm(10)
)

df2 <- tibble(
  grp = sample(2, 10, replace = TRUE),
  a = rnorm(10),
  b = rnorm(10),
  c = rnorm(10),
  d = rnorm(10)
)


```

```{r copy-code-chunk, child = system.file("child_documents/copy_button.Rmd", package = "tutorial.helpers")}
```

```{r info-section, child = system.file("child_documents/info_section.Rmd", package = "tutorial.helpers")}
```

<!-- Read chapter 26. -->

<!-- Avoid group_by() if possible. .by/by in the next function should work. -->

## Introduction
### 

This tutorial covers [Chapter 27: Iteration](https://r4ds.hadley.nz/iteration.html) from [*R for Data Science (2e)*](https://r4ds.hadley.nz/) by Hadley Wickham, Mine Çetinkaya-Rundel, and Garrett Grolemund. 


## Modifying multiple columns
### 

### Exercise 1

Load the **tidyverse** package with the `library()` command.

```{r modifying-multiple-c-1, exercise = TRUE}

```

```{r modifying-multiple-c-1-hint-1, eval = FALSE}
library(...)
```

### 

<!-- DK: Drop some knowledge. -->

### Exercise 2

Hit "Run Code" to create the `df` tibble.

```{r modifying-multiple-c-2, exercise = TRUE}
df <- tibble(
  a = rnorm(10),
  b = rnorm(10),
  c = rnorm(10),
  d = rnorm(10)
)
```

### 

<!-- DK: Drop some knowledge. -->


### Exercise 3

Type `df` and hit "Run Code".

```{r modifying-multiple-c-3, exercise = TRUE}

```

```{r modifying-multiple-c-3-hint-1, eval = FALSE}
df
```

### 

<!-- DK: Drop some knowledge. -->

### Exercise 4

Pipe `df` to `summarise()`.

```{r modifying-multiple-c-4, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r modifying-multiple-c-4-hint-1, eval = FALSE}
df |> 
  ...()
```

Note that this creates a tibble with one row and zero columns.

### 

<!-- DK: Drop some knowledge. -->


### Exercise 5

Add `n = n()` to the call to `summarise()`.

```{r modifying-multiple-c-5, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r modifying-multiple-c-5-hint-1, eval = FALSE}
df |> 
  summarise(n = ...)
```

### 

<!-- Knowledge -->

## Summary
### 

This tutorial covered [Chapter 27: Iteration](https://r4ds.hadley.nz/iteration.html) from [*R for Data Science (2e)*](https://r4ds.hadley.nz/) by Hadley Wickham, Mine Çetinkaya-Rundel, and Garrett Grolemund. 



```{r download-answers, child = system.file("child_documents/download_answers.Rmd", package = "tutorial.helpers")}
```
