---
title: Iteration
author: David Kane and Melissa Ban
tutorial:
  id: iteration
output:
  learnr::tutorial:
    progressive: yes
    allow_skip: yes
runtime: shiny_prerendered
description: 'Tutorial for Chapter 27: Iteration'
---

```{r setup, include = FALSE}
library(learnr)
library(tutorial.helpers)
library(tidyverse)

knitr::opts_chunk$set(echo = FALSE)
options(tutorial.exercise.timelimit = 60, 
        tutorial.storage = "local") 

df <- tibble(
  a = rnorm(10),
  b = rnorm(10),
  c = rnorm(10),
  d = rnorm(10)
)

df2 <- tibble(
  grp = sample(2, 10, replace = TRUE),
  a = rnorm(10),
  b = rnorm(10),
  c = rnorm(10),
  d = rnorm(10)
)

rnorm_na <- function(n, n_na, mean = 0, sd = 1) {
  sample(c(rnorm(n - n_na, mean = mean, sd = sd), rep(NA, n_na)))
}

df_miss <- tibble(
  a = rnorm_na(5, 1),
  b = rnorm_na(5, 1),
  c = rnorm_na(5, 2),
  d = rnorm(5)
)
```

```{r copy-code-chunk, child = system.file("child_documents/copy_button.Rmd", package = "tutorial.helpers")}
```

```{r info-section, child = system.file("child_documents/info_section.Rmd", package = "tutorial.helpers")}
```

<!-- Avoid group_by() if possible. .by/by in the next function should work. -->

<!-- Don't use the same variable name for the input and output columns! Use `med_b` for the median of the variable `b`. -->

## Introduction
### 

This tutorial covers [Chapter 27: Iteration](https://r4ds.hadley.nz/iteration.html) from [*R for Data Science (2e)*](https://r4ds.hadley.nz/) by Hadley Wickham, Mine Çetinkaya-Rundel, and Garrett Grolemund. 

<!-- finish this part later. -->

## Modifying multiple columns
### 

### Exercise 1

Load the **tidyverse** package with the `library()` command.

```{r modifying-multiple-c-1, exercise = TRUE}

```

```{r modifying-multiple-c-1-hint-1, eval = FALSE}
library(...)
```

### 

Recall that we have already learnt a number of tools that perform the same action for multiple “things”:

`facet_wrap()` and `facet_grid()` draws a plot for each subset.

`group_by()` plus `summarize()` computes summary statistics for each subset.

`unnest_wider()` and `unnest_longer()` create new rows and columns for each element of a list-column.

### Exercise 2

Hit "Run Code" to create the `df` tibble.

```{r modifying-multiple-c-2, exercise = TRUE}
df <- tibble(
  a = rnorm(10),
  b = rnorm(10),
  c = rnorm(10),
  d = rnorm(10)
)
```

### 

This is a simple tibble, now you wish to count the number of observations and compute the median of every column.

<!-- Add smth abt functional programming -->

### Exercise 3

Type `df` and hit "Run Code".

```{r modifying-multiple-c-3, exercise = TRUE}

```

```{r modifying-multiple-c-3-hint-1, eval = FALSE}
df
```

### Exercise 4

Pipe `df` to `summarise()`.

```{r modifying-multiple-c-4, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r modifying-multiple-c-4-hint-1, eval = FALSE}
df |> 
  summarise()
```

Note that this creates a tibble with one row and zero columns.

### 

The result above is due to absence of arguments and values in the function `summarise`. Next, we will add arguments to fix this problem. 

### Exercise 5

Add `n = n()` to the call to `summarise()`.

```{r modifying-multiple-c-5, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r modifying-multiple-c-5-hint-1, eval = FALSE}
df |> 
  summarise(n = ...)
```

### 

Recall that this adds a new column of a variable named `n` that represents the number of observations.

### Exercise 6

Add the median of a and median of b, `med_a = median (a)` and `med_b = median (b)`, to the call to `summarise()`.

```{r modifying-multiple-c-6, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r modifying-multiple-c-6-hint-1, eval = FALSE}
df |> 
  summarise(n = n(),
            med_a = median(a),
            med_b = median(b))
```

### 

Now you have a table of 3 columns (vairbale `n`, `a`, `b`) and their corresponding median. Next, we will add variable `c` and `d` to complete the code.


### Exercise 7

Copy and paste the previous code, then add `med_c = median(c)` and `med_d = median(d)` to the call to `summarize ()`.

```{r modifying-multiple-c-7, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r modifying-multiple-c-7-hint-1, eval = FALSE}
df |> ...
  a = ...,
  b = ...,
  c = median(c),
  d = median(c),
)
```

### 

Now we've completed the code. You should see a table of 5 columns (vairbale `n`, `med_a`, `med_b`, `med_c`, `med_d`) and their corresponding median.

### 

You may find the above process repetitive- we were simply copy-pasted `med_a = median(a)` 4 times- which, violates our rule of thumb to never copy and paste more than twice. 

This will get even more tedious if you have tens or even hundreds of columns!

### 

Solution? `across()` is here to make our lifes easier.

### Exercise 8

Run the code below.

```{r modifying-multiple-c-8, exercise = TRUE}
df |> 
  summarise(n = n())
```

### Exercise 9

Copy the previous code. Now, add `across(a:d, median)` to the call to `summarise()`. Hit "Run Code," see what happens.

```{r modifying-multiple-c-9, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r modifying-multiple-c-9-hint-1, eval = FALSE}
df |> 
  summarise(n = n(),
            across (a:d, ...))
```

You should get the exact same table as in Exercise 7. 

### 

We will discuss 3 important specifications of `across()` in the following sections. 

### 

The first specification is `.cols`, used to select columns.

### Exercise 10

Let's start by running the typed-in code. Hit "Run Code".

```{r modifying-multiple-c-10, exercise = TRUE}
df2 <- tibble(
  grp = sample(2, 10, replace = TRUE),
  a = rnorm(10),
  b = rnorm(10),
  c = rnorm(10),
  d = rnorm(10)
)
```

### Exercise 11

 Pipe `df2` into `.by (grp)`.

```{r modifying-multiple-c-11, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r modifying-multiple-c-11-hint-1, eval = FALSE}
df2 |> 
  .by(grp)
```

### 

Remember that when it comes to grouping, we always prefer `.by` instead of `group_by` where possible.

### Exercise 12

Pipe the code above into `summarise()`

```{r modifying-multiple-c-12, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r modifying-multiple-c-12-hint-1, eval = FALSE}
df2 |> 
  .by (grp) |>
  summarise ()
```

### 

A reminder that `.cols` function uses the same specifications as `select()`, so you can use functions like `starts_with()` and `ends_with()` to select columns based on their name.

### Exercise 13

Copy the previous code. Add `across (everything(), median)` to the call to `summarise ()`.

```{r modifying-multiple-c-13, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r modifying-multiple-c-13-hint-1, eval = FALSE}
df2 |> 
  .by (grp) |>
  summarize (across(everything(), ...))
```

### 

You should get a 2 x 5 table.

Note grouping columns (`grp` here) are not included in across(), because they’re automatically preserved by summarize().

### 

There are 2 additional selection techniques that are particularly useful for `across()`: `everything()` and `where()`.

### 

We have explored `everything()` in practice 12, it selects every (non-grouping) column.

### 

`where()` is a bit more complicated. It allows you to select columns based on their type:

`where(is.numeric)` selects all numeric columns.

`where(is.character)` selects all string columns.

`where(is.Date)` selects all date columns.

`where(is.POSIXct)` selects all date-time columns.

`where(is.logical)` selects all logical columns.

### 

You can combine these with Boolean algebra. For example:

`!where(is.numeric)` selects all non-numeric columns

`starts_with ("a") & where (is.logical)` selects all logical columns whose name starts with “a”.

### 

We will now explore `.fns`, the second specification of `across()`.

This argument defines how each column will be transformed. 

### 

What happens if we have some missing values in our data? We will use the exercises below to explore.

### Exercise 14

Hit "Run Code" to run the following code.

```{r modifying-multiple-c-14, exercise = TRUE}
df_miss <- tibble(
  a = rnorm_na(5, 1),
  b = rnorm_na(5, 1),
  c = rnorm_na(5, 2),
  d = rnorm(5)
)
```

### 

It’s important to note that we’re passing the function `.fns` to across(), so across() can call it; we’re not calling it ourselves. That means the function name should never be followed by (). 

ie. never write `median ()`.

If you forget, you’ll get an error.

### Exercise 15

Pipe `df_miss` into  `summarize()`

```{r modifying-multiple-c-15, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r modifying-multiple-c-15-hint-1, eval = FALSE}
df_miss |> 
  summarize ()
```

### Exercise 16

Copy the previous code. Add `across(a:d, median)` and `n = n()` to the call to `summarise ()`.

```{r modifying-multiple-c-16, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r modifying-multiple-c-16-hint-1, eval = FALSE}
df_miss |> 
  summarize (across(a:d, median), 
             n = n())
```

### 

You should get a 1 x 5 table with N/A as median for variables a, b, and c.

This is because `median()` propagates those missing values, giving us a suboptimal output (ie. get N/A as median).

### 

Solution? 

Recall that `na. rm = TRUE` can remove the NA values. Therefore, it would be nice if we could pass along `na.rm = TRUE` to `median()`.

### Exercise 17

Pipe `df_miss` into  `summarize()`

```{r modifying-multiple-c-17, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r modifying-multiple-c-17-hint-1, eval = FALSE}
df_miss |> 
  summarize()
```

### Exercise 18

Copy the previous code. Add `across()` and `n = n()` to the call to `summarise ()`.

```{r modifying-multiple-c-18, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r modifying-multiple-c-18-hint-1, eval = FALSE}
df_miss |> 
  summarise(across(),
            n = n())
```

### Exercise 19

Continue with the previous code. add argument `a:d` and `function(x) median(x, na.rm = TRUE)` to the function `across`.

```{r modifying-multiple-c-19, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r modifying-multiple-c-19-hint-1, eval = FALSE}
df_miss |> 
  summarise(across(a:d, 
                   function(x) median(x, na.rm = TRUE)),
            n = n())
```

### 

Now you should get a 1 x 5 table with all medians as numerical numbers. 

Notice here, instead of calling `median()` directly, we created a new function, `function(x)` that calls `median()` with the desired arguments. This is how you can avoid error. 

### Exercise 20

Use the previous code, but replace `function` with `\`
```{r modifying-multiple-c-20, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r modifying-multiple-c-20-hint-1, eval = FALSE}
df_miss |> 
  ...
    across(..., \(x) median(x, na.rm = TRUE)),
    ...
  )
```

### 

The above showcases a less verbose way by replacing `function` with `\`.

### 

To know how many values were removed from the `median()`, 2 other functions can be supplied to across(): one to compute the median, the other to count the missing values. 

### Exercise 21

Copy and paste your code from exercise 18.

```{r modifying-multiple-c-21, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r modifying-multiple-c-21-hint-1, eval = FALSE}
df_miss |> 
  summarise(across(),
            n = n())
```

### Exercise 22

Add a new argument, `a:d` and `list()`, to the function `across`.

```{r modifying-multiple-c-22, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r modifying-multiple-c-22-hint-1, eval = FALSE}
df_miss |> 
  summarise(across(a:d, 
                   list()),
            n = n())
```

### 

You supply multiple functions by using a named list to `.fns`.

### Exercise 23

Add arguments `median = \(x) median(x, na.rm = TRUE)` and `n_miss = \(x) sum(is.na(x))` to the call to `list()`.

```{r modifying-multiple-c-23, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r modifying-multiple-c-23-hint-1, eval = FALSE}
df_miss |> 
  summarize( across (a:d, 
                     list (median = \(x) median(x, na.rm = TRUE),
                           n_miss = \(x) sum(is.na(x)))),
             n = n())

```

### 

Recall that `sum(is.na(x))` gives you the total number of missing values in x. Here, we also use the new function `\(x)` that calls `sum()` with the desired arguments.

### 

The last specification of `across()` is `.names`. argument, which we will explore next.

### Exercise 24

<!-- add later -->

```{r modifying-multiple-c-24, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r modifying-multiple-c-24-hint-1, eval = FALSE}

```

### 

The result of across() is named according to the specification provided in the `.names` argument. We could specify our own if we wanted the name of the function to come first.

### Exercise 25

<!-- add later -->

```{r modifying-multiple-c-25, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r modifying-multiple-c-25-hint-1, eval = FALSE}

```

### 

### 

<!-- add this later: If you look carefully, you might intuit that the columns are named using a glue specification (Section 15.3.2) like {.col}_{.fn} where .col is the name of the original column and .fn is the name of the function. That’s not a coincidence! As you’ll learn in the next section, you can use .names argument to supply your own glue spec. -->

### 

Here's a quick summary of the abovementioned functions:

`.cols` specifies which columns you want to iterate over

`.fns` specifies what to do with each column

`.names` is used when you need additional control over the names of output columns, which is particularly important when you use `across()` with `mutate()`. 

`if_any()` and `if_all()` are two important variations which work with `filter()`.

## Summary
### 

This tutorial covered [Chapter 27: Iteration](https://r4ds.hadley.nz/iteration.html) from [*R for Data Science (2e)*](https://r4ds.hadley.nz/) by Hadley Wickham, Mine Çetinkaya-Rundel, and Garrett Grolemund. 



```{r download-answers, child = system.file("child_documents/download_answers.Rmd", package = "tutorial.helpers")}
```
