---
title: Logical vectors
author: David Kane and Zayan Farooq
tutorial:
  id: logical-vectors
output:
  learnr::tutorial:
    progressive: yes
    allow_skip: yes
runtime: shiny_prerendered
description: 'Tutorial for Chapter 13: Logical vectors'
---

```{r setup, include = FALSE}
library(learnr)
library(tutorial.helpers)
library(tidyverse)
library(nycflights13)
knitr::opts_chunk$set(echo = FALSE)
options(tutorial.exercise.timelimit = 60, 
        tutorial.storage = "local") 
```

```{r copy-code-chunk, child = system.file("child_documents/copy_button.Rmd", package = "tutorial.helpers")}
```

```{r info-section, child = system.file("child_documents/info_section.Rmd", package = "tutorial.helpers")}
```

## Introduction
### 

This tutorial covers [Chapter 13: Logical vectors](https://r4ds.hadley.nz/logical-vectors.html) from [*R for Data Science (2e)*](https://r4ds.hadley.nz/) by Hadley Wickham, Mine Çetinkaya-Rundel, and Garrett Grolemund. You will learn how to create logical vectors with `>`, `<`, `<=`, `=>`, `==`, `!=`, and `is.na()`, how to combine them with `!`, `&`, and `|`, and how to summarize them with `any()`, `all()`, `sum()`, and `mean()`. You will also learn about the powerful `if_else()` and `case_when()` functions that allow you to return values depending on the value of a logical vector.

## Comparisons
### 

Logical vectors are the simplest type of vector because each element can only be one of three possible values: `TRUE`, `FALSE`, and `NA`. It's relatively rare to find logical vectors in your raw data, but you'll create and manipulate them in the course of almost every analysis.

### Exercise 1

Load the **tidyverse** package

```{r comparisons-1, exercise = TRUE}

```

```{r comparisons-1-hint, eval = FALSE}
library(...)
```

### 

Most of the functions you'll learn about in this chapter are provided by base R, so we don't need the tidyverse, but we'll still load it so we can use `mutate()`, `filter()`, and friends to work with data frames.

### Exercise 2

Load the **nycflights13** package.

```{r comparisons-2, exercise = TRUE}

```

```{r comparisons-2-hint, eval = FALSE}
library(...)
```

### 

You can find more information about the package with `help(package = "nycflights13")`.

### Exercise 3

Type `flights` and hit "Run Code".

```{r comparisons-3, exercise = TRUE}

```

```{r comparisons-3-hint, eval = FALSE}
flights
```

### 

A very common way to create a logical vector is via a numeric comparison with `<`, `<=`, `>`, `>=`, `!=`, and `==`.

### Exercise 4

Pipe `flights` into the `filter()` function. Within `filter()`, use `dep_time > 600` to look at only flights scheduled to depart after 6:00 AM.

```{r comparisons-4, exercise = TRUE}

```

<button onclick="transfer_code(this)">Copy previous code</button>

```{r comparisons-4-hint, eval = FALSE}
flights |> 
  filter(...)
```

### 

As well as `&` and `|`, R also has `&&` and `||`. Don't use them in dplyr functions! These are called short-circuiting operators and only ever return a single `TRUE` or `FALSE`. They're important for programming, not data science.

### Exercise 5

Use the same code as above, add `& dep_time < 2000` to the call to `filter()`.This will look at flights scheduled to depart after 6:00 AM and begore 8:00 PM

```{r comparisons-5, exercise = TRUE}

```

<button onclick="transfer_code(this)">Copy previous code</button>

```{r comparisons-5-hint, eval = FALSE}
flights |> 
  filter(dep_time > 600 & ...)
```

### 

An easy way to avoid the problem of getting your `==`s and `|`s in the right order is to use `%in%`. `x %in% y` returns a logical vector the same length as `x` that is `TRUE` whenever a value in `x` is anywhere in `y` .

### Exercise 6

Use the same code as above and add `& abs(arr_delay) < 20` to the call to `filter()`. This will filter the data out so that the absolute value of the arrival delay is less then 20 minutes

```{r comparisons-6, exercise = TRUE}

```

<button onclick="transfer_code(this)">Copy previous code</button>

```{r comparisons-6-hint, eval = FALSE}
flights |> 
  filter(dep_time > 600 & dep_time < 2000 & ...)
```

### 

Good. It's useful to know that this is a shortcut and you can explicitly create the underlying logical variables with `mutate()`

### Exercise 7

Start a new pipe with `flights`. Pipe `flights` into the `mutate()` function. Within `mutate`, use `daytime = dep_time > 600 & dep_time < 2000`. This will create a new column called `daytime` which is `TRUE` when the `dep_time` is anytime after 6:00 AM and before 8:00 PM.

```{r comparisons-7, exercise = TRUE}

```

```{r comparisons-7-hint, eval = FALSE}
flights |> 
  mutate(...)
```

### 

Just remember that any manipulation we do to a free-floating vector, you can do to a variable inside a data frame with `mutate()` and friends.

### Exercise 8

Use the same code as above and add `, approx_ontime = abs(arr_delay) < 20` to the `mutate()` function. This will create a new column named `approx_ontime` which is true when the `arr_delay` is less than 20 minutes.

```{r comparisons-8, exercise = TRUE}

```

<button onclick="transfer_code(this)">Copy previous code</button>

```{r comparisons-8-hint, eval = FALSE}
flights |> 
  mutate(daytime = dep_time > 600 & dep_time < 2000, ...)
```

### 

As well as `&` and `|`, R also has `&&` and `||`. Don’t use them in **dplyr** functions!

### Exercise 9

Copy previous code and add `,.keep = "used"` to the `mutate()` function. This keeps only the new variables and the variables used to create that in the tibble and drops all the unused variables.

```{r comparisons-9, exercise = TRUE}

```

<button onclick="transfer_code(this)">Copy previous code</button>

```{r comparisons-9-hint, eval = FALSE}
flights |> 
  mutate(daytime = dep_time > 600 & dep_time < 2000,
         approx_ontime = abs(arr_delay) < 20,...)
```

### 

`&&` and `||` are called short-circuiting operators and only ever return a single `TRUE` or `FALSE`. They’re important for programming, not data science.

### Exercise 10

Copy previous code and pipe it into `filter()`. Within `filter()` add `daytime & approx_ontime`. This will keep only the rows that meet the specified conditions.

```{r comparisons-10, exercise = TRUE}

```

<button onclick="transfer_code(this)">Copy previous code</button>

```{r comparisons-10-hint, eval = FALSE}
flights |> 
  mutate(daytime = dep_time > 600 & dep_time < 2000,
         approx_ontime = abs(arr_delay) < 20,
         .keep = "used")|>
  filter(...)
```

### 

So far, we’ve mostly created logical variables transiently within `filter()` — they are computed, used, and then thrown away. For example, the previous `filter()` finds all daytime departures that arrive roughly on time.

### Exercise 11

Set `x` to `c(1 / 49 * 49, sqrt(2) ^ 2)` using `<-`. Click "Run Code".

```{r comparisons-11, exercise = TRUE}

```

```{r comparisons-11-hint, eval = FALSE}
x <- c(...)
```

### 

Beware of using `==` with numbers. For example, it looks like this vector contains the numbers `1` and `2`. But if you test them for equality, you get `FALSE`.

### Exercise 12

Copy previous code and run `x == c(1,2)`. It should say `FALSE`.

```{r comparisons-12, exercise = TRUE}

```

<button onclick="transfer_code(this)">Copy previous code</button>

```{r comparisons-12-hint, eval = FALSE}
x <- c(1 / 49 * 49, sqrt(2) ^ 2)
x <- c(...)
```

### 

What's going on? Computers store numbers with a fixed number of decimal places so there's no way to exactly represent `1/49` or `sqrt(2)` and subsequent computations will be very slightly off.

### Exercise 13

We can see the exact values by calling `print()` with the digits argument. Copy previous code and run `print(x, digits = 16)`. This will run the values of `x` with 16 digits.

```{r comparisons-13, exercise = TRUE}

```

<button onclick="transfer_code(this)">Copy previous code</button>

```{r comparisons-13-hint, eval = FALSE}
x <- c(1 / 49 * 49, sqrt(2) ^ 2)
x <- c(1,2)
print(...)
```

### 

You can see why R defaults to rounding these numbers; they really are very close to what you expect. Now that you've seen why `==`is failing, what can you do about it? One option is to use `dplyr::near()` which ignores small differences:

### Exercise 14

Copy previous code and type `near()` on the next line. Within `near()`, add `x, c(1,2)`. This should come out as `TRUE` for both.

```{r comparisons-14, exercise = TRUE}

```

<button onclick="transfer_code(this)">Copy previous code</button>

```{r comparisons-14-hint, eval = FALSE}
x <- c(1 / 49 * 49, sqrt(2) ^ 2)
x <- c(1,2)
print(x, digits = 16)
near(x,...)
```

### 

Missing values represent the unknown so they are "contagious": almost any operation involving an unknown value will also be unknown.

### Exercise 15

Run `NA > 5` and `10 == NA`. They should both come out as `NA`

```{r comparisons-15, exercise = TRUE}

```

<button onclick="transfer_code(this)">Copy previous code</button>

```{r comparisons-15-hint, eval = FALSE}
NA > 5
10 == NA
```

### 

R normally calls print for you (i.e. `x` is a shortcut for `print(x)`), but calling it explicitly is useful if you want to provide other arguments.

### Exercise 16

Now, run `NA == NA`.This should also come out as `NA`

```{r comparisons-16, exercise = TRUE}

```

<button onclick="transfer_code(this)">Copy previous code</button>

```{r comparisons-16-hint, eval = FALSE}
NA == NA
```

### 

That is the most confusing result. It's easiest to understand why this is true if we artificially supply a little more context.

### Exercise 17

Set `age_mary <- NA`, then set `age_john <- NA` on the next line. On a new line set `age_mary == age_john`.

```{r comparisons-17, exercise = TRUE}

```

<button onclick="transfer_code(this)">Copy previous code</button>

```{r comparisons-17-hint, eval = FALSE}
age_mary <- ...
... <- NA
age_john == ...
```

### 

This should come out as `NA` because if both of their ages are unknown, then we can't know if they are the same age.

### Exercise 18

Start a new pipe with `flights`. Pipe that into the `filter()` function. Within `filter()`, add `dep_time == NA`.This will attempt to find all the rows where `dep_time` is missing.

```{r comparisons-18, exercise = TRUE}

```

<button onclick="transfer_code(this)">Copy previous code</button>

```{r comparisons-18-hint, eval = FALSE}
flights|>
  filter(...)
```

### 

The following code doesn’t work because `dep_time == NA` will yield `NA` for every single row, and `filter()` automatically drops missing values. Instead we'll need a new tool: `is.na()`.

### Exercise 19

Type `is.na()` into the code chunk. Within `is.na()`, add `c(TRUE, NA, FALSE)`

```{r comparisons-19, exercise = TRUE}

```

```{r comparisons-19-hint, eval = FALSE}
is.na(c(TRUE, NA, FALSE))
```

### 

You will see the output `FALSE` for `TRUE` and `FALSE`,but `TRUE` for `NA`. This is because `is.na()` works with any type of vector and returns `TRUE` for missing values and `FALSE` for everything else.

### Exercise 20

Start a new pipe with `flights`. Pipe `flights` into `filter()`. Within `filter()`, add `is.na(dep_time)`. This will find all the rows with a missing `dep_time`

```{r comparisons-20, exercise = TRUE}

```

```{r comparisons-20-hint, eval = FALSE}
flights |> 
  filter(...)
```

### 

`is.na()` can also be useful in `arrange()`. `arrange()` usually puts all the missing values at the end but you can override this default by first sorting by `is.na()`.

### Exercise 21

Start a new pipe with `flights`. Pipe `flights` into `filter()`. Within `filter()`, add `month == 1, day == 1`.

```{r comparisons-21, exercise = TRUE}

```

```{r comparisons-21-hint, eval = FALSE}
flights|>
  filter(...)
```

### 

The `==` operator is a comparison operator in R that checks if two values are equal. It returns `TRUE` if the values are equal and `FALSE` otherwise

### Exercise 22

Continue the pipe with `arrange()`. Within `arrange()`, add `dep_time`. This should arrange `dep_time` from least to greatest.

```{r comparisons-22, exercise = TRUE}

```

<button onclick="transfer_code(this)">Copy previous code</button>

```{r comparisons-22-hint, eval = FALSE}
flights|>
  filter(month == 1, day == 1)|>
  arrange(...)
```

### 

The `arrange()` function is used to order rows in a data set based on a column. It allows you to sort the data set in either ascending or descending order.

### Exercise 23

Within `arrange()`, put `is.na(dep_time)`.

```{r comparisons-23, exercise = TRUE}

```

<button onclick="transfer_code(this)">Copy previous code</button>

```{r comparisons-23-hint, eval = FALSE}
flights|>
  filter(month == 1, day == 1)|>
  arrange(...(dep_time))
```

### 

This checks if the `dep_time` column has missing values (`NA`). It returns a logical vector with `TRUE` for rows where `dep_time` is `NA` and `FALSE` otherwise.

### Exercise 24

Within `arrange()`, put `desc(is.na(dep_time)), dep_time`

```{r comparisons-24, exercise = TRUE}

```

<button onclick="transfer_code(this)">Copy previous code</button>

```{r comparisons-24-hint, eval = FALSE}
flights|>
  filter(month == 1, day == 1)|>
  arrange(...(is.na(dep_time)), ...)
```

### 

The `desc()` function is used to create a descending order of the logical vector obtained from `is.na(dep_time)`. This means that rows with missing `dep_time` values (`NA`) will appear first in the data frame.



## Boolean algebra
### 

Once you have multiple logical vectors, you can combine them together using Boolean algebra. In R, `&` is “and”, `|` is “or”, `!` is “not”, and `xor()` is exclusive or^2. For example, `df |> filter(!is.na(x))` finds all rows where `x` is not missing and `df |> filter(x < -10 | x > 0)` finds all rows where `x` is smaller than `-10` or bigger than `0`. Figure 13.1 shows the complete set of Boolean operations and how they work.

### Exercise 1

The rules for missing values in Boolean algebra are a little tricky to explain because they seem inconsistent at first glance. Click "Run Code".

```{r boolean-algebra-1, exercise = TRUE}
df <- tibble(x = c(TRUE, FALSE, NA))

df |> 
  mutate(
    and = x & NA,
    or = x | NA
  )
```

###

To understand what’s going on, think about `NA | TRUE` (`NA` or `TRUE`). A missing value in a logical vector means that the value could either be `TRUE` or `FALSE`. `TRUE | TRUE` and `FALSE | TRUE` are both `TRUE` because at least one of them is `TRUE`. `NA | TRUE` must also be `TRUE` because `NA` can either be `TRUE` or `FALSE`. However, `NA | FALSE` is `NA` because we don’t know if `NA` is `TRUE` or `FALSE`. Similar reasoning applies with `NA & FALSE`.

### Exercise 2

Start a new pipe with `flights`. Pipe `flights` into `filter()`. Within `filter()`, add `month == 11 | month == 12`.

```{r boolean-algebra-2, exercise = TRUE}

```

```{r boolean-algebra-2-hint, eval = FALSE}
flights|>
  filter(...)
```

###

Note that the order of operations doesn’t work like English. Take the previous code that finds all flights that departed in November or December.You might be tempted to write it like you’d say in English: “Find all flights that departed in November or December.”.

### Exercise 3

Now, replace `month == 11 | month == 12` with `month == 11 | 12`. This should do the same thing as before. 

```{r boolean-algebra-3, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r boolean-algebra-3-hint, eval = FALSE}
flights|>
  filter(...)
```

###

This code doesn’t error but it also doesn’t seem to have worked. What’s going on? Here, R first evaluates `month == 11` creating a logical vector, which we call `nov`. It computes `nov | 12`. When you use a number with a logical operator it converts everything apart from `0` to `TRUE`, so this is equivalent to `nov | TRUE` which will always be `TRUE`, so every row will be selected.

### Exercise 4

Start a new pipe with `flights`. Pipe `flights` into the `mutate()` function. Within `mutate()`, set `nov = month == 11`

```{r boolean-algebra-4, exercise = TRUE}

```

```{r boolean-algebra-4-hint, eval = FALSE}
flights|>
  mutate(...)
```

###

`nov = month == 11` creates a new variable called `nov` and assigns it the value `TRUE` for rows where the month is equal to `11`, and `FALSE` otherwise. The expression `month == 11` checks the equality between the month column and the value `11`, resulting in a logical vector.

### Exercise 5

Copy previous code and add `final = nov | 12, .keep = "used"` to the `mutate()` function.

```{r boolean-algebra-5, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r boolean-algebra-5-hint, eval = FALSE}
flights|>
  mutate(nov = month == 12, ...)
```

###

`final = nov | 12` creates another new variable called `final`. It assigns the value `TRUE` to `final` if `nov` is `TRUE` or if the value `12` is considered as `TRUE`. In R, any non-zero numeric value is considered as `TRUE` in logical operations. Therefore, the expression `nov | 12` performs a logical OR operation between the `nov` column and the value `12`, resulting in a logical vector. The `.keep = "used"` argument keeps only the new variables, and the variables that were used to make them in the tibble.

### Exercise 6

`x %in% y` returns `TRUE` if anything in `x` is in `y`. For example, type `1:12 %in% c(1,5,11)`. This makes `x` any number `1` through `11`, and `y` `1`,`5`, and `11`. Click "Run Code"

```{r boolean-algebra-6, exercise = TRUE}

```

```{r boolean-algebra-6-hint, eval = FALSE}
... %in% c(...)
```

###

An easy way to avoid the problem of getting your `==`s and `|`s in the right order is to use `%in%`. `x %in% y` returns a logical vector the same length as `x` that is `TRUE` whenever a value in `x` is anywhere in `y` .

### Exercise 7

So to find all of the flights in November and December start a new pipe with `flights`. Pipe `flights` into `filter()`, and within `filter()` add `month %in% c(11,12)`.

```{r boolean-algebra-7, exercise = TRUE}

```

```{r boolean-algebra-7-hint, eval = FALSE}
flights|>
  filter(...)
```

###

Note that `%in%` obeys different rules for `NA` to `==`, as `NA %in% NA` is `TRUE`.

### Exercise 8

Type `c(1, 2, NA) == NA`. It should say `NA` for all 3. 

```{r boolean-algebra-8, exercise = TRUE}

```


```{r boolean-algebra-8-hint, eval = FALSE}
c(...) == NA
```

###

This should output `#> [1] NA NA NA` because when comparing any value, including `NA`, with `NA` using the `==` operator, yields `NA` as the result. This is because the value of `NA` represents unknown information, so the result of any comparison involving `NA` is also unknown.

### Exercise 9

Copy previous code but instead of using `==`, use `%in%`. This checks to see if any value on the left side (`x`), is on the rights side (`y`).

```{r boolean-algebra-9, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r boolean-algebra-9-hint, eval = FALSE}
c(...) %in% NA
```

###

This should output `#> [1] FALSE FALSE  TRUE` because `NA` is the only variable on both sides.

### Exercise 10

Start a new pipe with `flights`. Pipe `flights` into `filter()`. Within `filter()`, add `dep_time %in% c(NA, 0800)` to find all rows where `dep_time` is `NA`, of `0800`.

```{r boolean-algebra-10, exercise = TRUE}

```

```{r boolean-algebra-10-hint, eval = FALSE}
flights |> 
  filter(... %in% c(...))
```

###

This should result in a tibble where there is only `NA` and `0800` values in the `dep_time` column.

## Summaries
### 

The following sections describe some useful techniques for summarizing logical vectors. As well as functions that only work specifically with logical vectors, you can also use functions that work with numeric vectors.

### Exercise 1

Start a pipe with `flights`. Pipe it into `summarize()`. Within `summarize()`, add `all_delayed = all(dep_delay <= 60)`. This will make a new column named `all_delayed` that is `TRUE` when `dep_delay` is less than or equal to `60` minutes

```{r summaries-1, exercise = TRUE}

```


```{r summaries-1-hint, eval = FALSE}
flights|>
  summarize(... = all(...))
```

###

There are two main logical summaries: `any()` and `all()`. `any(x)` is the equivalent of `|`; it’ll return `TRUE` if there are any `TRUE`’s in `x`. `all(x)` is equivalent of `&`; it’ll return `TRUE` only if all values of `x` are `TRUE`’s.

### Exercise 2

Copy previous code. Within `all()`, add `na.rm = TRUE` separated with a `,`. When `na.rm` is set to `TRUE`, it removes all `NA` values. `na.rm` is short for `na.remove`.

```{r summaries-2, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r summaries-2-hint, eval = FALSE}
flights |> 
  summarize(
    all_delayed = all(dep_delay <= 60,...)
```

###

For example, we could use `all()` and `any()` to find out if every flight was delayed on departure by at most an hour or if any flights were delayed on arrival by five hours or more. And using `group_by()` allows us to do that by day:

### Exercise 3

Copy previous code. Within `summarize()`, add `any_long_delay = any(arr_delay >= 300)`. This will make a new column named `any_long_delay` that is `TRUE` when `arr_delay` is greater than or equal to `300` minutes.

```{r summaries-3, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r summaries-3-hint, eval = FALSE}
flights |> 
  summarize(
    all_delayed = all(dep_delay <= 60, na.rm = TRUE),
    any_long_delay = any(...)
```

###

Like all summary functions, `any()` and `all()` will return `NA` if there are any missing values present, and as usual you can make the missing values go away with `na.rm = TRUE`.

### Exercise 4

Copy previous code. Add the `na.rm = TRUE` argument in `any()`.

```{r summaries-4, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r summaries-4-hint, eval = FALSE}
flights |> 
  summarize(
    all_delayed = all(dep_delay <= 60, na.rm = TRUE),
    any_long_delay = any(arr_delay >= 300),...)
```

###

`>=` is used when you want to find a variable that that is greater than or equal to a number. You could also use `<=` to find a variable that that is less than or equal to a number and `==` to find a variable that is exactly equal to a number.

### Exercise 5

Copy previous code. To the `summarize()` function, add `.by = c(year, month)`. Make sure to separate these arguments with commas. `.by = c(year, month)` will make the tibble have one row for each variation of `month` and `year`

```{r summaries-5, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r summaries-5-hint, eval = FALSE}
flights |>
  summarize(
    all_delayed = all(dep_delay <= 60, na.rm = TRUE),
    any_long_delay = any(arr_delay >= 300, na.rm = TRUE),
    .by = c(..., ...))
```

###

In most cases, however, `any()` and `all()` are a little too crude, and it would be nice to be able to get a little more detail about how many values are `TRUE` or `FALSE`. That leads us to the numeric summaries.

### Exercise 6

Add `day` as a new argument in `.by()`. Make sure to separate the arguments with commas.

```{r summaries-6, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r summaries-6-hint, eval = FALSE}
flights |>
  summarize(
    all_delayed = all(dep_delay <= 60, na.rm = TRUE),
    any_long_delay = any(arr_delay >= 300, na.rm = TRUE),
    .by = c(year, month,...))
```

###

This allows us to find out if every flight was delayed on departure by at most an hour and the number of flights that were delayed on arrival by five hours or more. This will make the resulting tibble have one row for each combination of `year`, `month`, and `day`. Basically one row for each day of the year.

### Exercise 7

Make a new pipe with `flights`. Pipe `flights` into `summarize()`. Within `summarize()`, add `all_delayed = dep_delay <= 60`.

```{r summaries-7, exercise = TRUE}

```

```{r summaries-7-hint, eval = FALSE}
flights|>
  summarize(all_delayed = ...)
```

###

This should create a column called `all_delayed` that is `TRUE` when `dep_delay` is less than or equal to `60`.

### Exercise 8

Copy previous code. Put `dep_delay <= 60` within `mean()`.

```{r summaries-8, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r summaries-8-hint, eval = FALSE}
flights|>
  summarize(all_delayed = ...(dep_delay <= 60))
```

###

This should create a column called `all_delayed` that is the mean of the rows where `dep_delay` is less than `60`

### Exercise 9

Copy previous code. Add `na.rm = TRUE` within `mean()`

```{r summaries-9, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r summaries-9-hint, eval = FALSE}
flights|>
  summarize(all_delayed = mean(dep_delay <= 60,...))
```

###

Setting `na.rm` equal to `TRUE` removes all `NA` values.

### Exercise 10

Copy previous code. Add `any_long_delay = arr_delay >= 300` as a new argument in `summarize()`. Make sure to separate the arguments with commas.

```{r summaries-10, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r summaries-10-hint, eval = FALSE}
flights|>
  summarize(all_delayed = mean(dep_delay <= 60, na.rm = TRUE),
            any_long_delay = ...)
```

###

The `mean()` function gives the proportion of `TRUE`'s of whatever is inside it.

### Exercise 11

Copy previous code. Put `arr_delay >= 300` within `sum()`

```{r summaries-11, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r summaries-11-hint, eval = FALSE}
flights|>
  summarize(all_delayed = mean(dep_delay <= 60, na.rm = TRUE),
            any_long_delay = ...(arr_delay >= 300))
```

###

The `sum()` function gives the number of `TRUE`'s of whatever is inside it.

### Exercise 12

Copy previous code. Add `na.rm = TRUE` within `sum()`

```{r summaries-12, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r summaries-12-hint, eval = FALSE}
flights|>
  summarize(all_delayed = mean(dep_delay <= 60, na.rm = TRUE),
            any_long_delay = sum(arr_delay >= 300, ...))
```

###

When you use a logical vector in a numeric context, `TRUE` becomes `1` and `FALSE` becomes `0`. This makes `sum()` and `mean()` very useful with logical vectors because `sum(x)` gives the number of `TRUE`s and `mean(x)` gives the proportion of `TRUE`s (because `mean()` is just `sum()` divided by `length()`.

### Exercise 13

Copy previous code. Add `.by = c(year, month)` as a new argument within `summarize()`

```{r summaries-13, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r summaries-13-hint, eval = FALSE}
flights|>
  summarize(all_delayed = mean(dep_delay <= 60, na.rm = TRUE),
            any_long_delay = sum(arr_delay >= 300, na.rm = TRUE),
            .by = c(...))
```

###

Make sure to put lists of values within `c()`, otherwise the code will result in an error

### Exercise 14

Copy previous code and add `day` to the `.by()` argument within `summarize()` to get rows for every single day of the year. Make sure to separate the arguments with commas.

```{r summaries-14, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r summaries-14-hint, eval = FALSE}
flights |>  
  summarize(
    all_delayed = mean(dep_delay <= 60, na.rm = TRUE),
    any_long_delay = sum(arr_delay >= 300, na.rm = TRUE),
    .by = c(year, month, ...))
```

###

This allows for us to see the proportion of flights that were delayed on departure by at most 1 hour (`60` minutes) and the number of flights that were delayed on arrival by 5 hours (`300` minutes) or more

### Exercise 15

Start a new pipe with `flights`. Pipe it into `filter`. Within `filter()`, add `arr_delay > 0` as the argument.

```{r summaries-15, exercise = TRUE}

```

```{r summaries-15-hint, eval = FALSE}
flights |>
  filter(... > 0)
```

###

There’s one final use for logical vectors in summaries: you can use a logical vector to filter a single variable to a subset of interest. This makes use of the base `[` (pronounced subset) operator, which you’ll learn more about in Section 28.2.

### Exercise 16

Continue the pipe with `summarize()`. Within `summarize()`, add the argument `behind = mean(arr_delay)`.

```{r summaries-16, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r summaries-16-hint, eval = FALSE}
flights |>
  filter(arr_delay > 0)|>
  summarize(...)
```

###

This will look at the average delay just for flights that were actually delayed. We did this by first filtering the flights, then calculating the average delay.

### Exercise 17

Copy previous code. Add the argument `n = n()` within `summarize()`. Make sure to separate arguments within a function with comma's.

```{r summaries-17, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r summaries-17-hint, eval = FALSE}
flights |>
  filter(arr_delay > 0)|>
  summarize(behind = mean(arr_delay),
            n = ...)
```

###

`n = n()` is used to calculate the count of observations where `arr_delay` is greater than `0`.

### Exercise 18

Add `.by = c(year, month, day)` to have one row for each day of the year.

```{r summaries-18, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r summaries-18-hint, eval = FALSE}
flights |>
  filter(arr_delay > 0)|>
  summarize(behind = mean(arr_delay),
            n = n(),
            .by = c(...))
```

###

This looks at the average delay just for flights that were actually delayed. We did this by first filtering the flights and then calculating the average delay:

This works, but what if we wanted to also compute the average delay for flights that arrived early? We’d need to perform a separate filter step, and then figure out how to combine the two data frames together. Instead you could use `[` to perform an inline filtering: `arr_delay[arr_delay > 0]` will yield only the positive arrival delays.

### Exercise 19

Press "Run Code". 

```{r summaries-19, exercise = TRUE}
flights|> 
  summarize(
    behind = mean(arr_delay[arr_delay > 0], na.rm = TRUE),
    ahead = mean(arr_delay[arr_delay < 0], na.rm = TRUE),
    n = n(),
    .by = c(year, month, day)
  )
```

###

Also note the difference in the group size: in the first chunk `n()` gives the number of delayed flights per day; in the second, `n()` gives the total number of flights.

## Conditional transformations
### 

### Exercise 1

Let’s begin with a simple example of labeling a numeric vector as either `“+ve”` (positive) or `“-ve”` (negative). Press "Run Code"

```{r conditional-transfor-1, exercise = TRUE}
x <- c(-3:3, NA)
if_else(x > 0, "+ve", "-ve")
```

###

This should output `#> [1] "-ve" "-ve" "-ve" "-ve" "+ve" "+ve" "+ve" NA`

### Exercise 2

There’s an optional fourth argument, missing which will be used if the input is `NA`. Add the `"???"` argument to `if_else()`.

```{r conditional-transfor-2, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r conditional-transfor-2-hint, eval = FALSE}
x <- c(-3:3, NA)
if_else(x > 0, "+ve", "-ve", ...)
```

###

This output should be `#> [1] "-ve" "-ve" "-ve" "-ve" "+ve" "+ve" "+ve" "???"`

### Exercise 3

You can also use vectors for the the `TRUE` and `FALSE` arguments. For example, this allows us to create a minimal implementation of `abs()`. Click "Run Code"

```{r conditional-transfor-3, exercise = TRUE}
x <- c(-3:3, NA)
if_else(x < 0, -x, x)
```

###

This should output `#> [1]  3  2  1  0  1  2  3 NA`. So far all the arguments have used the same vectors, but you can of course mix and match.

### Exercise 4

For example, you could implement a simple version of `coalesce()` like this. Click "Run Code".

```{r conditional-transfor-4, exercise = TRUE}
x1 <- c(NA, 1, 2, NA)
y1 <- c(3, NA, 4, 6)
if_else(is.na(x1), y1, x1)
```

###

The output should be `#> [1] 3 1 2 6`. You might have noticed a small infelicity in our labeling example above: zero is neither positive nor negative. We could resolve this by adding an additional `if_else()`.

### Exercise 5

Click "Run Code".

```{r conditional-transfor-5, exercise = TRUE}
x <- c(-3:3, NA)
if_else(x == 0, "0", if_else(x < 0, "-ve", "+ve"), "???")
```

###

This is already a little hard to read, and you can imagine it would only get harder if you have more conditions. Instead, you can switch to `dplyr::case_when()`.

### Exercise 6

Below should have the same output as above, but instead of using `if_else`, we are using `case_when()`. Click "Run Code".

```{r conditional-transfor-6, exercise = TRUE}
x <- c(-3:3, NA)
case_when(
  x == 0   ~ "0",
  x < 0    ~ "-ve", 
  x > 0    ~ "+ve",
  is.na(x) ~ "???"
)
```

###

dplyr’s `case_when()` is inspired by SQL’s `CASE` statement and provides a flexible way of performing different computations for different conditions. It has a special syntax that unfortunately looks like nothing else you’ll use in the tidyverse. It takes pairs that look like `condition ~ output`. `condition` must be a logical vector; when it’s `TRUE`, output will be used.

### Exercise 7

To explain how `case_when()` works, let’s explore some simpler cases. If none of the cases match, the output gets an `NA`. Click "Run Code"

```{r conditional-transfor-7, exercise = TRUE}
x <- c(-3:3, NA)
case_when(
  x < 0 ~ "-ve",
  x > 0 ~ "+ve"
)
```

###

The output should be `#> [1] "-ve" "-ve" "-ve" NA    "+ve" "+ve" "+ve" NA`

### Exercise 8

If you want to create a “default”/catch all value, use TRUE on the left hand side. Copy the previous code and add `TRUE ~ "???"` as a new argument. Make sure to separate the arguments witha comma.

```{r conditional-transfor-8, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r conditional-transfor-8-hint, eval = FALSE}
x <- c(-3:3, NA)
case_when(
  x < 0 ~ "-ve",
  x > 0 ~ "+ve",
  ...
)
```

###

This should output `#> [1] "-ve" "-ve" "-ve" "???" "+ve" "+ve" "+ve" "???"`

### Exercise 9

Also, note that if multiple conditions match, only the first will be used. Click "Run Code" and see.

```{r conditional-transfor-9, exercise = TRUE}
x <- c(-3:3, NA)
case_when(
  x > 0 ~ "+ve",
  x > 2 ~ "big"
)
```

###

This should output `#> [1] NA    NA    NA    NA    "+ve" "+ve" "+ve" NA` because it used the first argument. Just like with `if_else()` you can use variables on both sides of the`~` and you can mix and match variables as needed for your problem.

### Exercise 10

For example, we could use `case_when()` to provide some human readable labels for the arrival delay. Click "Run Code".

```{r conditional-transfor-10, exercise = TRUE}
flights |> 
  mutate(
    status = case_when(
      is.na(arr_delay)      ~ "cancelled",
      arr_delay < -30       ~ "very early",
      arr_delay < -15       ~ "early",
      abs(arr_delay) <= 15  ~ "on time",
      arr_delay < 60        ~ "late",
      arr_delay < Inf       ~ "very late",
    ),
    .keep = "used"
  )

```

###

Be wary when writing this sort of complex case_when() statement; my first two attempts used a mix of < and > and I kept accidentally creating overlapping conditions.

### Exercise 11

Note that both `if_else()` and `case_when()` require compatible types in the output. If they’re not compatible, you’ll see errors like this. Click "Run Code"

```{r conditional-transfor-11, exercise = TRUE}
if_else(TRUE, "a", 1)
```

###

This should output `#> Error in if_else(): #> ! Can't combine true <character> and false <double>.`

## Summary
### 

This tutorial covered [Chapter 13: Logical vectors](https://r4ds.hadley.nz/logical-vectors.html) from [*R for Data Science (2e)*](https://r4ds.hadley.nz/) by Hadley Wickham, Mine Çetinkaya-Rundel, and Garrett Grolemund. You learned how to create logical vectors with `>`, `<`, `<=`, `=>`, `==`, `!=`, and `is.na()`, how to combine them with `!`, `&`, and `|`, and how to summarize them with `any()`, `all()`, `sum()`, and `mean()`. You also learnws about the powerful `if_else()` and `case_when()` functions that allow you to return values depending on the value of a logical vector.

```{r download-answers, child = system.file("child_documents/download_answers.Rmd", package = "tutorial.helpers")}
```
