---
title: Logical vectors
author: David Kane and Zayan Farooq
tutorial:
  id: logical-vectors
output:
  learnr::tutorial:
    progressive: yes
    allow_skip: yes
runtime: shiny_prerendered
description: 'Tutorial for Chapter 13: Logical vectors'
---

```{r setup, include = FALSE}
library(learnr)
library(tutorial.helpers)
library(tidyverse)
library(nycflights13)
knitr::opts_chunk$set(echo = FALSE)
options(tutorial.exercise.timelimit = 60, 
        tutorial.storage = "local") 
```

```{r copy-code-chunk, child = system.file("child_documents/copy_button.Rmd", package = "tutorial.helpers")}
```

```{r info-section, child = system.file("child_documents/info_section.Rmd", package = "tutorial.helpers")}
```

## Introduction
### 

<!-- DK: Check chapter link for correctness. -->

This tutorial covers [Chapter 13: Logical vectors](https://r4ds.hadley.nz/logical-vectors.html) from [*R for Data Science (2e)*](https://r4ds.hadley.nz/) by Hadley Wickham, Mine Çetinkaya-Rundel, and Garrett Grolemund. You will learn how to create logical vectors with `>`, `<`, `<=`, `=>`, `==`, `!=`, and `is.na()`, how to combine them with `!`, `&`, and `|`, and how to summarize them with `any()`, `all()`, `sum()`, and `mean()`. You will also learn about the powerful `if_else()` and `case_when()` functions that allow you to return values depending on the value of a logical vector.

## Comparisons
### 

Logical vectors are the simplest type of vector because each element can only be one of three possible values: TRUE, FALSE, and NA. It's relatively rare to find logical vectors in your raw data, but you'll create and manipulate them in the course of almost every analysis.

### Exercise 1

Load the **tidyverse** package

```{r comparisons-1, exercise = TRUE}

```

```{r comparisons-1-hint, eval = FALSE}
library(...)
```

### 

Most of the functions you'll learn about in this chapter are provided by base R, so we don't need the tidyverse, but we'll still load it so we can use `mutate()`, `filter()`, and friends to work with data frames.

### Exercise 2

Load the **nycflights13** package.

```{r comparisons-2, exercise = TRUE}

```

<button onclick="transfer_code(this)">Copy previous code</button>

```{r comparisons-2-hint, eval = FALSE}
library(...)
```

### 

You can find more information about the package with `help(package = "nycflights13")`.


### Exercise 3

Type `flights` and hit "Run Code."

```{r comparisons-3, exercise = TRUE}

```

```{r comparisons-3-hint, eval = FALSE}
flights
```

### 

A very common way to create a logical vector is via a numeric comparison with `<`, `<=`, `>`, `>=`, `!=`, and `==`.

### Exercise 4

Pipe `flights` into the `filter()` function. Within `filter()`, usw `dep_time > 600` to look at only flights scheduled to depart after 6:00 AM.

```{r comparisons-4, exercise = TRUE}

```

<button onclick="transfer_code(this)">Copy previous code</button>

```{r comparisons-4-hint, eval = FALSE}
flights |> 
  filter(dep_time > 600)
```

### 

<!-- DK: Drop some knowledge. -->

### Exercise 5

Use the same code as above, add `& dep_time < 2000` to the call to `filter()`.


```{r comparisons-5, exercise = TRUE}

```

<button onclick="transfer_code(this)">Copy previous code</button>

```{r comparisons-5-hint, eval = FALSE}
flights |> 
  filter(dep_time > 600 & dep_time < 2000)
```

### 

### Exercise 6

Within `filter()`, set `abs(arr_delay)` less than 20 using `&`

```{r comparisons-6, exercise = TRUE}

```

<button onclick="transfer_code(this)">Copy previous code</button>

```{r comparisons-6-hint, eval = FALSE}
flights |> 
  filter(dep_time > 600 & dep_time < 2000 & abs(arr_delay) < 20)
```

### 


## Boolean algebra
### 

## Summaries
### 

## Conditional transformations
### 

## Summary
### 

This tutorial covered [Chapter 13: Logical vectors](https://r4ds.hadley.nz/logical-vectors.html) from [*R for Data Science (2e)*](https://r4ds.hadley.nz/) by Hadley Wickham, Mine Çetinkaya-Rundel, and Garrett Grolemund. You learned how to create logical vectors with `>`, `<`, `<=`, `=>`, `==`, `!=`, and `is.na()`, how to combine them with `!`, `&`, and `|`, and how to summarize them with `any()`, `all()`, `sum()`, and `mean()`. You also learnws about the powerful `if_else()` and `case_when()` functions that allow you to return values depending on the value of a logical vector.

```{r download-answers, child = system.file("child_documents/download_answers.Rmd", package = "tutorial.helpers")}
```
