---
title: Using AI With R
author: Luke Li
tutorial:
  id: ai-tutorial-test
output:
  learnr::tutorial:
    progressive: yes
    allow_skip: yes
runtime: shiny_prerendered
description: 'Tutorial for R for Data Science Integrated with AI (test)'
---

```{r setup, include = FALSE}
library(learnr)
library(tutorial.helpers)
library(tidyverse)

knitr::opts_chunk$set(echo = FALSE)
options(tutorial.exercise.timelimit = 60, 
        tutorial.storage = "local") 
```

## Introduction
### 

This tutorial covers [R for Datas Science (2e)](https://r4ds.hadley.nz/) by Hadley Wickham, Mine Çetinkaya-Rundel, and Garrett Grolemund. You will learn how to use the [**lubridate**](https://lubridate.tidyverse.org/) on the `flights` data from the [**nyc13flights**](https://nycflights13.tidyverse.org/) package.

## Plotting Babynames
### 

In this section we are going to focus on generating code to create a graph based on our dataset called babynames. We will be using AI to generate the code for us and then we will break down the functions within the code and how they work.

### Exercise 1

Let's load the **tidyverse** library. 

```{r creating-date-times-through-im-1, exercise = TRUE}

```

```{r creating-date-times-through-im-1-hint-1, eval = FALSE}
library(...)
```

```{r creating-date-times-through-im-1-test, include = FALSE}
library(tidyverse)
```
