---
title: Regular expressions
author: David Kane
tutorial:
  id: regular-expressions
output:
  learnr::tutorial:
    progressive: yes
    allow_skip: yes
runtime: shiny_prerendered
description: 'Tutorial for Chapter 15: Strings'
---

```{r setup, include = FALSE}
library(learnr)
library(tutorial.helpers)
library(tidyverse)
knitr::opts_chunk$set(echo = FALSE)
options(tutorial.exercise.timelimit = 60, 
        tutorial.storage = "local") 
```

```{r copy-code-chunk, child = system.file("child_documents/copy_button.Rmd", package = "tutorial.helpers")}
```

```{r info-section, child = system.file("child_documents/info_section.Rmd", package = "tutorial.helpers")}
```

<!-- DK: Knowledge drop the two links from the summary. -->


## Introduction
### 

This tutorial covers [Chapter 16: Regular expressions](https://r4ds.hadley.nz/regexps.html) from [*R for Data Science (2e)*](https://r4ds.hadley.nz/) by Hadley Wickham, Mine Çetinkaya-Rundel, and Garrett Grolemund. With the help of the [**stringr**](https://stringr.tidyverse.org/index.html) package, we use regular expressions, a concise and powerful language for describing patterns within strings. 


## Pattern basics
###

The term "regular expression" is a bit of a mouthful, so most people abbreviate it to "regex" or "regexp."

### Exercise 1


```{r pattern-basics-1, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r pattern-basics-1-hint, eval = FALSE}

```

###

### Exercise 2


```{r pattern-basics-2, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r pattern-basics-2-hint, eval = FALSE}

```

###

### Exercise 3


```{r pattern-basics-3, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r pattern-basics-3-hint, eval = FALSE}

```

###



## Key functions
###

## Pattern details
###


## Pattern control
###

## Practice
###

## Regular expressions in other places
###

## Summary
### 

This tutorial covered [Chapter 16: Regular expressions](https://r4ds.hadley.nz/regexps.html)  from [*R for Data Science (2e)*](https://r4ds.hadley.nz/) by Hadley Wickham, Mine Çetinkaya-Rundel, and Garrett Grolemund. With the help of the [**stringr**](https://stringr.tidyverse.org/index.html) package, we used regular expressions, a concise and powerful language for describing patterns within strings. 

If you want to learn more, a good place to start is [`vignette("regular-expressions", package = "stringr")`](https://stringr.tidyverse.org/articles/regular-expressions.html): it documents the full set of syntax supported by the [**stringr**](https://stringr.tidyverse.org/index.html) package. Another useful reference is `https://www.regular-expressions.info/`. It’s not R specific, but you can use it to learn about the most advanced features of regexes and how they work under the hood.

```{r download-answers, child = system.file("child_documents/download_answers.Rmd", package = "tutorial.helpers")}
```
