---
title: Layers
author: David Kane
tutorial:
  id: layers
output:
  learnr::tutorial:
    progressive: yes
    allow_skip: yes
runtime: shiny_prerendered
description: 'Tutorial for Chapter 10: Layers'
---

```{r setup, include = FALSE}
library(learnr)
library(tutorial.helpers)
library(tidyverse)
library(ggridges)
library(maps)
knitr::opts_chunk$set(echo = FALSE)
options(tutorial.exercise.timelimit = 60, 
        tutorial.storage = "local") 

nz <- map_data("nz")

car_names <- c(`f` = "Front Wheel Drive",
               `4` = "4 Wheel Drive",
               `r` = "Rear Wheel Drive")

polling_data <- read_csv("data/vis4_polling_data.csv",
              col_types = cols(createddate = col_date(),
                               approval_type = col_character(),
                               rate = col_double()))
```

```{r copy-code-chunk, child = system.file("child_documents/copy_button.Rmd", package = "tutorial.helpers")}
```

```{r info-section, child = system.file("child_documents/info_section.Rmd", package = "tutorial.helpers")}
```

<!-- This tutorial is a bit of a mess. We need to add the code from the missing sections. And then we need to decide what to do with the extra plots. -->

<!-- Note that the later material sucks! There are not enough knowledge drops.  -->

<!-- Change elements of theme.. -->

<!-- Each section should begin with two questions looking at the data, even if they are the same questions each time. print, glimpse, etc -->

<!-- Make the plots cooler! Add links to  https://ggplot2-book.org/. And https://socviz.co/. Or maybe put these in case studies. Love this one: https://socviz.co/groupfacettx.html#histograms. -->



## Introduction
### 

*This is a ROUGH, UNFINISHED, draft.*

This tutorial covers [Chapter 10: Layers](https://r4ds.hadley.nz/layers.html) from [*R for Data Science (2e)*](https://r4ds.hadley.nz/) by Hadley Wickham, Mine Çetinkaya-Rundel, and Garrett Grolemund. You will learn how to use [**ggplot2**](https://ggplot2.tidyverse.org/) and related packages to create persuasive graphics. Key functions include [`geom_smooth()`](https://ggplot2.tidyverse.org/reference/geom_smooth.html),  [`geom_histogram()`](https://ggplot2.tidyverse.org/reference/geom_histogram.html), and  [`facet_wrap()`](https://ggplot2.tidyverse.org/reference/facet_wrap.html).


## Aesthetic mappings
### 

We’ll start with a deeper dive into aesthetic mappings. 

### Exercise 1

Load up the **tidyverse** library.


```{r aesthetic-mappings-1, exercise = TRUE}

```

```{r aesthetic-mappings-1-hint-1, eval = FALSE}
library(...)
```

```{r aesthetic-mappings-1-test, include = FALSE}
library(tidyverse)
```

### 

> “The greatest value of a picture is when it forces us to notice what we never expected to see.” — John Tukey

### Exercise 2

Run `glimpse()` on `mpg`.

```{r aesthetic-mappings-2, exercise = TRUE}

```

```{r aesthetic-mappings-2-hint-1, eval = FALSE}
glimpse(...)
```

```{r aesthetic-mappings-2-test, include = FALSE}
glimpse(mpg)
```

### 

The `mpg` tibble bundled with the **ggplot2** package contains 234 observations on 38 car models. The key variables for us include:

* **displ**: A car’s engine size, in liters. A numerical variable.

* **hwy**: A car’s fuel efficiency on the highway, in miles per gallon (mpg). A car with a low fuel efficiency consumes more fuel than a car with a high fuel efficiency when they travel the same distance. A numerical variable.

* **class**: Type of car. A categorical variable.


### Exercise 3

Run `ggplot()`, setting the first argument, `data`, equal to `mpg`.

```{r aesthetic-mappings-3, exercise = TRUE}

```

```{r aesthetic-mappings-3-hint-1, eval = FALSE}
ggplot(data = ...)
```

```{r aesthetic-mappings-3-test, include = FALSE}
ggplot(data = mpg)
```

### 

Note that this just creates an empty area in which a plot can be placed. Because `data` is the first argument, we don't need to include it. We get the same result with `ggplot(mpg)`.


### Exercise 4

Continue the code by setting the `mapping` argument equal to the `aes()` function, setting the `x` argument to `displ` and the `y` argument to equal `hwy`.  

```{r aesthetic-mappings-4, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r aesthetic-mappings-4-hint-1, eval = FALSE}
ggplot(mpg, aes(x = ..., 
                ... = hwy))
```

```{r aesthetic-mappings-4-test, include = FALSE}
ggplot(mpg, aes(x = displ, 
                y = hwy))
```

### 

We want to visualize the relationship between the car's engine size and fuel efficiency on the highway. Because we have not added a geom, no data is plotted. But the axes are now labelled because R knows the range of the `displ` and `hwy` variables.


### Exercise 5

Add `geom_point()` --- don't forget the `+` after the call to `ggplot()` --- to see the actual data.

```{r aesthetic-mappings-5, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r aesthetic-mappings-5-hint-1, eval = FALSE}
ggplot(mpg, aes(x = displ, 
                y = hwy)) +
  ...()
```

```{r aesthetic-mappings-5-test, include = FALSE}
ggplot(mpg, aes(x = displ, 
                y = hwy)) +
  geom_point()
```

### 

A scatter plot like this is the most common way to look at data. If there is a lot of overlap among the data points, then you may want to use [`geom_jitter()`](https://ggplot2.tidyverse.org/reference/geom_jitter.html).


### Exercise 6

Set the `color` argument within `aes()` to `class`. Each point will have a color based on the `class` of the car.

```{r aesthetic-mappings-6, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r aesthetic-mappings-6-hint-1, eval = FALSE}
ggplot(mpg, aes(x = displ, 
                y = hwy, 
                color = ...)) +
  geom_point()
```

```{r aesthetic-mappings-6-test, include = FALSE}
ggplot(mpg, aes(x = displ, 
                y = hwy,
                color = class)) +
  geom_point()
```

### 

The plot is much richer, even though it isn't any larger, because we have used color to incorporate information about the type of car. Notice how the SUVs cluster in the lower right. Their engines are (mostly) bigger, and they all get more gas mileage.

### Exercise 7

Replace the `color` argument with the `shape` argument within `aes()`. See the value to `class`, as before. Each point will have a shape based on the `class` of the car.

```{r aesthetic-mappings-7, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r aesthetic-mappings-7-hint-1, eval = FALSE}
ggplot(mpg, aes(x = displ, 
                y = hwy, 
                shape = ...)) +
  geom_point()
```

```{r aesthetic-mappings-7-test, include = FALSE}
ggplot(mpg, aes(x = displ, 
                y = hwy,
                shape = class)) +
  geom_point()
```

When class is mapped to shape, we get two warnings:

````
1: The shape palette can deal with a maximum of 6 discrete values because more than 6 becomes difficult to discriminate; you have 7. Consider specifying shapes manually if you must have them.

2: Removed 62 rows containing missing values (geom_point()).
````
 
### 

Since `ggplot()` will only use six shapes at a time, by default, additional groups will go unplotted when you use the shape aesthetic. The second warning is related – there are 62 SUVs in the dataset and they’re not plotted.

### Exercise 8

Replace the `color` argument with the `size` argument within `aes()`. Set the value to `class`, as before. Each point will be sized based on the `class` of the car.

```{r aesthetic-mappings-8, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r aesthetic-mappings-8-hint-1, eval = FALSE}
ggplot(mpg, aes(x = displ, 
                y = hwy, 
                size = ...)) +
  geom_point()
```

```{r aesthetic-mappings-8-test, include = FALSE}
ggplot(mpg, aes(x = displ, 
                y = hwy,
                size = class)) +
  geom_point()
```

You should see a warning message:

````
Using size for a discrete variable is not advised. 
````

### 

Mapping an unordered discrete (categorical) variable, like `class`, to an ordered aesthetic, like `size` is generally not a good idea because it implies a ranking that does not in fact exist, hence the warning.


### Exercise 9

Replace the `color` argument with the `alpha` argument within `aes()`. Set the value to `class`, as before. Each point will have a transparency based on the `class` of the car.

```{r aesthetic-mappings-9, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r aesthetic-mappings-9-hint-1, eval = FALSE}
ggplot(mpg, aes(x = displ, 
                y = hwy, 
                alpha = ...)) +
  geom_point()
```

```{r aesthetic-mappings-9-test, include = FALSE}
ggplot(mpg, aes(x = displ, 
                y = hwy,
                alpha = class)) +
  geom_point()
```

You should see a warning message, as above

````
Using alpha for a discrete variable is not advised. 
````

### 

Once you map an aesthetic, ggplot2 takes care of the rest. It selects a reasonable scale to use with the aesthetic, and it constructs a legend that explains the mapping between levels and values. For x and y aesthetics, ggplot2 does not create a legend, but it creates an axis line with tick marks and a label. The axis line provides the same information as a legend; it explains the mapping between locations and values.


### Exercise 10

You can also set the visual properties of your geom manually as an argument of your geom function (outside of aes()) instead of relying on a variable mapping to determine the appearance. Remove the `alpha` argument from your code and add `color = "blue"` **within** the class to `geom_point()`.

```{r aesthetic-mappings-10, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r aesthetic-mappings-10-hint-1, eval = FALSE}
ggplot(mpg, aes(x = ..., 
                y = hwy)) + 
  geom_point(color = "...")
```

```{r aesthetic-mappings-10-test, include = FALSE}
ggplot(mpg, aes(x = displ, 
                y = hwy)) + 
  geom_point(color = "blue")
```

### 

Here, the color doesn’t convey information about a variable, but only changes the appearance of the plot. You’ll need to pick a value that makes sense for that aesthetic:

* The name of a color as a character string, e.g., color = "blue".
* The size of a point in mm, e.g., size = 1.
* The shape of a point as a number, e.g, shape = 1. More details [here](https://ggplot2.tidyverse.org/articles/ggplot2-specs.html?q=shape#sec:shape-spec).

## Geometric objects
### 

So far we have discussed aesthetics that we can map or set in a scatterplot, when using a point geom. You can learn more about all possible aesthetic mappings in the aesthetic specifications [vignette](https://ggplot2.tidyverse.org/articles/ggplot2-specs.html).

### Exercise 1

To change the geom in your plot, change the geom function that you add to `ggplot()`. Let's start with the plot which we were just working with, but with `color = "blue"` removed from the call to `geom_point()`. 

```{r geometric-objects-1, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r geometric-objects-1-hint-1, eval = FALSE}
ggplot(..., aes(x = displ, 
                y = hwy)) + 
  geom_point()
```

```{r geometric-objects-1-test, include = FALSE}
ggplot(mpg, aes(x = displ, 
                y = hwy)) + 
  geom_point()
```

### 

Every geom function in **ggplot2** takes a mapping argument, either defined locally in the geom layer or globally in the `ggplot()` layer. 

### Exercise 2

Replace `geom_point()` with `geom_smooth()`. Both plots contain the same `x` variable, the same `y` variable, and both describe the same data. But the plots are not identical. Each plot uses a different geometric object, geom, to represent the data.

```{r geometric-objects-2, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r geometric-objects-2-hint-1, eval = FALSE}
ggplot(mpg, aes(x = displ, 
                y = hwy)) + 
  ...()
```

```{r geometric-objects-2-test, include = FALSE}
ggplot(mpg, aes(x = displ, 
                y = hwy)) + 
  geom_smooth()
```

### 

Not every aesthetic works with every geom. You could set the shape of a point, but you couldn’t set the “shape” of a line. If you try, **ggplot2** will silently ignore that aesthetic mapping. 


### Exercise 3

You can set the linetype of a line. `geom_smooth()` will draw a different line, with a different linetype, for each unique value of the variable that you map to linetype. Within the call to `aes()`, add `linetype = drv`.

```{r geometric-objects-3, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r geometric-objects-3-hint-1, eval = FALSE}
ggplot(mpg, aes(x = displ, 
                y = hwy, 
                linetype = ...)) + 
  geom_smooth()
```

```{r geometric-objects-3-test, include = FALSE}
ggplot(mpg, aes(x = displ, 
                y = hwy, 
                linetype = drv)) + 
  geom_smooth()
```

### 

Here, `geom_smooth()` separates the cars into three lines based on their drv value, which describes a car’s drive train. One line describes all of the points that have a `4` value, one line describes all of the points that have an `f` value, and one line describes all of the points that have an `r` value. Here, `4` stands for four-wheel drive, `f` for front-wheel drive, and `r` for rear-wheel drive.

### Exercise 4

Replace `linetype = drv` with `color = drv` within `aes()`.


```{r geometric-objects-4, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r geometric-objects-4-hint-1, eval = FALSE}
ggplot(mpg, aes(x = displ, 
                y = hwy, 
                ... = drv)) + 
  geom_smooth()
```

```{r geometric-objects-4-test, include = FALSE}
ggplot(mpg, aes(x = displ, 
                y = hwy, 
                color = drv)) + 
  geom_smooth()
```

### 

The `geom_smooth()` geom is applied to each category of color separately. With one call to a geom, we get three different plots, all on the same graph.

### Exercise 5

If this sounds strange, we can make it clearer by overlaying the lines on top of the raw data. To do this, simply add a call to `geom_point()` in between `ggplot()` and `geom_smooth()`. Don't forget to separate different lines with `+`.


```{r geometric-objects-5, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r geometric-objects-5-hint-1, eval = FALSE}
ggplot(mpg, aes(x = displ, 
                y = hwy, 
                color = drv)) + 
  ...() +
  geom_smooth()
```

```{r geometric-objects-5-test, include = FALSE}
ggplot(mpg, aes(x = displ, 
                y = hwy, 
                color = drv)) + 
  geom_point() +
  geom_smooth()
```

### 

We are still get the same warning as before:

> `geom_smooth()` using method = 'loess' and formula = 'y ~ x'

In essence, R is reminding you are the default values it is using for the `method` and `formula` arguments. You should probably provide those values youself, if only to assure R (and future readers of your code) that this is what you want.


### Exercise 6

The `aes()` values cascade across all the geom's, unless and until we change them ouselves. To see this, add `aes(linetype = drv)` to the call to `geom_smooth()`.

```{r geometric-objects-6, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r geometric-objects-6-hint-1, eval = FALSE}
ggplot(mpg, aes(x = displ, 
                y = hwy, 
                color = drv)) + 
  ... +
  geom_smooth(...(linetype = ...))
```

```{r geometric-objects-6-test, include = FALSE}
ggplot(mpg, aes(x = displ, 
                y = hwy, 
                color = drv)) + 
  geom_point() +
  geom_smooth(aes(linetype = drv))
```

### 

Many geoms, like `geom_smooth()`, use a single geometric object to display multiple rows of data. For these geoms, you can set the `group` aesthetic to a categorical variable to draw multiple objects. 

### Exercise 7

The explore the use of the `group` aesthetic, begin with our standard plot, which runs `geom_smooth()` with `aes(x = displ, y = hwy)` and no other arguments.

```{r geometric-objects-7, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r geometric-objects-7-hint-1, eval = FALSE}
ggplot(mpg, aes(x = displ, 
                y = hwy)) +
  ...()
```

```{r geometric-objects-7-test, include = FALSE}
ggplot(mpg, aes(x = displ, 
                y = hwy)) +
  geom_smooth()
```

### 

### Exercise 8

ggplot2 will draw a separate object for each unique value of the grouping variable. In practice, ggplot2 will automatically group the data for these geoms whenever you map an aesthetic to a discrete variable (as in the linetype example).

```{r geometric-objects-8, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r geometric-objects-8-hint-1, eval = FALSE}

```

```{r geometric-objects-8-test, include = FALSE}
ggplot(mpg, aes(x = displ, y = hwy)) +
  geom_smooth(aes(group = drv))
```

### 

 It is convenient to rely on this feature because the group aesthetic by itself does not add a legend or distinguishing features to the geoms.

### Exercise 9

```{r geometric-objects-9, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r geometric-objects-9-hint-1, eval = FALSE}

```

```{r geometric-objects-9-test, include = FALSE}
ggplot(mpg, aes(x = displ, y = hwy)) +
  geom_smooth(aes(color = drv), show.legend = FALSE)
```

### 

### Exercise 10

```{r geometric-objects-10, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r geometric-objects-10-hint-1, eval = FALSE}

```

```{r geometric-objects-10-test, include = FALSE}
ggplot(mpg, aes(x = displ, y = hwy)) + 
  geom_point(aes(color = class)) + 
  geom_smooth()
```

### 

### Exercise 11

To explore some of these concepts, let's construct this plot:

```{r}
geom_p <- ggplot(mpg, aes(x = displ, y = hwy)) + 
  geom_point() + 
  geom_point(
    data = mpg |> filter(class == "2seater"), 
    color = "red"
  ) +
  geom_point(
    data = mpg |> filter(class == "2seater"), 
    shape = "circle open", size = 3, color = "red"
  )

geom_p
```

Begin with


```{r geometric-objects-11, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r geometric-objects-11-hint-1, eval = FALSE}

```

```{r geometric-objects-11-test, include = FALSE}
ggplot(mpg, aes(x = displ, y = hwy)) + 
  geom_point()
```

### 

### Exercise 12

```{r geometric-objects-12, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r geometric-objects-12-hint-1, eval = FALSE}

```

```{r geometric-objects-12-test, include = FALSE}
ggplot(mpg, aes(x = displ, y = hwy)) + 
  geom_point() + 
  geom_point(
    data = mpg |> filter(class == "2seater"), 
    color = "red"
  )
```

### 

### Exercise 13

```{r geometric-objects-13, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r geometric-objects-13-hint-1, eval = FALSE}

```

```{r geometric-objects-13-test, include = FALSE}
ggplot(mpg, aes(x = displ, y = hwy)) + 
  geom_point() + 
  geom_point(
    data = mpg |> filter(class == "2seater"), 
    color = "red"
  ) +
  geom_point(
    data = mpg |> filter(class == "2seater"), 
    shape = "circle open", size = 3, color = "red"
  )
```

### 

Geoms are the fundamental building blocks of ggplot2. You can completely transform the look of your plot by changing its geom, and different geoms can reveal different features of your data. 

### Exercise 14

```{r geometric-objects-14, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r geometric-objects-14-hint-1, eval = FALSE}

```

```{r geometric-objects-14-test, include = FALSE}

```

### 

## Facets
### 

Facets, created with functions like `facet_wrap()` and `facet_grid()`, are the easiest way to create plot "multiples," copies of the same plot for each value of a third variable.

### Exercise 1

Start with our usual basic scatter plot, `displ` versus `hwy` from the `mpg` tibble.

```{r facets-1, exercise = TRUE}

```

```{r facets-1-hint-1, eval = FALSE}
ggplot(mpg, aes(... = displ, 
                y = ...)) + 
  ...()
```

```{r facets-1-test, include = FALSE}
ggplot(mpg, aes(x = displ, 
                y = hwy)) + 
  geom_point()
```

### 

To facet your plot by a single variable, use `facet_wrap()`. The first argument of `facet_wrap()` is a formula, which you create with `~` followed by a variable name. The variable that you pass to `facet_wrap()` should be categorical, i.e., either character or factor.


### Exercise 2

Add `facet_wrap(~cyl)` to the end of your plot code. Don't forget the `+` after `geom_point()`.


```{r facets-2, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r facets-2-hint-1, eval = FALSE}
ggplot(mpg, aes(x = displ, 
                y = hwy)) + 
  geom_point() + 
  facet_wrap(~ ...)
```

```{r facets-2-test, include = FALSE}
ggplot(mpg, aes(x = displ, 
                y = hwy)) + 
  geom_point() + 
  facet_wrap(~cyl)
```

### 

[`facet_wrap()`](https://ggplot2.tidyverse.org/reference/facet_wrap.html) wraps a 1d sequence of panels into 2d. This is generally a better use of screen space than `facet_grid()` because most displays are roughly rectangular.

### Exercise 3

To facet your plot with the combination of two variables, switch from `facet_wrap()` to `facet_grid()`. The first argument of `facet_grid()` is also a formula, but now it’s a double sided formula: rows ~ cols. 

Replace `facet_wrap(~cyl)` with `facet_grid(drv ~ cyl)`.


```{r facets-3, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r facets-3-hint-1, eval = FALSE}
ggplot(mpg, aes(x = displ, 
                y = hwy)) + 
  geom_point() + 
  facet_...(... ~ cyl)
```

```{r facets-3-test, include = FALSE}
ggplot(mpg, aes(x = displ, 
                y = hwy)) + 
  geom_point() + 
  facet_grid(drv ~ cyl)
```

### 

By default each of the facets share the same scale and range for x and y axes. This is useful when you want to compare data across facets but it can be limiting when you want to visualize the relationship within each facet better. 

### Exercise 4

Setting the scales argument in a faceting function to "free" will allow for different axis scales across both rows and columns, "free_x" will allow for different scales across rows, and "free_y" will allow for different scales across columns.

Add `scales = "free_y"` to the call to `facet_grid()`. Don't forget the comma after `drv ~ cyl`.

```{r facets-4, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r facets-4-hint-1, eval = FALSE}
ggplot(mpg, aes(x = displ, y = hwy)) + 
  geom_point() + 
  facet_grid(drv ~ cyl, ... = ...)
```

```{r facets-4-test, include = FALSE}
ggplot(mpg, aes(x = displ, y = hwy)) + 
  geom_point() + 
  facet_grid(drv ~ cyl, scales = "free_y")
```

### 

[`facet_wrap()`](https://ggplot2.tidyverse.org/reference/facet_wrap.html) and 
[`facet_grid()`](https://ggplot2.tidyverse.org/reference/facet_grid.html) include lots of useful arguments, but note that the names of those arguments are not always consistent, for historical reasons. Consider `nrow` from `facet_wrap()` versus `rows` from `facet_grid()`.

## Statistical transformations
### 

### Exercise 1

```{r statistical-transformations-1, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r statistical-transformations-1-hint-1, eval = FALSE}

```

```{r statistical-transformations-1-test, include = FALSE}
diamonds
```

### 

### Exercise 2

```{r statistical-transformations-2, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r statistical-transformations-2-hint-1, eval = FALSE}

```

```{r statistical-transformations-2-test, include = FALSE}
ggplot(diamonds, aes(x = cut)) + 
  geom_bar()
```

### 

### Exercise 3

```{r statistical-transformations-3, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r statistical-transformations-3-hint-1, eval = FALSE}

```

```{r statistical-transformations-3-test, include = FALSE}
diamonds |>
  count(cut)
```

### 

### Exercise 4

```{r statistical-transformations-4, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r statistical-transformations-4-hint-1, eval = FALSE}

```

```{r statistical-transformations-4-test, include = FALSE}
diamonds |>
  count(cut) |>
  ggplot(aes(x = cut, y = n)) +
  geom_bar(stat = "identity")
```

### 

### Exercise 5

```{r statistical-transformations-5, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r statistical-transformations-5-hint-1, eval = FALSE}

```

```{r statistical-transformations-5-test, include = FALSE}
ggplot(diamonds, aes(x = cut, y = after_stat(prop), group = 1)) + 
  geom_bar()
```

### 

<!-- ## Position adjustments -->
<!-- ###  -->

## Coordinate Systems
### 

Coordinate systems are probably the most complicated part of ggplot2. The default coordinate system is the Cartesian coordinate system where the x and y positions act independently to determine the location of each point. We are going to create one of these systems with the US.

### Exercise 1

Load the `maps` package with `library()`.

```{r coordinate-systems-1, exercise = TRUE}

```

```{r coordinate-systems-1-hint-1, eval = FALSE}
library(...)
```

### 

The `maps` package includes the area of each matching region in the map is computed, and regions which match the same element of regions have their areas combined. Each region is assumed planar, with vertices specified by the x and y components of the map object.

### Exercise 2

Run `summary()` on `nz` to view its contents.

```{r coordinate-systems-2, exercise = TRUE}

```

```{r coordinate-systems-2-hint-1, eval = FALSE}
nz <- map_data("nz")
```

### Exercise 3

We now want to plot our `nz` data with `ggplot()`

```{r coordinate-systems-3, exercise = TRUE}

```

```{r coordinate-systems-3-hint-1, eval = FALSE}
ggplot(...)
```

### 

We don't have to use `data = nz` because it is in the variable spot for data, meaning we don't have to specify `data = nz`.

### Exercise 4

Set x-axis to `long` and y-axis to `lat` and group to `group`

```{r coordinate-systems-4, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r coordinate-systems-4-hint-1, eval = FALSE}
ggplot(nz, aes(x = ..., 
               y = ...,
               group = ...))
```

### Exercise 5

Now we want to add the layer of `geom_polygon()`.

```{r coordinate-systems-5, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r coordinate-systems-5-hint-1, eval = FALSE}
... +
  ...()
```

### 

`geom_polygon()` is used to create filled polygons or regions on a plot. Commonly used to visualize geographic data, such as maps, choropleth plots, or any other data that can be represented as polygons.

### Exercise 6

Now we want to add color the item that we have plotted. Set `fill` to white and `color` to black

```{r coordinate-systems-6, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r coordinate-systems-6-hint-1, eval = FALSE}
... +
  geom_polygon(fill = ..., 
               color = ...)
```

### Exercise 7

Finally let us use the `coord_quickmap()` which sets the aspect ratio correctly for geographic maps. This is very important because we are plotting spatial data with `ggplot2`

```{r coordinate-systems-7, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r coordinate-systems-7-hint-1, eval = FALSE}
... +
  coord_quickmap()
```

### 

Now you should see the New Zealand data plotted out based on the spatial data. We can visualize more data through `coord_polar()`. Take a look at the use of [`coord_polar()`](https://r4ds.hadley.nz/layers.html#coordinate-systems)


## The layered grammar of graphics
### 

<!-- DK: There is more text, but no code, in this chapter. Maybe add some simple copy/paste questions and then include that extra text? Or maybe take one of the examples and use it here, while discussing the grammar of graphics? -->

Consider this overview:

````
ggplot(data = <DATA>) + 
  <GEOM_FUNCTION>(
     mapping = aes(<MAPPINGS>),
     stat = <STAT>, 
     position = <POSITION>
  ) +
  <COORDINATE_FUNCTION> +
  <FACET_FUNCTION>
````

Our new template takes seven parameters, the bracketed words that appear in the template. In practice, you rarely need to supply all seven parameters to make a graph because ggplot2 will provide useful defaults for everything except the data, the mappings, and the geom function.


### 

The seven parameters in the template compose the grammar of graphics, a formal system for building plots. The [grammar of graphics](https://vita.had.co.nz/papers/layered-grammar.pdf) is based on the insight that you can uniquely describe any plot as a combination of a dataset, a geom, a set of mappings, a stat, a position adjustment, a coordinate system, a faceting scheme, and a theme.

<!-- DK: Maybe turn these into a graduation exercise with one repo, one qmd which contains the three plots, then publish them? -->

## Plot 1
### 

Let's make this scatterplot using `geom_point()`.

```{r aesthetic-mappings-advanced-0}
point_p2 <- ggplot(data = diamonds,
                   mapping = aes(x = price,
                                 y = carat,
                                 color = clarity)) +
  geom_point(size = .25, alpha = .5) +
  scale_x_log10() +
  labs(title = "Diamond Carat and Price",
       subtitle = "3+ carat diamonds are rare and of low quality",
       x = "Price",
       y =  "Carat",
       color = "Clarity",
       caption = "Source: R datasets")

point_p2
```

### Exercise 1

Type `?diamonds` to look at the help page for the `diamonds` tibble. If you cannot run it in your code chunk below, run it in your Console. Remember to load `library(tidyverse)` beforehand.

```{r plot-1-1, exercise = TRUE}

```

```{r plot-1-1-hint-1, eval = FALSE}
?diamonds
```

### 

The `diamonds` dataset is a collection of data that contains information about diamonds, including their carat weight, cut, color, clarity, and price for analysis and modeling purposes.

### Exercise 2

Explore the `diamonds` dataset. Since we are looking at the "carats" and "clarity" categories, copy and paste the definitions for "carats" and "clarity" below.

```{r plot-1-2}
  question_text(NULL,
    answer(NULL, correct = TRUE),
    allow_retry = TRUE,
    try_again_button = "Edit Answer",
    incorrect = NULL,
    rows = 3)
```

### Exercise 3

Start a new pipe with `diamonds`. Select the columns `carat`, `price`, and `clarity`.

```{r plot-1-3, exercise = TRUE}

```

```{r plot-1-3-hint-1, eval = FALSE}
diamonds |>
  select(..., ..., ...)
```

### Exercise 4

Continue your pipe and add `ggplot()`. Map `price` to the x-axis and `carat` to the y-axis.

```{r plot-1-4, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r plot-1-4-hint-1, eval = FALSE}
... |>
  ggplot(mapping = aes(x = ...,
                       y = ...))
```

### 

If you want to add any themes to your plots check out: [**ggplot2**](https://ggplot2.tidyverse.org/articles/ggplot2-specs.html) contains all aesthetic specifications that are possible for the data. In addition, [**ggplot2: elegant graphics**](https://ggplot2-book.org/) provides all possible aesthetics that any user will use.

### Exercise 5

Add the `geom_point()` layer. Within `aes()` set `clarity` to `color`.

```{r plot-1-5, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r plot-1-5-hint-1, eval = FALSE}
geom_point() is a layer so use + to add it
```

```{r plot-1-5-hint-2, eval = FALSE}
... |>
  ggplot(mapping = aes(x = ...,
                       y = ...,
                       color = ...)) +
  geom_point()
```

### 

`ggplot()` contains two specific uses: color and shape. We often use color when we are dealing with a vast amount of data. When compared to color, shape can only use up to 6 discrete values, meaning that we cannot utilize shape to its fullest with more data.

### Exercise 6

There are several ways to minimize overplotting. One such way is through the use of `alpha`. In `geom_point()` set `alpha` to 0.5.

```{r plot-1-6, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r plot-1-6-hint-1, eval = FALSE}
... +
  geom_point(alpha = ...)
```

### 

With the use of alpha we can control the transparency of the dots.

### Exercise 7

Now, within `geom_point()` set `size` to .25.

```{r plot-1-7, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r plot-1-7-hint-1, eval = FALSE}
... +
  geom_point(alpha = ...,
             size = ...)
```

### 

Size controls the diameter of the points. The default diameter size is 1.

### Exercise 8

Use the layer `scale_x_log10()` function to make the scale of the plot logarithmic.

```{r plot-1-8, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r plot-1-8-hint-1, eval = FALSE}
Remembder you are adding a layers here so use +
```

```{r plot-1-8-hint-2, eval = FALSE}
... +
  scale_x_log10()
```

### 

Remember: a logarithmic scale means that each increment increases by a factor of the base of the logarithm. In our case, that's base 10. We want to implement the use of a logarithmic scale here because we want to be able to visualize all of our data without having any of it too close to each other.

### Exercise 9

To finish your plot, use `labs()` to give the graph a title, subtitle, axis labels, legend label, and caption of your choosing. To set the legend label, use the argument `color`.

Reminder: This is what your plot should look like.

```{r}
point_p2
```

```{r plot-1-9, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r plot-1-9-hint-1, eval = FALSE}
... +
  labs(...,
       color = "Clarity",
       ...)
```

### 

Now we know how to minimize the over plotting, it will come in handy in future graphs. Another way to minimize the over plotting is through `geom_jitter()`.

## Plot 2
### 

In this exercise, we will map Joe Biden's approval rating until June 30th, the first ~6 months of his term. We will explore different types of `geom_smooth()` in this section.

### 

We will be attempting to recreate this plot below.

<!-- DK: Too long subtitle. Did not explain use of "\n". Don't mention year this is from. -->

```{r}
biden_p <- ggplot(data = polling_data, 
       mapping = aes(x = createddate,
                     y = rate,
                     color = approval_type)) +
  geom_point(alpha = 0.5) + 
  geom_smooth(method = "gam",
              formula = y ~ s(x, bs = "cs"),
              se = TRUE) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
  theme(legend.position="bottom") +
  labs(x = "Month",
       y = "Percent",
       title = "Biden's Approval Rating From Jan 22 to Jun 30",
       subtitle = "While Biden's approval rating has stayed relatively static, his disapproval rating \nhas been rising.",
       color = "",
       caption = "Source: Five Thirty Eight Data")

biden_p
```

### Exercise 1

We will first explore the file that contains this data. Run `summary()` on `polling_data` in the console below.

```{r plot-2-1, exercise = TRUE}

```

```{r plot-2-1-hint-1, eval = FALSE}
summary(...)
```

### 

This data was taken from the fivethirtyeight website. 

### Exercise 2

Write down the name of the three variables.

```{r plot-2-2}
question_text(NULL,
	message = "createddate, rate, approval_type",
	answer(NULL,
	correct = TRUE),
	allow_retry = FALSE,
	incorrect = NULL,
	rows = 6)
```

### 

`createddate` stands for the day the poll was finished. `rate` is the % of approval or disapproval, and `approval_type` indicates whether or not the rate is approval or disapproval.

### Exercise 3

We will get started building the plot. Start the function `ggplot()`, and set the argument `data` to `polling_data`.

```{r plot-2-3, exercise = TRUE}

```

```{r plot-2-3-hint-1, eval = FALSE}
ggplot(data = ...)
```

### 

Remember: you should not see a plot here as you have not mapped anything, nor do you have a geom layer. 

### Exercise 4

Now, map `createddate` to the x-axis, and map `rate` to the y-axis.

```{r plot-2-4, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r plot-2-4-hint-1, eval = FALSE}
ggplot(data = polling_data,
       mapping = aes(x = ...,
                     y = ...))
```

### 

`createddate` is a date column, and `rate` is a double column.

### Exercise 5

We want to be able to see both Joe Biden's approval and disapproval rate. Therefore, we'll change the color of the dots depending on whether the dot represents an approval/disapproval rating. Set `color` to `approval_type`.

```{r plot-2-5, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r plot-2-5-hint-1, eval = FALSE}
ggplot(data = polling_data,
       mapping = aes(x = createddate,
                     y = rate,
                     color = ...))
```

### 

Every geom function in ggplot2 takes a mapping argument, either defined locally in the geom layer or globally in the `ggplot()` layer. However, not every aesthetic works with every geom. Now we're ready to start building the plot.

### Exercise 6

Next, we will add the points to the graph. Add the layer `geom_point()`.

```{r plot-2-6, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r plot-2-6-hint-1, eval = FALSE}
ggplot(data = polling_data,
       mapping = aes(x = createddate,
                     y = rate,
                     color = approval_type)) +
  ...()
```

### 

Currently, we have a bit of overplotting. We will rectify this in the next code chunk.

### Exercise 7

In the `geom_point()` layer set the argument `alpha` to 0.5.

```{r plot-2-7, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r plot-2-7-hint-1, eval = FALSE}
ggplot(data = ...,
       mapping = aes(x = ...,
                     y = ...,
                     color = ...)) +
  geom_point(alpha = ...)
```

### 

Good! Now you should be able to see more distinct points.

### Exercise 8

We want to see the trend line for Biden's approval/disapproval ratings throughout his term. Add the layer of `geom_smooth().`

```{r plot-2-8, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r plot-2-8-hint-1, eval = FALSE}
... + 
  ...()
```

### 

If you just add this layer, you will get a few warning messages. We will fix this in the following exercises. If you place mappings in a geom function, `ggplot2` will treat them as local mappings for the layer. It will use these mappings to extend or overwrite the global mappings for that layer only. This makes it possible to display different aesthetics in different layers.

### Exercise 9

Within `geom_smooth()`, set the `method` argument to "gam". We use "gam" since there is a large quantity of data points.

```{r plot-2-9, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r plot-2-9-hint-1, eval = FALSE}
... + 
  geom_smooth(method = ...)
```

### 

"gam" stands for **generalized additive model**. It is a modified form of a linear model. If you are curious to learn more, [GAMs in R](https://noamross.github.io/gams-in-r-course/chapter1) is a good resource.

### Exercise 10

Within `geom_smooth()` set `formula` to `y ~ s(x, bs = "cs")`.

```{r plot-2-10, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r plot-2-10-hint-1, eval = FALSE}
... + 
  geom_smooth(method = "gam",
              formula = ...)
```

### 

Just like `y ~ x` indicates a relationship between x and y, `y ~ s(x, bs = "cs")` indicates a "generalized additive model" relationship. With a generalized additive model we are able to handle large data.

### Exercise 11

Set `se = TRUE`.

```{r plot-2-11, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r plot-2-11-hint-1, eval = FALSE}
... + 
  geom_smooth(method = "gam",
              formula = y ~ s(x, bs = "cs"),
              se = ...)
```

### 

With `se` we are able to create a margin of error that makes our data more accurate.

### Exercise 12

Next, we want to change the labels on the graph so they are in percentage format. Use scale_y_continuous to do so. Set the `labels` argument to `scales::percent_format(accuracy = 1)`.

```{r plot-2-12, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r plot-2-12-hint-1, eval = FALSE}
...+
  scale_y_continuous(labels = ...)
```

### 

We use scale_y_continuous when we want to modify the y-axis for continuous data. [This documentation here](https://ggplot2.tidyverse.org/reference/scale_continuous.html) will tell you more, and all the possible arguments you can use with `scale_y_continuous`

### Exercise 13

Currently our legend is to the right of the graph. Let's change the way that we see our legend. Add `theme()` to the pipe and set `legend.position` to "bottom".

```{r plot-2-13, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r plot-2-13-hint-1, eval = FALSE}
+ 
  theme(...="bottom")
```

### 

With the help of `legend.position`. We have the ability to set it to the `left`, `right`, `top`, and `bottom` based on how we would like to see it within our graph.

### Exercise 14

Finally, use `labs()` to add an appropriate heading, subtitle, caption, etc. 

Reminder - your graph should look like this:

```{r}
biden_p
```

```{r plot-2-14, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r plot-2-14-hint-1, eval = FALSE}
... +
  labs(x = ...,
       y = ...,
       title = ...,
       subtitle = ...,
       color = ...,
       caption = ...)
```

### 

Good! Now you have a plot of Biden's approval rating for the first six months of his term.

## Plot 3
### 

Let's create the following histogram.

```{r}
hist_p <- mpg |> 
  mutate(year = as.character(year)) |> 
  ggplot(aes(cty, fill = year)) +
  geom_histogram(bins = 6,
                 color = "black") +
  labs(title = "Histogram of City Fuel Efficiency in 1999 and 2008 Models", 
       subtitle = "1999 had more models of high city fuel efficiency", 
       x = "City MPG", 
       y = "Count",
       fill = "Year",
       caption = "Source: EPA") 

hist_p
```

### Exercise 1

Run `?mpg` in the code chunk below. 

```{r plot-3-1, exercise = TRUE}

```

This dataset is about fuel economy of cars from the Environmental Protection Agency.

### Exercise 2

Run `glimpse()` on the mpg data set. Take a look at the different variables available.

```{r plot-3-2, exercise = TRUE}

```

```{r plot-3-2-hint-1, eval = FALSE}
glimpse(...)
```

### 

The variables that we care about are `year` and `cty`. Year indicates the year of manufacture, and `cty` indicates the city miles per gallon.

### Exercise 3

Use the `$` to extract the column `year` as a vector.

```{r plot-3-3, exercise = TRUE}

```

```{r plot-3-3-hint-1, eval = FALSE}
mpg$...
```

### 

Notice that there are only two different years, 1999 and 2008.

### Exercise 4

Start a new pipe with `mpg`. Use `mutate()` and the function `as.character()` to change the `year` column from a numeric variable to a character variable.

```{r plot-3-4, exercise = TRUE}

```

```{r plot-3-4-hint-1, eval = FALSE}
...  |> 
  mutate(... = as.character(...))
```

### 

We make this change so that we can use year as a fill color when we plot. We use the `as.character()` function to convert an object into its character representation for textual manipulation or analysis.

### Exercise 5

Select the columns `year` and `cty`.

```{r plot-3-5, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r plot-3-5-hint-1, eval = FALSE}
...  |> 
  select(..., ...)
```

### Exercise 6

Continue your pipe and add `ggplot()`. Map `cty` to the x-axis and `year` to `fill`.

```{r plot-3-6, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r plot-3-6-hint-1, eval = FALSE}
... |> 
  ggplot(aes(x = ..., fill = ...))
```

### 

We don't have to use `mapping = aes()` because it is in the variable spot for mapping, meaning we don't have to specify `mapping = aes()`. Again, you should not see anything. The `geom` layer is what displays objects on the plot.

### Exercise 7

Now, we will add a geom layer. Add the layer `geom_histogram()`. 

```{r plot-3-7, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r plot-3-7-hint-1, eval = FALSE}
geom_histogram() is a layer so use + to add it
```

```{r plot-3-7-hint-2, eval = FALSE}
...|> 
  ... +
  ...()
```

### 

You may notice that there are too many "bins" for our liking. There can be too many bins if the number chosen for a histogram or other visualizations is too high compared to the size of data, which results in loss of clarity. We'll change this now.

### Exercise 8

Within `geom_histogram()` set the number of `bins` to 6.

```{r plot-3-8, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r plot-3-8-hint-1, eval = FALSE}
... +
  geom_histogram(bins = ...)
```

You can change the width of each bin with the `binwidth` argument. We won't be doing that now, but you can do that in future plots. Read more about [**binwidth**](https://r-graphics.org/recipe-quick-histogram) and how it can be used.

### Exercise 9

Within `geom_histogram()` set  `color` to `"black"`. 

```{r plot-3-9, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r plot-3-9-hint-1, eval = FALSE}
... +
  geom_histogram(bins = 6, 
                 color = ...)
```

Remember, `color` only changes the outline of the bars. If you want to change the fill, use `fill`. 

### Exercise 10

To finish your plot, use `labs()` to give the graph a title, subtitle, axis labels, legend labels, and caption of your choosing. To change the legend labels use the argument `fill`. 

```{r plot-3-10, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r plot-3-10-hint-1, eval = FALSE}
... +
  labs(..., 
       fill = "Year",
       ...)
```

Reminder: This is what your graph should look like. 

```{r show-hist-p}
hist_p
```

<!-- DK: Fix above. -->

### 

In the previous graph we used `color` to change the legend labels, while here we used `fill`. The argument within labels matches the argument you use within `ggplot()`. `color` is used for objects like lines and points. `fill` is used for larger shapes. 

Are you satisfied with the way we have made the bin sizes? Do the axes labels make this easy to understand? We will do better later. 

[**Histograms**](https://socviz.co/groupfacettx.html#histograms) contain many more vast resources with items that contain comparing distributions in a new way.

### Exercise 11

Let's try the same data for before but with a `geom_boxplot()` instead of a `geom_histogram()`. Box plots compute the five-number summary of the distribution and then display that summary as a specially formatted box.

```{r plot-3-11, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r plot-3-11-hint-1, eval = FALSE}
... +
  ...()
```

### 

ggplot2 provides a vast amount of opportunities, however, there are a countless number of plots that can be made through different geoms and extensions. Check out [**possible options**](https://exts.ggplot2.tidyverse.org/gallery/) for any future implementations or possibilities. 

## Plot 4
### 

Let's make the same plot, using a related function, `geom_col()`. Unlike `geom_bar()`, you have to specify the value for the y-axis when using this function. 

```{r make-col-p}
col_p <- mpg |> 
  select(drv, fl, class) |>
  summarize(total = n(),
            .by = c(drv, fl, class)) |> 
  ggplot(aes(x = fl, 
             y = total, 
             fill = class)) +
    geom_col() +
    facet_wrap(~ drv,
               labeller = as_labeller(car_names)) +
    labs(title = "The Fuel Type Of Cars Based On Their Class Type",
         subtitle = "Higher Proportion of 4 Wheel and Front Wheel Drive Participants",
         x = "Fuel Type",
         y = "Number of Participants",
         fill = "Class")
col_p
```

### Exercise 1

Start a pipe with `mpg`. Select the columns `drv`, `fl` and `class`. 

```{r plot-4-1, exercise = TRUE}

```

```{r plot-4-1-hint-1, eval = FALSE}
mpg |> 
  select(..., ..., ...)
```

### Exercise 2

Copy/paste the previous code. Use `summarize()` to create a new table with the column `total` which is equal to the function `n()`. 

```{r plot-4-2, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r plot-4-2-hint-1, eval = FALSE}
... |> 
  summarize(total = ...)
```

```{r plot-4-2-hint-2, eval = FALSE}
... |> 
  summarize(total = n())
```

### Exercise 3

Add a `.by` argument to `summarize()` with the values set to `drv`, `fl`, and `class`. You need to place these three values with a vector using `c()`. 

```{r plot-4-3, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r plot-4-3-hint-1, eval = FALSE}
... |> 
  summarize(total = n(),
            .by = c(..., ..., ...))  
```

### Exercise 4

Continue your pipe with `ggplot()`. Map the variable `fl` to the x-axis. Map `total` to the `y` axis. Add the layer `geom_col()`.

```{r plot-4-4, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r plot-4-4-hint-1, eval = FALSE}
ggplot(aes(x = ..., 
           y = ...) +
  ...()
```

### 

`geom_col()` will automatically take the three to four values for each `fl` in the `total` column and stack them to form one bar. This behavior is from the fact that the `position` argument for `geom_col()` is set by default to "stack". 

### Exercise 5

Now, map `class` to `fill`.

```{r plot-4-5, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r plot-4-5-hint-1, eval = FALSE}
ggplot(data = ... , 
          aes(x = ..., 
              y = ...,
              fill = ...) +
  geom_col()
```

### Exercise 6

Add the layer `facet_wrap()` and facet by the variable `drv`. 

```{r plot-4-6, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r plot-4-6-hint-1, eval = FALSE}
... + 
  facet_wrap(~ ...)
```

### 

The `facet_wrap()` function arranges the panels in a grid, wrapping them onto multiple rows and columns. It is useful for visualizing and comparing subsets of data simultaneously.

### Exercise 7

<!-- DK: Check that this works. -->

We want to change the labels of the facet_wrap that we just created. We can do this by setting `labeller` to `as_labeller(car_names)`. Note: `car_names` has already been created for you.

```{r plot-4-7, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r plot-4-7-hint-1, eval = FALSE}
facet_wrap(~ drv,
           ... = as_labeller(...)) +
```

### 

Through the `as_labeller()` function we are able to change the labels of `facet_wrap()`.

### Exercise 8

To finish your plot, use `labs()` to give the graph a title, subtitle, axes labels, legend labels, and caption of your choosing.

Reminder: This is what your plot should look like.

```{r}
col_p
```

```{r plot-4-8, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r plot-4-8-hint-1, eval = FALSE}
... +
  labs(..., 
       fill = "...",
       ...)
```

### 

Through this we are able to view our data with a `facet_wrap` which splits a plot into subplots that each display one subset of the data based on a categorical variable.

## Plot 5
### 

Let's create this graph using `geom_bar()`.

```{r}
bar_p <- ggplot(data = midwest, 
                   mapping = aes(x = state, 
                                 fill = category)) + 
  geom_bar(alpha = 1/5, 
           position = "dodge") + 
  labs(title = "Categories found in each state",
       subtitle = "The 5 states in the midwest with each category present.",
       x = "State",
       y = "Count",
       fill = "Category")

bar_p
```

### Exercise 1

Run `glimpse()` on the `midwest` data set. 

```{r plot-5-1, exercise = TRUE}

```

```{r plot-5-1-hint-1, eval = FALSE}
...(midwest)
```

### 

### Exercise 2

Add `ggplot()` and map `state` to the x axis of the `midwest` data and set `fill` to `category`.

```{r plot-5-2, exercise = TRUE}

```

```{r plot-5-2-hint-1, eval = FALSE}
ggplot(data = ..., 
       mapping = aes(x =..., 
                     fill =...))
```

### 

Here we have used `state` data on the x axis and the count on the y axis is automatically implemented. Within a bar graph, we only input the x variable due to the y variable automatically being created usually as count. Through this we can see count within each `state` in the `midwest` data.

### Exercise 3

Add the layer `geom_bar()`.

```{r plot-5-3, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r plot-5-3-hint-1, eval = FALSE}
... +
  geom_bar()
```

### 

`geom_bar()` displays the frequency or count of a categorical variable as bars of equal width. By default, `geom_bar()` counts the number of occurrences of each category and maps them to the height of the bars.

### Exercise 4

Within `geom_bar()`, set `alpha` to 1/5, `position` to `"dodge"`.

```{r plot-5-4, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r plot-5-4-hint-1, eval = FALSE}
... + geom_bar(alpha = ..., 
               position = ...)
```

### Exercise 5

Finally, use `labs()` to give the graph a title, subtitle, axis labels, legend labels, and caption of your choosing.

```{r plot-5-5, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r plot-5-5-hint-1, eval = FALSE}
... +
  labs(..., 
       fill = "...",
       ...)
```

Reminder: This is what your plot should look like.

```{r make-bar-p}
bar_p
```

### 

We like using bar charts here because they bin your data and then plot bin counts, the number of points that fall in each bin.

### 

For more information on other stat_summary plots besides bar graphs check out [Statistical transformations](https://r4ds.hadley.nz/layers.html#statistical-transformations) from [Chapter 10: Layers](https://r4ds.hadley.nz/layers.html) in [*R for Data Science (2e)*](https://r4ds.hadley.nz/).


## Summary
### 

This tutorial covered [Chapter 10: Layers](https://r4ds.hadley.nz/layers.html) from [*R for Data Science (2e)*](https://r4ds.hadley.nz/) by Hadley Wickham, Mine Çetinkaya-Rundel, and Garrett Grolemund. You learned how to use [**ggplot2**](https://ggplot2.tidyverse.org/) and related packages to create persuasive graphics. Key functions included [`geom_smooth()`](https://ggplot2.tidyverse.org/reference/geom_smooth.html),  [`geom_histogram()`](https://ggplot2.tidyverse.org/reference/geom_histogram.html), and  [`facet_wrap()`](https://ggplot2.tidyverse.org/reference/facet_wrap.html).

```{r download-answers, child = system.file("child_documents/download_answers.Rmd", package = "tutorial.helpers")}
```
