---
title: Iteration
author: David Kane and Melissa Ban
tutorial:
  id: iteration
output:
  learnr::tutorial:
    progressive: yes
    allow_skip: yes
runtime: shiny_prerendered
description: 'Tutorial for Chapter 26: Iteration'
---

```{r setup, include = FALSE}
library(learnr)
library(tutorial.helpers)
library(tidyverse)
library(readxl)
library(duckdb)

knitr::opts_chunk$set(echo = FALSE)
options(tutorial.exercise.timelimit = 60, 
        tutorial.storage = "local") 

df <- tibble(
  a = rnorm(10),
  b = rnorm(10),
  c = rnorm(10),
  d = rnorm(10)
)

df2 <- tibble(
  grp = sample(1:2, 10, replace = TRUE),
  a = rnorm(10),
  b = rnorm(10),
  c = rnorm(10),
  d = rnorm(10)
)

rnorm_na <- function(n, n_na, mean = 0, sd = 1) {
  sample(c(rnorm(n - n_na, mean = mean, sd = sd), rep(NA, n_na)))
}

df_miss <- tibble(
  a = rnorm_na(5, 1),
  b = rnorm_na(5, 1),
  c = rnorm_na(5, 2),
  d = rnorm(5)
)

expand_dates <- function(df) {
  df |> 
    mutate(
      across(where(is.Date), list(year = year, month = month, day = mday))
    )
}

df_date <- tibble(
  name = c("Amy", "Bob"),
  date = ymd(c("2009-08-03", "2010-01-16"))
)

summarize_means <- function(df, summary_vars = where(is.numeric)) {
  df |> 
    summarize(
      across({{ summary_vars }}, \(x) mean(x, na.rm = TRUE)),
      n = n()
    )
}

long <- df |> 
  pivot_longer(a:d) |> 
  group_by(name) |> 
  summarize(
    median = median(value),
    mean = mean(value))

df_paired <- tibble(
  a_val = rnorm(10),
  a_wts = runif(10),
  b_val = rnorm(10),
  b_wts = runif(10),
  c_val = rnorm(10),
  c_wts = runif(10),
  d_val = rnorm(10),
  d_wts = runif(10)
)



gapminder_1952 <- readxl::read_excel("data/gapminder/1952.xlsx")
gapminder_1957 <- readxl::read_excel("data/gapminder/1957.xlsx")
gapminder_1962 <- readxl::read_excel("data/gapminder/1962.xlsx")
gapminder_1952 <- readxl::read_excel("data/gapminder/1967.xlsx")
gapminder_1957 <- readxl::read_excel("data/gapminder/1972.xlsx")
gapminder_1962 <- readxl::read_excel("data/gapminder/1977.xlsx")
gapminder_1952 <- readxl::read_excel("data/gapminder/1982.xlsx")
gapminder_1957 <- readxl::read_excel("data/gapminder/1987.xlsx")
gapminder_1962 <- readxl::read_excel("data/gapminder/1992.xlsx")
gapminder_1952 <- readxl::read_excel("data/gapminder/1997.xlsx")
gapminder_1957 <- readxl::read_excel("data/gapminder/2002.xlsx")
gapminder_2007 <- readxl::read_excel("data/gapminder/2007.xlsx")

paths <- list.files("data/gapminder", 
                    pattern = "[.]xlsx$", 
                    full.names = TRUE)

# Saving Multiple Outputs
template <- readxl::read_excel(paths[[1]]) |> 
  mutate(year = 1952)


# Not sure why we need to specify purrr. See chapter 27 for discussion.

files <- purrr::map(paths, readxl::read_excel)

failed <- map_vec(files, is.null)

# DK: Had to comment out the creation of this object since it was causing a
# weird error during Github checks. Let's look at it next time. Why does this
# code work with Run Document but not with R CMD check.

# gapminder <- paths |>
#   set_names(basename) |>
#   map(readxl::read_excel) |>
#   list_rbind(names_to = "year") |>
#   mutate(year = parse_number(year))

process_file <- function(path) {
  df <- read_csv(path)
  
  df |> 
    filter(!is.na(id)) |> 
    mutate(id = tolower(id)) |> 
    pivot_longer(jan:dec, names_to = "month")
}

df_types <- function(df) {
  tibble(
    col_name = names(df), 
    col_type = map_chr(df, vctrs::vec_ptype_full),
    n_miss = map_int(df, \(x) sum(is.na(x))))}

by_clarity <- 
  diamonds |> 
    group_nest(clarity) |> 
    mutate(path = str_glue("diamonds-{clarity}.csv"))

carat_histogram <- function(df) {
  ggplot(df, aes(x = carat)) + geom_histogram(binwidth = 0.1)}
```

```{r copy-code-chunk, child = system.file("child_documents/copy_button.Rmd", package = "tutorial.helpers")}
```

```{r info-section, child = system.file("child_documents/info_section.Rmd", package = "tutorial.helpers")}
```



## Introduction
### 

This tutorial covers [Chapter 26: Iteration](https://r4ds.hadley.nz/iteration.html) from [*R for Data Science (2e)*](https://r4ds.hadley.nz/) by Hadley Wickham, Mine Çetinkaya-Rundel, and Garrett Grolemund. We use the [**purrr**](https://purrr.tidyverse.org/) package to learn tools for iteration, repeatedly performing the same action on different objects. Key functions include  [`across()`](https://dplyr.tidyverse.org/reference/across.html),  [`map()`](https://purrr.tidyverse.org/reference/map.html), and user-created anonymous functions.



## Modifying multiple columns
### 

In this section, we will learn some more general tools, often called functional programming, because they are built around functions that take other functions as inputs. 

### Exercise 1

Load the **tidyverse** package with the `library()` command.

```{r modifying-multiple-columns-1, exercise = TRUE}

```

```{r modifying-multiple-columns-1-hint-1, eval = FALSE}
library(...)
```

```{r modifying-multiple-columns-1-test, include = FALSE}
library(tidyverse)
```

### 

Recall that we have already learnt a number of tools that perform the same action for multiple "things." For example, `facet_wrap()` and `facet_grid()` draw a plot for each subset of the data.


### Exercise 2

Hit "Run Code" to create the `df` tibble.

```{r modifying-multiple-columns-2, exercise = TRUE}
df <- tibble(
  a = rnorm(10),
  b = rnorm(10),
  c = rnorm(10),
  d = rnorm(10)
)
```

### 

Other examples of functions which perform the same action for multiple "things" are `group_by()` plus `summarize()`, which compute summary statistics for each subset.

### Exercise 3

Type `df` and hit "Run Code".

```{r modifying-multiple-columns-3, exercise = TRUE}

```

```{r modifying-multiple-columns-3-hint-1, eval = FALSE}
df
```

```{r modifying-multiple-columns-3-test, include = FALSE}
df
```

### 

Our last examples of functions which perform the same action for multiple "things" are `unnest_wider()` and `unnest_longer()` which create new rows and columns for each element of a list-column.


### Exercise 4

Pipe `df` to `summarise()`.

```{r modifying-multiple-columns-4, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r modifying-multiple-columns-4-hint-1, eval = FALSE}
df |> 
  ...()
```

```{r modifying-multiple-columns-4-test, include = FALSE}
df |> 
  summarise()
```

### 

Note that this creates a tibble with one row and zero columns, because you have not told summarize to create a new variable.

### Exercise 5

Add `n = n()` to the call to `summarise()`.

```{r modifying-multiple-columns-5, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r modifying-multiple-columns-5-hint-1, eval = FALSE}
df |> 
  summarise(n = ...)
```

```{r modifying-multiple-columns-5-test, include = FALSE}
df |> 
  summarise(n = n())
```

### 

Recall that this create a new column named `n` which records the number of rows in `df`.

### Exercise 6

Copy previous code. Add the median of `a` and median of `b`, `med_a = median (a)` and `med_b = median (b)`, to the call to `summarise()`. Remember the arguments to summarise must be separated by commas.

```{r modifying-multiple-columns-6, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r modifying-multiple-columns-6-hint-1, eval = FALSE}
df |> 
  summarise(n = n(),
            ... = median(a),
            med_b = ...(b))
```

```{r modifying-multiple-columns-6-test, include = FALSE}
df |> 
  summarise(n = n(),
            med_a = median(a),
            med_b = median(b))
```

### 

Now you have a tibble with 3 columns: `n`, `med_a`, and `med_b`. 

### Exercise 7

Copy previous code, then add `med_c = median(c)` and `med_d = median(d)` to the call to `summarise ()`.

```{r modifying-multiple-columns-7, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r modifying-multiple-columns-7-hint-1, eval = FALSE}
df |> 
  summarise(n = n(),
            ... = ...length(),
            med_b = median(b),
            ... = ...,
            med_d = median(...))
```

```{r modifying-multiple-columns-7-test, include = FALSE}
df |> 
  summarise(n = n(),
            med_a = median(a),
            med_b = median(b),
            med_c = median(c),
            med_d = median(d))
```

### 

You should see a tibble with five columns: `n`, `med_a`, `med_b`, `med_c`, and `med_d`.

### 

The above process is repetitive -- we were simply copy-pasted `med_a = median(a)` 4 times and then changing the variable name. This violates our rule of thumb to never copy and paste more than twice.

This would get even more tedious if you have tens or even hundreds of columns! Solution? `across()` is here to make our lives easier.

### Exercise 8

Start a new pipe by sending `df` to `summarise()`. Add `n = n()` to `summarise()`.

```{r modifying-multiple-columns-8, exercise = TRUE}

```

```{r modifying-multiple-columns-8-hint-1, eval = FALSE}
df |> 
  ...(n = n())
```

```{r modifying-multiple-columns-8-test, include = FALSE}
df |> 
  summarise(n = n())
```

### 

The `across()` function is the key trick 


### Exercise 9

Copy previous code. Add `across(a:d, median)` to the call to `summarise()`. 

```{r modifying-multiple-columns-9, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r modifying-multiple-columns-9-hint-1, eval = FALSE}
df |> 
  summarise(n = n(),
            across(..., median))
```

```{r modifying-multiple-columns-9-test, include = FALSE}
df |> 
  summarise(n = n(),
            across(a:d, median))
```

### 

`across()` makes it easy to apply the same transformation to multiple columns, allowing you to use `select()` semantics inside in "data-masking" functions like `summarise()` and `mutate()`.  Note how you get the variable names `a` through `d` "for free," without having to specify them as when using `mutate()`.


## Selecting columns with .cols
### 

We will discuss 3 important specifications of `across()` in the following sections. The first specification is `.cols`, used to select columns.

### Exercise 1

Hit "Run Code".

```{r selecting-columns-with-cols-1, exercise = TRUE}
df2 <- tibble(
  grp = sample(1:2, 10, replace = TRUE),
  a = rnorm(10),
  b = rnorm(10),
  c = rnorm(10),
  d = rnorm(10)
)
```

`grp = sample(1:2, 10, replace = TRUE)` produces a vector of length 10, containing 1s and 2s at random. 

### Exercise 2

Pipe `df2` into `summarise()`.

```{r selecting-columns-with-cols-2, exercise = TRUE}

```

```{r selecting-columns-with-cols-2-hint-1, eval = FALSE}
df2 |> 
  ...()
```

```{r selecting-columns-with-cols-2-test, include = FALSE}
df2 |> 
  summarise()
```

### 

This should output 0 rows.

Remember that when it comes to grouping, we always prefer `.by` instead of `group_by()` where possible.

### Exercise 3

Copy previous code. Add `.by = grp` to the call to `summarise()`

```{r selecting-columns-with-cols-3, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r selecting-columns-with-cols-3-hint-1, eval = FALSE}
df2 |> 
  summarise (.by = ...)
```

```{r selecting-columns-with-cols-3-test, include = FALSE}
df2 |> 
  summarise (.by = grp)
```

### 

Because `.cols` function uses the same specifications as `select()`, so you can use functions like `starts_with()` and `ends_with()` to select columns based on their name.

### Exercise 4

Copy the previous code. Add `across (everything(), median)` to the call to `summarise ()`. You should have a comma between arguments to summarize.

```{r selecting-columns-with-cols-4, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r selecting-columns-with-cols-4-hint-1, eval = FALSE}
... |> 
  summarise (.by = grp,
             across(everything(), ...))
```

```{r selecting-columns-with-cols-4-test, include = FALSE}
df2 |> 
  summarise (.by = grp,
             across(everything(), median))
```

### 

Grouping columns (`grp` here) are not included in `across()`, because they're automatically reserved by `summarise()`. There are 2 selection techniques that are particularly useful for `across()`: `everything()` and `where()`.

`everything()` selects every (non-grouping) column.

<!-- DK: Too much material. Should be spread out, or more exercises added. -->

`where()` is a bit more complicated. It allows you to select columns based on their type:

`where(is.numeric)` selects all numeric columns.

`where(is.character)` selects all string columns.

`where(is.Date)` selects all date columns.

`where(is.POSIXct)` selects all date-time columns.

`where(is.logical)` selects all logical columns.

### 

You can combine these with Boolean algebra. For example:

`!where(is.numeric)` selects all non-numeric columns

`starts_with ("a") & where (is.logical)` selects all logical columns whose name starts with "a".

### Exercise 5

We will now explore `.fns`, the second specification of `across()`. This argument defines how each column will be transformed.

Hit "Run Code".

```{r selecting-columns-with-cols-5, exercise = TRUE}
rnorm_na <- function(n, n_na, mean = 0, sd = 1) {
  sample(c(rnorm(n - n_na, mean = mean, sd = sd), rep(NA, n_na)))
}

df_miss <- tibble(
  a = rnorm_na(5, 1),
  b = rnorm_na(5, 1),
  c = rnorm_na(5, 2),
  d = rnorm(5)
)

df_miss
```

### 

<!-- DK: Since we don't use this again, maybe just define the tibble directly? -->

We created and then used the `rnorm_na()` function in order to more easily create a tibble with variables with lots of missing values.


### Exercise 6

Pipe `df_miss` into `summarise()`

```{r selecting-columns-with-cols-6, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r selecting-columns-with-cols-6-hint-1, eval = FALSE}
df_miss |> 
  ...()
```

```{r selecting-columns-with-cols-6-test, include = FALSE}
df_miss |> 
  summarise()
```

### 

It's important to note that we're passing the function `.fns` to `across()`, so `across()` can call it; we're not calling it ourselves. That means the function name should never be followed by `()`. If you forget, you'll get an error.


### Exercise 7

Copy previous code. Add `across(a:d, median)` and `n = n()` to the call to `summarise ()`.

```{r selecting-columns-with-cols-7, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r selecting-columns-with-cols-7-hint-1, eval = FALSE}
df_miss |> 
  summarise(across(..., median), 
             n = ...())
```

```{r selecting-columns-with-cols-7-test, include = FALSE}
df_miss |> 
  summarise(across(a:d, median), 
             n = n())
```

### 

Several of the results are NA because `median()` propagates the missing values in `df_miss`. If we want to ignore missing values, then we need to use the `na. rm = TRUE` argument to `median()`.


### Exercise 8

Copy previous code and replace `median` with `function(x) median(x, na.rm = TRUE)`.

```{r selecting-columns-with-cols-8, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r selecting-columns-with-cols-8-hint-1, eval = FALSE}
... |> 
  summarise(across(a:d, 
                   function(x) median(x, na.rm = ...)),
            n = n())
```

```{r selecting-columns-with-cols-8-test, include = FALSE}
df_miss |> 
  summarise(across(a:d, 
                   function(x) median(x, na.rm = TRUE)),
            n = n())
```

### 

Instead of calling `median()` directly, we created a new (anonymous) function that calls `median()` with the desired arguments. 

### Exercise 9

Use the previous code, but replace `function` with `\`

```{r selecting-columns-with-cols-9, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r selecting-columns-with-cols-9-hint-1, eval = FALSE}
df_miss |> 
  summarise
    across(a:d, ...(x) median(x, na.rm = TRUE)),
    ... = n()
  )
```

```{r selecting-columns-with-cols-9-test, include = FALSE}
df_miss |> 
  summarise(across(a:d, 
                   \(x) median(x, na.rm = TRUE)),
            n = n())
```

### 

Using the full word `function` when creating temporary (anonymous) functions is somewhat verbose, hence the option to replace `function` with `\`.


### Exercise 10

Use the previous code. Replace `\(x) median(x, na.rm = TRUE))` with `list()`.

```{r selecting-columns-with-cols-10, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r selecting-columns-with-cols-10-hint-1, eval = FALSE}
df_miss |> 
  summarise(across(a:d, ...),
            n = n())
```

```{r selecting-columns-with-cols-10-test, include = FALSE}
df_miss |> 
  summarise(across(a:d, list()),
            n = n())
```

### 

`list()` does not do anything yet. Instead, it provides a location in which you can supply multiple functions by using a named list.

### Exercise 11

Add the argument `med = \(x) median(x, na.rm = TRUE)` to the call inside of `list()`.

```{r selecting-columns-with-cols-11, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r selecting-columns-with-cols-11-hint-1, eval = FALSE}
df_miss |> 
  summarize(across(a:d, 
                   list(...)),
             n = n())
```

```{r selecting-columns-with-cols-11-test, include = FALSE}
df_miss |> 
  summarize(across(a:d, 
                   list(med = \(x) median(x, na.rm = TRUE))),
             n = n())
```

### 

R applies the function in `list()` to every variable in the first argument to `across()`. But, because the argument to `.fns` --- the second argument to `across` --- is a list, R creates new variable names by combining the each variable with the name of the function, which is `med`, not `median`. 

### Exercise 12

Copy previous code. Add the argument ``n_miss = \(x) sum(is.na(x))` to the call inside of `list()`.

```{r selecting-columns-with-cols-12, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r selecting-columns-with-cols-12-hint-1, eval = FALSE}
df_miss |> 
  summarize(across(a:d, 
                   list(med = \(x) median(x, na.rm = TRUE),
                   ...)),
             n = n())
```

```{r selecting-columns-with-cols-12-test, include = FALSE}
df_miss |> 
  summarize(across(a:d, 
                   list(med = \(x) median(x, na.rm = TRUE),
                        n_miss = \(x) sum(is.na(x)))),
             n = n())
```

### 

There is a lot going on here. We have four variables: `a` through `d`. We have two functions: `med` and `n_miss`. The magic of `across()` is that it applies every function to eac variable, creating new names for the output columns.


## Column names
### 

The last specification of `across()` is the `.names` argument, which we will explore next.

### Exercise 1

Hit `Run Code`.

```{r column-names-1, exercise = TRUE}
df_miss |> 
  summarize(across(a:d, 
                   list(med = \(x) median(x, na.rm = TRUE),
                        n_miss = \(x) sum(is.na(x)))),
             n = n())
```

### 

The result of `across()` is named according to the specification provided in the `.names` argument. The default (NULL) is equivalent to "{.col}" for the single function case and "{.col}_{.fn}" for the case where a list is used for `.fns`.


### Exercise 2

Add `.names = "{.fn}_{.col}"` to the call to `across()`.

```{r column-names-2, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r column-names-2-hint-1, eval = FALSE}
df_miss |> 
  summarize(across(a:d, 
                   list(med = \(x) median(x, na.rm = TRUE),
                        n_miss = \(x) sum(is.na(x))),
                    .names = ...),
             n = n())
```

```{r column-names-2-test, include = FALSE}
df_miss |> 
  summarize(across(a:d, 
                   list(med = \(x) median(x, na.rm = TRUE),
                        n_miss = \(x) sum(is.na(x))),
                     .names = "{.fn}_{.col}"),
             n = n())
```

### 

By default, columns are named using a glue specification like {.col}\_{.fn}. `.col` is the name of the original column (ie. `a`, `b`) and `.fn` is the name of the function (ie. `med`). 

### Exercise 3

For this practice, pipe `df_miss` into `mutate()`.

```{r column-names-3, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r column-names-3-hint-1, eval = FALSE}
df_miss |> 
  mutate()
```

```{r column-names-3-test, include = FALSE}
df_miss |> 
  mutate()
```

### 

The `.names` argument is particularly important when you use `across()` with `mutate()`.

### Exercise 4

Add `across(a:d, \(x) coalesce(x, 0))` to the call to `mutate()`.

```{r column-names-4, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r column-names-4-hint-1, eval = FALSE}
df_miss |> 
  mutate(...(a:d, 
             \(x) ...(x, 0)))
```

```{r column-names-4-test, include = FALSE}
df_miss |> 
  mutate(across(a:d, 
                \(x) coalesce(x, 0)))
```

### 

Recall that function `coalesce()` replaces NAs with 0.

Note that by default, the output of `across()` is given the same names as the inputs. This means that `across()` inside of `mutate()` will replace any existing columns with the same names. The new `a` is the same as the old `a`, but with any NA replaced by a zero.

To create new columns, we need to specify `.names`.


### Exercise 5

Within the call to `across()`, replace `\(x) coalesce(x, 0)` with `\(x) abs(x)`. Also, add `.names = "{.col}_abs"`. Make sure to separate each argument with comma.

```{r column-names-5, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r column-names-5-hint-1, eval = FALSE}
df_miss |> 
  mutate(
    across(a:d, \(x) ...(x), 
           ... = "{.col}_abs"))
```

```{r column-names-5-test, include = FALSE}
df_miss |> 
  mutate(
    across(a:d, \(x) abs(x), 
           .names = "{.col}_abs"))
```

### 

Using `.names` allows you to keep the original variables and also add new ones. In this case, `a_abs` is just the absolute value of `a`, and so on for `b`, `c`, and `d`.

### Exercise 6

Create a new pipe by sending `df_miss` into `filter()`.

```{r column-names-6, exercise = TRUE}

```

```{r column-names-6-hint-1, eval = FALSE}
df_miss |> 
  ...()
```

```{r column-names-6-test, include = FALSE}
df_miss |> 
  filter()
```

### 

`across()` is a great match for `summarize()` and `mutate()`, but it's more awkward to use with `filter()`. We will use two variants of `across()` called `if_any()` and `if_all()` with `filter()`.

### Exercise 7

Add `if_any()` to the call to `filter()`

```{r column-names-7, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r column-names-7-hint-1, eval = FALSE}
df_miss |> 
  filter(...())
```

```{r column-names-7-test, include = FALSE}
df_miss |> 
  filter(if_any())
```

### 

`if_any()` is the same as using `|` in-between the arguments. Right now, it has no effect because we haven't provided any arguments.

### Exercise 8

Add the arguments `a:d` and `is.na`, separated by a comma, to the function `if_any()`.

```{r column-names-8, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r column-names-8-hint-1, eval = FALSE}
df_miss |> 
  filter(if_any(..., is.na))
```

```{r column-names-8-test, include = FALSE}
df_miss |> 
  filter(if_any(a:d, is.na))
```

### 

The code above is the same as `df_miss |> filter(is.na(a) | is.na(b) | is.na(c) | is.na(d))`.

### Exercise 9

The other alternative, `if_all()`, is the same as using `&`.

Copy and past the previous code. Change `if_any` to `if_all`, see what happens.

```{r column-names-9, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r column-names-9-hint-1, eval = FALSE}
df_miss |> 
  filter(if_all(...))
```

```{r column-names-9-test, include = FALSE}
df_miss |> 
  filter(if_all(a:d, is.na))
```


You will get an empty tibble because none of the rows have all the values for `a` through `d` equal to NA.


## across() in functions
### 

`across()` is particularly useful to program with because it allows you to operate on multiple columns.

### Exercise 1

Hit `Run Code`.

```{r across-in-functions-1, exercise = TRUE}
expand_dates <- function(df) {
  df |> 
    mutate(
      across(where(is.Date), list(year = year, month = month, day = mday))
    )
}

df_date <- tibble(
  name = c("Amy", "Bob"),
  date = ymd(c("2009-08-03", "2010-01-16"))
)
```


This should not output a tibble.

### 

`expand_dates()` uses **lubridate** functions to expand all date columns within a tibble into year, month, and day columns.

### Exercise 2

Pipe `df_date` into `expand_dates()`.

```{r across-in-functions-2, exercise = TRUE}

```

```{r across-in-functions-2-hint-1, eval = FALSE}
df_date |> 
  ...()
```

```{r across-in-functions-2-test, include = FALSE}
df_date |> 
  expand_dates()
```

### 

Instead of having to specify the name of the column ahead of time --- a name which we might not even know --- we use `where(is.Date)` to automatically pick out any date columns.


### Exercise 3

Hit `Run Code` to define the following function.

```{r across-in-functions-3, exercise = TRUE}
summarize_means <- function(df, summary_vars = where(is.numeric)) {
  df |> 
    summarize(
      across({{ summary_vars }}, \(x) mean(x, na.rm = TRUE)),
      n = n()
    )
}
```

### 

<!-- DK: Have students already been exposed to {{}}? -->

`across()` makes it easy to supply multiple columns in a single argument. We would not expect you to be able to create a function like this, given the current state of your R knowledge. But you should be able to read and understand this code.

In this case, `summary_vars` will be all the numeric variables in `df`. The `{{ }}` syntax is used for embracing the `summary_vars` argument, allowing it to be passed as a variable or expression.

### Exercise 4

Pipe `diamonds` into `group_by(cut)`, then pipe into `summarize_means()`.

```{r across-in-functions-4, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r across-in-functions-4-hint-1, eval = FALSE}
diamonds |> 
  group_by(...) |> 
  ...()
```

```{r across-in-functions-4-test, include = FALSE}
diamonds |> 
  group_by(cut) |> 
  summarize_means()
```

### 

We mentioned that where possible, `.by` should be sued instead of `group_by()`. However, this is a situation where it is not possible to use `.by`.


### Exercise 5

Now, add argument `c(carat, x:z)` to the function `summarize_means()`.

```{r across-in-functions-5, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r across-in-functions-5-hint-1, eval = FALSE}
diamonds |> 
  group_by(cut) |> 
  summarize_means(...(carat,...))
```

```{r across-in-functions-5-test, include = FALSE}
diamonds |> 
  group_by(cut) |> 
  summarize_means(c(carat, x:z))
```

### 

Another interesting function associated with `across()` is `pivot_longer()`, which makes data sets longer by increasing the number of rows and decreasing the number of columns.

### Exercise 6

Hit `Run Code`. Pay attention to the table generated

```{r across-in-functions-6, exercise = TRUE}
df |> 
  summarize(across(a:d, 
                   list(median = median, mean = mean)))
```

```{r across-in-functions-6-test, include = FALSE}

```

### 

What did you notice about this table? It's too wide and hence hard to compare values.

There's a way to compute the same values by pivoting longer and then summarizing.

### Exercise 7

Hit `Run Code` to create the following tibble and set it to `df_paired`.

```{r across-in-functions-7, exercise = TRUE}
df_paired <- tibble(
  a_val = rnorm(10),
  a_wts = runif(10),
  b_val = rnorm(10),
  b_wts = runif(10),
  c_val = rnorm(10),
  c_wts = runif(10),
  d_val = rnorm(10),
  d_wts = runif(10)
)
```

### Exercise 8

Now type in `df_paired` and hit `Run Code`. 

```{r across-in-functions-8, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r across-in-functions-8-hint-1, eval = FALSE}
df_paired
```

```{r across-in-functions-8-test, include = FALSE}
df_paired
```

### 

Now you should see the tibble you have created!


### Exercise 9

`pivot_longer()` makes datasets longer by increasing the number of rows and decreasing the number of columns.

Pipe `df_paired` into function `pivot_longer()`.

```{r across-in-functions-9, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r across-in-functions-9-hint-1, eval = FALSE}
df_paired |> 
  ...()
```

```{r across-in-functions-9-test, include = FALSE}
df_paired |> 
   pivot_longer()
```

You will receive an error message because `pivot_longer` requires a value for the `cols` argument.

### Exercise 10

Copy the previous code. Add the argument `everything()` to the function `pivot_longer()`. 

```{r across-in-functions-10, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r across-in-functions-10-hint-1, eval = FALSE}
df_paired |> 
  pivot_longer(...())
```

```{r across-in-functions-10-test, include = FALSE}
df_paired |> 
  pivot_longer(everything())
```

### 

Remember that the `cols` argument specifies which columns to pivot and `everything()` specifies that we want to pivot every column.

### Exercise 11

Add two more arguments to `pivot_longer()`: `names_to = c("group", ".value")` and `names_sep = "_"`.


```{r across-in-functions-11, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r across-in-functions-11-hint-1, eval = FALSE}
df_paired |> 
  pivot_longer(everything(),
               ... = c("group", ".value"),
               names_sep = ...)
```

```{r across-in-functions-11-test, include = FALSE}
df_paired |> 
  pivot_longer(everything(),
               names_to = c("group", ".value"),
               names_sep = "_")
```

### 

The `names_sep` argument tells `pivot_longer()` what separates a `group` value from a  `.value` in each of the column names. Here, that's an underscore `_`.

### 

Here's a quick summary of the above mentioned functions:

`.cols` specifies which columns you want to iterate over

`.fns` specifies what to do with each column

`.names` is used when you need additional control over the names of output columns, which is particularly important when you use `across()` with `mutate()`.

`if_any()` and `if_all()` are two important variations which work with `filter()`.

## Reading multiple files
### v

In this section, you'll learn how to use `purrr::map()` to do something to every file in a directory. Imagine you have a directory full of excel spreadsheets you want to read. You could do it with copy and paste:

```
data2019 <- readxl::read_excel("data/y2019.xlsx")
data2020 <- readxl::read_excel("data/y2020.xlsx")
data2021 <- readxl::read_excel("data/y2021.xlsx")
data2022 <- readxl::read_excel("data/y2022.xlsx")
```

And then use `dplyr::bind_rows()` to combine them all together:

```
data <- bind_rows(data2019, data2020, data2021, data2022)
```

You can imagine that this would get tedious quickly, especially if you had hundreds of files, not just four. 


### Exercise 1

Hit `Run Code`.

```{r reading-multiple-files-1, exercise = TRUE}
paths <- list.files("data/gapminder", pattern = "[.]xlsx$", full.names = TRUE)

paths
```

```{r reading-multiple-files-1-test, include = FALSE}
paths <- list.files("data/gapminder", pattern = "[.]xlsx$", full.names = TRUE)

paths
```

### 

Notice that we used function `list.files()` here, which basically lists the files in a directory.


<!-- DK: Ask new question. Type ?list.files in the Console and copy/paste the Description. This is all handed much better in the next chapter. Make them the same. -->

### 

This is the code from last exercise: `paths <- list.files("data/gapminder", pattern = "[.]xlsx$", full.names = TRUE)`

The first argument, `path`, is the directory to look in. In this case, it is `"data/gapminder"`.

`pattern` is a regular expression used to filter the file names. The most common pattern is something like `[.]xlsx$` or `[.]csv$` to find all files with a specified extension.

`full.names` determines whether or not the directory name should be included in the output. You almost always want this to be `TRUE`.

Based on the 12 paths created from the last exercise, call `read_excel()` 12 times to get 12 data frames. 

Your code should look something like this:

```
gapminder_1952 <- readxl::read_excel("data/gapminder/1952.xlsx")
gapminder_1957 <- readxl::read_excel("data/gapminder/1957.xlsx")
gapminder_1962 <- readxl::read_excel("data/gapminder/1962.xlsx")
gapminder_1952 <- readxl::read_excel("data/gapminder/1967.xlsx")
gapminder_1957 <- readxl::read_excel("data/gapminder/1972.xlsx")
gapminder_1962 <- readxl::read_excel("data/gapminder/1977.xlsx")
gapminder_1952 <- readxl::read_excel("data/gapminder/1982.xlsx")
gapminder_1957 <- readxl::read_excel("data/gapminder/1987.xlsx")
gapminder_1962 <- readxl::read_excel("data/gapminder/1992.xlsx")
gapminder_1952 <- readxl::read_excel("data/gapminder/1997.xlsx")
gapminder_1957 <- readxl::read_excel("data/gapminder/2002.xlsx")
gapminder_2007 <- readxl::read_excel("data/gapminder/2007.xlsx")
```

### 

Separate excel sheets are always easier to work with once we put them into a single object. You can use `list()` to combine them together, then assign the list to a new variable, `files`.

your code should look like this:

```
files <- list(
  readxl::read_excel("data/gapminder/1952.xlsx"),
  readxl::read_excel("data/gapminder/1957.xlsx"),
  readxl::read_excel("data/gapminder/1962.xlsx"),
  ...,
  readxl::read_excel("data/gapminder/2007.xlsx")
)

```

### Exercise 2

Type `files[[3]]` and hit `Run Code`, see what happens. Note that we are using `[[]]` here instead of `()`.

```{r reading-multiple-files-2, exercise = TRUE}

```

```{r reading-multiple-files-2-hint-1, eval = FALSE}
...[[3]]
```

```{r reading-multiple-files-2-test, include = FALSE}
files[[3]]
```

### 

Remember that after you have data frames in a list, you can use `files[[i]]` to extract the `i`th element in the list.


### Exercise 3

Type `map()` in the code chunk, then hit `Run Code`.

```{r reading-multiple-files-3, exercise = TRUE}

```

```{r reading-multiple-files-3-hint-1, eval = FALSE}
map()
```

```{r reading-multiple-files-3-test, include = FALSE}

```


This will give you an error because we haven't added any arguments yet.

### 

The `map()` functions transform their input by applying a function to each element of a list or atomic vector and returning an object of the same length as the input.

### Exercise 4

Add argument `paths` and `readxl::read_excel` to the call to `map()`.

```{r reading-multiple-files-4, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r reading-multiple-files-4-hint-1, eval = FALSE}
map(..., readxl::read_excel)
```

<!-- DK: This code gives a problem, but only with R CMD check. If you just Run Document the tutorial, everything works fine. But, with check, you get some weird message. Is there something about the paths not existing the built file because of something wrong in .Rbuildignore. Would be nice to fix this.  -->

```{r include = FALSE}
# map(paths, readxl::read_excel)
```

### 

The `map()` vector takes the format fo `map(x, f)`, it is a way to make good use of the `path` vector. 

### Exercise 5

Type `length()`, then hit `Run Code`.

```{r reading-multiple-files-5, exercise = TRUE}

```

```{r reading-multiple-files-5-hint-1, eval = FALSE}
length()
```

```{r reading-multiple-files-5-test, include = FALSE}
# length()
```

This will give you an error because we haven't added any arguments yet. Function `length()` can get or set the length of vectors (including lists) and factors, and of any other R object for which a method has been defined.

### Exercise 6

Add `files`, the collection of tibbles we just created, to the function `length()`.

```{r reading-multiple-files-6, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r reading-multiple-files-6-hint-1, eval = FALSE}
length(files)
```

```{r reading-multiple-files-6-test, include = FALSE}

```

### 

Now you know how many files are in `files`.

### Exercise 7

Run the code below to see what happens.

```{r reading-multiple-files-7, exercise = TRUE}
files[[1]]
```

```{r reading-multiple-files-7-test, include = FALSE}

```

### 

This is another data structure that doesn’t display particularly compactly with `str()` so you might want to load it into RStudio and inspect it with `View()`.

### Exercise 8

Type `list_rbind()` in the space below. Hit `Run Code`.

```{r reading-multiple-files-8, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r reading-multiple-files-8-hint-1, eval = FALSE}
list_rbind() 
```

```{r reading-multiple-files-8-test, include = FALSE}

```

### 

`list_rbind()` combines elements into a data frame by row-binding them together with `vctrs::vec_rbind()`.

### Exercise 9

Add `files` to the call to `list_rbind()`.

```{r reading-multiple-files-9, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r reading-multiple-files-9-hint-1, eval = FALSE}
list_rbind(files)
```

```{r reading-multiple-files-9-test, include = FALSE}

```

### 

There are two more functions that combine list elements into a single data structure than just `list_rbind()`.

1. `list_c()` combines elements into a vector by **concatenating them together** with vctrs::vec_c().

2. `list_rbind()` combines elements into a data frame by **row-binding them together** with vctrs::vec_rbind().

3. `list_cbind()` combines elements into a data frame by **column-binding them together** with vctrs::vec_cbind().

### Exercise 10

Pipe `paths` into fucntion `map()`.

```{r reading-multiple-files-10, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r reading-multiple-files-10-hint-1, eval = FALSE}
paths |> 
  map
```

```{r reading-multiple-files-10-test, include = FALSE}

```

### Exercise 11

Add `\(path) readxl::read_excel(path, n_max = 1)` into the function `map()`.

```{r reading-multiple-files-11, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r reading-multiple-files-11-hint-1, eval = FALSE}
paths |> 
  map(\(path) readxl::read_excel(path, n_max = 1))
```

```{r reading-multiple-files-11-test, include = FALSE}

```

### 

`n_max = 1` allows us to peak at the first few rows of the data.

### Exercise 12

Pipe the code above into function `list_rbind()`.

```{r reading-multiple-files-12, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r reading-multiple-files-12-hint-1, eval = FALSE}
paths |> 
  map(\(path) readxl::read_excel(path, n_max = 1)) |> 
  list_rbind()
```

```{r reading-multiple-files-12-test, include = FALSE}

```

### 

This combination of `map()` and `list_rbind()` allows all sporeasheet to be viewed at once and combined into one. Like what we did above with `n_max = 1`, you can also customize how many rows you want to see.

### 

Another important observation is that there's no "year" column indicating which row of data belong to which spreadsheet. 

### Exercise 13

Pipe `paths` into `set_names(basename)`

```{r reading-multiple-files-13, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r reading-multiple-files-13-hint-1, eval = FALSE}
paths |> set_names(basename) 
```

```{r reading-multiple-files-13-test, include = FALSE}

```

### 

Using `basename()` in the `set_names()` function, we can extract just the file name from the full path.

### Exercise 14

Pipe the previous code into `map(readxl::read_excel)`

```{r reading-multiple-files-14, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r reading-multiple-files-14-hint-1, eval = FALSE}
paths |> 
  set_names(basename) |> 
  map(readxl::read_excel)
```

```{r reading-multiple-files-14-test, include = FALSE}

```

### 

The names created in the last exercise automatically carried along by all the map functions, so the list of data frames will have those same names.

### Exercise 15

Assign a new variable `file` to the previous code.

```{r reading-multiple-files-15, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r reading-multiple-files-15-hint-1, eval = FALSE}
file <- paths |> 
  set_names(basename) |> 
  map(readxl::read_excel)
```

```{r reading-multiple-files-15-test, include = FALSE}

```

### 

What we've created above is actually shorthand for:

```
files <- list(
  "1952.xlsx" = readxl::read_excel("data/gapminder/1952.xlsx"),
  "1957.xlsx" = readxl::read_excel("data/gapminder/1957.xlsx"),
  "1962.xlsx" = readxl::read_excel("data/gapminder/1962.xlsx"),
  ...,
  "2007.xlsx" = readxl::read_excel("data/gapminder/2007.xlsx"))
```

Imagine how much time we would save!

### Exercise 16

Type `files[[""]]`, with any `year.xlsx` between the `""`

```{r reading-multiple-files-16, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r reading-multiple-files-16-hint-1, eval = FALSE}
files[["1962.xlsx"]]
```

```{r reading-multiple-files-16-test, include = FALSE}

```

### 

You should get the file of the respective year you chose! 

### Exercise 17

Copy and paste code from Exercise 18 to the space below. Hit `Run Code`.

```{r reading-multiple-files-17, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r reading-multiple-files-17-hint-1, eval = FALSE}
paths |> 
  set_names(basename) |> 
  map(readxl::read_excel)
```

```{r reading-multiple-files-17-test, include = FALSE}

```

### Exercise 18

Pipe the code above into `list_rbind() `, then add argument `names_to = "year"` to this function.

```{r reading-multiple-files-18, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r reading-multiple-files-18-hint-1, eval = FALSE}
paths |> 
  set_names(basename) |> 
  map(readxl::read_excel) |> 
  ...(names_to = "year")
```

```{r reading-multiple-files-18-test, include = FALSE}

```

### 

The argument `names_to` gives the name of the variable that will be created from the data stored in the column names

### Exercise 19

Pipe the previous code into `mutate()` to create a new column, with the argument `year = parse_number(year)`.

```{r reading-multiple-files-19, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r reading-multiple-files-19-hint-1, eval = FALSE}
paths |> 
  set_names(basename) |> 
  map(readxl::read_excel) |> 
  list_rbind(names_to = "year") |> 
  mutate(year = parse_number(year))
```

```{r reading-multiple-files-19-test, include = FALSE}

```

### 

`parse_number` parses the first number it finds, dropping any non-numeric characters before the first number and all characters after the first number. The grouping mark specified by the locale is ignored inside the number.

### 

Overall, we used the  `names_to` argument to `list_rbind()` to tell it to save the names into a new column called year then use `readr::parse_number()` to extract the number from the string.


### Exercise 20

Assign the previous code to the variable `gapminder`

```{r reading-multiple-files-20, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r reading-multiple-files-20-hint-1, eval = FALSE}
gapminder <- paths |> 
  set_names(basename) |> 
  map(readxl::read_excel) |> 
  list_rbind(n...) |> 
  mutate(...)
```

```{r reading-multiple-files-20-test, include = FALSE}

```

## Save your work
### 

### Exercise 1

Type `write_csv()` in the empty code chunk below, hit `Run Code`.

```{r save-your-work-1, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r save-your-work-1-hint-1, eval = FALSE}
write_csv()
```

### 

In brief, `write_csv()` is used to write a data frame to a delimited file.

### Exercise 2

Add arguments `gapminder` and `"gapminder.csv"` to the call to `write_csv()`

```{r save-your-work-2, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r save-your-work-2-hint-1, eval = FALSE}
write_csv(gapminder, "gapminder.csv")
```

### 

This is how we save the work we have previously done, a single csv file is now created and ready for future use.

### Exercise 3

Run the following code.

```{r save-your-work-3, exercise = TRUE}
df <- read_csv(paths)
```

### Exercise 4

Pipe the newly created variable `df` into `filter()`, with argument `!is.na(id)`.

```{r save-your-work-4, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r save-your-work-4-hint-1, eval = FALSE}
df |> 
    filter(!is.na(id))
```

### 

`is.na()` is used to deal with missing values in the dataset or data frame.

### Exercise 5

Pipe the code above into `mutate()` with argument `id = tolower(id)`.
```{r save-your-work-5, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r save-your-work-5-hint-1, eval = FALSE}
  df |> 
    filter(!is.na(id)) |> 
    mutate(...)
```

### Exercise 6

Then use `pivot_longer()` and add arguments `jan:dec` and `names_to = "month"` to the call. Make sure to add a `,` between the two arguments.

```{r save-your-work-6, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r save-your-work-6-hint-1, eval = FALSE}
  df |> 
    filter(!is.na(id)) |> 
    mutate(id = tolower(id)) |> 
    pivot_longer(...)
```

### 

Just a recap, `pivot_longer()` makes datasets longer by increasing the number of rows and decreasing the number of columns.

### Exercise 7

Assigned variable name `process_file` to <- `function(path)`,followed by a `{}` with the previous code. 

```{r save-your-work-7, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r save-your-work-7-hint-1, eval = FALSE}
process_file <- function(path) {
  df <- read_csv(path)
  
  df |> 
    filter(!is.na(id)) |> 
    mutate(id = tolower(id)) |> 
    pivot_longer(jan:dec, names_to = "month")
}
```

### 

In this step, we write a function that takes a file and does all the modifications (in this case, we read a bunch of files, filter out missing values, pivot, and then combine).

### Exercise 8

Pipe `paths` into `map()`.

```{r save-your-work-8, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r save-your-work-8-hint-1, eval = FALSE}
paths |> 
  map()
```

### Exercise 9

Add `process_file` as argument to `map()`.

```{r save-your-work-9, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r save-your-work-9-hint-1, eval = FALSE}
paths |> 
  map(process_file)
```

### Exercise 10

Pipe the above into `list_rbind()`.

```{r save-your-work-10, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r save-your-work-10-hint-1, eval = FALSE}
paths |> 
  map(process_file) |> 
  list_rbind()
```

### 

This is an example of doing one round of iteration with a complex function.

### Exercise 11

Pipe `paths` into `map(readxl::read_excel)`

```{r save-your-work-11, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r save-your-work-11-hint-1, eval = FALSE}
paths |> 
  map(readxl::read_excel)
```

### 

Recall that this code helps to load all files needed.

With a homogeneous data, we are able to go from `map()` straight to `list_rbind()`. However, when the data frames are heterogeneous, `list_rbind()` either fails or yields a data frame that’s not very useful. In this case, we can create a handy `df_types` function that returns a tibble with one row for each column.

### Exercise 12

Type `tibble()` and hit `Run Code`.

```{r save-your-work-12, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r save-your-work-12-hint-1, eval = FALSE}
tibble()
```

### Exercise 13

Let's add the argument `col_name` with value `names(df)` to the call to `tibble()`.

```{r save-your-work-13, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r save-your-work-13-hint-1, eval = FALSE}
tibble(
    col_name = names(df))
```

### 

Recall that this argument `col_name` retrieve/set column names.

### Exercise 14

Then, add the second argument, `col_type`, then set its value to be `map_chr(df, vctrs::vec_ptype_full)`.

```{r save-your-work-14, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r save-your-work-14-hint-1, eval = FALSE}
tibble(
    col_name = names(df), 
    col_type = map_chr(df, vctrs::vec_ptype_full))
```

### 

`col_type` specifies which types the columns in your imported data frame should have.

`map_chr()` returns a character vector, which is then specified by `vctrs::vec_ptype_full`. 

### 

`vec_ptype_full()` displays the full type of the vector.

### Exercise 15

Add a new argument `n_miss` to the previous code, with value `map_int(df, \(x) sum(is.na(x)))`.

```{r save-your-work-15, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r save-your-work-15-hint-1, eval = FALSE}
tibble(
    col_name = names(df), 
    col_type = map_chr(df, vctrs::vec_ptype_full),
    n_miss = map_int(df, \(x) sum(is.na(x))))
```

### 

<!-- add some knowledge drop abt `map_int(df, \(x) sum(is.na(x)))`... my brain is numb at this point but i'll come back later -->

### Exercise 16

To make the above code into a function, type `function(df) {}` in front of `tibble()`, remember to put`tibble()` inside `{}`.

```{r save-your-work-16, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r save-your-work-16-hint-1, eval = FALSE}
function(df) {
  tibble(
    col_name = names(df), 
    col_type = map_chr(df, vctrs::vec_ptype_full),
    n_miss = map_int(df, \(x) sum(is.na(x))))}
```

### 

The previous chapter gave a detailed break down of how a new function is defined.

### Exercise 17

Assign the new function a name as `df_types`.

```{r save-your-work-17, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r save-your-work-17-hint-1, eval = FALSE}
... <- function(df) {
  tibble(
    col_name = names(df), 
    col_type = map_chr(df, vctrs::vec_ptype_full),
    n_miss = map_int(df, \(x) sum(is.na(x))))}
```

### 

This assignment will make life easier for us later on!

### Exercise 18

Type `df_types()` and an argument `gapminder`, hit `Run Code`, see what happens.

```{r save-your-work-18, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r save-your-work-18-hint-1, eval = FALSE}
df_types(gapminder)
```

### 

You should have a 6x3 tibble!

### 

There are two additiolnal functions worth mentioning.

`map_if()` allows you to selectively modify elements of a list based on their values.

`map_at()` allows you to selectively modify elements based on their names.

### Exercise 19

Pipe `paths` into `map()`.

```{r save-your-work-19, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r save-your-work-19-hint-1, eval = FALSE}
paths |> 
  map()
```

### 

`map()` has a downside: it succeeds or fails as a whole. It will either successfully read all of the files in a directory or fail with an error, reading zero files. 

### Exercise 20

Add an argument `possibly()` to the call to `map()`.

```{r save-your-work-20, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r save-your-work-20-hint-1, eval = FALSE}
paths |> 
  map(possibly())
```

### 

`possibly()` is a function operator: it takes a function and returns a function with modified behavior. 

In particular, `possibly()` changes a function from erroring to returning a value that you specify。

### Exercise 21

Add arguments `\(path) readxl::read_excel(path)` and `NULL` to the call to `possibly()`.

```{r save-your-work-21, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r save-your-work-21-hint-1, eval = FALSE}
paths |> 
  map(possibly(\(path) readxl::read_excel(path), NULL))
```

### 

`possibly()` consists of two main values: `.f` and `otherwise`.

'.f` is a function to modify.

`otherwise` is a default value to use when an error occurs.

### Exercise 22

Assign the above code to variable `files`.

```{r save-your-work-22, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r save-your-work-22-hint-1, eval = FALSE}
... <- paths |> 
  map(possibly(\(path) readxl::read_excel(path), NULL))
```

### Exercise 23

Pipe `files` into `list_rbind()`.

```{r save-your-work-23, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r save-your-work-23-hint-1, eval = FALSE}
files |> list_rbind()
```

### 

This works particularly well here because `list_rbind()`, like many tidyverse functions, automatically ignores `NULLs`.

### Exercise 24

With function `map_vec()`, add arguments `files` and `is.null` to the call. 

```{r save-your-work-24, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r save-your-work-24-hint-1, eval = FALSE}
map_vec(files, is.null)
```

### Exercise 25

Assign `failed` as a new variable of the previous code.

```{r save-your-work-25, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r save-your-work-25-hint-1, eval = FALSE}
failed <- map_vec(files, is.null)
```

### Exercise 26

Hit `Run Code`.

```{r save-your-work-26, exercise = TRUE}
paths[failed]
```

### 

You should see paths that failed.

Then call the import function again for each failure and figure out what went wrong.

### 

<!-- End of Section knowledge drop -->

Just a recap, there are three basic steps:

use `list.files()` to list all the files in a directory, then use `purrr::map()` to read each of them into a list, then use `purrr::list_rbind()` to combine them into a single data frame. We'll then discuss how you can handle situations of increasing heterogeneity, where you can't do exactly the same thing to every file.

## Saving Multiple Outputs
### 

<!-- Biggest annoyance is that we need to include all code in each Exercise because the `con` connection object disappears after each Exercise.  -->

In the last section, you learned about `map()`, which is useful for reading multiple files into a single object. In this and the following sections, we’ll now explore sort of the opposite problem: how can you take one or more R objects and save it to one or more files? We’ll explore this challenge using three examples:

* Saving multiple data frames into one database.
* Saving multiple data frames into multiple .csv files.
* Saving multiple plots to multiple .png files.

### Exercise 1

Sometimes when working with many files, it’s not possible to fit all your data into memory at once, and you can’t do `map(files, read_csv)`. One approach to deal with this problem is to load your data into a database. Run the code in the code chunk below. 

```{r saving-multiple-outputs-1, exercise = TRUE}
con <- DBI::dbConnect(duckdb::duckdb())
```

```{r saving-multiple-outputs-1-test, include = FALSE}
con <- DBI::dbConnect(duckdb::duckdb())
dbDisconnect(con, shutdown=TRUE)
```

### 

`con` is a connection to a database created by the [**duckdb**](https://r.duckdb.org/) package. Note that because of limits on **learnr** tutorials, we will need to run this command --- and any other commands which modify the database --- in each Exercise block for this tutorial. Apologies for the repetition!

### Exercise 2

We need to start by creating a table that we will fill in with data. The easiest way to do this is by creating a template, a dummy data frame that contains all the columns we want, but only a sampling of the data. Assign `template` to `readxl::read_excel(paths[[1]])`.

```{r saving-multiple-outputs-2, exercise = TRUE}

```

```{r saving-multiple-outputs-2-hint-1, eval = FALSE}
template <- readxl::...(paths[[1]])
```

```{r saving-multiple-outputs-2-test, include = FALSE}
template <- readxl::read_excel(paths[[1]])
```

### 

Other databases might allow us to load Excel files directly. With **duckdb**, however, we’re going to have to do it “by hand”. Learning to do it by hand will also help you when you have a bunch of CSVs and the database that you’re working with doesn’t have one function that will load them all in.

### Exercise 3

Take the assignment from the previous question, add a pipe, and then use `mutate()` to add a new variable, `year`, and assign it a value of 1952.

```{r saving-multiple-outputs-3, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r saving-multiple-outputs-3-hint-1, eval = FALSE}
template <- readxl::read_excel(paths[[1]]) |> 
  mutate(... = 1952)
```

```{r saving-multiple-outputs-3-test, include = FALSE}
template <- readxl::read_excel(paths[[1]]) |> 
  mutate(year = 1952)
```

### 

We are about to add in data from a bunch of different files, each from a different year. So, we need to ensure that there is a `year` column in the template in which we can store that year information.


### Exercise 4

Run this code.

```{r saving-multiple-outputs-4, exercise = TRUE}
con <- DBI::dbConnect(duckdb::duckdb())
DBI::dbCreateTable(con, "gapminder", template)
```

```{r saving-multiple-outputs-4-test, include = FALSE}
con <- DBI::dbConnect(duckdb::duckdb())
DBI::dbCreateTable(con, "gapminder", template)
dbDisconnect(con, shutdown=TRUE)
```

### 

Note that `dbCreateTable()` doesn’t use the data in template, just the variable names and types. 

### Exercise 5

Now, let's test it by pipe `con` into `tbl("gapminder")`. (We provide the creation code. You add the pipe below it.)

```{r saving-multiple-outputs-5, exercise = TRUE}
con <- DBI::dbConnect(duckdb::duckdb())
DBI::dbCreateTable(con, "gapminder", template)

```

```{r saving-multiple-outputs-5-hint-1, eval = FALSE}
con <- DBI::dbConnect(duckdb::duckdb())
DBI::dbCreateTable(con, "gapminder", template)

con |> 
  tbl("gapminder")
```

```{r saving-multiple-outputs-5-test, include = FALSE}
con <- DBI::dbConnect(duckdb::duckdb())
DBI::dbCreateTable(con, "gapminder", template)

con |> 
  tbl("gapminder")
dbDisconnect(con, shutdown=TRUE)
```

### 

The `gapminder` table is empty now but it has the variables we need with the types we expect. You can see this explicitly if you add `colnames()` to the pipe and re-run the code.

### Exercise 6

We need a function which will accept a path to a correctly formatted Excel file, read the file into a tibble, add the appropriate year variable, and then append that tibble to the `gapminder` table we have created in our **duckdb** database.

Run this code.

```{r saving-multiple-outputs-6, exercise = TRUE}
append_file <- function(path) {
  df <- readxl::read_excel(path)
  df$year <- parse_number(basename(path))
  DBI::dbAppendTable(con, "gapminder", df)
  }
```

```{r saving-multiple-outputs-6-test, include = FALSE}
append_file <- function(path) {
  df <- readxl::read_excel(path)
  df$year <- parse_number(basename(path))
  DBI::dbAppendTable(con, "gapminder", df)
  }
```

### 

Before creating a function like this, we often write the code which accomplishes what we want for a single case. With that script as a guide, we can create a function which is often just a "wrap" around that code.

### Exercise 7

Pipe `paths` into `map(append_file)`.

```{r saving-multiple-outputs-7, exercise = TRUE}
con <- DBI::dbConnect(duckdb::duckdb())
DBI::dbCreateTable(con, "gapminder", template)

append_file <- function(path) {
  df <- readxl::read_excel(path)
  df$year <- parse_number(basename(path))
  DBI::dbAppendTable(con, "gapminder", df)
  }

```

```{r saving-multiple-outputs-7-hint-1, eval = FALSE}
con <- DBI::dbConnect(duckdb::duckdb())
DBI::dbCreateTable(con, "gapminder", template)

append_file <- function(path) {
  df <- readxl::read_excel(path)
  df$year <- parse_number(basename(path))
  DBI::dbAppendTable(con, "gapminder", df)
  }

paths |> 
  map(...)
```

```{r saving-multiple-outputs-7-test, include = FALSE}
con <- DBI::dbConnect(duckdb::duckdb())
DBI::dbCreateTable(con, "gapminder", template)

append_file <- function(path) {
  df <- readxl::read_excel(path)
  df$year <- parse_number(basename(path))
  DBI::dbAppendTable(con, "gapminder", df)
}

# DK: This causes a failure with R CMD check. Why?

# paths |>
#   map(append_file)

dbDisconnect(con, shutdown=TRUE)
```

### 

Again, it is annoying that we need to repeat so much code each Exercise.

### Exercise 8

But we don’t care about the output of append_file(), so instead of map() it’s slightly nicer to use walk(). walk() does exactly the same thing as map() but throws the output away. Pipe `paths` into `walk(append_file)`.

```{r saving-multiple-outputs-8, exercise = TRUE}
con <- DBI::dbConnect(duckdb::duckdb())
DBI::dbCreateTable(con, "gapminder", template)

append_file <- function(path) {
  df <- readxl::read_excel(path)
  df$year <- parse_number(basename(path))
  DBI::dbAppendTable(con, "gapminder", df)
}

```

```{r saving-multiple-outputs-8-hint-1, eval = FALSE}
con <- DBI::dbConnect(duckdb::duckdb())
DBI::dbCreateTable(con, "gapminder", template)

append_file <- function(path) {
  df <- readxl::read_excel(path)
  df$year <- parse_number(basename(path))
  DBI::dbAppendTable(con, "gapminder", df)
  }

paths |> 
  ...(append_file)
```

```{r saving-multiple-outputs-8-test, include = FALSE}
con <- DBI::dbConnect(duckdb::duckdb())
DBI::dbCreateTable(con, "gapminder", template)

append_file <- function(path) {
  df <- readxl::read_excel(path)
  df$year <- parse_number(basename(path))
  DBI::dbAppendTable(con, "gapminder", df)
  }

paths |> 
  walk(append_file)
dbDisconnect(con, shutdown=TRUE)
```

### 

### Exercise 9

Now we can check if we have all data in our table. Start a new line at the end of this code, piping `con` into `tbl("gapminder")`.

```{r saving-multiple-outputs-9, exercise = TRUE}
con <- DBI::dbConnect(duckdb::duckdb())
DBI::dbCreateTable(con, "gapminder", template)

append_file <- function(path) {
  df <- readxl::read_excel(path)
  df$year <- parse_number(basename(path))
  DBI::dbAppendTable(con, "gapminder", df)
  }

paths |> 
  walk(append_file)

```

```{r saving-multiple-outputs-9-hint-1, eval = FALSE}
...

con |> 
  tbl("gapminder")
```

```{r saving-multiple-outputs-9-test, include = FALSE}
con <- DBI::dbConnect(duckdb::duckdb())
DBI::dbCreateTable(con, "gapminder", template)

append_file <- function(path) {
  df <- readxl::read_excel(path)
  df$year <- parse_number(basename(path))
  DBI::dbAppendTable(con, "gapminder", df)
  }

paths |> 
  walk(append_file)

con |> 
  tbl("gapminder")
dbDisconnect(con, shutdown=TRUE)
```

### 

We have successfully loaded a dozen Excel files into a database table.

### Exercise 10

Pipe the code above into `count(year)`.

```{r saving-multiple-outputs-10, exercise = TRUE}
con <- DBI::dbConnect(duckdb::duckdb())
DBI::dbCreateTable(con, "gapminder", template)

append_file <- function(path) {
  df <- readxl::read_excel(path)
  df$year <- parse_number(basename(path))
  DBI::dbAppendTable(con, "gapminder", df)
  }

paths |> 
  walk(append_file)

con |> 
  tbl("gapminder")

```

```{r saving-multiple-outputs-10-hint-1, eval = FALSE}
  ... |> 
  count(year)
```

```{r saving-multiple-outputs-10-test, include = FALSE}
con <- DBI::dbConnect(duckdb::duckdb())
DBI::dbCreateTable(con, "gapminder", template)

append_file <- function(path) {
  df <- readxl::read_excel(path)
  df$year <- parse_number(basename(path))
  DBI::dbAppendTable(con, "gapminder", df)
  }

paths |> 
  walk(append_file)

con |> 
  tbl("gapminder") |> 
  count(year)
dbDisconnect(con, shutdown=TRUE)
```

### 

The gapminder data is so small that we can just work with it in memory. But, for larger amounts of data, we need to use a database like **duckdb**.

## Writing csv files
### 

<!-- This would be much easier without intermediary objects. Why not just a pipe? Answer seems to be that walk2() is not really designed to work with pipes. -->

The same basic principle applies if we want to write multiple CSV files, one for each group. Let’s imagine that we want to take the `ggplot2::diamonds` data and save one CSV file for each value of `clarity`.

### Exercise 1

Pipe `diamonds` to `group_nest()`.

```{r writing-csv-files-1, exercise = TRUE}

```

```{r writing-csv-files-1-hint-1, eval = FALSE}
diamonds |> 
  ...()
```

```{r writing-csv-files-1-test, include = FALSE}
diamonds |> 
  group_nest()
```

### 

`group_nest()` is used to nest a tibble using a grouping specification. The output is a tibble with as many rows as there are values for the grouping variable. Because we did not specify a grouping variable, the output is a tibble with one row. 

### Exercise 2

Add the value `clarity` to the call to `group_nest()`.

```{r writing-csv-files-2, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r writing-csv-files-2-hint-1, eval = FALSE}
diamonds |> 
  group_nest(...)
```

```{r writing-csv-files-2-test, include = FALSE}
diamonds |> 
  group_nest(clarity)
```

### 

The first argument to `group_nest()`, `.key`, is the most important. The resulting tibble has two columns. The first is the same name as the grouping variable and includes the individual values. The second is always called `data`. It is a list column in which each element is a tibble with all the rows from `diamonds` which have the specified value of `clarity`.

### Exercise 3

Add a call to `mutate()` with argument `path = str_glue("diamonds-{clarity}.csv"` to the pipe.

```{r writing-csv-files-3, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r writing-csv-files-3-hint-1, eval = FALSE}
... |> 
  ...(path = str_glue("diamonds-{clarity}.csv"))
```

```{r writing-csv-files-3-test, include = FALSE}
diamonds |> 
  group_nest(clarity) |> 
  mutate(path = str_glue("diamonds-{clarity}.csv"))
```

### 

`str_glue()` is a glue function. You now have a 8 x 3 tibble, with a new column that gives the name of an output file.


### Exercise 4

Assign the result of this pipe to `by_clarity`, a new object.

```{r writing-csv-files-4, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r writing-csv-files-4-hint-1, eval = FALSE}
... <- 
  diamonds |> 
    group_nest(clarity) |> 
    mutate(path = str_glue("diamonds-{clarity}.csv"))
```

```{r writing-csv-files-4-test, include = FALSE}
by_clarity <- 
  diamonds |> 
    group_nest(clarity) |> 
    mutate(path = str_glue("diamonds-{clarity}.csv"))
```

Note how we first get some code which does what we want and then, once it is working, we assign the result to a new object. This is the typical way of doing data science because it is so useful, until the very end, for your code to just "spit out" whatever the current result is.

### Exercise 5

Type `by_clarity` and hit `Run Code`.
  
```{r writing-csv-files-5, exercise = TRUE}

```

```{r writing-csv-files-5-hint-1, eval = FALSE}
by_clarity
```

```{r writing-csv-files-5-test, include = FALSE}
by_clarity
```

### 

Our new tibble has eight rows and three columns.

### Exercise 6

Run the following code.

```{r writing-csv-files-6, exercise = TRUE}
by_clarity$data[[1]]
```

```{r writing-csv-files-6-test, include = FALSE}
by_clarity$data[[1]]
```

### 

The `$` pulls out the `data` column, which is a list column. The `[[1]]` pulls out the first element in that list column, which is a tibble. Remember that `clarity` is our grouping variable and `data` is a list-column containing one tibble for each unique value of clarity. 

### Exercise 7

If we were going to save these data frames by hand, we might write something like:

```
write_csv(by_clarity$data[[1]], by_clarity$path[[1]])
write_csv(by_clarity$data[[2]], by_clarity$path[[2]])
write_csv(by_clarity$data[[3]], by_clarity$path[[3]])
...
write_csv(by_clarity$by_clarity[[8]], by_clarity$path[[8]])
```

But that would be annoying, especially if there were hundreds of values for `clarity`. Type `?walk2` to the space below. hit `Run Code`.

```{r writing-csv-files-7, exercise = TRUE}

```

```{r writing-csv-files-7-hint-1, eval = FALSE}
?walk2
```

```{r writing-csv-files-7-test, include = FALSE}
?walk2
```

### 

This is a little different to our previous uses of `map()` because there are two arguments that are changing, not just one. 


### Exercise 8

Call `walk2()` with values `by_clarity$data`, `by_clarity$path`, and `write_csv`.

```{r writing-csv-files-8, exercise = TRUE}

```

```{r writing-csv-files-8-hint-1, eval = FALSE}
walk2(..., by_clarity$path, ...)
```

```{r writing-csv-files-8-test, include = FALSE}
walk2(by_clarity$data, by_clarity$path, write_csv)
```

### 

We needed a new function, `map2()`, which varies both the first and second arguments. Because we again don’t care about the output, we want `walk2()` rather than `map2()`.

### Exercise 9

Run `list.files()` with the `pattern` argument set to  `"*.csv"`. 

```{r writing-csv-files-9, exercise = TRUE}

```

```{r writing-csv-files-9-hint-1, eval = FALSE}
list.files(pattern = ...)
```

```{r writing-csv-files-9-test, include = FALSE}
list.files(pattern = "*.csv")
```

### 

This should produce the 8 CSV files which you just created. Make sure that it does not return any files which you care about since, in the next question, we will be deleting all these files.


### Exercise 10

Always clean up the files which you create, unless you want to keep them. Pipe `list.files(pattern = "*.csv")` to `file.remove()`.

```{r writing-csv-files-10, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r writing-csv-files-10-hint-1, eval = FALSE}
list.files(pattern = "*.csv") |> 
  ...()
```

```{r writing-csv-files-10-test, include = FALSE}
list.files(pattern = "*.csv") |> 
  file.remove()
```

### 

This removes all the files you created. It should return a logical vector of length 8, which each element equal to `TRUE`. This is what `file.remove()` returns if the removal is successful.


## Saving plots
### 

We can take the same basic approach in order to create many plots.

### Exercise 1

Below is a new function that draws the plot we want. Hit `Run Code`.

```{r saving-plots-1, exercise = TRUE}
carat_histogram <- function(df) {
  ggplot(df, aes(x = carat)) + geom_histogram(binwidth = 0.1)  
}
```

### 

This function will only work if the tibble which you pass in as the value for the `df` argument includes a column called `carot`. If it doesn't, you will get an error like "object 'carat' not found".


### Exercise 2

Now we can use `map()` to create a list of many plots and their eventual file paths. 

Pipe `by_clarity` into `mutate()`. Add arguments `plot = map(data, carat_histogram)` and `path = str_glue("clarity-{clarity}.png")` to the function.


```{r saving-plots-2, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r saving-plots-2-hint-1, eval = FALSE}
by_clarity |> 
  mutate(
    plot = map(data, carat_histogram),
    path = str_glue("clarity-{clarity}.png")
  )
```

### 

This step is similar to what we did earlier. 

### Exercise 3

Assign name `by_clarity` to the code above.

```{r saving-plots-3, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r saving-plots-3-hint-1, eval = FALSE}
by_clarity <- by_clarity |> 
  mutate(
    plot = map(data, carat_histogram),
    path = str_glue("clarity-{clarity}.png")
  )
```

### 

### Exercise 4

Type `walk2()` then add arguments `by_clarity$path` and `by_clarity$plot`.

```{r saving-plots-4, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r saving-plots-4-hint-1, eval = FALSE}
walk2(
  by_clarity$path,
  by_clarity$plot)
```

### 

### Exercise 5

Add `\(path, plot)` and `ggsave()` to the call to `walk2()`.

```{r saving-plots-5, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r saving-plots-5-hint-1, eval = FALSE}
walk2(
  by_clarity$path,
  by_clarity$plot,
  \(path, plot) ggsave())
```

### 

### Exercise 6

Add arguments `path, plot, width = 6, height = 6` to `ggsave()`.

```{r saving-plots-6, exercise = TRUE}

```

<button onclick = "transfer_code(this)">Copy previous code</button>

```{r saving-plots-6-hint-1, eval = FALSE}
walk2(
  by_clarity$path,
  by_clarity$plot,
  \(path, plot) ggsave(path, plot, width = 6, height = 6)
```

### 

Remember that `ggsave()` is a convenient function for saving a plot. It defaults to saving the last plot that you displayed, using the size of the current graphics device. It also guesses the type of graphics device from the extension.


## Summary
### 

This tutorial covered [Chapter 26: Iteration](https://r4ds.hadley.nz/iteration.html) from [*R for Data Science (2e)*](https://r4ds.hadley.nz/) by Hadley Wickham, Mine Çetinkaya-Rundel, and Garrett Grolemund. We used the [**purrr**](https://purrr.tidyverse.org/) package to learn tools for iteration, repeatedly performing the same action on different objects. Key functions included  [`across()`](https://dplyr.tidyverse.org/reference/across.html),  [`map()`](https://purrr.tidyverse.org/reference/map.html), and user-created anonymous functions.


```{r download-answers, child = system.file("child_documents/download_answers.Rmd", package = "tutorial.helpers")}
```
